### 2017-09-28 article/articleApplied.tex
first commit

\documentclass[twocolumn,a4paper,10pt]{article}

\input{../packageEN}
\input{../packageStyle}

\title{Estimation of the traffic sound level from measurements made in an urban environment}

\author{
    Jean-Rémy GLOAGUEN\\
    Arnaud Can\\
    LAE\\
    Ifsttar\\
    jean-remy.gloaguen@ifsttar.fr
  \and
    Mathieu Lagrange\\
	Jean-François Petiot \\
    LS2N\\
    Central School of Nantes\\
}
\date{}
\begin{document}

\maketitle

\section*{Abstract}

\section{Introduction}

The road traffic noise is one of the main annoyance perceived by the urban dweller. This boredom is not harmless and becomes a public problem because too long exposure to high sound levels can generate health problems  \cite{who_burden_2017}.  Accordingly, European directive 2002/EC/49 \cite{directive} has been developed to determine the impact of this noise on the population. For cities with a population of more than 100,000, it requires the production of traffic noise maps, which consist in calculating the sound level of traffic noise in the streets. These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and using the acoustic propagation models. The maps express the average sound levels over 24 hours, $L_ {DEN}$ (\textit{Day-Evening-Night}), and for the night over a period of 8h, $L_N$ (\textit{Night}). These maps make it possible to estimate the number of city dwellers impacted by the high noise levels and to draw up action plans to reduce them. However, these maps have been criticized because a lot of uncertainty can be generated because of the errors produced by the  numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. In addition, these maps offer restrictive information with only two calculated global sound levels that need to be updated every five years while road traffic varies throughout the day. Since the creation of the traffic noise maps is a long process, it is inconceivable to update them every hour, the use of measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed on fixed points spreading all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which allow to get the long-term evolution of the traffic noise levels, or can be performed with  mobile points \cite{can_exploring_2012} \cite{manvell2004sadmam}, which have the advantage of being able to cover a larger area with fewer sensors. Also, the clustering between mobile and fixed measurements have been studied in \cite{can_measurement_2014}. Currently, we are seeing the diffusion of sensor networks in cities for multiple applications (air quality assessment, measurement of meteorological parameters), including assessment of urban noise levels. Various projects exist to study the establishment and feasibility of such installations as the DYNAMAP \cite{dynamap_2016} project. It focuses on sensor installations on specific roads in all cities of Milan and Rome \cite{bellucci_life_2017}. The principle is, since the measurements made on these roads, to correct sound emissions on similar roads that are supposed to follow the same evolution than the measured ones. In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} plains to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modelling, by using data assimilation techniques. In addition to improving the traffic noise maps, the measurements will be used to lead perceptual studies on the urban soundscapes and to develop sound recognition tools. If these installations, from their measurements, allow to have a more precise evaluation of the traffic noise in cities than the simulated maps, the issue of the correct estimation from measurement of the traffic sound level is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds both from a temporal and a frequency point of view (car passages, voices, bird's whistles, car horn \dots) that can overlap. In consequence, the traffic sound level estimation is not trivial task. Our study proposed to applied a source separation tool, Non-Negative Matrix Factorization, on simulated sound mixtures to succeed on the estimation of the traffic sound level. Part \ref{part:stateOfArt} sums up on the different approaches in signal processing to isolate the traffic component of a monaural signal. Part \ref{part:nmf} details the technical aspect of the chosen method. Part \ref{part:protocol} is focused on the experimental protocol set up. Then part \ref{part:results} reveals the results obtained during the parametric study.

\section{State of the art in the detection and separation tasks for environmental sounds}\label{part:stateOfArt}

Many references have focused on the classification of the sound environments \cite{allegro_automatic_2001}, on the recognition \cite{defreville_automatic_2006} or one the detection of the  environmental sound events \cite{heittola_sound_2011}. This category of sound is not specific to the urban environment but includes all sound sources that are not related to music or speech. The algorithms are first learned on a known database to evaluate the performance of the algorithm then is applied to a second database simulating its application in real cases. To achieve these tasks, many approaches have the same generic pattern: a first extraction phase followed by a classification phase. In the first phase, the parameters of the characteristic are applied to an audio file to describe it, from the simplest, such as the sound level, the spectrum Gravity Centrums, the harmonicity to the most complex (Mel-Frequency Cepstral Coefficient, Perceptual Linear Prediction). In the second step, the characteristics are classified according to their evolution making it possible to distinguish the different categories of sounds. Common classification techniques are Hidden Markov Models, Artificial Neural Network, Gaussian Mixture Models and Support Vector Machines. Cowling and Sitte \cite{cowling_comparison_2003} summarize multiples of these extraction features and classification techniques that are used in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}. Even if each of them tested their algorithm on their own isolated sound database, specific database have been proposed to test and compare the performances of the algorithms as the DCASE database \cite{giannoulis_database_2013} or the TUT database \cite{mesaros_tut_2016}. Currently, the interest is focused on the use of the Neural Network in the case of the detection of acoustics events \cite{genaro_neural_2010} \cite{parascandolo_recurrent_2016}.

However, many of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context. Although in some places, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where there is little traffic and overlaps with other sound sources. Consequently, the only detection of the traffic component does not make it possible to determine precisely its noise level. In consequence, to effective on all the different sound environment, it seems more suited to consider the issue as the one of blind source separation. One of the first and the most widely used techniques is the Independent Component Analysis \cite{comon_independent_1994}. The principle is 
to decompose $N$ recorded signals in a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.  However, ICA is limited to only over determined cases ($N > P$). Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to install multiples sensors on one point to realize the source separation which is not feasible. In consequence, the use of Non-negative Matrix Factorization (NMF) is choosen has it allows to take into account easily the overlapping between the different sound sources and the time-frequency evolution of the sound sources by its spectrogram representation. This method has been widely used in the sound domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been empoyed for the source separation task dedicated for speech \cite{wilson_speech_2008} and musical content \cite{wang_musical_2005} on monaural signals.

\section{Non-negative Matrix Factorization}\label{part:nmf}
\subsection{Supervised NMF}
Non-negative Matrix Factorization (NMF) is an approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which estimates the spectrogram of an audio file, $\mathbf{V}$, with $F \times N $ dimension and which belong in $\mathbb{R}^+$ as :

\begin{equation}\label{eq:nmf}
\mathbf{V} = \mathbf{\tilde{V}} \approx \mathbf{WH}
\end{equation}

where $\mathbf{W}_{F \times K} \in \mathbb{R}^+$ is the \textit{basis} matrix composed of audio spectrum and $\mathbf{H}_{K \times N} \in \mathbb{R}^+$ is the \textit{activation} matrix which sum up the temporal evolution of each element of $\mathbf{W}$ (fig.  \ref{fig:example_NMF}).

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.9\linewidth]{../image/illustration_NMF.PNG}
\caption{Example of a simple NMF  for a musical content \cite{bertin_les_2009}}
\label{fig:example_NMF}
\end{figure}

The choice of the dimensions is done in a such way that $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used

\begin{equation}\label{eq:min-D-WH}
\min D\left(\mathbf{V} \vert \vert \mathbf{\tilde{V}}\right) \quad  \textrm{w.r.t.} \quad \mathbf{H} \geq 0, \mathbf{W} \geq 0
\end{equation}

The operator $D(x\vert y)$ is a divergence calculation such as 
\begin{equation}
D\left(\textbf{V} \vert\vert \mathbf{\tilde{V}} \right) = \sum_{f = 1}^{F} \sum_{n = 1}^{N} d_{\beta} 
\left(\textbf{V}_{fn} \vert \left[ \textbf{WH} \right]_{fn} \right)
\end{equation}

and usually belongs to the $\beta-$divergence class \cite{fevotte_nonnegative_2009} in which the well known Euclidean distance (eq. \ref{eq:def_distEUC}) and the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) belong

\begin{subequations}\label{eq:divBetaGenerale}
\begin{numcases}{d_{\beta}(x\vert y) =}
    \frac{1}{2}(x-y)^2, & $\beta = 2$, \label{eq:def_distEUC}\\
    x\log \dfrac{x}{y} - x + y, & $\beta = 1$.\label{eq:def_divKL}
\end{numcases}
\end{subequations}

The prior knowledges on the content can be adjust with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function \ref{eq:min-D-WH}.

NMF can be employed in an unsupervised case where nor $\mathbf{W}$ or $\mathbf{H}$ is known. The \textit{dictionary} is learned first during the training phase with different sound samples, the \textit{basis} got are not retained. $\mathbf{W}$ is afterwards used on the test phase where a new \textit{basis} is got. For the supervised case, $\mathbf{W}$ is already known. In the urban environments, a lot of different sound sources present in the cities are known and their spectrum can be obtained. 
Multiple methods exist to solve the minimization problem (\ref{eq:min-D-WH}) (Alternating Least Square Method \cite{cichocki_regularized_2007}, Projected Gradient \cite{lin_projected_2007} \dots). Here, the multiplicative update is the chosen method \cite{lee_algorithms_2000} as it assures to have a non-negative results where the convergence of the results have been proved \cite{fevotte_algorithms_2011}. The minimization problem (\ref{eq:min-D-WH}) is solved iteratively. As it is the supervised method that is consider first, only the $\mathbf{H}$ is updated. In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic  algorithm

\begin{equation}
\textbf{H}^{(i+1)} \leftarrow \textbf{H}^{(i)}.\left(\frac{\textbf{W}^T \left[\left(\textbf{WH}^{(i)} \right)^{(\beta-2)}.\textbf{V} \right]}{\textbf{W}^T \left[\textbf{WH}^{(i)} \right]^{(\beta-1)}}\right)^{\gamma(\beta)}
\end{equation}

with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$. The product $A.B$ and $A/B$ symbolized the Hadamard product.

The source separation is made after $I$ iterations by generating a filtering mask, $M_s$,

\begin{equation} \label{eq:definitionMask}
\mathbf{M_s} = \frac{\left[\mathbf{W} \mathbf{H}\right]_{traffic}}{\sum_{k = 1}^K\left[\mathbf{W H}\right]}, 
\end{equation}

which allows to determine the spectrogram of the traffic component, $ \mathbf{\tilde{V}}_{traffic}$

\begin{equation}\label{eq:separation}
\mathbf{\tilde{V}}_{traffic} = \mathbf{M_s}.\mathbf{V}.
\end{equation}

For the environmental sounds, the method has been used for geo-localisation and classification of the sound environment like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded. For the source separation, it has been used by Innami and Kasai in the unsupervised case. They proposed a source separation in two step  by separating the sound background from the events first and by separating the events between them. They used the MFCC features to composed the dictionary by adding a time variant gain features (a calculation of variance and a ratio to nearly zero values on the \textit{basis} vectors). The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing). If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.

\subsection{NMF Semi-supervised}

One of the main issue with the supervised approach is this is not always adapted to different sound environments. Because in the supervised case, the fixed dictionary has to sum up all the sound sources present in the different sound environment. In consequence, as the dictionary dimension is limited and cannot include all the urban sounds, to offer more flexibility, semi-supervised NMF has been proposed \cite{lee_semi-supervised_2010}. This method consists in composing the dictionary $\mathbf{W}_{F \times K+J}$ with a fixed part $\mathbf{W_s}_{F \times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r}_{F \times J}$ that is updated. The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic. The problem (\ref{eq:nmf}) become

\begin{equation}
\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}
\end{equation}

In a similar way as to solve the equation \ref{eq:min-D-WH}, $\mathbf{W_r}$, $\mathbf{H_r}$ and $\mathbf{H_s}$ are successively updated with the relations (\ref{eq:WH-SSupdate}):

{\scriptsize
\begin{subequations}\label{eq:WH-SSupdate}
\begin{align}
\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}.\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}\label{eq:W_r_SS}\\
\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}.\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}\label{eq:H_r_SS}\\
\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}.\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}\label{eq:H_s_SS}
\end{align}
\end{subequations}}

This approach has been used TO COMPLETE.

\section{Experimental protocol}\label{part:protocol}

NMF are not tested on urban sound recordings: the traffic component is here unknown. Therefore, the estimation of the traffic sound level cannot be compare. In consequence, NMF is applied on simulated sound mixtures as it offers a controlled framework where the contribution of each sound source (and especially the traffic) is known and offers an exact solution to compare with the estimated one.

\subsection{Environmental sound scene corpus}

A first corpus is design with the \textit{simScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}. \textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a .wav format by superposing audio samples that come from an isolated sound database. This database is divided in two categories: the \textit{event} category  which are the brief sounds (from 1 to 20 seconds) that are considered as salient whereas all the sounds that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category. Inside each category, the sound samples are grouped in sound classes (\textit{bird, car, foot steps} \dots), each of them being composed of multiples samples (bird01.wav, bird02.wav \dots).

The software allows the user to control some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the presence of a fade in and a fade out \dots. Each parameter is completed with a standard deviation that may brings some random behavior between the scenes. Furthermore, with the global scene mixture, an audio for each sound class presented in the scene can be generated that allow to know its exact contribution and a text file is created that sum up the time presence of all the event.\\

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{../image/exampleSimScene.pdf} 
\label{fig:exampleSimScene}
\caption{Example of a sound mixture composed of 3 sound classes (car, bird, car horn)}
\end{figure}

With this software, a sound database has been built including 245 sound event samples (with the sound classes \textit{ringing bell, birds, sweeping bropom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and 154 sound background samples (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffci noise, ventilation, wind}). The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway  on different speeds with multiple gear ratio \cite{gloaguen_creation_2017}. The other audio files are sampled at 44.1 kHz and have been found online (\textit{freesound.org}) and with the help of the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}. A sound mixture corpus is designed composed of 6 sub-corpus of 25 30 seconds audio files ($N = 25$). Each sub-corpus is characterized by a specific generic sound class : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noise) and \textit{transportation} (train, tramway and plane). In each file, traffic component is present as sound background and event and is mixed with on or multiple sound classes. The sound classes that are not related to the traffic component are summed up as the \textit{perturbator} (n'existe pas comme terme, disruptive ?) class. To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound mixture, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{perturbator} class, $L_{p,perturbator}$ following the relation (\ref{eq:tpr}).

\begin{equation}\label{eq:tpr}
TPR = L_{p,traffic}-L_{p,perturbator}
\end{equation}

with the \textit{Traffic Perturbator Ratio} $TPR = \left[-12, -6, 0, 6, 12\right]$. When $TPR = -12$, the traffic component is then less present than when $TPR = 12$ where it is predominant on the \textit{perturbator} class. The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is too calculated (figure \ref{fig:exampleScene}). Finally, the number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TPR values).

\begin{figure*}
\centering
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/spectrogramExample.pdf} 
   \end{minipage} 
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/animals_10-pianoRoll.png} 
   \end{minipage} 
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/evolutionLpExample.pdf}

   \end{minipage}
\caption{Example of a scene of the \textit{animals} sub corpus. Spectrogram (on left), \textit{Piano Roll} of the different sound classes (on the middle) and 1-s equivalent sound level of the traffic, $Lp_{1s,traffic}$ and of the global sound mixture, $Lp_{1s,global}$(on right)}
\label{fig:exampleScene}
\end{figure*}

\subsection{Experiment}

The experiment consists in applied NMF on the environmental sound corpus to estimate the sound level of the traffic road. This method is compared to a simple approach which is a low-frequency filter. This method considers, as the traffic component is mostly composed of low-frequencies, that the sound level can be calculated from the band-pass filter. Similarly, the low-pass filter is applied with NMF with the aim to focus the reconstruction of the signal on the frequencies where the traffic components are. The figure \ref{fig:block_diagram_protocol} sum up the different step of the process depending on the chosen method.\\

\begin{figure*}[!t]
\centering
	\begin{minipage}[t]{.48\textwidth}
		\centering
		\includegraphics[width=.9\linewidth]{../image/bloc_diagram_filtrage_EN.pdf} 
	\end{minipage}
	\begin{minipage}[t]{.48\textwidth}
		\centering
		\includegraphics[width=.9\linewidth]{../image/bloc_diagram_NMF_EN.pdf} 
	\end{minipage}
	\caption{Block diagrams summed up the different step of the precess. On top, for the low-pass filter, on bottom, for NMF.}
	\label{fig:block_diagram_protocol}
\end{figure*}

3 calculation steps are set up : 
\begin{itemize}
\item the dictionary building for NMF
\item the estimation of the traffic sound level according the chosen method (NMF or low-pass filter)
\item calculation of the error between the exact and the estimated sound levels.\\
\end{itemize}

\subsubsection{Dictionary building}

The dictionary is built from a sound database dedicated specially to this task. It is composed of 53 audio files of car passages. Multiple version of dictionaries are tested. First, for each audio file, its spectrogram is calculated with fixed parameters ($w = 2^{14}$, $noverlap = 75 \%$, $nfft = 2^{14}$). Then a temporal rectangular window of 0.5 second is applied without overlapping on the spectrogram. In each window, the root mean square value are calculated. This window allow to get multiples spectrum from an audio file.  An example, to illustrate the process, can be found on figure with a large window of 4 seconds.

\begin{figure}[!t]
\centering
	\begin{minipage}[t]{.24\textwidth}
		\centering
		\includegraphics[width=.9\linewidth]{../image/extractionDictionary1Example.pdf} 
	\end{minipage}
	\begin{minipage}[t]{.24\textwidth}
		\centering
		\includegraphics[width=.9\linewidth]{../image/extractionDictionary2Example.pdf} 
	\end{minipage}
	\caption{Example of the extraction of the spectrum with a 4 seconds window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
	\label{fig:dictionaryExtraction}
\end{figure}

As the number of spectrum got on all the sound database does not respect the constraint imposed about the dimension of NMF, a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$. A special case is added where the root mean square of the spectrum is applied on all the spectrogram. In consequence, the audio file is summed up in one spectra. 
% It consists in the estimation of the $K$-means and then to calculate the euclidean distance between each centroid with all the spectrum. The spectrum having the smaller distance with the centroid is then included in the dictionary $\mathbf{W}$.

\subsubsection{Estimation of the traffic sound level}
\subsubsection{Low-pass filter}
The first estimator to determine the traffic sound level is a basic low-pass filter with the cut-off frequencies $f_c = \left[ 500, 1k, 2k, 5k, 10k, 20k\right]$ Hz. As the road traffic is mainly composed of low frequencies (< 5000 Hz), this method considers that all the components located on the pass-band can be considered as road traffic. 
\subsubsection{NMF}
The second estimator is NMF with the supervised and the semi-supervised approaches (see part \ref{part:nmf}). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands. Theses two methods enables to compare a fine approach (the linear scale) with a degraded one (the third octave bands) as it reduces the number of frequency bins (27 bands) and allows to reduces the number of bands in the high frequencies where the traffic component is less present. Furthermore, in the case of the linear frequency scale, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins. The $\beta$-divergence is reduced to the Kullback-Leibler divergence ($\beta = 1$) and to the Euclidean distance ($\beta = 2$). 400 iterations are performed. 
The source separation is performed following equations (\ref{eq:definitionMask}) and (\ref{eq:separation}) according two ways: based on the filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) or on the entire spectrogram and dictionary ($\mathbf{V}$ and $\mathbf{W}$). Then for each spectrogram, the estimated equivalent traffic sound level in dB of the entire scene, $\tilde{L}_{p,traffic}$, is calculated as well as the 1 second equivalent sound pressure level, $\tilde{p}_{1s,traffic}$.

\subsubsection{Metric}
The performances of the two estimators to estimate the sound level of the traffic correctly are estimated through the calculation of two metrics based on the Root Mean Square Error (RMSE) :

\begin{itemize}
\item the global RMSE, \textit{gRMSE}, which consists in calculate for each combination of factors the error between the exact and estimated global traffic sound level in dB of the $N$ scenes.

\begin{equation}
gRMSE = \sqrt{\frac{1}{N}\sum_{n = 1}^N \left(L^n_{p,traffic}-\tilde{L}^n_{p,traffic}\right)^2}
\end{equation}

This metric expresses the quality of the long-term reconstruction of the signal. 
\item The normalized short-term RMSE, $nRMSE$, consists in calculate, for each sound mixture, the error between the 1-s equivalent sound level exact and estimated of the traffic of each file normalized by the 1-s equivalent sound level of the global mixture in the linear scale,

\begin{equation}
nRMSE = \sqrt{\frac{1}{T}\sum_{t = 1}^T \left(\frac{p^t_{1s,traffic}-\tilde{p}^t_{1s,traffic}}{p^t_{1s,global}}\right)^2}
\end{equation}

with $T$ is the number of temporal bin in the signal. The linear scale is here more relevant than the dB scale as it is more sensitive to the error on the high sound levels. Then for one combination of factors, the $N$ $nRMSE$ calculated are averaged.\\
\end{itemize}
The calculation are performed in 3 steps. The first set up the dictionary with two modalities : $K$ value (25 50 100) and description of the audio files (in one fft or with a temporal window of 0.5 second). The second step include  modalities : the sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)), the estimator (low-pass filter, NMF), the cut-off frequency $f_c$ ([500 1k 2k 5k 10k 20k] Hz), the $TPR$ ([-12 -6 0 6 12] dB), $\beta = [1~2]$, the approach NMF  (supervised or semi-supervised) and the frequency scale (linear or third octave). In all, 5940 settings are performed.

\section{Results}\label{part:results}

In a first time, the low-pass filter estimator is performed on all the scene. The table \ref{tab:results_filter} sums up the mean error on the totality of the 750 scenes to find the cut-off frequency the most efficient on all the cases.

\begin{table}[h]
\centering
\begin{tabular}{lll}
$f_c$ (Hz) & $nRMSE$ & $gRMSE$ (dB) \\ \hline
\textbf{500}  &\textbf{1.48 $\pm$ 1.10}&\textbf{5.96 $\pm$ 6.70}\\ \hline
1k &  1.52 $\pm$ 1.34 & 6.52 $\pm$ 7.73\\ \hline
2k & 1.57 $\pm$ 1.54 & 6.80 $\pm$ 8.32\\ \hline
5k & 1.82 $\pm$ 1.65 & 7.44 $\pm$ 8.89       \\ \hline
10k  & 1.89 $\pm$ 1.70 & 7.56 $\pm$ 8.99       \\ \hline
20k  & 1.89 $\pm$ 1.72 & 7.59 $\pm$ 9.02      
\end{tabular}
\caption{RMSE error for the low pass filter averaged on all the TPR and sub-classes}
\label{tab:results_filter}
\end{table}

According the table, the cut-off frequency at 500 Hz is the most efficient on all the TPR and all the sub-classes. The error for a 500 Hz cut-off frequency according to the sub-classes and the TPR is sums up in the figure \ref{fig:filterAmbiance}.

\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{../image/filterAmbianceBar.pdf}
\caption{Bar plot for the low-pass filter according to the sub-classes and the TPR at 500 Hz cut-off frequency}
\label{fig:filterAmbiance}
\end{figure}

The error for all the sub-corpus is important for the low TPR (-12 and -6) due to the confusion of the \textit{perturbator} class as the traffic component whereas  it is not. The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while for the \textit{climate} and \textit{transportation} the error is far more important. For the other TPR, the traffic is more present and become dominating on the \textit{perturbator} class. Then the error is due to the suppression of the traffic energy by the low-pass filter.

The first results averaged on all the TPR and the sub-classes for the supervised NMF is sums up on table \ref{tab:}, the frequency domain is, for now, limited to \textit{spectra}

\begin{table}[h]
\centering
\begin{tabular}{lll}

\end{tabular}
\caption{RMSE error for the low pass filter averaged on all the TPR and sub-classes}
\label{tab:results_filter}
\end{table}

Résultat global pour le NMF supervisée en tableau puis en bar pour chaque ambiance et chaque TPR
Tracer l'évolution 1s du signal sonore pour 2 ambiances et 2 TPR
résultat semi-supervisée
beta = 2, = 1
%

\section{Conclusion}
\footnotesize
\bibliographystyle{unsrt}
\bibliography{bibliographie_applied}

\end{document}

### 2017-10-04 article/articleApplied.tex
redaction pursuit + add images

@@
 
-The road traffic noise is one of the main annoyance perceived by the urban dweller.
-This boredom is not harmless and becomes a public problem because too long exposure to high sound levels can generate health problems  \cite{who_burden_2017}.
-Accordingly, European directive 2002/EC/49 \cite{directive} has been developed to determine the impact of this noise on the population.
-For cities with a population of more than 100,000, it requires the production of traffic noise maps, which consist in calculating the sound level of traffic noise in the streets.
-These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and using the acoustic propagation models.
-The maps express the average sound levels over 24 hours, $L_ {DEN}$ (\textit{Day-Evening-Night}), and for the night over a period of 8h, $L_N$ (\textit{Night}).
-These maps make it possible to estimate the number of city dwellers impacted by the high noise levels and to draw up action plans to reduce them.
-However, these maps have been criticized because a lot of uncertainty can be generated because of the errors produced by the  numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
-In addition, these maps offer restrictive information with only two calculated global sound levels that need to be updated every five years while road traffic varies throughout the day.
-Since the creation of the traffic noise maps is a long process, it is inconceivable to update them every hour, the use of measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
-These measurements can be performed on fixed points spreading all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which allow to get the long-term evolution of the traffic noise levels, or can be performed with  mobile points \cite{can_exploring_2012} \cite{manvell2004sadmam}, which have the advantage of being able to cover a larger area with fewer sensors.
-Also, the clustering between mobile and fixed measurements have been studied in \cite{can_measurement_2014}.
+Road traffic noise is one of the main annoyance perceived by urban dwellers and a major societal issue because too long exposures to high sound levels can generate health problems  \cite{who_burden_2017}.
+Accordingly, the European Directive 2002/EC/49 \cite{directive} has been developed to determine the exposuer of the populations.
+Cities with a population of more than 100 000 must produce road traffic noise maps.
+These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and calculating sound propagation.
+The maps express $L_ {DEN}$ and $L_N$, respectively, \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels.
+These maps make it possible to estimate the number of city dwellers exposed to highly annoyant noise levels and to draw up action plans to reduce them.
+However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
+In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day.
+Since the creation of road traffic noise maps entails long data collection and calculation times, the use of measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
+These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which allow to get the long-term evolution of the traffic noise levels, or can be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time perdiods.
+The clustering between mobile and fixed measurements have been studied in \cite{can_measurement_2014}.
+
 Currently, we are seeing the diffusion of sensor networks in cities for multiple applications (air quality assessment, measurement of meteorological parameters), including assessment of urban noise levels.
@@
 It focuses on sensor installations on specific roads in all cities of Milan and Rome \cite{bellucci_life_2017}.
-The principle is, since the measurements made on these roads, to correct sound emissions on similar roads that are supposed to follow the same evolution than the measured ones.
-In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} plains to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modelling, by using data assimilation techniques.
-In addition to improving the traffic noise maps, the measurements will be used to lead perceptual studies on the urban soundscapes and to develop sound recognition tools.
+In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} plains to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, by using data assimilation techniques.
+
 If these installations, from their measurements, allow to have a more precise evaluation of the traffic noise in cities than the simulated maps, the issue of the correct estimation from measurement of the traffic sound level is still unsolved \cite{Mioduszewski}.
@@
 In consequence, the traffic sound level estimation is not trivial task.
-Our study proposed to applied a source separation tool, Non-Negative Matrix Factorization, on simulated sound mixtures to succeed on the estimation of the traffic sound level.
-Part \ref{part:stateOfArt} sums up on the different approaches in signal processing to isolate the traffic component of a monaural signal.
-Part \ref{part:nmf} details the technical aspect of the chosen method.
+
+Many works have been focused on the detection  \cite{heittola_sound_2011} or on recognition \cite{defreville_automatic_2006} task of environmental sounds.
+These methods use a two step process : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classified them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
+A description of there features and classifiers can be found on \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+However, many of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
+Although in some places, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where there is little traffic and overlaps with other sound sources.
+The only detection of the traffic component does not make it possible to determine precisely its noise level.
+In consequence, to be effective on all the different sound environments, it seems more suited to consider the issue as the one of blind source separation.
+
+One of the first and the most widely used techniques is the Independent Component Analysis \cite{comon_independent_1994}.
+The principle is to decompose $N$ recorded signals in a sum of $P$ independent sound sources weighted by linear relations.
+This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.
+However, ICA is limited to only over determined cases ($N > P$).
+Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to install multiples sensors on one point to realize the source separation which is not feasible.
+A second method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximate the non-negative spectrogram of an audio file from the product of two matrices.
+It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of a monaural signal included speech and musical contents  \cite{wang_musical_2005} \cite{wilson_speech_2008}.
+This method has the advantage to easily deals with the overlap of the sound sources.
+For the environmental sounds, the method has been used for geo-localisation and classification of the sound environment like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded.
+For the source separation, it has been used by Innami and Kasai in the unsupervised case.
+They proposed a source separation in two step  by separating the sound background from the events first and by separating the events between them.
+The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
+If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
+
+Our study proposed to applied Non-Negative Matrix Factorization on simulated sound mixtures to succeed on the estimation of the traffic sound level.
+The use of simulated sound mixtures is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component what recordings do not allow.
+Part \ref{part:nmf} details the technical aspect of NMF.
 Part \ref{part:protocol} is focused on the experimental protocol set up.
@@
 
-\section{State of the art in the detection and separation tasks for environmental sounds}\label{part:stateOfArt}
-
-Many references have focused on the classification of the sound environments \cite{allegro_automatic_2001}, on the recognition \cite{defreville_automatic_2006} or one the detection of the  environmental sound events \cite{heittola_sound_2011}.
-This category of sound is not specific to the urban environment but includes all sound sources that are not related to music or speech.
-The algorithms are first learned on a known database to evaluate the performance of the algorithm then is applied to a second database simulating its application in real cases.
-To achieve these tasks, many approaches have the same generic pattern: a first extraction phase followed by a classification phase.
-In the first phase, the parameters of the characteristic are applied to an audio file to describe it, from the simplest, such as the sound level, the spectrum Gravity Centrums, the harmonicity to the most complex (Mel-Frequency Cepstral Coefficient, Perceptual Linear Prediction).
-In the second step, the characteristics are classified according to their evolution making it possible to distinguish the different categories of sounds.
-Common classification techniques are Hidden Markov Models, Artificial Neural Network, Gaussian Mixture Models and Support Vector Machines.
-Cowling and Sitte \cite{cowling_comparison_2003} summarize multiples of these extraction features and classification techniques that are used in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
-Even if each of them tested their algorithm on their own isolated sound database, specific database have been proposed to test and compare the performances of the algorithms as the DCASE database \cite{giannoulis_database_2013} or the TUT database \cite{mesaros_tut_2016}.
-Currently, the interest is focused on the use of the Neural Network in the case of the detection of acoustics events \cite{genaro_neural_2010} \cite{parascandolo_recurrent_2016}.
-
-However, many of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
-Although in some places, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where there is little traffic and overlaps with other sound sources.
-Consequently, the only detection of the traffic component does not make it possible to determine precisely its noise level.
-In consequence, to effective on all the different sound environment, it seems more suited to consider the issue as the one of blind source separation.
-One of the first and the most widely used techniques is the Independent Component Analysis \cite{comon_independent_1994}.
-The principle is
-to decompose $N$ recorded signals in a sum of $P$ independent sound sources weighted by linear relations.
-This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.
-However, ICA is limited to only over determined cases ($N > P$).
-Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to install multiples sensors on one point to realize the source separation which is not feasible.
-In consequence, the use of Non-negative Matrix Factorization (NMF) is choosen has it allows to take into account easily the overlapping between the different sound sources and the time-frequency evolution of the sound sources by its spectrogram representation.
-This method has been widely used in the sound domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been empoyed for the source separation task dedicated for speech \cite{wilson_speech_2008} and musical content \cite{wang_musical_2005} on monaural signals.
-
 \section{Non-negative Matrix Factorization}\label{part:nmf}
 \subsection{Supervised NMF}
-Non-negative Matrix Factorization (NMF) is an approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which estimates the spectrogram of an audio file, $\mathbf{V}$, with $F \times N $ dimension and which belong in $\mathbb{R}^+$ as :
+Non-negative Matrix Factorization (NMF) is an approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which estimates the spectrogram (get from a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
 
@@
 
-where $\mathbf{W}_{F \times K} \in \mathbb{R}^+$ is the \textit{basis} matrix composed of audio spectrum and $\mathbf{H}_{K \times N} \in \mathbb{R}^+$ is the \textit{activation} matrix which sum up the temporal evolution of each element of $\mathbf{W}$ (fig.
+where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{basis} matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$ (fig.
 \ref{fig:example_NMF}).
@@
 \begin{equation}\label{eq:min-D-WH}
-\min D\left(\mathbf{V} \vert \vert \mathbf{\tilde{V}}\right) \quad  \textrm{w.r.t.} \quad \mathbf{H} \geq 0, \mathbf{W} \geq 0
+\underset{\mathbf{H} \geq 0, \mathbf{W} \geq 0}{\min} D\left(\mathbf{V} \vert \vert \mathbf{\tilde{V}}\right)
 \end{equation}
@@
 
-The prior knowledges on the content can be adjust with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function \ref{eq:min-D-WH}.
+The prior knowledges on the content can be adjust with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})).
 
-NMF can be employed in an unsupervised case where nor $\mathbf{W}$ or $\mathbf{H}$ is known.
-The \textit{dictionary} is learned first during the training phase with different sound samples, the \textit{basis} got are not retained.
-$\mathbf{W}$ is afterwards used on the test phase where a new \textit{basis} is got.
-For the supervised case, $\mathbf{W}$ is already known.
-In the urban environments, a lot of different sound sources present in the cities are known and their spectrum can be obtained.
 Multiple methods exist to solve the minimization problem (\ref{eq:min-D-WH}) (Alternating Least Square Method \cite{cichocki_regularized_2007}, Projected Gradient \cite{lin_projected_2007} \dots).
@@
 The minimization problem (\ref{eq:min-D-WH}) is solved iteratively.
-As it is the supervised method that is consider first, only the $\mathbf{H}$ is updated.
+NMF is employed here in the supervised case: the \textit{dictionary} is already known as in the urban environments, a lot of different sound sources present are known and their spectrum can be obtained.
+The \textit{basis} are the unknown to find.
 In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic  algorithm
@@
 with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$.
-The product $A.B$ and $A/B$ symbolized the Hadamard product.
+The product $A.B$ and $A/B$ symbolized the Hadamard product and ratio.
+The source separation is made after $I$ iterations either by extracting the dictionary and basis elements related to the traffic
 
-The source separation is made after $I$ iterations by generating a filtering mask, $M_s$,
-
-\begin{equation} \label{eq:definitionMask}
-\mathbf{M_s} = \frac{\left[\mathbf{W} \mathbf{H}\right]_{traffic}}{\sum_{k = 1}^K\left[\mathbf{W H}\right]}, 
+\begin{equation}\label{eq:separationExtraction}
+\mathbf{\tilde{V}}_{traffic} = \left[ \mathbf{WH} \right]_{traffic}
 \end{equation}
 
-which allows to determine the spectrogram of the traffic component, $ \mathbf{\tilde{V}}_{traffic}$
+%or by generating a soft mask, $\mathbf{M_s}$, 
+% 
+%\begin{equation} \label{eq:definitionMask}
+%\mathbf{M_s} = \frac{\left[\mathbf{W} \mathbf{H}\right]_{traffic}}{\mathbf{W H}}, 
+%\end{equation}
+%
+%which allows to determine the spectrogram of the traffic component, $ \mathbf{\tilde{V}}_{traffic}$
+%
+%\begin{equation}\label{eq:separation}
+%\mathbf{\tilde{V}}_{traffic} = \mathbf{M_s}.\mathbf{V}.
+%\end{equation}
 
-\begin{equation}\label{eq:separation}
-\mathbf{\tilde{V}}_{traffic} = \mathbf{M_s}.\mathbf{V}.
-\end{equation}
-
-For the environmental sounds, the method has been used for geo-localisation and classification of the sound environment like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded.
-For the source separation, it has been used by Innami and Kasai in the unsupervised case.
-They proposed a source separation in two step  by separating the sound background from the events first and by separating the events between them.
-They used the MFCC features to composed the dictionary by adding a time variant gain features (a calculation of variance and a ratio to nearly zero values on the \textit{basis} vectors).
-The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
-If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
-
-\subsection{NMF Semi-supervised}
+\subsection{Semi-supervised NMF}
 
@@
 In consequence, as the dictionary dimension is limited and cannot include all the urban sounds, to offer more flexibility, semi-supervised NMF has been proposed \cite{lee_semi-supervised_2010}.
-This method consists in composing the dictionary $\mathbf{W}_{F \times K+J}$ with a fixed part $\mathbf{W_s}_{F \times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r}_{F \times J}$ that is updated.
+This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ that is updated.
 The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic.
@@
 
-NMF are not tested on urban sound recordings: the traffic component is here unknown.
-Therefore, the estimation of the traffic sound level cannot be compare.
-In consequence, NMF is applied on simulated sound mixtures as it offers a controlled framework where the contribution of each sound source (and especially the traffic) is known and offers an exact solution to compare with the estimated one.
+To assess the performance of NMF, simulated sound mixtures are used as it offers a controlled framework to design specific sound environments where all the traffic component is known.
+The estimation of the traffic sound level can be compare to an exact solution.
 
@@
 Each parameter is completed with a standard deviation that may brings some random behavior between the scenes.
-Furthermore, with the global scene mixture, an audio for each sound class presented in the scene can be generated that allow to know its exact contribution and a text file is created that sum up the time presence of all the event.\\
+Furthermore, with the global scene mixture, an audio for each sound class presented in the scene can be generated that allow to know its exact contribution and a text file is created that summarizes the time presence of all the events.\\
 
@@
 \centering
-\includegraphics[width=\linewidth]{../image/exampleSimScene.pdf} 
+\includegraphics[width=\linewidth]{../image/exampleSimScene2.pdf} 
 \label{fig:exampleSimScene}
@@
 The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway on different speeds with multiple gear ratio \cite{gloaguen_creation_2017}.
-The other audio files are sampled at 44.1 kHz and have been found online (\textit{freesound.org}) and with the help of the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
+The other audio files have been found online (\textit{freesound.org}) and with the help of the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
 A sound mixture corpus is designed composed of 6 sub-corpus of 25 30 seconds audio files ($N = 25$).
@@
 In each file, traffic component is present as sound background and event and is mixed with on or multiple sound classes.
-The sound classes that are not related to the traffic component are summed up as the \textit{perturbator} (n'existe pas comme terme, disruptive ?) class.
+The sound classes that are not related to the traffic component are summed up as the \textit{perturbator} (n'existe pas comme terme, disruptive à la place ?) class.
 To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound mixture, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{perturbator} class, $L_{p,perturbator}$ following the relation (\ref{eq:tpr}).
@@
 This method considers, as the traffic component is mostly composed of low-frequencies, that the sound level can be calculated from the band-pass filter.
-Similarly, the low-pass filter is applied with NMF with the aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
-The figure \ref{fig:block_diagram_protocol} sum up the different step of the process depending on the chosen method.\\
+Similarly, for NMF, the low-pass filter is applied on $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ with the aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
+The figure \ref{fig:block_diagram_protocol} summarizes the different steps of the process depending on the chosen method.\\
 
@@
 		\centering
-		\includegraphics[width=.9\linewidth]{../image/extractionDictionary1Example.pdf} 
+		\includegraphics[width=\linewidth]{../image/extractionDictionary1Example.pdf} 
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=.9\linewidth]{../image/extractionDictionary2Example.pdf} 
+		\includegraphics[width=\linewidth]{../image/extractionDictionary2Example.pdf} 
 	\end{minipage}
@@
 As the number of spectrum got on all the sound database does not respect the constraint imposed about the dimension of NMF, a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
-A special case is added where the root mean square of the spectrum is applied on all the spectrogram.
-In consequence, the audio file is summed up in one spectra.
+A special case is added where the root mean square of the entire spectrum is applied on all the spectrogram of the audio file.
+In consequence, one spectra is generated by audio file.
 % It consists in the estimation of the $K$-means and then to calculate the euclidean distance between each centroid with all the spectrum.
@@
 \subsubsection{Estimation of the traffic sound level}
-\subsubsection{Low-pass filter}
-The first estimator to determine the traffic sound level is a basic low-pass filter with the cut-off frequencies $f_c = \left[ 500, 1k, 2k, 5k, 10k, 20k\right]$ Hz.
+The traffic sound level is estimated according to the choice of the estimator (low-pass filter or NMF).
+These methods are applied on the totality of the corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)), for all the cut-off frequencies $f_c$ ([500 1k 2k 5k 10k 20k] Hz) and for 5 $TPR$ ([-12 -6 0 6 12] dB).
+
+\paragraph{Low-pass filter}\mbox{} \\
+The first estimator to determine the traffic sound level is a basic low-pass filter which depend only on the cut-off frequencies $f_c$.
 As the road traffic is mainly composed of low frequencies (< 5000 Hz), this method considers that all the components located on the pass-band can be considered as road traffic.
-\subsubsection{NMF}
+
+\paragraph{NMF}\mbox{} \\
 The second estimator is NMF with the supervised and the semi-supervised approaches (see part \ref{part:nmf}).
@@
 400 iterations are performed.
-The source separation is performed following equations (\ref{eq:definitionMask}) and (\ref{eq:separation}) according two ways: based on the filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) or on the entire spectrogram and dictionary ($\mathbf{V}$ and $\mathbf{W}$).
+As only one sound source is considered in the dictionary, the source separation is performed following the equation  (\ref{eq:separationExtraction}).
+Also, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is made with the original dictionary $\mathbf{W}$, as
+
+\begin{equation}
+\mathbf{\tilde{V}}_{traffic} = \left[\mathbf{WH}_{f_c}\right]_{traffic}
+\end{equation}
+
 Then for each spectrogram, the estimated equivalent traffic sound level in dB of the entire scene, $\tilde{L}_{p,traffic}$, is calculated as well as the 1 second equivalent sound pressure level, $\tilde{p}_{1s,traffic}$.
@@
 \begin{itemize}
+
+\item the Mean Absolute Error, $MAE$, which consists in calculate for each combination of factors the absolute difference between the exact and estimated global traffic sound level in dB of the $N$ scenes.
+
+\begin{equation}
+MAE = \frac{\sum_{n = 1}^N\vert L^n_{p,traffic}-\tilde{L}^n_{p,traffic} \vert}{N}
+\end{equation}
+
+This metric expresses the quality of the long-term reconstruction of the signal.
+
 \item the global RMSE, \textit{gRMSE}, which consists in calculate for each combination of factors the error between the exact and estimated global traffic sound level in dB of the $N$ scenes.
@@
 \end{itemize}
-The calculation are performed in 3 steps.
-The first set up the dictionary with two modalities : $K$ value (25 50 100) and description of the audio files (in one fft or with a temporal window of 0.5 second).
-The second step include  modalities : the sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)), the estimator (low-pass filter, NMF), the cut-off frequency $f_c$ ([500 1k 2k 5k 10k 20k] Hz), the $TPR$ ([-12 -6 0 6 12] dB), $\beta = [1~2]$, the approach NMF  (supervised or semi-supervised) and the frequency scale (linear or third octave).
+
 In all, 5940 settings are performed.
@@
 
+\subsection{Low-pass filter results}
 In a first time, the low-pass filter estimator is performed on all the scene.
 The table \ref{tab:results_filter} sums up the mean error on the totality of the 750 scenes to find the cut-off frequency the most efficient on all the cases.
+\\
 
@@
 \centering
-\begin{tabular}{lll}
-$f_c$ (Hz) & $nRMSE$ & $gRMSE$ (dB) \\ \hline
-\textbf{500}  &\textbf{1.48 $\pm$ 1.10}&\textbf{5.96 $\pm$ 6.70}\\ \hline
-1k &  1.52 $\pm$ 1.34 & 6.52 $\pm$ 7.73\\ \hline
-2k & 1.57 $\pm$ 1.54 & 6.80 $\pm$ 8.32\\ \hline
-5k & 1.82 $\pm$ 1.65 & 7.44 $\pm$ 8.89       \\ \hline
-10k  & 1.89 $\pm$ 1.70 & 7.56 $\pm$ 8.99       \\ \hline
-20k  & 1.89 $\pm$ 1.72 & 7.59 $\pm$ 9.02      
+\begin{tabular}{llll}
+$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB) \\ \hline
+ 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} & \textbf{\textcolor{red}{5.96 $\pm$6.70}} \\ 
+ 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} & \textbf{6.52 $\pm$7.73} \\ 
+ 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} & \textbf{6.80 $\pm$8.32} \\ 
+ 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 & \textbf{7.44 $\pm$8.89} \\ 
+10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 & \textbf{7.56 $\pm$8.99} \\ 
+20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 & \textbf{7.59 $\pm$9.02} \\ 
 \end{tabular} 
@@
 According the table, the cut-off frequency at 500 Hz is the most efficient on all the TPR and all the sub-classes.
-The error for a 500 Hz cut-off frequency according to the sub-classes and the TPR is sums up in the figure \ref{fig:filterAmbiance}.
+The error for a 500 Hz cut-off frequency according to the sub-classes and the TPR is sums up in the figure \ref{fig:filterAmbiance}.\\
 
@@
 
-The first results averaged on all the TPR and the sub-classes for the supervised NMF is sums up on table \ref{tab:}, the frequency domain is, for now, limited to \textit{spectra}
+\subsection{Supervised NMF results}
 
-\begin{table}[h]
+As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ are presented in table \ref{tab:results_supervised}.
+The bar plot in figure \ref{fig:nmfSupervisedAmbiance} display the $gRMSE$ error for the best combination.
+
+\begin{table*}[h]
 \centering
-\begin{tabular}{lll}
+\begin{tabular}{lllllccc} 
+$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB)\\ 
+\hline 
+ 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} & \textbf{6.36 $\pm$7.55} \\ 
+ 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} & \textbf{\textcolor{red}{5.66 $\pm$6.38}} \\ 
+ 25 & 0 & thirdOctave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34 & \textbf{6.88 $\pm$8.15} \\ 
+ 25 & 500 & thirdOctave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} & \textbf{6.36 $\pm$7.73} \\ 
+\end{tabular} 
+\caption{RMSE error for supervised NMF averaged on all the TPR and sub-classes}
+\label{tab:results_supervised}
+\end{table*}
 
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=\linewidth]{../image/nmfAmbianceBar.pdf}
+\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = )}
+\label{fig:nmfSupervisedAmbiance}
+\end{figure}
+
+The best combination for the supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, the temporal window of 500 ms and $f_c = 500$ Hz.
+The errors produced by NMF is smaller than the filter approach for all the metrics.
+The figure \ref{fig:nmfSupervisedAmbiance} allows to see that the supervised NMF improves the estimation of the road traffic for $TPR = \left[-12~6\right]$ whereas the error for the higher $TPR$ is superior.\\
+
+The evolution of the error for 3 sub classes (\textit{alert}, \textit{climate} and \textit{transportation}) and 2 \textit{TPR} (-6 and 6) are displayed in figures \ref{} and \ref{}.\\
+
+\begin{figure}[!t]
+\centering
+	\begin{minipage}[t]{.24\textwidth}
+		\centering
+		\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6.pdf} 
+	\end{minipage}
+	\begin{minipage}[t]{.24\textwidth}
+		\centering
+		\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6.pdf} 
+	\end{minipage}
+	\caption{Evolution of the $gRMSE$ for 3 sub-classes for 2 \textit{TPR}}
+	\label{fig:RMSEevolution}
+\end{figure}
+
+We observe that the error if growing for TPR = -6, with an minimum in . The phenomena of early stop is present with few iteration . 
+In contrary, for \textit{TPR} = 6, the  error is decreasing constantly.
+
+\subsection{Semi-supervised NMF results}
+The error produced for the semi-supervised approach is summarize in table \ref{tab:results_semi_supervised}.
+
+\begin{table*}
+\centering
+\begin{tabular}{cclllccc} 
+$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB) \\ 
+\hline 
+ &  &  & spectra & 1 &  &  &  \\ 
+100 & 0 & 20000 & spectra & 2 & \textbf{4.02 $\pm$3.48} & \textbf{\textcolor{red}{1.14 $\pm$0.54}} & \textbf{\textcolor{red}{4.41 $\pm$3.69}} \\ 
+50 & 0 & 20000 & third octave & 1 & \textbf{\textcolor{red}{3.88 $\pm$2.33}} & 1.29 $\pm$0.69 & \textbf{4.43 $\pm$2.71} \\ 
+25 & 0 & 20000 & third octave & 2 & \textbf{4.29 $\pm$3.58} & 1.21 $\pm$0.61 & \textbf{4.74 $\pm$3.84} \\ 
 \end{tabular} 
-\caption{RMSE error for the low pass filter averaged on all the TPR and sub-classes}
-\label{tab:results_filter}
-\end{table}
+\caption{RMSE error for semi-supervised NMF averaged on all the TPR and sub-classes}
+\label{tab:results_semi_supervised}
+\end{table*}
 
-Résultat global pour le NMF supervisée en tableau puis en bar pour chaque ambiance et chaque TPR
-Tracer l'évolution 1s du signal sonore pour 2 ambiances et 2 TPR
-résultat semi-supervisée
-beta = 2, = 1
-%
+The best combination is for the third octave domain with $f_c$ = 20 kHz, $K$ = 50 and a temporal window null.
+The error,  compared to the supervised approach,  is lower.
+Furthermore, the standard deviation is lower too meaning that the semi-supervised approach offer a much more stable error than the previous method.
+The figure \ref{} summarize the error for the best scenario expanded for the sub-classes and the TPR.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=\linewidth]{../image/nmfAmbianceBar.pdf}
+\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 20 kHz)}
+\label{fig:nmfSemiSupervisedAmbiance}
+\end{figure}
+
+The mobile part of the dictionary, $\mathbf{W_r}$, is displayed in two particular sub-classes (\textit{alert}, which include harmonic sound sources, and \textit{climate}, composed of low frequencies) for two \textit{TPR} (-6 and 6).
+
+For \textit{TPR} = -6, for the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
+Whereas, in \textit{climate} sub-class, the \textit{perturbator}sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
+This class might then be describe with the help of $\mathbf{W_s}$.
+
+In the opposite, for the high \textit{TPR}, for the \textit{climate} sub-class, the harmonic content disappears in $\mathbf{W_r}$ and is replace by low frequencies content that might be traffic component.
+In this approach, as the objective function of to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
+This freedom deteriorates the reconstruction of the traffic signal.
+In the supervised approach, as $\mathbf{W}$ is only limited to the traffic spectrum.
+In consequence, NMF is constrained to use them to approximate $\mathbf{V}$, which improve the estimation of the traffic sound level for the high TPR.
+
+PLOT COST
+PLOT RMSE
+PLOT Lp 1s pour 2 ambiance 2 TPR
 
@@
 \end{document}
+
+%\section{State of the art in the detection and separation tasks for environmental sounds}\label{part:stateOfArt}
+%
+%Many references have focused on the classification of the sound environments \cite{allegro_automatic_2001}, on the recognition \cite{defreville_automatic_2006} or one the detection of the  environmental sound events \cite{heittola_sound_2011}.
+This category of sound is not specific to the urban environment but includes all sound sources that are not related to music or speech.
+The algorithms are first learned on a known database to evaluate the performance of the algorithm then is applied to a second database simulating its application in real cases.
+To achieve these tasks, many approaches have the same generic pattern: a first extraction phase followed by a classification phase.
+In the first phase, the parameters of the characteristic are applied to an audio file to describe it, from the simplest, such as the sound level, the spectrum Gravity Centrums, the harmonicity to the most complex (Mel-Frequency Cepstral Coefficient, Perceptual Linear Prediction).
+In the second step, the characteristics are classified according to their evolution making it possible to distinguish the different categories of sounds.
+Common classification techniques are Hidden Markov Models, Artificial Neural Network, Gaussian Mixture Models and Support Vector Machines.
+Cowling and Sitte \cite{cowling_comparison_2003} summarize multiples of these extraction features and classification techniques that are used in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+Even if each of them tested their algorithm on their own isolated sound database, specific database have been proposed to test and compare the performances of the algorithms as the DCASE database \cite{giannoulis_database_2013} or the TUT database \cite{mesaros_tut_2016}.
+Currently, the interest is focused on the use of the Neural Network in the case of the detection of acoustics events \cite{genaro_neural_2010} \cite{parascandolo_recurrent_2016}.

### 2017-10-17 article/articleApplied.tex
corrections AC part 1-3

@@
 
-\title{Estimation of the traffic sound level from measurements made in an urban environment}
+\title{Estimation of road traffic sound levels on simulated sound scenes with the Non-Negative Matrix Factorization.}
 
@@
 \section{Introduction}
-
-Road traffic noise is one of the main annoyance perceived by urban dwellers and a major societal issue because too long exposures to high sound levels can generate health problems  \cite{who_burden_2017}.
-Accordingly, the European Directive 2002/EC/49 \cite{directive} has been developed to determine the exposuer of the populations.
-Cities with a population of more than 100 000 must produce road traffic noise maps.
-These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and calculating sound propagation.
-The maps express $L_ {DEN}$ and $L_N$, respectively, \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels.
-These maps make it possible to estimate the number of city dwellers exposed to highly annoyant noise levels and to draw up action plans to reduce them.
+Since the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
+It allows to estimate the number of city dwellers exposed to highly noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and a sound propagation calculation.
+They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
 However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
@@
 These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which allow to get the long-term evolution of the traffic noise levels, or can be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time perdiods.
-The clustering between mobile and fixed measurements have been studied in \cite{can_measurement_2014}.
+The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}.
 
-Currently, we are seeing the diffusion of sensor networks in cities for multiple applications (air quality assessment, measurement of meteorological parameters), including assessment of urban noise levels.
-Various projects exist to study the establishment and feasibility of such installations as the DYNAMAP \cite{dynamap_2016} project.
-It focuses on sensor installations on specific roads in all cities of Milan and Rome \cite{bellucci_life_2017}.
-In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} plains to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, by using data assimilation techniques.
+Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters), including assessment of urban noise levels.
+DYNAMAP project \cite{dynamap_2016} studied the establishment and feasibility of such installations.
+It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
+In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
 
-If these installations, from their measurements, allow to have a more precise evaluation of the traffic noise in cities than the simulated maps, the issue of the correct estimation from measurement of the traffic sound level is still unsolved \cite{Mioduszewski}.
-Indeed, the urban sound environment is a complex environment gathering lots of different sounds both from a temporal and a frequency point of view (car passages, voices, bird's whistles, car horn \dots) that can overlap.
-In consequence, the traffic sound level estimation is not trivial task.
+If sensors networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}.
+Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap.
+In consequence, the traffic sound level estimation based on measurements is not trivial task.
 
-Many works have been focused on the detection  \cite{heittola_sound_2011} or on recognition \cite{defreville_automatic_2006} task of environmental sounds.
-These methods use a two step process : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classified them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
-A description of there features and classifiers can be found on \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
-However, many of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
-Although in some places, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where there is little traffic and overlaps with other sound sources.
-The only detection of the traffic component does not make it possible to determine precisely its noise level.
+Many works have been focused on the detection  or recognition tasks of environmental sounds \cite{heittola_sound_2011},  \cite{defreville_automatic_2006},  \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
+A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
+A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+However, most of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
+Although, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels.
+In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level.
 In consequence, to be effective on all the different sound environments, it seems more suited to consider the issue as the one of blind source separation.
@@
 One of the first and the most widely used techniques is the Independent Component Analysis \cite{comon_independent_1994}.
-The principle is to decompose $N$ recorded signals in a sum of $P$ independent sound sources weighted by linear relations.
+The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
 This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.
@@
 Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to install multiples sensors on one point to realize the source separation which is not feasible.
-A second method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximate the non-negative spectrogram of an audio file from the product of two matrices.
-It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of a monaural signal included speech and musical contents  \cite{wang_musical_2005} \cite{wilson_speech_2008}.
-This method has the advantage to easily deals with the overlap of the sound sources.
-For the environmental sounds, the method has been used for geo-localisation and classification of the sound environment like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded.
-For the source separation, it has been used by Innami and Kasai in the unsupervised case.
-They proposed a source separation in two step  by separating the sound background from the events first and by separating the events between them.
+A second method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the non-negative spectrogram of an audio file from the product of two matrices.
+It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
+This method has the advantage to easily deal with the overlap of the sound sources.
+For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded.
+For the source separation, it has been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012}.
+They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them.
 The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
@@
 
-Our study proposed to applied Non-Negative Matrix Factorization on simulated sound mixtures to succeed on the estimation of the traffic sound level.
-The use of simulated sound mixtures is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component what recordings do not allow.
+Our study proposed to applied Non-Negative Matrix Factorization on simulated sound scenes to estimate the traffic sound level.
+The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component (what recordings do not allow).
 Part \ref{part:nmf} details the technical aspect of NMF.
-Part \ref{part:protocol} is focused on the experimental protocol set up.
+Part \ref{part:protocol} described on the experimental protocol set up.
 Then part \ref{part:results} reveals the results obtained during the parametric study.
@@
 \section{Non-negative Matrix Factorization}\label{part:nmf}
-\subsection{Supervised NMF}
+\subsection{Description of NMF}
 Non-negative Matrix Factorization (NMF) is an approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which estimates the spectrogram (get from a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
@@
 
-where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{basis} matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$ (fig.
+where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$ (fig.
 \ref{fig:example_NMF}).
@@
 
-The choice of the dimensions is done in a such way that $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used
+The choice of the dimensions is imposes such as $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used
 
@@
 
-The operator $D(x\vert y)$ is a divergence calculation such as 
+The operator $D(x\vert y)$ is a divergence calculation such as:
 \begin{equation}
@@
 
-The prior knowledges on the content can be adjust with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})).
+The prior knowledges on the content can be adjusted with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})) to better take account prior knowledge of the sources.
 
-Multiple methods exist to solve the minimization problem (\ref{eq:min-D-WH}) (Alternating Least Square Method \cite{cichocki_regularized_2007}, Projected Gradient \cite{lin_projected_2007} \dots).
-Here, the multiplicative update is the chosen method \cite{lee_algorithms_2000} as it assures to have a non-negative results where the convergence of the results have been proved \cite{fevotte_algorithms_2011}.
-The minimization problem (\ref{eq:min-D-WH}) is solved iteratively.
-NMF is employed here in the supervised case: the \textit{dictionary} is already known as in the urban environments, a lot of different sound sources present are known and their spectrum can be obtained.
-The \textit{basis} are the unknown to find.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative upatde, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen \cite{lee_algorithms_2000} as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
+
+\subsection{Supervised NMF} 
+First, supervised NMF is used: the \textit{dictionary} includes audio spectrum of urban sound sources as, in the urban environments, a lot of different sound sources present are known and their spectrum can be obtained.
+The \textit{basis} are then the unknown to find.
 In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm
@@
 The product $A.B$ and $A/B$ symbolized the Hadamard product and ratio.
-The source separation is made after $I$ iterations either by extracting the dictionary and basis elements related to the traffic
+The source separation is made by extracting the dictionary and basis elements related to the traffic
 
@@
 
-One of the main issue with the supervised approach is this is not always adapted to different sound environments.
-Because in the supervised case, the fixed dictionary has to sum up all the sound sources present in the different sound environment.
+One of the main issue with the supervised approach is this is not always adapted to different sound environments as the fixed dictionary has to sum up all the sound sources that can be present.
 In consequence, as the dictionary dimension is limited and cannot include all the urban sounds, to offer more flexibility, semi-supervised NMF has been proposed \cite{lee_semi-supervised_2010}.
-This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ that is updated.
+This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
+Here, $J = 2$.
 The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic.
@@
 
-To assess the performance of NMF, simulated sound mixtures are used as it offers a controlled framework to design specific sound environments where all the traffic component is known.
-The estimation of the traffic sound level can be compare to an exact solution.
+Simulated sound scenes are used to assess the performance of NMF.
+This offers a controlled framework to design specific sound environments in which all the traffic component is known.
+Then, the road traffic sound levels estimated with the method can be compared to the real ones, introduced within each simulated sound scene.
 
@@
 
-A first corpus is design with the \textit{simScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}.
-\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a .wav format by superposing audio samples that come from an isolated sound database.
-This database is divided in two categories: the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient whereas all the sounds that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category.
+A corpus is designed with the \textit{simScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}.
+\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by superposing audio samples that come from an isolated sound database.
+This database is divided in two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient, $ii)$ the \textit{background} category includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
 Inside each category, the sound samples are grouped in sound classes (\textit{bird, car, foot steps} \dots), each of them being composed of multiples samples (bird01.wav, bird02.wav \dots).
 
-The software allows the user to control some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the presence of a fade in and a fade out \dots.
-Each parameter is completed with a standard deviation that may brings some random behavior between the scenes.
-Furthermore, with the global scene mixture, an audio for each sound class presented in the scene can be generated that allow to know its exact contribution and a text file is created that summarizes the time presence of all the events.\\
+The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may brings some random behavior between the scenes.
+Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
 
@@
 \label{fig:exampleSimScene}
-\caption{Example of a sound mixture composed of 3 sound classes (car, bird, car horn)}
+\caption{Example of a sound scene composed of 3 sound classes (car, bird, car horn)}
 \end{figure}
 
-With this software, a sound database has been built including 245 sound event samples (with the sound classes \textit{ringing bell, birds, sweeping bropom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and 154 sound background samples (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffci noise, ventilation, wind}).
-The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway on different speeds with multiple gear ratio \cite{gloaguen_creation_2017}.
+A sound database has been built including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and 154 sound background samples divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
+The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway on different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and with the help of the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-A sound mixture corpus is designed composed of 6 sub-corpus of 25 30 seconds audio files ($N = 25$).
+This database enables creating realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
+A sound mixing corpus is composed of 6 sub-corpuses of 25 audio files each lasting 30 seconds.
 Each sub-corpus is characterized by a specific generic sound class : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noise) and \textit{transportation} (train, tramway and plane).
-In each file, traffic component is present as sound background and event and is mixed with on or multiple sound classes.
+In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the sound classes.
 The sound classes that are not related to the traffic component are summed up as the \textit{perturbator} (n'existe pas comme terme, disruptive à la place ?) class.
-To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound mixture, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{perturbator} class, $L_{p,perturbator}$ following the relation (\ref{eq:tpr}).
+To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{perturbator} class, $L_{p,perturbator}$ following the relation (\ref{eq:tpr}).
 
@@
 When $TPR = -12$, the traffic component is then less present than when $TPR = 12$ where it is predominant on the \textit{perturbator} class.
-The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is too calculated (figure \ref{fig:exampleScene}).
+The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated (figure \ref{fig:exampleScene}).
 Finally, the number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TPR values).
@@
 \caption{Example of a scene of the \textit{animals} sub corpus.
-Spectrogram (on left), \textit{Piano Roll} of the different sound classes (on the middle) and 1-s equivalent sound level of the traffic, $Lp_{1s,traffic}$ and of the global sound mixture, $Lp_{1s,global}$(on right)}
+Spectrogram (on left), \textit{Piano Roll} of the different sound classes (on the middle) and 1-s equivalent sound level of the traffic, $Lp_{1s,traffic}$ and of the global sound scene, $Lp_{1s,global}$(on right)}
 \label{fig:exampleScene}
@@
 
-The experiment consists in applied NMF on the environmental sound corpus to estimate the sound level of the traffic road.
-This method is compared to a simple approach which is a low-frequency filter.
-This method considers, as the traffic component is mostly composed of low-frequencies, that the sound level can be calculated from the band-pass filter.
-Similarly, for NMF, the low-pass filter is applied on $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ with the aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
-The figure \ref{fig:block_diagram_protocol} summarizes the different steps of the process depending on the chosen method.\\
+The experiment consists in estimating the traffic road sound level on the environmental sound corpus.
+NMF is used and compared to a simple approach which consist in a frequency low-pass filter (this second method relies on the fact that road traffic is mainly composed of low frequencies, compared to main of the confounded sounds (voices, birds, etc.)) with the cut-off frequency $f_c$.
+Similarly, for NMF, the frequency low-pass filter is applied on the spectrogram $\mathbf{V}$ and on the dictionary $\mathbf{W}$ renamed $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ respectively.
+The aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
+The figure \ref{fig:block_diagram_protocol} summarizes the different steps of the process depending on the chosen method, which consist of 3 steps:
+
+\begin{itemize}
+\item the dictionary building for NMF,
+\item the estimation of the traffic sound level according the chosen method (NMF or frequency low-pass filter),
+\item the calculation of the error between the exact and the estimated sound levels.\\
+\end{itemize}
 
@@
 	\end{minipage}
-	\caption{Block diagrams summed up the different step of the precess. On top, for the low-pass filter, on bottom, for NMF.}
+	\caption{Block diagrams summed up the different step of the precess. On top, for the frequency low-pass filter, on bottom, for NMF.}
 	\label{fig:block_diagram_protocol}
 \end{figure*}
-
-3 calculation steps are set up : 
-\begin{itemize}
-\item the dictionary building for NMF
-\item the estimation of the traffic sound level according the chosen method (NMF or low-pass filter)
-\item calculation of the error between the exact and the estimated sound levels.\\
-\end{itemize}
 
@@
 It is composed of 53 audio files of car passages.
-Multiple version of dictionaries are tested.
-First, for each audio file, its spectrogram is calculated with fixed parameters ($w = 2^{14}$, $noverlap = 75 \%$, $nfft = 2^{14}$).
-Then a temporal rectangular window of 0.5 second is applied without overlapping on the spectrogram.
-In each window, the root mean square value are calculated.
-This window allow to get multiples spectrum from an audio file.
-An example, to illustrate the process, can be found on figure with a large window of 4 seconds.
+First, for each audio file, its spectrogram is calculated with fixed parameters (window size $w = 2^{14}$ with a 75 $\%$ overlapping and a number of point $nfft = 2^{14}$).
+Then a temporal rectangular window is applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
+The size of the window is set up at 0.5 and 1 second.
+In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra.
+An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} in the ease of a large window of 4 seconds.
 
@@
 
-As the number of spectrum got on all the sound database does not respect the constraint imposed about the dimension of NMF, a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
-A special case is added where the root mean square of the entire spectrum is applied on all the spectrogram of the audio file.
-In consequence, one spectra is generated by audio file.
-% It consists in the estimation of the $K$-means and then to calculate the euclidean distance between each centroid with all the spectrum.
-The spectrum having the smaller distance with the centroid is then included in the dictionary $\mathbf{W}$.
+As the number of elements got on all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
+A special case is added where the root mean square of the entire spectrogram is applied.
+One spectra is then generated by audio file.
 
 \subsubsection{Estimation of the traffic sound level}
-The traffic sound level is estimated according to the choice of the estimator (low-pass filter or NMF).
+The traffic sound level is estimated according to the choice of the estimator (a frequency low-pass filter or NMF).
 These methods are applied on the totality of the corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)), for all the cut-off frequencies $f_c$ ([500 1k 2k 5k 10k 20k] Hz) and for 5 $TPR$ ([-12 -6 0 6 12] dB).
 
-\paragraph{Low-pass filter}\mbox{} \\
-The first estimator to determine the traffic sound level is a basic low-pass filter which depend only on the cut-off frequencies $f_c$.
-As the road traffic is mainly composed of low frequencies (< 5000 Hz), this method considers that all the components located on the pass-band can be considered as road traffic.
+\paragraph{Frequency low-pass filter}\mbox{} \\
+The first estimator to determine the traffic sound level is a basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$.
 
 \paragraph{NMF}\mbox{} \\
-The second estimator is NMF with the supervised and the semi-supervised approaches (see part \ref{part:nmf}).
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands.
-Theses two methods enables to compare a fine approach (the linear scale) with a degraded one (the third octave bands) as it reduces the number of frequency bins (27 bands) and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
+The second estimator is supervised and semi-supervised NMF (see part \ref{part:nmf}).
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands (27 bands).
+Theses two methods enable to compare a fine approach (the linear scale) with a degraded one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
+NMF is performed on 400 iterations which is sufficiently enough to get a stabilized reconstruction.
 Furthermore, in the case of the linear frequency scale, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
-The $\beta$-divergence is reduced to the Kullback-Leibler divergence ($\beta = 1$) and to the Euclidean distance ($\beta = 2$).
-400 iterations are performed.
-As only one sound source is considered in the dictionary, the source separation is performed following the equation  (\ref{eq:separationExtraction}).
-Also, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is made with the original dictionary $\mathbf{W}$, as
+In consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is made with the original dictionary $\mathbf{W}$, as
 
@@
 \subsubsection{Metric}
-The performances of the two estimators to estimate the sound level of the traffic correctly are estimated through the calculation of two metrics based on the Root Mean Square Error (RMSE) :
+The performances of the two estimators of the road traffic sound level are assessed through the calculation of two metrics.
 
@@
 
-\item the Mean Absolute Error, $MAE$, which consists in calculate for each combination of factors the absolute difference between the exact and estimated global traffic sound level in dB of the $N$ scenes.
+\item The Mean Absolute Error, $MAE$, expresses the quality of the long-term reconstruction of the signal.
+It consists in the average over the $N$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
 
 \begin{equation}
-MAE = \frac{\sum_{n = 1}^N\vert L^n_{p,traffic}-\tilde{L}^n_{p,traffic} \vert}{N}
+MAE = \frac{\sum_{n = 1}^N\vert L^n_{p,traffic}-\tilde{L}^n_{p,traffic} \vert}{N}.
 \end{equation}
 
-This metric expresses the quality of the long-term reconstruction of the signal.
+%\item the global RMSE, \textit{gRMSE}, which consists in calculate for each combination of factors the error between the exact and estimated global traffic sound level in dB of the $N$ scenes.
+%
+%\begin{equation}
+%gRMSE = \sqrt{\frac{1}{N}\sum_{n = 1}^N \left(L^n_{p,traffic}-\tilde{L}^n_{p,traffic}\right)^2}
+%\end{equation}
+%
+%This metric expresses the quality of the long-term reconstruction of the signal.
 
-\item the global RMSE, \textit{gRMSE}, which consists in calculate for each combination of factors the error between the exact and estimated global traffic sound level in dB of the $N$ scenes.
-
-\begin{equation}
-gRMSE = \sqrt{\frac{1}{N}\sum_{n = 1}^N \left(L^n_{p,traffic}-\tilde{L}^n_{p,traffic}\right)^2}
-\end{equation}
-
-This metric expresses the quality of the long-term reconstruction of the signal. 
-\item The normalized short-term RMSE, $nRMSE$, consists in calculate, for each sound mixture, the error between the 1-s equivalent sound level exact and estimated of the traffic of each file normalized by the 1-s equivalent sound level of the global mixture in the linear scale,
+\item The normalized short-term Root Mean Square Error, $nRMSE$, calculates, for each sound scene, the error between the exact and estimated road traffic 1-s equivalent sound level of each file normalized by the 1-s equivalent sound level of the global scene, $p^t_{1s,global}$, in the linear scale,
 
@@
 with $T$ is the number of temporal bin in the signal.
-The linear scale is here more relevant than the dB scale as it is more sensitive to the error on the high sound levels.
+The linear scale is here more relevant as it is more sensitive to the error on the high sound levels than the dB scale.
 Then for one combination of factors, the $N$ $nRMSE$ calculated are averaged.\\
@@
 
-\subsection{Low-pass filter results}
-In a first time, the low-pass filter estimator is performed on all the scene.
-The table \ref{tab:results_filter} sums up the mean error on the totality of the 750 scenes to find the cut-off frequency the most efficient on all the cases.
+\subsection{Frequency low-pass filter results}
+In a first time, the frequency low-pass filter estimator is performed on all the scene.
+The table \ref{tab:results_filter} summarized the mean error on the totality of the 750 scenes to find the most efficient cut-off frequency.
 \\
@@
 \begin{tabular}{llll}
-$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB) \\ \hline
- 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} & \textbf{\textcolor{red}{5.96 $\pm$6.70}} \\ 
- 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} & \textbf{6.52 $\pm$7.73} \\ 
- 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} & \textbf{6.80 $\pm$8.32} \\ 
- 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 & \textbf{7.44 $\pm$8.89} \\ 
-10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 & \textbf{7.56 $\pm$8.99} \\ 
-20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 & \textbf{7.59 $\pm$9.02} \\ 
+$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$  \\ \hline
+ 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} \\ 
+ 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} \\ 
+ 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} \\ 
+ 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 \\ 
+10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 \\ 
+20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 \\ 
 \end{tabular} 
@@
 
-According the table, the cut-off frequency at 500 Hz is the most efficient on all the TPR and all the sub-classes.
-The error for a 500 Hz cut-off frequency according to the sub-classes and the TPR is sums up in the figure \ref{fig:filterAmbiance}.\\
+According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TPR and all the sub-classes.
+The corresponding errors according to the sub-classes and the TPR are summarized in the figure \ref{fig:filterAmbiance}.\\
 
@@
 \centering
-\includegraphics[width=\linewidth]{../image/filterAmbianceBar.pdf}
-\caption{Bar plot for the low-pass filter according to the sub-classes and the TPR at 500 Hz cut-off frequency}
+\includegraphics[width=\linewidth]{../image/AmbianceFilter.pdf}
+\caption{Bar plot for the filter according to the sub-classes and the TPR at $f_c$ = 500 Hz cut-off frequency}
 \label{fig:filterAmbiance}
@@
 
-The error for all the sub-corpus is important for the low TPR (-12 and -6) due to the confusion of the \textit{perturbator} class as the traffic component whereas  it is not.
-The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while for the \textit{climate} and \textit{transportation} the error is far more important.
-For the other TPR, the traffic is more present and become dominating on the \textit{perturbator} class.
-Then the error is due to the suppression of the traffic energy by the low-pass filter.
+The error for all the sub-corpus is important for the low TPR (-12 and -6) due to the confusion between the \textit{perturbator} class and the \textit{traffic} class.
+The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while, for other sub-classes, the error are far more important.
+For the higher \textit{TPR}, the traffic is more present and become dominating on the \textit{perturbator} class.
+The lower error, here, is due to the suppression of the traffic energy by the filter.
+Finally, the error for $f_c = 20$ kHz is equivalent to the one produced when the \textit{traffic} and the \textit{perturbator} classes are take into account together without distinction.
 
@@
 
-As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ are presented in table \ref{tab:results_supervised}.
-The bar plot in figure \ref{fig:nmfSupervisedAmbiance} display the $gRMSE$ error for the best combination.
+As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ at the $\nth{400}$ iteration are presented in table \ref{tab:results_supervised}.
+The bar plot in figure \ref{fig:nmfSupervisedAmbiance} displays the $MAE$ error for the best combination.\\
 
-\begin{table*}[h]
+\begin{table*}[t]
 \centering
-\begin{tabular}{lllllccc} 
-$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB)\\ 
+\begin{tabular}{lllllcc} 
+$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\ 
 \hline 
- 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} & \textbf{6.36 $\pm$7.55} \\ 
- 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} & \textbf{\textcolor{red}{5.66 $\pm$6.38}} \\ 
- 25 & 0 & thirdOctave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34 & \textbf{6.88 $\pm$8.15} \\ 
- 25 & 500 & thirdOctave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} & \textbf{6.36 $\pm$7.73} \\ 
+ 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} \\ 
+ 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} \\ 
+ 25 & 0 & third octave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34\\ 
+ 25 & 500 & third octave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} \\ 
 \end{tabular} 
@@
 
-\begin{figure}[hbtp]
+\begin{figure}[h]
 \centering
-\includegraphics[width=\linewidth]{../image/nmfAmbianceBar.pdf}
-\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = )}
+\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised.pdf}
+\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 500 Hz )}
 \label{fig:nmfSupervisedAmbiance}
@@
 
-The best combination for the supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, the temporal window of 500 ms and $f_c = 500$ Hz.
-The errors produced by NMF is smaller than the filter approach for all the metrics.
-The figure \ref{fig:nmfSupervisedAmbiance} allows to see that the supervised NMF improves the estimation of the road traffic for $TPR = \left[-12~6\right]$ whereas the error for the higher $TPR$ is superior.\\
+\begin{table}[h]
+\centering
+\begin{tabular}{ccc}
+    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
+TPR & filtre & supervised NMF \\ \hline
+ -12 & 17.20 $\pm$ 4.09 & 15.76 $\pm$ 5.82 \\ 
+ -6 &  6.91 $\pm$ 2.82 & 6.32 $\pm$ 3.24\\ 
+ 0 & 1.09 $\pm$ 0.16 & 1.18 $\pm$ 0.19\\
+  6 &  1.47 $\pm$ 0.12 & 1.59 $\pm$ 0.13\\ 
+ 12 &  1.60 $\pm$ 0.16 & 1.77 $\pm$ 0.07
+\end{tabular}
+\caption{Comparison of the $MAE$ error for the filter at 500 Hz and the best combinaison of the supervised NMF}
+\label{tab:resultsComparison}
+\end{table}
 
-The evolution of the error for 3 sub classes (\textit{alert}, \textit{climate} and \textit{transportation}) and 2 \textit{TPR} (-6 and 6) are displayed in figures \ref{} and \ref{}.\\
+The best combination for supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, a temporal window of 500 ms and $f_c = 500$ Hz.
+The errors produced by NMF, on the totality of the scenes, is smaller than the filter approach for all the metrics.
+For the higher \textit{TPR}, both methods have similar performances (table \ref{tab:resultsComparison}).
+For the lower \textit{TPR}, TPR, supervised NMF reduces significantly the error.
+By distinguishing each class, it is lower for the \textit{alert} and \textit{animals} than the rest sub-classes.
+\\
 
-\begin{figure}[!t]
+This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TPR} (-6 and 6) until the $\nth {100}$ iterations.\\
+
+\begin{figure}
     \centering
-	\begin{minipage}[t]{.24\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6.pdf} 
-	\end{minipage}
-	\begin{minipage}[t]{.24\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6.pdf} 
-	\end{minipage}
-	\caption{Evolution of the $gRMSE$ for 3 sub-classes for 2 \textit{TPR}}
-	\label{fig:RMSEevolution}
+    \subfigure[]{\label{fig:EarlyStop1}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6.pdf}}
+    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:EarlyStop1}) and TPR = 6 (fig. \ref{fig:EarlyStop2})}
+    \label{fig:EarlyStop12}
 \end{figure}
 
-We observe that the error if growing for TPR = -6, with an minimum in . The phenomena of early stop is present with few iteration . 
-In contrary, for \textit{TPR} = 6, the  error is decreasing constantly.
+\begin{figure}
+    \centering
+    \subfigure[]{\label{fig:costSup1}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6Cost.pdf}}
+    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6Cost.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:costSup1}) and TPR = 6 (fig. \ref{fig:costSup2})}
+    \label{fig:costSup12}
+\end{figure}
+
+For the TPR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
+Here as the $\mathbf{W}$ is only composed of traffic elements and the sound scenes are composed of a predominant traffic, the reconstruction of the signal is easier.
+In opposite, for the case where \textit{TPR} = -6, for the \textit{alert} signal, the cost function and the $MAE$ error are decreasing too.
+But for the \textit{climate} and \textit{transportation} sub-classes, the $MAE$ error is increasing.
+Even if the cost function decreases, the quality of the traffic signal rebuilt is not improved.
+This difference can be explained by the type of sound present in these different sub-classes: for \textit{alert} and \textit{animals}, it includes harmonic sounds which belong in the frequency range $\left[2000-5000\right]$ Hz while the other sub-classes include lower frequency sounds.
+As $\mathbf{W}$ is composed of traffic spectrum, located in low frequencies too, it is more difficult to recompose correctly the traffic signal when the \textit{perturbator} sound is predominant and composed of low frequencies too.
+As NMF minimized the cost function (equation \ref{eq:min-D-WH}), \textit{traffic} spectrum then are used to synthesized the global mixture at the expense of the traffic signal.
 
 \subsection{Semi-supervised NMF results}
-The error produced for the semi-supervised approach is summarize in table \ref{tab:results_semi_supervised}.
+
+The errors produced for the semi-supervised approach is summarized in table \ref{tab:results_semi_supervised} according to $\beta$ and the frequency domain (spectral or third octave).\\
 
@@
 \begin{tabular}{cclllccc} 
-$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ & $gRMSE$ (dB) \\ 
+$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\ 
 \hline 
- &  &  & spectra & 1 &  &  &  \\ 
-100 & 0 & 20000 & spectra & 2 & \textbf{4.02 $\pm$3.48} & \textbf{\textcolor{red}{1.14 $\pm$0.54}} & \textbf{\textcolor{red}{4.41 $\pm$3.69}} \\ 
-50 & 0 & 20000 & third octave & 1 & \textbf{\textcolor{red}{3.88 $\pm$2.33}} & 1.29 $\pm$0.69 & \textbf{4.43 $\pm$2.71} \\ 
-25 & 0 & 20000 & third octave & 2 & \textbf{4.29 $\pm$3.58} & 1.21 $\pm$0.61 & \textbf{4.74 $\pm$3.84} \\ 
+ 50 & 0 & spectra &  1000 & 1 & \textbf{4.50 $\pm$3.05} & 1.45 $\pm$0.76 \\
+100 & 0 & spectra & 20000 & 2 & \textbf{4.05 $\pm$3.43} & \textbf{\textcolor{red}{1.16 $\pm$0.53}} \\ 
+ 25 & 0 & third octave & 20000 & 1 & \textbf{\textcolor{red}{3.96 $\pm$2.18}} & 1.34 $\pm$0.72 \\ 
+ 25 & 0 & third octave & 20000 & 2 & \textbf{4.29 $\pm$3.58} & \textbf{1.21 $\pm$0.61}
 \end{tabular} 
@@
 
-The best combination is for the third octave domain with $f_c$ = 20 kHz, $K$ = 50 and a temporal window null.
+The best combination is got in the third octave domain with $\beta = 1$, $f_c$ = 20 kHz, $K$ = 25 and a temporal window null.
 The error, compared to the supervised approach, is lower.
 Furthermore, the standard deviation is lower too meaning that the semi-supervised approach offer a much more stable error than the previous method.
-The figure \ref{} summarize the error for the best scenario expanded for the sub-classes and the TPR.
+The figure \ref{fig:nmfSemiSupervisedAmbiance} displayed the error for the best scenario averaged on all the sub-classes and the \textit{TPR}.
+The table \ref{tab:comparisonFilterSemiSupervised} summarizes the error expanded to the \textit{TPR}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{../image/nmfAmbianceBar.pdf}
+\includegraphics[width=\linewidth]{../image/AmbianceNmfSemiSupervised.pdf}
 \caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 20 kHz)}
@@
 
-The mobile part of the dictionary, $\mathbf{W_r}$, is displayed in two particular sub-classes (\textit{alert}, which include harmonic sound sources, and \textit{climate}, composed of low frequencies) for two \textit{TPR} (-6 and 6).
+\begin{table}
+\centering
+\begin{tabular}{cccc} 
+    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
+TPR & filter & semi-Supervised NMF \\ \hline 
+ -12 & 17.20 $\pm$4.09 &  7.46 $\pm$2.88 \\ 
+ -6 &  6.91 $\pm$2.82 & 2.15 $\pm$0.70\\ 
+ 0 & 1.09 $\pm$0.16 &  2.20 $\pm$0.45\\ 
+ 6 &  1.47 $\pm$0.12 &  3.61 $\pm$0.29\\ 
+ 12 &  1.60 $\pm$0.16 &  4.02 $\pm$0.33\\ 
+\end{tabular} 
+\caption{Comparison of the baseline and semi-supervised NMF with the best combination}
+\label{tab:comparisonFilterSemiSupervised}
+\end{table}
 
-For \textit{TPR} = -6, for the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
-Whereas, in \textit{climate} sub-class, the \textit{perturbator}sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
-This class might then be describe with the help of $\mathbf{W_s}$.
+The error for low TPR is much lower than the filter or supervised NMF.
+The add of the mobile part makes NMF less constraint and therefore allows to adapt NMF to low traffic contents.
+Figure display the mobile part of the dictionary, $\mathbf{W_r}$ for two particular sub-classes (\textit{alert} and \textit{climate} and for \textit{TPR} = -6.\\
 
-In the opposite, for the high \textit{TPR}, for the \textit{climate} sub-class, the harmonic content disappears in $\mathbf{W_r}$ and is replace by low frequencies content that might be traffic component.
-In this approach, as the objective function of to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
-This freedom deteriorates the reconstruction of the traffic signal.
-In the supervised approach, as $\mathbf{W}$ is only limited to the traffic spectrum.
-In consequence, NMF is constrained to use them to approximate $\mathbf{V}$, which improve the estimation of the traffic sound level for the high TPR.
+\textbf{figure $W_r$ for alert and climate}\\
 
-PLOT COST
-PLOT RMSE
-PLOT Lp 1s pour 2 ambiance 2 TPR
+Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TPR}.
+Here, NMF uses $\mathbf{W_r}$ in order to minimize the cost function.
+As the mobile part is not constraint, traffic content is included in it, decreasing the quality of the traffic signal reconstruction.
+Figure display the 2 basis of $\mathbf{W_r}$.
+\\
+
+\textbf{figure $W_r$ for alert and climate}\\
+
+These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TPR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
+For the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
+Whereas, in \textit{climate} sub-class, the \textit{perturbator} sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
+This class might be describe with the help of $\mathbf{W_s}$.
+
+For \textit{TPR} = 6, the error is increasing for all sub-classes.
+As the objective function is to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
+This degree of freedom deteriorates the reconstruction of the traffic signal.
+\\
+
+\begin{figure}
+    \centering
+    \subfigure[]{\label{fig:EarlyStopSemi1}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR-6.pdf}}
+    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR6.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TPR = 6 (fig. \ref{fig:EarlyStopSemi2})}
+    \label{fig:EarlyStopSemi12}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfigure[]{\label{fig:costSemiSup1}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR-6Cost.pdf}}
+    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR6Cost.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:costSemiSup1}) and TPR = 6 (fig. \ref{fig:costSemiSup2})}
+    \label{fig:costSemiSup12}
+\end{figure}
+
+\textbf{miss the evolution of the Lp for the 25 scenes, estimate and exact + 2/3 scenes to see the evolution during the 30 seconds of the scenes ?}
 
@@
 \end{document}
-
-%\section{State of the art in the detection and separation tasks for environmental sounds}\label{part:stateOfArt}
-%
-%Many references have focused on the classification of the sound environments \cite{allegro_automatic_2001}, on the recognition \cite{defreville_automatic_2006} or one the detection of the  environmental sound events \cite{heittola_sound_2011}.
-This category of sound is not specific to the urban environment but includes all sound sources that are not related to music or speech.
-The algorithms are first learned on a known database to evaluate the performance of the algorithm then is applied to a second database simulating its application in real cases.
-To achieve these tasks, many approaches have the same generic pattern: a first extraction phase followed by a classification phase.
-In the first phase, the parameters of the characteristic are applied to an audio file to describe it, from the simplest, such as the sound level, the spectrum Gravity Centrums, the harmonicity to the most complex (Mel-Frequency Cepstral Coefficient, Perceptual Linear Prediction).
-In the second step, the characteristics are classified according to their evolution making it possible to distinguish the different categories of sounds.
-Common classification techniques are Hidden Markov Models, Artificial Neural Network, Gaussian Mixture Models and Support Vector Machines.
-Cowling and Sitte \cite{cowling_comparison_2003} summarize multiples of these extraction features and classification techniques that are used in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
-Even if each of them tested their algorithm on their own isolated sound database, specific database have been proposed to test and compare the performances of the algorithms as the DCASE database \cite{giannoulis_database_2013} or the TUT database \cite{mesaros_tut_2016}.
-Currently, the interest is focused on the use of the Neural Network in the case of the detection of acoustics events \cite{genaro_neural_2010} \cite{parascandolo_recurrent_2016}.

### 2017-10-18 article/articleApplied.tex
updates (tab with factors mainly)

@@
 
-This approach has been used TO COMPLETE.
+%This approach has been used TO COMPLETE.
 
@@
 NMF is used and compared to a simple approach which consist in a frequency low-pass filter (this second method relies on the fact that road traffic is mainly composed of low frequencies, compared to main of the confounded sounds (voices, birds, etc.)) with the cut-off frequency $f_c$.
-Similarly, for NMF, the frequency low-pass filter is applied on the spectrogram $\mathbf{V}$ and on the dictionary $\mathbf{W}$ renamed $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ respectively.
+The spectrogram $\mathbf{V}$ is built with a window size $w = 2^{14}$ with a 75 $\%$ overlapping and a number of point $nfft = 2^{14}$.
+Therefore, the dimensions of $V$ are $F$ = 8192 and $N$ = 772.
+Similarly, the frequency low-pass filter is applied on the spectrogram $\mathbf{V}$ and on the dictionary $\mathbf{W}$ ( renamed $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ respectively).
 The aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
@@
 It is composed of 53 audio files of car passages.
-First, for each audio file, its spectrogram is calculated with fixed parameters (window size $w = 2^{14}$ with a 75 $\%$ overlapping and a number of point $nfft = 2^{14}$).
+First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 75 $\%$ overlap, $nfft$).
 Then a temporal rectangular window is applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
@@
 	\end{minipage}
-	\caption{Example of the extraction of the spectrum with a 4 seconds window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
+	\caption{Example of the extraction of the spectrum with a 1 second window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
 	\label{fig:dictionaryExtraction}
@@
 
-As the number of elements got on all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
-A special case is added where the root mean square of the entire spectrogram is applied.
+However, the number of elements got on all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$).
+With a 1 second window, $K$ = 1003.
+In consequence, a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
+A special case is added where the root mean square of \textit{all} the spectrogram is applied.
 One spectra is then generated by audio file.
+Each element is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ is the $\ell-1$ norm.
+Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{cccc}
+$K$ & 25  & 50 & 100 \\ \hline
+\begin{tabular}[c]{@{}c@{}}\textbf{temporal} \\ \textbf{window (s)}\end{tabular} & 0.5 & 1  & \textit{all}
+\end{tabular} 
+\caption{Summary of the dictionary parameters}
+\label{tab:dictionary_factors}
+\end{table}
 
@@
 
-Then for each spectrogram, the estimated equivalent traffic sound level in dB of the entire scene, $\tilde{L}_{p,traffic}$, is calculated as well as the 1 second equivalent sound pressure level, $\tilde{p}_{1s,traffic}$.
+For the third octave frequency scale, only the case $f_c = 20$ kHz is applied. 
+During the iteration process, the estimated equivalent traffic sound level in dB of the entire scene, $\tilde{L}_{p,traffic}$, is calculated as well as the 1 second equivalent sound pressure level, $\tilde{p}_{1s,traffic}$.
+Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
+
+\begin{table*}[]
+\centering
+\begin{tabular}{lcccccc}
+$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
+$\mathbf{TPR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
+\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
+\textbf{method} & \multicolumn{2}{c}{filter} & \multicolumn{2}{c}{supervised NMF} & \multicolumn{2}{c}{semi-supervised NMF} \\ \hline
+$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
+\begin{tabular}[c]{@{}c@{}}\textbf{frequency} \\ \textbf{scale}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}
+\end{tabular}
+\caption{Summary of the different parameters and their values for the estimation of the traffic sound level}
+\label{tab:estimation_factors}
+\end{table*}
 
@@
 
-%\item the global RMSE, \textit{gRMSE}, which consists in calculate for each combination of factors the error between the exact and estimated global traffic sound level in dB of the $N$ scenes.
-%
-%\begin{equation}
-%gRMSE = \sqrt{\frac{1}{N}\sum_{n = 1}^N \left(L^n_{p,traffic}-\tilde{L}^n_{p,traffic}\right)^2}
-%\end{equation}
-%
-%This metric expresses the quality of the long-term reconstruction of the signal.
-
 \item The normalized short-term Root Mean Square Error, $nRMSE$, calculates, for each sound scene, the error between the exact and estimated road traffic 1-s equivalent sound level of each file normalized by the 1-s equivalent sound level of the global scene, $p^t_{1s,global}$, in the linear scale,
@@
 
-In all, 5940 settings are performed.
+In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 6848 settings are performed between the different dictionary $\mathbf{W}$ and the multiple parameters of the estimation step.
 

### 2017-10-19 article/articleApplied.tex
done with intro

@@
 
-\title{Estimation of road traffic sound levels on simulated sound scenes with the Non-Negative Matrix Factorization.}
+\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
+
+\title{Estimation of the road traffic sound levels based on the Semi-Supervised Non-Negative Matrix Factorization of Magnitude Spectrograms}
 
@@
 \section{Introduction}
-Since the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-It allows to estimate the number of city dwellers exposed to highly noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
-These maps come from a simulation process resulting from the estimation of the traffic density on the main roads and a sound propagation calculation.
+With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
+It allows \ml{who ?}  to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
 They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
@@
 In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day.
-Since the creation of road traffic noise maps entails long data collection and calculation times, the use of measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
-These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which allow to get the long-term evolution of the traffic noise levels, or can be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time perdiods.
+Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
+These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels.
+It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time perdiods.
 The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}.
+\ml{je ne vois pas l'intéret de cette dernière phrase}
 
-Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters), including assessment of urban noise levels.
+Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels.
 DYNAMAP project \cite{dynamap_2016} studied the establishment and feasibility of such installations.
@@
 In consequence, the traffic sound level estimation based on measurements is not trivial task.
+\ml{parler ici de l'approche des espagnols.
+ref par email}
 
-Many works have been focused on the detection  or recognition tasks of environmental sounds \cite{heittola_sound_2011},  \cite{defreville_automatic_2006},  \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
+Many recent works have focused on the detection  or recognition tasks of environmental sounds \cite{heittola_sound_2011},  \cite{defreville_automatic_2006},  \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
 A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
@@
 However, most of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
-Although, near major roads or ring roads, traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels.
+Although near major roads or ring roads traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels.
 In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level.
-In consequence, to be effective on all the different sound environments, it seems more suited to consider the issue as the one of blind source separation.
+In consequence, to be effective on a wide range of sound environments, we propose in this paper to follow the blind source separation paradigm.
+That is, separating the contribution of the traffic from the other sources within a polyphonic scene.
 
-One of the first and the most widely used techniques is the Independent Component Analysis \cite{comon_independent_1994}.
+One of the first and the most widely used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}.
 The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
@@
 However, ICA is limited to only over determined cases ($N > P$).
-Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to install multiples sensors on one point to realize the source separation which is not feasible.
-A second method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the non-negative spectrogram of an audio file from the product of two matrices.
+Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation.
+A more convenient method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices.
 It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
-This method has the advantage to easily deal with the overlap of the sound sources.
-For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF allows to classify the audio files according to the 10 cities where they have been recorded.
-For the source separation, it has been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012}.
+By design, this method deals with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane.
+For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.
+It has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012} for source separation.
 They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them.
@@
 
-Our study proposed to applied Non-Negative Matrix Factorization on simulated sound scenes to estimate the traffic sound level.
-The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component (what recordings do not allow).
-Part \ref{part:nmf} details the technical aspect of NMF.
-Part \ref{part:protocol} described on the experimental protocol set up.
-Then part \ref{part:results} reveals the results obtained during the parametric study.
+We propose in this paper a method based on the Non-Negative Matrix Factorization technique to estimate the traffic sound level.
+To validate the approach, we consider a corpus of simulated (artifically created) scenes.
+The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of a urban scene.
+
+The remaining of the paper is organized as follows.
+Section \ref{part:nmf} details the technical aspect of NMF.
+Section \ref{part:protocol} described on the experimental protocol set up.
+Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
 

### 2017-10-19 article/articleApplied.tex
done up to experiments

@@
 Section \ref{part:nmf} details the technical aspect of NMF.
-Section \ref{part:protocol} described on the experimental protocol set up.
+Section \ref{part:protocol} described on the experimental protocol setup.
 Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
+
+\ml{faire une figure avec un diagrame de bloc et une section proposed approach avec des sub temps/frequence nmfsup nmfsemi}
 
@@
 \subsection{Description of NMF}
-Non-negative Matrix Factorization (NMF) is an approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which estimates the spectrogram (get from a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
+Non-negative Matrix Factorization (NMF) is a matrix approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
 
@@
 
-The choice of the dimensions is imposes such as $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used
+The choice of the dimensions should be such that $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used
 
@@
 
-The prior knowledges on the content can be adjusted with the add of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})) to better take account prior knowledge of the sources.
+Prior knowledge on the content can be adjusted with the addition of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})) to better take account prior knowledge of the sources.
 
-Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative upatde, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen \cite{lee_algorithms_2000} as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen \cite{lee_algorithms_2000} as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
 
@@
 First, supervised NMF is used: the \textit{dictionary} includes audio spectrum of urban sound sources as, in the urban environments, a lot of different sound sources present are known and their spectrum can be obtained.
-The \textit{basis} are then the unknown to find.
+The \textit{basis} are then the unknown to estimate.
 In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm
@@
 
-One of the main issue with the supervised approach is this is not always adapted to different sound environments as the fixed dictionary has to sum up all the sound sources that can be present.
-In consequence, as the dictionary dimension is limited and cannot include all the urban sounds, to offer more flexibility, semi-supervised NMF has been proposed \cite{lee_semi-supervised_2010}.
-This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$ (composed here of traffic audio spectrum) and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
+One of the main issue with the supervised approach is To better acount for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010}  to offer more flexibility.
+This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
 Here, $J = 2$.
@@
 
-Simulated sound scenes are used to assess the performance of NMF.
+In order to vzlidate the usefulness of considering the NMF framewrok to estimate the road traffic noise level, one need to have a reference level.
+It can hardly de measured or even annotated from real life recordings.
+Thus,  simulated sound scenes are used to assess the performance of the propsed NMF.
 This offers a controlled framework to design specific sound environments in which all the traffic component is known.
@@
 A corpus is designed with the \textit{simScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}.
-\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by superposing audio samples that come from an isolated sound database.
+\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by summing audio samples that come from an isolated sound database.
 This database is divided in two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient, $ii)$ the \textit{background} category includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
@@
 The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway on different speeds with multiple gear ratio.
-The other audio files have been found online (\textit{freesound.org}) and with the help of the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
+The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
 This database enables creating realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
 A sound mixing corpus is composed of 6 sub-corpuses of 25 audio files each lasting 30 seconds.
-Each sub-corpus is characterized by a specific generic sound class : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noise) and \textit{transportation} (train, tramway and plane).
+Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult.
+The classes are : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noise) and \textit{transportation} (train, tramway and plane).
 In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the sound classes.
-The sound classes that are not related to the traffic component are summed up as the \textit{perturbator} (n'existe pas comme terme, disruptive à la place ?) class.
-To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{perturbator} class, $L_{p,perturbator}$ following the relation (\ref{eq:tpr}).
+The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class.
+To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$ following the relation (\ref{eq:tpr}).
 
 \begin{equation}\label{eq:tpr}
-TPR = L_{p,traffic}-L_{p,perturbator}
+TPR = L_{p,traffic}-L_{p,interfering}
 \end{equation}
 
-with the \textit{Traffic Perturbator Ratio} $TPR = \left[-12, -6, 0, 6, 12\right]$.
-When $TPR = -12$, the traffic component is then less present than when $TPR = 12$ where it is predominant on the \textit{perturbator} class.
+with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$.
+When $TIR = -12$, the traffic component is then less present than when $TIR = 12$ where it is predominant on the \textit{perturbator} class.
 The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated (figure \ref{fig:exampleScene}).

### 2017-10-19 article/articleApplied.tex
starting results

@@
 
-The experiment consists in estimating the traffic road sound level on the environmental sound corpus.
-NMF is used and compared to a simple approach which consist in a frequency low-pass filter (this second method relies on the fact that road traffic is mainly composed of low frequencies, compared to main of the confounded sounds (voices, birds, etc.)) with the cut-off frequency $f_c$.
+The experiment consists in estimating the traffic road sound level on the environmental sound corpus built using the simulation tool.
+NMF is used and compared to a baseli,e approach which consist in a frequency low-pass filter as road traffic is mainly composed of low frequencies, compared to main of the intefereing.
+this baseline has one parameter being the cut-off frequency $f_c$.
+
 The spectrogram $\mathbf{V}$ is built with a window size $w = 2^{14}$ with a 75 $\%$ overlapping and a number of point $nfft = 2^{14}$.
@@
 Similarly, the frequency low-pass filter is applied on the spectrogram $\mathbf{V}$ and on the dictionary $\mathbf{W}$ ( renamed $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ respectively).
-The aim to focus the reconstruction of the signal on the frequencies where the traffic components are.
+This to focus the NMF estimation of the signal on the frequencies where the traffic components are most present.
 The figure \ref{fig:block_diagram_protocol} summarizes the different steps of the process depending on the chosen method, which consist of 3 steps:
@@
 
-The dictionary is built from a sound database dedicated specially to this task.
-It is composed of 53 audio files of car passages.
+The dictionary is built from a sound database dedicated specifically to this task.
+It is composed of 53 audio files of passing cars.
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 75 $\%$ overlap, $nfft$).
-Then a temporal rectangular window is applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
+Then a temporal rectangular window \ml{pourquoi pas de fenetrage ?} is applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
 The size of the window is set up at 0.5 and 1 second.
@@
 	\end{minipage}
-	\caption{Example of the extraction of the spectrum with a 1 second window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
+	\caption{Example of the computation of the spectrum with a 1 second window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
 	\label{fig:dictionaryExtraction}
@@
 
-However, the number of elements got on all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$).
-With a 1 second window, $K$ = 1003.
-In consequence, a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
+Since the number of elements given by porcessing all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
 A special case is added where the root mean square of \textit{all} the spectrogram is applied.
 One spectra is then generated by audio file.
+\ml{pas clair}
 Each element is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ is the $\ell-1$ norm.
@@
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands (27 bands).
-Theses two methods enable to compare a fine approach (the linear scale) with a degraded one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
-NMF is performed on 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
+NMF inference is performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
 Furthermore, in the case of the linear frequency scale, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
-In consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is made with the original dictionary $\mathbf{W}$, as
+As a consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
 
@@
 
-\subsubsection{Metric}
+\subsubsection{Metrics}
 The performances of the two estimators of the road traffic sound level are assessed through the calculation of two metrics.
@@
 
-with $T$ is the number of temporal bin in the signal.
+where $T$ is the number of temporal bin in the signal.
 The linear scale is here more relevant as it is more sensitive to the error on the high sound levels than the dB scale.
@@
 \section{Results}\label{part:results}
+
+\ml{la section resultats disctue des resultats, ne pas etre lineaire, genre on fait  puis on fait ca...}
 

### 2017-10-25 article/articleApplied.tex
Corrections ML

@@
 \section{Introduction}
+\subsection{Related work}
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-It allows \ml{who ?}  to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+Theses maps allow to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
 These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
@@
 These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels.
-It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time perdiods.
-The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}.
+It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
+%The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}.
 \ml{je ne vois pas l'intéret de cette dernière phrase}
+
+\begin{figure*}
+    \centering
+    \subfigure[]{\label{fig:blocDiagramNMF}
+    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
+    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
+    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
+    \label{fig:blocDiagram}
+\end{figure*}
 
@@
 Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap.
-In consequence, the traffic sound level estimation based on measurements is not trivial task.
+Consequently, the traffic sound level estimation based on measurements is not trivial task.
 \ml{parler ici de l'approche des espagnols.
@@
 
-We propose in this paper a method based on the Non-Negative Matrix Factorization technique to estimate the traffic sound level.
-To validate the approach, we consider a corpus of simulated (artifically created) scenes.
-The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of a urban scene.
+\subsection{Proposed approach}
+
+We propose in this paper a method based on the Non-Negative Matrix Factorization (NMF) technique to estimate the global, $\tilde{L}_{p,traffic}$, and the 1-s equivalent, $\tilde{p}_{1s,traffic}$ sound level of the traffic through the supervised and the semi-supervised approaches.
+To validate this approach, we consider a corpus of simulated scenes artificially created with the simulator software \textit{simScene}.
+The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of an urban scene ($L_{p,traffic}$ and $p_{1s,traffic}$).
+Both the sound scene simulation and NMF require the creation of two sound databases (figure  \ref{fig:blocDiagramNMF}).
+In parallel, a baseline method built from a frequency low-pass filter is computed.
+This method considers that road traffic is mainly composed of low frequencies and therefore can be filtered by a low-pass filter at the cut-off frequency $f_c$ (figure \ref{fig:blocDiagramFilter}).
+The performance of the frequency low-pass filter and NMF are then compared with the calculation of two metrics (Mean Absolute Error, normalized Root Mean Square Error).
 
@@
 Section \ref{part:nmf} details the technical aspect of NMF.
-Section \ref{part:protocol} described on the experimental protocol setup.
+Section \ref{part:protocol} described on the design of the environmental sound scene corpus and the experimental protocol setup.
 Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
 
-\ml{faire une figure avec un diagrame de bloc et une section proposed approach avec des sub temps/frequence nmfsup nmfsemi}
+%\ml{faire une figure avec un diagrame de bloc et une section proposed approach avec des sub temps/frequence nmfsup nmfsemi}
 
@@
 \subsection{Description of NMF}
-Non-negative Matrix Factorization (NMF) is a matrix approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
+Non-negative Matrix Factorization is a matrix approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
 
@@
 
-\begin{figure}[hbtp]
+\begin{figure}[h]
 \centering
@@
 
-In order to vzlidate the usefulness of considering the NMF framewrok to estimate the road traffic noise level, one need to have a reference level.
-It can hardly de measured or even annotated from real life recordings.
-Thus,  simulated sound scenes are used to assess the performance of the propsed NMF.
+In order to validate the usefulness of considering NMF framework to estimate the road traffic noise level, one need to have a reference level.
+It can hardly be measured or even annotated from real life recordings.
+Thus,  simulated sound scenes are used to assess the performance of the proposed NMF.
 This offers a controlled framework to design specific sound environments in which all the traffic component is known.
@@
 \textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by summing audio samples that come from an isolated sound database.
-This database is divided in two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient, $ii)$ the \textit{background} category includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
-Inside each category, the sound samples are grouped in sound classes (\textit{bird, car, foot steps} \dots), each of them being composed of multiples samples (bird01.wav, bird02.wav \dots).
 
+This database is divided in two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
+154 sound samples belong to this category divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
+The sound class \textit{car passages} comes from 60 recordings of 2 cars made on the Ifsttar's runway on different speeds with multiple gear ratio.
+The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
+Each sound classes are composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
 The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may brings some random behavior between the scenes.
@@
 
-A sound database has been built including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and 154 sound background samples divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
-The sound class \textit{car passages} comes from recordings of 4 cars made on the Ifsttar's runway on different speeds with multiple gear ratio.
-The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
 This database enables creating realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
-A sound mixing corpus is composed of 6 sub-corpuses of 25 audio files each lasting 30 seconds.
+A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds.
 Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult.
-The classes are : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noise) and \textit{transportation} (train, tramway and plane).
+The classes are : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
 In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the sound classes.
 The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class.
-To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$ following the relation (\ref{eq:tpr}).
+To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$ following the relation (\ref{eq:tir}).
 
-\begin{equation}\label{eq:tpr}
-TPR = L_{p,traffic}-L_{p,interfering}
+\begin{equation}\label{eq:tir}
+TIR = L_{p,traffic}-L_{p,interfering}
 \end{equation}
@@
 with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$.
-When $TIR = -12$, the traffic component is then less present than when $TIR = 12$ where it is predominant on the \textit{perturbator} class.
+When $TIR = -12$, the traffic component is then less present than when $TIR = 12$ where it is predominant on the \textit{interfering} class.
 The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated (figure \ref{fig:exampleScene}).
@@
 
-The experiment consists in estimating the traffic road sound level on the environmental sound corpus built using the simulation tool.
-NMF is used and compared to a baseli,e approach which consist in a frequency low-pass filter as road traffic is mainly composed of low frequencies, compared to main of the intefereing.
-this baseline has one parameter being the cut-off frequency $f_c$.
+The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
+First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{14}$ with a 75 $\%$ overlap and a number of point $nfft = 2^{14}$.
+Therefore, the dimensions of $V$ are $F$ = 8192 and $N$ = 772.
 
-The spectrogram $\mathbf{V}$ is built with a window size $w = 2^{14}$ with a 75 $\%$ overlapping and a number of point $nfft = 2^{14}$.
-Therefore, the dimensions of $V$ are $F$ = 8192 and $N$ = 772.
-Similarly, the frequency low-pass filter is applied on the spectrogram $\mathbf{V}$ and on the dictionary $\mathbf{W}$ ( renamed $\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$ respectively).
-This to focus the NMF estimation of the signal on the frequencies where the traffic components are most present.
-The figure \ref{fig:block_diagram_protocol} summarizes the different steps of the process depending on the chosen method, which consist of 3 steps:
+The first estimator to determine the traffic sound level is the basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
+The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (equation \ref{eq:v_tr_filtered}).
 
-\begin{itemize}
-\item the dictionary building for NMF,
-\item the estimation of the traffic sound level according the chosen method (NMF or frequency low-pass filter),
-\item the calculation of the error between the exact and the estimated sound levels.\\
-\end{itemize}
+\begin{equation}\label{eq:v_tr_filtered}
+\mathbf{V}_{traffic} = \mathbf{V}_{f_c}
+\end{equation}
 
-\begin{figure*}[!t]
-\centering
-	\begin{minipage}[t]{.48\textwidth}
-		\centering
-		\includegraphics[width=.9\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}
-	\end{minipage}
-	\begin{minipage}[t]{.48\textwidth}
-		\centering
-		\includegraphics[width=.9\linewidth]{../image/bloc_diagram_NMF_EN.pdf}
-	\end{minipage}
-	\caption{Block diagrams summed up the different step of the precess. On top, for the frequency low-pass filter, on bottom, for NMF.}
-	\label{fig:block_diagram_protocol}
-\end{figure*}
+The second estimator is NMF. Multiples parameters are involved here between the building of the dictionary and the performed NMF.
 
 \subsubsection{Dictionary building}
+The dictionary is built from a second sound database dedicated specifically to this task.
+It is composed of 53 audio files of passing cars.
+These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{simScene} database but with two different cars ().
+First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 75 $\%$ overlap, $nfft$).
+Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
+In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra.
 
-The dictionary is built from a sound database dedicated specifically to this task.
-It is composed of 53 audio files of passing cars.
-First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 75 $\%$ overlap, $nfft$).
-Then a temporal rectangular window \ml{pourquoi pas de fenetrage ?} is applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
-The size of the window is set up at 0.5 and 1 second.
-In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra.
-An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} in the ease of a large window of 4 seconds.
-
-\begin{figure}[!t]
-\centering
-	\begin{minipage}[t]{.24\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{../image/extractionDictionary1Example.pdf}
-	\end{minipage}
-	\begin{minipage}[t]{.24\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{../image/extractionDictionary2Example.pdf}
-	\end{minipage}
-	\caption{Example of the computation of the spectrum with a 1 second window. On left, the original spectrogram with the temporal window, on the left, the 4 spectrum obtained (dB scale)}
-	\label{fig:dictionaryExtraction}
-\end{figure}
-
-Since the number of elements given by porcessing all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
+Since the number of elements given by processing all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
 A special case is added where the root mean square of \textit{all} the spectrogram is applied.
@@
 \ml{pas clair}
-Each element is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ is the $\ell-1$ norm.
+An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} in the cases where $w_t$ = [1 \textit{all}].
+
+\begin{figure}
+    \centering
+\begin{minipage}[t]{\linewidth}
+    \subfigure[]{\label{fig:specW}
+    \includegraphics[width=.45\linewidth]{../image/extractionDictionary1Example.pdf}}
+    \subfigure[]{\label{fig:ElementW}
+    \includegraphics[width=.45\linewidth]{../image/extractionDictionary2Example.pdf}}
+
+    \caption{Example of the creation of the dictionary from the spectrogram of a bird's whistle sample (fig.
+\ref{fig:specW}).
+2 time/frequency window :  with $w_t = 1$ s (fig.
+\ref{fig:ElementW} on left), 3 elements are composed while with $w_t = all$ (fig.
+\ref{fig:ElementW} on right), only one element is generated.}
+    \label{fig:dictionaryExtraction}
+    \end{minipage}
+\end{figure}
+
+Each $k$ element of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ is the $\ell$-1 norm.
 Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.
@@
 \begin{tabular}{cccc}
+Parameter &  \multicolumn{3}{c}{value}\\ \hline
 $K$ & 25  & 50 & 100 \\ \hline
-\begin{tabular}[c]{@{}c@{}}\textbf{temporal} \\ \textbf{window (s)}\end{tabular} & 0.5 & 1  & \textit{all}
+$\mathbf{w_t}$& 0.5 & 1  & \textit{all}
 \end{tabular}
@@
 
-\subsubsection{Estimation of the traffic sound level}
-The traffic sound level is estimated according to the choice of the estimator (a frequency low-pass filter or NMF).
-These methods are applied on the totality of the corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)), for all the cut-off frequencies $f_c$ ([500 1k 2k 5k 10k 20k] Hz) and for 5 $TPR$ ([-12 -6 0 6 12] dB).
+\subsubsection{Performed NMF}
 
-\paragraph{Frequency low-pass filter}\mbox{} \\
-The first estimator to determine the traffic sound level is a basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$.
-
-\paragraph{NMF}\mbox{} \\
-The second estimator is supervised and semi-supervised NMF (see part \ref{part:nmf}).
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands (27 bands).
+NMF inference is performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands (29 bands).
 Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
-NMF inference is performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
 Furthermore, in the case of the linear frequency scale, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
-As a consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
+As a consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$.
+The traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
 
@@
 For the third octave frequency scale, only the case $f_c = 20$ kHz is applied.
-During the iteration process, the estimated equivalent traffic sound level in dB of the entire scene, $\tilde{L}_{p,traffic}$, is calculated as well as the 1 second equivalent sound pressure level, $\tilde{p}_{1s,traffic}$.
-Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
 
-\begin{table*}[]
-\centering
-\begin{tabular}{lcccccc}
-$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
-$\mathbf{TPR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
-\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
-\textbf{method} & \multicolumn{2}{c}{filter} & \multicolumn{2}{c}{supervised NMF} & \multicolumn{2}{c}{semi-supervised NMF} \\ \hline
-$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
-\begin{tabular}[c]{@{}c@{}}\textbf{frequency} \\ \textbf{scale}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}
-\end{tabular}
-\caption{Summary of the different parameters and their values for the estimation of the traffic sound level}
-\label{tab:estimation_factors}
-\end{table*}
-
-\subsubsection{Metrics}
+\subsection{Metrics}
 The performances of the two estimators of the road traffic sound level are assessed through the calculation of two metrics.
@@
 
-In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 6848 settings are performed between the different dictionary $\mathbf{W}$ and the multiple parameters of the estimation step.
+In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 6848 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple parameters taken into account by NMF.
+Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
+
+\begin{table*}[]
+\centering
+\begin{tabular}{lcccccc}
+Parameter &  \multicolumn{6}{c}{value}\\ \hline
+$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
+$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
+\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
+\textbf{method} & \multicolumn{2}{c}{filter} & \multicolumn{2}{c}{supervised NMF} & \multicolumn{2}{c}{semi-supervised NMF} \\ \hline
+$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
+\begin{tabular}[c]{@{}c@{}}\textbf{frequency} \\ \textbf{scale}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}
+\end{tabular}
+\caption{Summary of the different parameters and their values for the estimation of the traffic sound level}
+\label{tab:estimation_factors}
+\end{table*}
 
@@
 \end{tabular}
-\caption{RMSE error for the low pass filter averaged on all the TPR and sub-classes}
+\caption{RMSE error for the low pass filter averaged on all the TIR and sub-classes}
 \label{tab:results_filter}
@@
 
-According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TPR and all the sub-classes.
-The corresponding errors according to the sub-classes and the TPR are summarized in the figure \ref{fig:filterAmbiance}.\\
+According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TIR and all the sub-classes.
+The corresponding errors according to the sub-classes and the TIR are summarized in the figure \ref{fig:filterAmbiance}.\\
 
@@
 \includegraphics[width=\linewidth]{../image/AmbianceFilter.pdf}
-\caption{Bar plot for the filter according to the sub-classes and the TPR at $f_c$ = 500 Hz cut-off frequency}
+\caption{Bar plot for the filter according to the sub-classes and the TIR at $f_c$ = 500 Hz cut-off frequency}
 \label{fig:filterAmbiance}
@@
 
-The error for all the sub-corpus is important for the low TPR (-12 and -6) due to the confusion between the \textit{perturbator} class and the \textit{traffic} class.
+The error for all the sub-corpus is important for the low TIR (-12 and -6) due to the confusion between the \textit{interfering} class and the \textit{traffic} class.
 The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while, for other sub-classes, the error are far more important.
-For the higher \textit{TPR}, the traffic is more present and become dominating on the \textit{perturbator} class.
+For the higher \textit{TIR}, the traffic is more present and become dominating on the \textit{interfering} class.
 The lower error, here, is due to the suppression of the traffic energy by the filter.
-Finally, the error for $f_c = 20$ kHz is equivalent to the one produced when the \textit{traffic} and the \textit{perturbator} classes are take into account together without distinction.
+Finally, the error for $f_c = 20$ kHz is equivalent to the one produced when the \textit{traffic} and the \textit{interfering} classes are take into account together without distinction.
 
@@
 \begin{tabular}{lllllcc}
-$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
+$K$ & \shortstack{temporal\\window (ms)} & domain & $f_c$ (Hz) & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
 \hline
@@
 \end{tabular}
-\caption{RMSE error for supervised NMF averaged on all the TPR and sub-classes}
+\caption{RMSE error for supervised NMF averaged on all the TIR and sub-classes}
 \label{tab:results_supervised}
@@
     & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-TPR & filtre & supervised NMF \\ \hline
+TIR & filtre & supervised NMF \\ \hline
  -12 & 17.20 $\pm$ 4.09 & 15.76 $\pm$ 5.82 \\
@@
 The errors produced by NMF, on the totality of the scenes, is smaller than the filter approach for all the metrics.
-For the higher \textit{TPR}, both methods have similar performances (table \ref{tab:resultsComparison}).
-For the lower \textit{TPR}, TPR, supervised NMF reduces significantly the error.
+For the higher \textit{TIR}, both methods have similar performances (table \ref{tab:resultsComparison}).
+For the lower \textit{TIR}, TIR, supervised NMF reduces significantly the error.
 By distinguishing each class, it is lower for the \textit{alert} and \textit{animals} than the rest sub-classes.
@@
 
-This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TPR} (-6 and 6) until the $\nth {100}$ iterations.\\
+This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TIR} (-6 and 6) until the $\nth {100}$ iterations.\\
 
@@
     \subfigure[]{\label{fig:EarlyStop1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6.pdf}}
-    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:EarlyStop1}) and TPR = 6 (fig. \ref{fig:EarlyStop2})}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6.pdf}}
+    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStop1}) and TIR = 6 (fig. \ref{fig:EarlyStop2})}
     \label{fig:EarlyStop12}
@@
     \subfigure[]{\label{fig:costSup1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR-6Cost.pdf}}
-    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TPR6Cost.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:costSup1}) and TPR = 6 (fig. \ref{fig:costSup2})}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6Cost.pdf}}
+    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6Cost.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSup1}) and TIR = 6 (fig. \ref{fig:costSup2})}
     \label{fig:costSup12}
@@
 
-For the TPR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
+For the TIR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
 Here as the $\mathbf{W}$ is only composed of traffic elements and the sound scenes are composed of a predominant traffic, the reconstruction of the signal is easier.
-In opposite, for the case where \textit{TPR} = -6, for the \textit{alert} signal, the cost function and the $MAE$ error are decreasing too.
+In opposite, for the case where \textit{TIR} = -6, for the \textit{alert} signal, the cost function and the $MAE$ error are decreasing too.
 But for the \textit{climate} and \textit{transportation} sub-classes, the $MAE$ error is increasing.
@@
 This difference can be explained by the type of sound present in these different sub-classes: for \textit{alert} and \textit{animals}, it includes harmonic sounds which belong in the frequency range $\left[2000-5000\right]$ Hz while the other sub-classes include lower frequency sounds.
-As $\mathbf{W}$ is composed of traffic spectrum, located in low frequencies too, it is more difficult to recompose correctly the traffic signal when the \textit{perturbator} sound is predominant and composed of low frequencies too.
+As $\mathbf{W}$ is composed of traffic spectrum, located in low frequencies too, it is more difficult to recompose correctly the traffic signal when the \textit{interfering} sound is predominant and composed of low frequencies too.
 As NMF minimized the cost function (equation \ref{eq:min-D-WH}), \textit{traffic} spectrum then are used to synthesized the global mixture at the expense of the traffic signal.
@@
 \end{tabular}
-\caption{RMSE error for semi-supervised NMF averaged on all the TPR and sub-classes}
+\caption{RMSE error for semi-supervised NMF averaged on all the TIR and sub-classes}
 \label{tab:results_semi_supervised}
@@
 Furthermore, the standard deviation is lower too meaning that the semi-supervised approach offer a much more stable error than the previous method.
-The figure \ref{fig:nmfSemiSupervisedAmbiance} displayed the error for the best scenario averaged on all the sub-classes and the \textit{TPR}.
-The table \ref{tab:comparisonFilterSemiSupervised} summarizes the error expanded to the \textit{TPR}.
+The figure \ref{fig:nmfSemiSupervisedAmbiance} displayed the error for the best scenario averaged on all the sub-classes and the \textit{TIR}.
+The table \ref{tab:comparisonFilterSemiSupervised} summarizes the error expanded to the \textit{TIR}.
 
@@
     & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-TPR & filter & semi-Supervised NMF \\ \hline
+TIR & filter & semi-Supervised NMF \\ \hline
  -12 & 17.20 $\pm$4.09 &  7.46 $\pm$2.88 \\
@@
 
-The error for low TPR is much lower than the filter or supervised NMF.
+The error for low TIR is much lower than the filter or supervised NMF.
 The add of the mobile part makes NMF less constraint and therefore allows to adapt NMF to low traffic contents.
-Figure display the mobile part of the dictionary, $\mathbf{W_r}$ for two particular sub-classes (\textit{alert} and \textit{climate} and for \textit{TPR} = -6.\\
+Figure display the mobile part of the dictionary, $\mathbf{W_r}$ for two particular sub-classes (\textit{alert} and \textit{climate} and for \textit{TIR} = -6.\\
 
@@
 
-Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TPR}.
+Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TIR}.
 Here, NMF uses $\mathbf{W_r}$ in order to minimize the cost function.
@@
 
-These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TPR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
+These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TIR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
 For the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
-Whereas, in \textit{climate} sub-class, the \textit{perturbator} sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
+Whereas, in \textit{climate} sub-class, the \textit{interfering} sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
 This class might be describe with the help of $\mathbf{W_s}$.
 
-For \textit{TPR} = 6, the error is increasing for all sub-classes.
+For \textit{TIR} = 6, the error is increasing for all sub-classes.
 As the objective function is to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
@@
     \subfigure[]{\label{fig:EarlyStopSemi1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR-6.pdf}}
-    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR6.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TPR = 6 (fig. \ref{fig:EarlyStopSemi2})}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6.pdf}}
+    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TIR = 6 (fig. \ref{fig:EarlyStopSemi2})}
     \label{fig:EarlyStopSemi12}
@@
     \subfigure[]{\label{fig:costSemiSup1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR-6Cost.pdf}}
-    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TPR6Cost.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TPR = -6 (fig.  \ref{fig:costSemiSup1}) and TPR = 6 (fig. \ref{fig:costSemiSup2})}
+    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6Cost.pdf}}
+    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6Cost.pdf}}
+    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSemiSup1}) and TIR = 6 (fig. \ref{fig:costSemiSup2})}
     \label{fig:costSemiSup12}

### 2017-12-15 article/articleApplied.tex
Add THC NMF + start improvements in results part

@@
 
-\title{Estimation of the road traffic sound levels based on the Semi-Supervised Non-Negative Matrix Factorization of Magnitude Spectrograms}
+\title{Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization}
 
@@
 Consequently, the traffic sound level estimation based on measurements is not trivial task.
-\ml{parler ici de l'approche des espagnols.
-ref par email}
-
-Many recent works have focused on the detection  or recognition tasks of environmental sounds \cite{heittola_sound_2011},  \cite{defreville_automatic_2006},  \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
+Many recent works have focused on the detection or recognition tasks of environmental sounds without distinction between them\cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
 A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
 A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
-However, most of these results in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
+Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
+If the detection of the road traffic noise is good, the detection of theses anomalous noise events stay weak and no information on the improvement on the estimation of the traffic nous level are presented.
+
+Furthermore, this work and as well as the other works in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
 Although near major roads or ring roads traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels.
@@
 \begin{equation}\label{eq:nmf}
-\mathbf{V} = \mathbf{\tilde{V}} \approx \mathbf{WH}
+\mathbf{V} \approx \mathbf{\tilde{V}} = \mathbf{WH}
 \end{equation}
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.9\linewidth]{../image/illustration_NMF.PNG}
+\includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
 \caption{Example of a simple NMF  for a musical content \cite{bertin_les_2009}}
@@
 \begin{equation}\label{eq:min-D-WH}
-\underset{\mathbf{H} \geq 0, \mathbf{W} \geq 0}{\min} D\left(\mathbf{V} \vert \vert \mathbf{\tilde{V}}\right)
+\underset{\mathbf{H} \geq 0, \mathbf{W} \geq 0}{\min} D\left(\mathbf{V} \vert \vert \mathbf{\tilde{V}}\right).
 \end{equation}
@@
 
-Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen \cite{lee_algorithms_2000} as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
 
@@
 
-\begin{equation}
+\begin{equation}\label{eq:updateH_Sup}
 \textbf{H}^{(i+1)} \leftarrow \textbf{H}^{(i)}.\left(\frac{\textbf{W}^T \left[\left(\textbf{WH}^{(i)} \right)^{(\beta-2)}.\textbf{V} \right]}{\textbf{W}^T \left[\textbf{WH}^{(i)} \right]^{(\beta-1)}}\right)^{\gamma(\beta)}
@@
 The product $A.B$ and $A/B$ symbolized the Hadamard product and ratio.
-The source separation is made by extracting the dictionary and basis elements related to the traffic
+As in the supervised approach, the position in $\mathbf{W}$ of traffic component is known, the source separation of this sound source is made by extracting the dictionary and basis elements related,
 
 \begin{equation}\label{eq:separationExtraction}
-\mathbf{\tilde{V}}_{traffic} = \left[ \mathbf{WH} \right]_{traffic}
+\mathbf{\tilde{V}}_{traffic} = \left[ \mathbf{WH} \right]_{traffic}.
 \end{equation}
@@
 
-One of the main issue with the supervised approach is To better acount for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010}  to offer more flexibility.
+One of the main issue with the supervised approach is
+
+To better take account for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility.
 This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
@@
 \begin{equation}
-\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}
+\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}.
 \end{equation}
@@
 \begin{align}
-\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}.\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}\label{eq:W_r_SS}\\
-\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}.\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}\label{eq:H_r_SS}\\
-\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}.\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}\label{eq:H_s_SS}
+\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}.\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}, \label{eq:W_r_SS}\\
+\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}.\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}, \label{eq:H_r_SS}\\
+\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}.\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}.\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}.\label{eq:H_s_SS}
 \end{align}
@@
 %This approach has been used TO COMPLETE.
+
+\subsection{Thresholded constrained NMF}
+
+A last approach is tested based on unsupervised NMF.
+Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly.
+Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain .
+Then NMF is performed where $\mathbf{W}$ (eq.
+\ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.
+\ref{eq:updateH_Sup}) are updated iteratively.
+
+\begin{equation}\label{eq:updateW_unsup}
+\textbf{W}^{(i+1)} \leftarrow \mathbf{W}^{(i)}.\left(\frac{\left[\left(\mathbf{W}^{(i)}\mathbf{H} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H}^T}{\left[\mathbf{W}^{(i)}\mathbf{H} \right]^{(\beta-1)}\mathbf{H}^T}\right)^{\gamma(\beta)}
+\end{equation}
+
+After $N$ iterations, a measure of similarity $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ between $\mathbf{W_0}$ and the get dictionary $\mathbf{W}$ for each element $k$ is computed through a cosine similarity,
+
+\begin{equation}
+\cos \theta = \frac{\mathbf{W}.\mathbf{W_0}}{\vert \vert \mathbf{W}  \vert \vert . \vert \vert \mathbf{W_0} \vert \vert}.
+\end{equation}
+
+$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are diametrically opposed.
+This measure allows a normalized and an invariant scale estimation of the similarity.
+It is then sort in descending order.
+The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_THC_NMF}).
+It is defined as :
+
+\begin{equation}
+\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W}_{0,k} \vert \vert \mathbf{W}_{k} \right) > t
+\end{equation}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.8\linewidth]{../image/distanceCosLinDisplay.pdf}
+\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
+\label{fig:W_THC_NMF}
+\end{figure}
+
+Other thresholding methods as the \textit{soft} and the \textit{firm} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
+A fast parametric study has revealed that the \textit{hard} tresholding method  with a linear representation of the similarity according to $K$ (see fig.
+\ref{fig:W_THC_NMF}) was the best way to get better performances.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
-First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{14}$ with a 75 $\%$ overlap and a number of point $nfft = 2^{14}$.
-Therefore, the dimensions of $V$ are $F$ = 8192 and $N$ = 772.
+First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$.
+Therefore, the dimensions of $V$ are $F$ = 2049 and $N$ = 664.
 
 The first estimator to determine the traffic sound level is the basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
-The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (equation \ref{eq:v_tr_filtered}).
+The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq.
+\ref{eq:v_tr_filtered}).
 
 \begin{equation}\label{eq:v_tr_filtered}
-\mathbf{V}_{traffic} = \mathbf{V}_{f_c}
+\mathbf{\tilde{V}}_{traffic} = \mathbf{V}_{f_c}.
 \end{equation}
@@
 \subsubsection{Dictionary building}
+
 The dictionary is built from a second sound database dedicated specifically to this task.
@@
 These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{simScene} database but with two different cars ().
-First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 75 $\%$ overlap, $nfft$).
+First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
 Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
+$w_t$ is fixed at $wt = [0.5 1]$ second.
 In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra.
-
-Since the number of elements given by processing all the sound database does not respect the constraint imposed about the dimension of NMF ($F \times K + K \times N < F \times N$), a $K$-medoid clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
+Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
 A special case is added where the root mean square of \textit{all} the spectrogram is applied.
-One spectra is then generated by audio file.
-\ml{pas clair}
-An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} in the cases where $w_t$ = [1 \textit{all}].
+Each audio file generates one element $k$ of $\mathbf{W}$.
+An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig.
+\ref{fig:specW}).
+In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig.
+\ref{fig:ElementW}).
 
@@
     \subfigure[]{\label{fig:specW}
-    \includegraphics[width=.45\linewidth]{../image/extractionDictionary1Example.pdf}}
+    \includegraphics[width=.45\linewidth]{../image/extractionDictionary3Example.pdf}}
     \subfigure[]{\label{fig:ElementW}
-    \includegraphics[width=.45\linewidth]{../image/extractionDictionary2Example.pdf}}
+    \includegraphics[width=.45\linewidth]{../image/extractionDictionary4Example.pdf}}
 
-    \caption{Example of the creation of the dictionary from the spectrogram of a bird's whistle sample (fig.
-\ref{fig:specW}).
-2 time/frequency window :  with $w_t = 1$ s (fig.
-\ref{fig:ElementW} on left), 3 elements are composed while with $w_t = all$ (fig.
-\ref{fig:ElementW} on right), only one element is generated.}
+    \caption{Example of the creation of the dictionary from the spectrogram of a three seconds sample of car passage with two temporal windows ($w_t$ = 1 and $w_t = all$) expressed in a logarithmic scale.}
     \label{fig:dictionaryExtraction}
@@
 
-Each $k$ element of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ is the $\ell$-1 norm.
+Each $k$ element of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.\\
 Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.
@@
 $K$ & 25  & 50 & 100 \\ \hline
-$\mathbf{w_t}$& 0.5 & 1  & \textit{all}
+$\mathbf{w_t}$ (s)& 0.5 & 1  & \textit{all}
 \end{tabular}
@@
 
-NMF inference is performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 2.8$ Hz) and with third octave bands (29 bands).
+Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+THC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary.
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
 Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
-Furthermore, in the case of the linear frequency scale, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
-As a consequence, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$.
-The traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
+Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
+Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
 
 \begin{equation}
-\mathbf{\tilde{V}}_{traffic} = \left[\mathbf{WH}_{f_c}\right]_{traffic}
+\mathbf{\tilde{V}}_{traffic} = \left[\mathbf{WH}_{f_c}\right]_{traffic}.
 \end{equation}
 
-For the third octave frequency scale, only the case $f_c = 20$ kHz is applied.
+With the third octave frequency scale, only the case $f_c = 20$ kHz is applied.
+Finally, for the THC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
+Table \ref{tab:estimation_factors} summarizes the parameters and the related values.
 
-\subsection{Metrics}
+\begin{table*}[]
+\centering
+\begin{tabular}{lcccccc}
+Parameter &  \multicolumn{6}{c}{value}\\ \hline
+$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
+\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
+\begin{tabular}[l]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}\\ \hline
+$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
+$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
+\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & THC NMF \\ \hline
+$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
+\end{tabular}
+\caption{Summary of the different parameters taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
+\label{tab:estimation_factors}
+\end{table*}
+
+\subsubsection{Metrics}
 The performances of the two estimators of the road traffic sound level are assessed through the calculation of two metrics.
@@
 
-In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 6848 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple parameters taken into account by NMF.
-Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
+In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 14040 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF.
+
+\section{Results}\label{part:results}
+
+Table \ref{tab:results} summarized, according to the 3 main parameters (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$.\\
 
@@
 \centering
-\begin{tabular}{lcccccc}
-Parameter &  \multicolumn{6}{c}{value}\\ \hline
-$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
-$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
-\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
-\textbf{method} & \multicolumn{2}{c}{filter} & \multicolumn{2}{c}{supervised NMF} & \multicolumn{2}{c}{semi-supervised NMF} \\ \hline
-$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
-\begin{tabular}[c]{@{}c@{}}\textbf{frequency} \\ \textbf{scale}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}
+\begin{tabular}{cccccccc}
+$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \hline
+ &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
+ &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
+ & & 2 & spectra & 1 & supervised NMF &  & \\ 
+50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
+50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
+50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
+100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
+100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
+100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
+100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
+ &  &  & spectra & 1 & THC NMF &  &  \\ 
+100 & all & 20 & spectra & 2 & THC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
+100 & all & 20 & third octave & 1 & THC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
+100 & all & 20 & third octave & 2 & THC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}
 \end{tabular}
-\caption{Summary of the different parameters and their values for the estimation of the traffic sound level}
-\label{tab:estimation_factors}
+\caption{Best results according to $\beta$, method and  spectral representation}
+\label{tab:results}
 \end{table*}
 
-\section{Results}\label{part:results}
+The first line resumes the error produced by the filter.
+$f_c = 20 $kHz is equivalent to consider all the sound source of the sound mixtures without distinguish the traffic from the others sound sources.
+$f_c = 500$ Hz is the cut-off frequency with the lower error.
+It is then the baseline to refer to compare the performance of NMF.
 
-\ml{la section resultats disctue des resultats, ne pas etre lineaire, genre on fait  puis on fait ca...}
+Then the different version of NMF is summarized with, in bold, the NMF errors lower than the baseline.
+With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
+However, the best combination is got with a THC NMF in a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+As the standard deviation is lower, the case where $\beta = 1$ is selected too.
 
-\subsection{Frequency low-pass filter results}
-In a first time, the frequency low-pass filter estimator is performed on all the scene.
-The table \ref{tab:results_filter} summarized the mean error on the totality of the 750 scenes to find the most efficient cut-off frequency.
-\\
+For each method, the best scenario is detailed according the sub-classes and the $TIR$ (Table \ref{tab:filtre_error}, \ref{tab:Sup_error}, \ref{tab:Semi_error} and \ref{tab:THC_error}).
 
-\begin{table}[h]
+\begin{table*}[]
 \centering
-\begin{tabular}{llll}
-$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$  \\ \hline
- 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} \\
- 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} \\
- 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} \\
- 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 \\
-10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 \\
-20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 \\
+\begin{tabular}{ccccccccc}
+\textbf{method} & $\mathbf{\beta}$ & \textbf{\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular}} & $\mathbf{f_c}$ (kHz) & \textbf{-12} & \textbf{-6} & \textbf{0} & \textbf{6} & \textbf{12} \\ \hline
+filter &  &  & 20 &  &  &  &  &  \\ \hline
+filter &  &  & 0.5 &  &  &  &  &  \\ \hline
+Supervised NMF &  &  &  &  &  &  &  &  \\ \hline
+Semi-supervised NMF &  &  &  &  &  &  &  &  \\ \hline
+THC NMF &  &  &  &  &  &  &  & 
 \end{tabular}
-\caption{RMSE error for the low pass filter averaged on all the TIR and sub-classes}
-\label{tab:results_filter}
-\end{table}
-
-According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TIR and all the sub-classes.
-The corresponding errors according to the sub-classes and the TIR are summarized in the figure \ref{fig:filterAmbiance}.\\
+\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
+\label{ta:results_TIR}
+\end{table*}
 
@@
 \centering
-\includegraphics[width=\linewidth]{../image/AmbianceFilter.pdf}
-\caption{Bar plot for the filter according to the sub-classes and the TIR at $f_c$ = 500 Hz cut-off frequency}
-\label{fig:filterAmbiance}
+\includegraphics[width=\linewidth]{../image/filter_bar.pdf}
+\caption{MAE error for the best cut-off frequency for the filter method according each sub-classes and $TIR$}
+\label{tab:filtre_error}
 \end{figure}
 
-The error for all the sub-corpus is important for the low TIR (-12 and -6) due to the confusion between the \textit{interfering} class and the \textit{traffic} class.
+\begin{figure}[hbtp]
+\centering
+
+\caption{MAE error for the best combination parameters for supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
+\label{tab:Sup_error}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
+\caption{MAE error for the best combination parameters for semi-supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
+\label{tab:Semi_error}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=\linewidth]{../image/THC_bar.pdf}
+\caption{MAE error for the best combination of parameters for THC NMF ($\beta = 2$, third octave) according each sub-classes and $TIR$}
+\label{tab:THC_error}
+\end{figure}
+
+The first approach consider is the frequency low-pass filter.
+With $f_c = 20 kHz$, the error is naturally higher for low $TIR$ (-12,-6) as the traffic is less present.
+In the opposite, for high $TIR$ (6 and 12), as the traffic become preponderant, the traffic  sound level estimation is better as this method conserve all the traffic energy.
+A second cut-off frequency, $f_c = 500 Hz$, allow to have a balance as it put aside the high frequencies and to keep low frequencies.
+It results that for low $TIR$ with high frequencies class sound (\textit{alert}, \textit{animals}), the errors can be low but are high as soon as sound classes are composed of low frequencies (\textit{climate}, \textit{humans}, \textit{transport} and \textit{mechanics}).
+
+Semi-supervised NMF is efficient particularly for low $TIR$ as the mobile part, $W_r$, allows to take into account the other predominant sound sources while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$.
+Nevertheless, these degrees of freedom are restrictive for high $TIR$.
+Indeed, semi-supervised NMF is free to include traffic components in $W_r$.
+The traffic sound level estimation is then less right.
+In the opposite, for $TIR = [6 , 12]$, supervised NMF has to use the traffic element present in $W$.
+In this case, the performances are then better.
+
+Finally, THC NMF with a threshold fixed at $t = 0.54$ is the one where for low and high $TIR$, the performance are lower than the baseline's.
+The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../image/dist_-12_12.pdf}    
+    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
+    \label{fig:dist_-12_12}
+\end{figure}
+
+%\subsection{Frequency low-pass filter results}
+%In a first time, the frequency low-pass filter estimator is performed on all the scene.
+The table \ref{tab:results_filter} summarized the mean error on the totality of the 750 scenes to find the most efficient cut-off frequency.
+\\
+%
+%\begin{table}[h]
+%\centering
+%\begin{tabular}{llll}
+%$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$  \\ \hline
+% 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} \\
+% 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} \\
+% 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} \\
+% 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 \\
+%10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 \\
+%20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 \\
+%\end{tabular}
+%\caption{RMSE error for the low pass filter averaged on all the TIR and sub-classes}
+%\label{tab:results_filter}
+%\end{table}
+%
+%According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TIR and all the sub-classes.
+The corresponding errors according to the sub-classes and the TIR are summarized in the figure \ref{fig:filterAmbiance}.\\
+%
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=\linewidth]{../image/AmbianceFilter.pdf}
+%\caption{Bar plot for the filter according to the sub-classes and the TIR at $f_c$ = 500 Hz cut-off frequency}
+%\label{fig:filterAmbiance}
+%\end{figure}
+%
+%The error for all the sub-corpus is important for the low TIR (-12 and -6) due to the confusion between the \textit{interfering} class and the \textit{traffic} class.
 The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while, for other sub-classes, the error are far more important.
@@
 Finally, the error for $f_c = 20$ kHz is equivalent to the one produced when the \textit{traffic} and the \textit{interfering} classes are take into account together without distinction.
-
-\subsection{Supervised NMF results}
-
-As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ at the $\nth{400}$ iteration are presented in table \ref{tab:results_supervised}.
+%
+%\subsection{Supervised NMF results}
+%
+%As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ at the $\nth{400}$ iteration are presented in table \ref{tab:results_supervised}.
 The bar plot in figure \ref{fig:nmfSupervisedAmbiance} displays the $MAE$ error for the best combination.\\
-
-\begin{table*}[t]
-\centering
-\begin{tabular}{lllllcc}
-$K$ & \shortstack{temporal\\window (ms)} & domain & $f_c$ (Hz) & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
-\hline
- 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} \\
- 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} \\
- 25 & 0 & third octave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34\\
- 25 & 500 & third octave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} \\
-\end{tabular}
-\caption{RMSE error for supervised NMF averaged on all the TIR and sub-classes}
-\label{tab:results_supervised}
-\end{table*}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised.pdf}
-\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 500 Hz )}
-\label{fig:nmfSupervisedAmbiance}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{ccc}
-    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-TIR & filtre & supervised NMF \\ \hline
- -12 & 17.20 $\pm$ 4.09 & 15.76 $\pm$ 5.82 \\
- -6 &  6.91 $\pm$ 2.82 & 6.32 $\pm$ 3.24\\
- 0 & 1.09 $\pm$ 0.16 & 1.18 $\pm$ 0.19\\
-  6 &  1.47 $\pm$ 0.12 & 1.59 $\pm$ 0.13\\
- 12 &  1.60 $\pm$ 0.16 & 1.77 $\pm$ 0.07
-\end{tabular}
-\caption{Comparison of the $MAE$ error for the filter at 500 Hz and the best combinaison of the supervised NMF}
-\label{tab:resultsComparison}
-\end{table}
-
-The best combination for supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, a temporal window of 500 ms and $f_c = 500$ Hz.
+%
+%\begin{table*}[t]
+%\centering
+%\begin{tabular}{lllllcc}
+%$K$ & \shortstack{temporal\\window (ms)} & domain & $f_c$ (Hz) & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
+%\hline
+% 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} \\
+% 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} \\
+% 25 & 0 & third octave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34\\
+% 25 & 500 & third octave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} \\
+%\end{tabular}
+%\caption{RMSE error for supervised NMF averaged on all the TIR and sub-classes}
+%\label{tab:results_supervised}
+%\end{table*}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised.pdf}
+%\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 500 Hz )}
+%\label{fig:nmfSupervisedAmbiance}
+%\end{figure}
+%
+%\begin{table}[h]
+%\centering
+%\begin{tabular}{ccc}
+%    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
+%TIR & filtre & supervised NMF \\ \hline
+% -12 & 17.20 $\pm$ 4.09 & 15.76 $\pm$ 5.82 \\
+% -6 &  6.91 $\pm$ 2.82 & 6.32 $\pm$ 3.24\\
+% 0 & 1.09 $\pm$ 0.16 & 1.18 $\pm$ 0.19\\
+%  6 &  1.47 $\pm$ 0.12 & 1.59 $\pm$ 0.13\\
+% 12 &  1.60 $\pm$ 0.16 & 1.77 $\pm$ 0.07
+%\end{tabular}
+%\caption{Comparison of the $MAE$ error for the filter at 500 Hz and the best combinaison of the supervised NMF}
+%\label{tab:resultsComparison}
+%\end{table}
+%
+%
+%The best combination for supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, a temporal window of 500 ms and $f_c = 500$ Hz.
 The errors produced by NMF, on the totality of the scenes, is smaller than the filter approach for all the metrics.
@@
 \\
-
-This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TIR} (-6 and 6) until the $\nth {100}$ iterations.\\
-
-\begin{figure}
-    \centering
-    \subfigure[]{\label{fig:EarlyStop1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6.pdf}}
-    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStop1}) and TIR = 6 (fig. \ref{fig:EarlyStop2})}
-    \label{fig:EarlyStop12}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfigure[]{\label{fig:costSup1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6Cost.pdf}}
-    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6Cost.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSup1}) and TIR = 6 (fig. \ref{fig:costSup2})}
-    \label{fig:costSup12}
-\end{figure}
-
-For the TIR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
+%
+%This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TIR} (-6 and 6) until the $\nth {100}$ iterations.\\
+%
+%\begin{figure}
+%    \centering
+%    \subfigure[]{\label{fig:EarlyStop1}
+%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6.pdf}}
+%    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6.pdf}}
+%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStop1}) and TIR = 6 (fig. \ref{fig:EarlyStop2})}
+%    \label{fig:EarlyStop12}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \subfigure[]{\label{fig:costSup1}
+%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6Cost.pdf}}
+%    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6Cost.pdf}}
+%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSup1}) and TIR = 6 (fig. \ref{fig:costSup2})}
+%    \label{fig:costSup12}
+%\end{figure}
+%
+%For the TIR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
 Here as the $\mathbf{W}$ is only composed of traffic elements and the sound scenes are composed of a predominant traffic, the reconstruction of the signal is easier.
@@
 As NMF minimized the cost function (equation \ref{eq:min-D-WH}), \textit{traffic} spectrum then are used to synthesized the global mixture at the expense of the traffic signal.
-
-\subsection{Semi-supervised NMF results}
-
-The errors produced for the semi-supervised approach is summarized in table \ref{tab:results_semi_supervised} according to $\beta$ and the frequency domain (spectral or third octave).\\
-
-\begin{table*}
-\centering
-\begin{tabular}{cclllccc}
-$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
-\hline
- 50 & 0 & spectra &  1000 & 1 & \textbf{4.50 $\pm$3.05} & 1.45 $\pm$0.76 \\
-100 & 0 & spectra & 20000 & 2 & \textbf{4.05 $\pm$3.43} & \textbf{\textcolor{red}{1.16 $\pm$0.53}} \\
- 25 & 0 & third octave & 20000 & 1 & \textbf{\textcolor{red}{3.96 $\pm$2.18}} & 1.34 $\pm$0.72 \\
- 25 & 0 & third octave & 20000 & 2 & \textbf{4.29 $\pm$3.58} & \textbf{1.21 $\pm$0.61}
-\end{tabular}
-\caption{RMSE error for semi-supervised NMF averaged on all the TIR and sub-classes}
-\label{tab:results_semi_supervised}
-\end{table*}
-
-The best combination is got in the third octave domain with $\beta = 1$, $f_c$ = 20 kHz, $K$ = 25 and a temporal window null.
+%
+%\subsection{Semi-supervised NMF results}
+%
+%The errors produced for the semi-supervised approach is summarized in table \ref{tab:results_semi_supervised} according to $\beta$ and the frequency domain (spectral or third octave).\\
+%
+%\begin{table*}
+%\centering
+%\begin{tabular}{cclllccc}
+%$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
+%\hline
+% 50 & 0 & spectra &  1000 & 1 & \textbf{4.50 $\pm$3.05} & 1.45 $\pm$0.76 \\
+%100 & 0 & spectra & 20000 & 2 & \textbf{4.05 $\pm$3.43} & \textbf{\textcolor{red}{1.16 $\pm$0.53}} \\
+% 25 & 0 & third octave & 20000 & 1 & \textbf{\textcolor{red}{3.96 $\pm$2.18}} & 1.34 $\pm$0.72 \\
+% 25 & 0 & third octave & 20000 & 2 & \textbf{4.29 $\pm$3.58} & \textbf{1.21 $\pm$0.61}
+%\end{tabular}
+%\caption{RMSE error for semi-supervised NMF averaged on all the TIR and sub-classes}
+%\label{tab:results_semi_supervised}
+%\end{table*}
+%
+%The best combination is got in the third octave domain with $\beta = 1$, $f_c$ = 20 kHz, $K$ = 25 and a temporal window null.
 The error, compared to the supervised approach, is lower.
@@
 The table \ref{tab:comparisonFilterSemiSupervised} summarizes the error expanded to the \textit{TIR}.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=\linewidth]{../image/AmbianceNmfSemiSupervised.pdf}
-\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 20 kHz)}
-\label{fig:nmfSemiSupervisedAmbiance}
-\end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}{cccc}
-    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-TIR & filter & semi-Supervised NMF \\ \hline
- -12 & 17.20 $\pm$4.09 &  7.46 $\pm$2.88 \\
- -6 &  6.91 $\pm$2.82 & 2.15 $\pm$0.70\\
- 0 & 1.09 $\pm$0.16 &  2.20 $\pm$0.45\\
- 6 &  1.47 $\pm$0.12 &  3.61 $\pm$0.29\\
- 12 &  1.60 $\pm$0.16 &  4.02 $\pm$0.33\\
-\end{tabular}
-\caption{Comparison of the baseline and semi-supervised NMF with the best combination}
-\label{tab:comparisonFilterSemiSupervised}
-\end{table}
-
-The error for low TIR is much lower than the filter or supervised NMF.
+%
+%\begin{figure}[hbtp]
+%\centering
+%\includegraphics[width=\linewidth]{../image/AmbianceNmfSemiSupervised.pdf}
+%\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 20 kHz)}
+%\label{fig:nmfSemiSupervisedAmbiance}
+%\end{figure}
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{cccc}
+%    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
+%TIR & filter & semi-Supervised NMF \\ \hline
+% -12 & 17.20 $\pm$4.09 &  7.46 $\pm$2.88 \\
+% -6 &  6.91 $\pm$2.82 & 2.15 $\pm$0.70\\
+% 0 & 1.09 $\pm$0.16 &  2.20 $\pm$0.45\\
+% 6 &  1.47 $\pm$0.12 &  3.61 $\pm$0.29\\
+% 12 &  1.60 $\pm$0.16 &  4.02 $\pm$0.33\\
+%\end{tabular}
+%\caption{Comparison of the baseline and semi-supervised NMF with the best combination}
+%\label{tab:comparisonFilterSemiSupervised}
+%\end{table}
+%
+%The error for low TIR is much lower than the filter or supervised NMF.
 The add of the mobile part makes NMF less constraint and therefore allows to adapt NMF to low traffic contents.
 Figure display the mobile part of the dictionary, $\mathbf{W_r}$ for two particular sub-classes (\textit{alert} and \textit{climate} and for \textit{TIR} = -6.\\
-
-\textbf{figure $W_r$ for alert and climate}\\
-
-Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TIR}.
+%
+%
+%\textbf{figure $W_r$ for alert and climate}\\
+%
+%Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TIR}.
 Here, NMF uses $\mathbf{W_r}$ in order to minimize the cost function.
@@
 \\
-
-\textbf{figure $W_r$ for alert and climate}\\
-
-These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TIR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
+%
+%\textbf{figure $W_r$ for alert and climate}\\
+%
+%These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TIR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
 For the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
@@
 This class might be describe with the help of $\mathbf{W_s}$.
-
-For \textit{TIR} = 6, the error is increasing for all sub-classes.
+%
+%For \textit{TIR} = 6, the error is increasing for all sub-classes.
 As the objective function is to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
@@
 \\
-
-\begin{figure}
-    \centering
-    \subfigure[]{\label{fig:EarlyStopSemi1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6.pdf}}
-    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TIR = 6 (fig. \ref{fig:EarlyStopSemi2})}
-    \label{fig:EarlyStopSemi12}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfigure[]{\label{fig:costSemiSup1}
-    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6Cost.pdf}}
-    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6Cost.pdf}}
-    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSemiSup1}) and TIR = 6 (fig. \ref{fig:costSemiSup2})}
-    \label{fig:costSemiSup12}
-\end{figure}
-
-\textbf{miss the evolution of the Lp for the 25 scenes, estimate and exact + 2/3 scenes to see the evolution during the 30 seconds of the scenes ?}
+%
+%\begin{figure}
+%    \centering
+%    \subfigure[]{\label{fig:EarlyStopSemi1}
+%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6.pdf}}
+%    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6.pdf}}
+%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TIR = 6 (fig. \ref{fig:EarlyStopSemi2})}
+%    \label{fig:EarlyStopSemi12}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \subfigure[]{\label{fig:costSemiSup1}
+%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6Cost.pdf}}
+%    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6Cost.pdf}}
+%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSemiSup1}) and TIR = 6 (fig. \ref{fig:costSemiSup2})}
+%    \label{fig:costSemiSup12}
+%\end{figure}
+%
+%\textbf{miss the evolution of the Lp for the 25 scenes, estimate and exact + 2/3 scenes to see the evolution during the 30 seconds of the scenes ?}
 

### 2017-12-19 article/articleApplied.tex
Following of the results part

@@
 
-\begin{figure*}
-    \centering
-    \subfigure[]{\label{fig:blocDiagramNMF}
-    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
-    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
-    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
-    \label{fig:blocDiagram}
-\end{figure*}
+%\begin{figure*}
+%    \centering
+%    \subfigure[]{\label{fig:blocDiagramNMF}
+%    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
+%    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
+%    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
+%    \label{fig:blocDiagram}
+%\end{figure*}
 
@@
 \includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
-\caption{Example of a simple NMF  for a musical content \cite{bertin_les_2009}}
+\caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
 \label{fig:example_NMF}
@@
 
-The choice of the dimensions should be such that $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used
+The choice of the dimensions is often made as that $F\times K + K \times N < F \times N$.
+NMF is then considered as a low rank approximation method.
+However, this constraint is not essential.
+To estimate the quality of the approximation, an objective function is used
 
@@
 
-One of the main issue with the supervised approach is
-
+One of the main issue with the supervised approach is the generalization issue: how to be adapted to different sound mixtures with a fixed dictionary ?
 To better take account for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility.
@@
 
-%This approach has been used TO COMPLETE.
-
 \subsection{Thresholded constrained NMF}
@@
 Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly.
-Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain .
+Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain (see part \ref{part:dictionary_learning}).
 Then NMF is performed where $\mathbf{W}$ (eq.
 \ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.
-\ref{eq:updateH_Sup}) are updated iteratively.
+\ref{eq:updateH_Sup}) are updated alternatively.
 
@@
 
-$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are diametrically opposed.
-This measure allows a normalized and an invariant scale estimation of the similarity.
-It is then sort in descending order.
-The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_THC_NMF}).
+$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are significantly different.
+This measure allows a bound between 1 and 0 and is an invariant scale estimation of the similarity.
+Then, the similarity is sorted in descending order.
+The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_ThC_NMF}).
 It is defined as :
@@
 \caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
-\label{fig:W_THC_NMF}
+\label{fig:W_ThC_NMF}
 \end{figure}
 
-Other thresholding methods as the \textit{soft} and the \textit{firm} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
+Other thresholding methods as the \textit{soft} \cite{donoho1995noising}and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
 A fast parametric study has revealed that the \textit{hard} tresholding method  with a linear representation of the similarity according to $K$ (see fig.
-\ref{fig:W_THC_NMF}) was the best way to get better performances.
+\ref{fig:W_ThC_NMF}) was the best way to get better performances.
 
@@
 First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$.
-Therefore, the dimensions of $V$ are $F$ = 2049 and $N$ = 664.
+Therefore, the dimensions of $\mathbf{V}$ are $F$ = 2049 and $N$ = 664.
 
@@
 
-\subsubsection{Dictionary building}
+\subsubsection{Dictionary building}\label{part:dictionary_learning}
 
@@
 It is composed of 53 audio files of passing cars.
-These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{simScene} database but with two different cars ().
+These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars ().
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
@@
 
-\begin{figure}
+\begin{figure}[t!]
     \centering
-\begin{minipage}[t]{\linewidth}
-    \subfigure[]{\label{fig:specW}
-    \includegraphics[width=.45\linewidth]{../image/extractionDictionary3Example.pdf}}
-    \subfigure[]{\label{fig:ElementW}
-    \includegraphics[width=.45\linewidth]{../image/extractionDictionary4Example.pdf}}
-
-    \caption{Example of the creation of the dictionary from the spectrogram of a three seconds sample of car passage with two temporal windows ($w_t$ = 1 and $w_t = all$) expressed in a logarithmic scale.}
-    \label{fig:dictionaryExtraction}
-    \end{minipage}
+    \begin{subfigure}[t]{0.47\linewidth}
+        \centering
+       \includegraphics[width=\linewidth]{../image/dictionary1.pdf}
+        \caption{Sample of 3 seconds of a car passage}
+        \label{fig:TIR_class_filter}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.47\linewidth}
+        \centering
+       \includegraphics[width=\linewidth]{../image/dictionary2.pdf}
+        \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
+        \label{fig:TIR_class_filter}
+    \end{subfigure}
 \end{figure}
@@
 Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
-THC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary.
+ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary.
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
@@
 With the third octave frequency scale, only the case $f_c = 20$ kHz is applied.
-Finally, for the THC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
+Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
 Table \ref{tab:estimation_factors} summarizes the parameters and the related values.
@@
 $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
-\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & THC NMF \\ \hline
+\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & ThC NMF \\ \hline
 $\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
@@
 \centering
-\begin{tabular}{cccccccc}
-$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \hline
+\begin{tabular}{@{}cccccccc@{}}
+\toprule
+$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
  &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
  &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
- & & 2 & spectra & 1 & supervised NMF &  & \\ 
+25 & all & 0.5 & spectra & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
 50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
@@
 100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
- &  &  & spectra & 1 & THC NMF &  &  \\ 
-100 & all & 20 & spectra & 2 & THC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
-100 & all & 20 & third octave & 1 & THC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
-100 & all & 20 & third octave & 2 & THC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}
+ &  &  & spectra & 1 & ThC NMF &  &  \\ 
+100 & all & 20 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
+100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
+100 & all & 20 & third octave & 2 & ThC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
 \end{tabular}
@@
 
-The first line resumes the error produced by the filter.
-$f_c = 20 $kHz is equivalent to consider all the sound source of the sound mixtures without distinguish the traffic from the others sound sources.
-$f_c = 500$ Hz is the cut-off frequency with the lower error.
+The two first lines resume the error produced by the filter.
+$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguish the traffic from the others sound sources.
+Consequently , in low $TIR$ (-12 and -6), where traffic component is discreet, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant.
+$f_c = 500$ Hz is the cut-off frequency with the lower mean error.
 It is then the baseline to refer to compare the performance of NMF.
+In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the impact of the filter is quite interesting as it suppress theses frequency components whereas for the other sub-classes where low frequencies are presents, the error as higher as the filter cannot dissociate the traffic element from the other sound sources.
+The sounds levels are then overestimate (figure \ref{fig:alert_-12}).
+In opposite, in high $TIR$, the low-pass filter removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\
 
-Then the different version of NMF is summarized with, in bold, the NMF errors lower than the baseline.
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_-12.pdf}
+        \caption{}
+        \label{fig:alert_-12}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_12.pdf}
+        \caption{}
+        \label{fig:alert_12}
+    \end{subfigure}
+
+    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the sub-classes \textit{alert}: at $TIR = -12$ (\ref{fig:alert_-12}) and at $TIR$ (\ref{fig:alert_12}).}
+    \label{fig:dictionaryExtraction}
+\end{figure*}
+
+The results of different versions of NMF are summarized next to the filter results with, in bold, the NMF errors lower than the baseline.
 With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
-However, the best combination is got with a THC NMF in a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
-As the standard deviation is lower, the case where $\beta = 1$ is selected too.
+However, the best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).
 
-For each method, the best scenario is detailed according the sub-classes and the $TIR$ (Table \ref{tab:filtre_error}, \ref{tab:Sup_error}, \ref{tab:Semi_error} and \ref{tab:THC_error}).
-
-\begin{table*}[]
+\begin{table*}[t]
 \centering
-\begin{tabular}{ccccccccc}
-\textbf{method} & $\mathbf{\beta}$ & \textbf{\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular}} & $\mathbf{f_c}$ (kHz) & \textbf{-12} & \textbf{-6} & \textbf{0} & \textbf{6} & \textbf{12} \\ \hline
-filter &  &  & 20 &  &  &  &  &  \\ \hline
-filter &  &  & 0.5 &  &  &  &  &  \\ \hline
-Supervised NMF &  &  &  &  &  &  &  &  \\ \hline
-Semi-supervised NMF &  &  &  &  &  &  &  &  \\ \hline
-THC NMF &  &  &  &  &  &  &  & 
+\begin{tabular}{@{}cccccc@{}}
+\toprule
+method & filter & filter & supervised NMF & semi-supervised NMF & ThC NMF \\ \midrule
+$\beta$ &  &  & 2 & 2 & 2 \\
+\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular} &  &  & spectra & spectra & third octave \\
+$f_c$ (kHz) & 20 & 0.5 & 0.5 & 20 & 20 \\ \hline
+-12 & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & \textbf{6.23 ($\pm$ 3.19)} &  & \textbf{5.11 ($\pm$ 3.10)} \\
+-6 & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & \textbf{3.00 ($\pm$ 1.39)} &  & \textbf{2.87 ($\pm$ 1.55)} \\
+0 & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
+6 & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & \textbf{1.01 ($\pm$ 0.28)} &  & \textbf{0.70 ($\pm$ 0.32)} \\
+12 & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & \textbf{1.41 ($\pm$ 0.10)} &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
 \end{tabular}
@@
 
-\begin{figure}[hbtp]
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
         \centering
         \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
-\caption{MAE error for the best cut-off frequency for the filter method according each sub-classes and $TIR$}
-\label{tab:filtre_error}
+        \caption{filter method with $f_c$ = 500 Hz}
+        \label{fig:TIR_class_filter}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
+        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
+                \label{fig:TIR_class_sup}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
+        \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
+                \label{fig:TIR_class_semi}
+    \end{subfigure}%
+    \hfill 
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/THC_bar}
+        \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
+        \label{fig:TIR_class_thc}
+    \end{subfigure}
+    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter and each method (supervised, semi-supervised and ThC}
+\end{figure*}
+
+In the case of supervised NMF, on all the $TIR$, if the mean score is lower than the baseline method, it offers similar performances as for $TIR = [0, 6, 12]$, the mean score are nearly close.
+The performance of supervised NMF is mainly visible for $TIR = -12$ where the error with the baseline decrease by more than 1 dB.
+
+The gain of the semi-supervised approach is most of all significant for these low $TIR$ where the add of a mobile part, $\mathbf{W_r}$, into $\mathbf{W}$ allows to take into account the other predominant sound sources (example in Figure \ref{fig:Y_alert-12}) while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$.
+This improvement, according to supervised NMF results, can be seen in Figure \ref{fig:TIR_class_semi} where the error on each sub class for $TIR = -12$ and $TIR$ = -6 are much lower.
+Nevertheless, these degrees of freedom are restrictive for high $TIR$.
+Indeed, without constraint semi-supervised NMF is free to include traffic components in $W_r$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}).
+The error are then increasing for theses $TIR$.
+In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances.
+
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
+        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = -12}
+        \label{fig:Y_alert-12}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
+        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = 12}
+		\label{fig:Y_alert_12}
+    \end{subfigure}
 \end{figure}
 
-\begin{figure}[hbtp]
-\centering
-
-\caption{MAE error for the best combination parameters for supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
-\label{tab:Sup_error}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
-\caption{MAE error for the best combination parameters for semi-supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
-\label{tab:Semi_error}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=\linewidth]{../image/THC_bar.pdf}
-\caption{MAE error for the best combination of parameters for THC NMF ($\beta = 2$, third octave) according each sub-classes and $TIR$}
-\label{tab:THC_error}
-\end{figure}
-
-The first approach consider is the frequency low-pass filter.
-With $f_c = 20 kHz$, the error is naturally higher for low $TIR$ (-12,-6) as the traffic is less present.
-In the opposite, for high $TIR$ (6 and 12), as the traffic become preponderant, the traffic  sound level estimation is better as this method conserve all the traffic energy.
-A second cut-off frequency, $f_c = 500 Hz$, allow to have a balance as it put aside the high frequencies and to keep low frequencies.
-It results that for low $TIR$ with high frequencies class sound (\textit{alert}, \textit{animals}), the errors can be low but are high as soon as sound classes are composed of low frequencies (\textit{climate}, \textit{humans}, \textit{transport} and \textit{mechanics}).
-
-Semi-supervised NMF is efficient particularly for low $TIR$ as the mobile part, $W_r$, allows to take into account the other predominant sound sources while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$.
-Nevertheless, these degrees of freedom are restrictive for high $TIR$.
-Indeed, semi-supervised NMF is free to include traffic components in $W_r$.
-The traffic sound level estimation is then less right.
-In the opposite, for $TIR = [6 , 12]$, supervised NMF has to use the traffic element present in $W$.
-In this case, the performances are then better.
-
-Finally, THC NMF with a threshold fixed at $t = 0.54$ is the one where for low and high $TIR$, the performance are lower than the baseline's.
+Finally, ThC NMF with a threshold fixed at $t = 0.54$ offers the most lower results.
 The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.
@@
 
-%\subsection{Frequency low-pass filter results}
-%In a first time, the frequency low-pass filter estimator is performed on all the scene.
-The table \ref{tab:results_filter} summarized the mean error on the totality of the 750 scenes to find the most efficient cut-off frequency.
-\\
-%
-%\begin{table}[h]
-%\centering
-%\begin{tabular}{llll}
-%$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$  \\ \hline
-% 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} \\
-% 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} \\
-% 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} \\
-% 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 \\
-%10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 \\
-%20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 \\
-%\end{tabular}
-%\caption{RMSE error for the low pass filter averaged on all the TIR and sub-classes}
-%\label{tab:results_filter}
-%\end{table}
-%
-%According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TIR and all the sub-classes.
-The corresponding errors according to the sub-classes and the TIR are summarized in the figure \ref{fig:filterAmbiance}.\\
-%
-%\begin{figure}[t]
-%\centering
-%\includegraphics[width=\linewidth]{../image/AmbianceFilter.pdf}
-%\caption{Bar plot for the filter according to the sub-classes and the TIR at $f_c$ = 500 Hz cut-off frequency}
-%\label{fig:filterAmbiance}
-%\end{figure}
-%
-%The error for all the sub-corpus is important for the low TIR (-12 and -6) due to the confusion between the \textit{interfering} class and the \textit{traffic} class.
-The error is less for the \textit{alert} and \textit{humans} sub-corpus as it is composed of higher frequencies while, for other sub-classes, the error are far more important.
-For the higher \textit{TIR}, the traffic is more present and become dominating on the \textit{interfering} class.
-The lower error, here, is due to the suppression of the traffic energy by the filter.
-Finally, the error for $f_c = 20$ kHz is equivalent to the one produced when the \textit{traffic} and the \textit{interfering} classes are take into account together without distinction.
-%
-%\subsection{Supervised NMF results}
-%
-%As it is not possible to summarize all the parameter combinations, just the best results according to the domain and $\beta$ at the $\nth{400}$ iteration are presented in table \ref{tab:results_supervised}.
-The bar plot in figure \ref{fig:nmfSupervisedAmbiance} displays the $MAE$ error for the best combination.\\
-%
-%\begin{table*}[t]
-%\centering
-%\begin{tabular}{lllllcc}
-%$K$ & \shortstack{temporal\\window (ms)} & domain & $f_c$ (Hz) & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
-%\hline
-% 25 & 500 & spectra &  2000 & 1 & \textbf{6.18 $\pm$7.49} & \textbf{1.48 $\pm$1.20} \\
-% 25 & 500 & spectra &   500 & 2 & \textbf{\textcolor{red}{5.32 $\pm$6.28}} & \textbf{1.40 $\pm$0.93} \\
-% 25 & 0 & third octave & 20000 & 1 & \textbf{6.82 $\pm$8.16} & 1.60 $\pm$1.34\\
-% 25 & 500 & third octave & 20000 & 2 & \textbf{6.26 $\pm$7.69} & \textbf{\textcolor{red}{1.36 $\pm$1.19}} \\
-%\end{tabular}
-%\caption{RMSE error for supervised NMF averaged on all the TIR and sub-classes}
-%\label{tab:results_supervised}
-%\end{table*}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=\linewidth]{../image/AmbianceNmfSupervised.pdf}
-%\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 500 Hz )}
-%\label{fig:nmfSupervisedAmbiance}
-%\end{figure}
-%
-%\begin{table}[h]
-%\centering
-%\begin{tabular}{ccc}
-%    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-%TIR & filtre & supervised NMF \\ \hline
-% -12 & 17.20 $\pm$ 4.09 & 15.76 $\pm$ 5.82 \\
-% -6 &  6.91 $\pm$ 2.82 & 6.32 $\pm$ 3.24\\
-% 0 & 1.09 $\pm$ 0.16 & 1.18 $\pm$ 0.19\\
-%  6 &  1.47 $\pm$ 0.12 & 1.59 $\pm$ 0.13\\
-% 12 &  1.60 $\pm$ 0.16 & 1.77 $\pm$ 0.07
-%\end{tabular}
-%\caption{Comparison of the $MAE$ error for the filter at 500 Hz and the best combinaison of the supervised NMF}
-%\label{tab:resultsComparison}
-%\end{table}
-%
-%
-%The best combination for supervised NMF is the one with described in the \textit{spectra} domain and for the euclidean distance with $K = 25$, a temporal window of 500 ms and $f_c = 500$ Hz.
-The errors produced by NMF, on the totality of the scenes, is smaller than the filter approach for all the metrics.
-For the higher \textit{TIR}, both methods have similar performances (table \ref{tab:resultsComparison}).
-For the lower \textit{TIR}, TIR, supervised NMF reduces significantly the error.
-By distinguishing each class, it is lower for the \textit{alert} and \textit{animals} than the rest sub-classes.
-\\
-%
-%This difference can be illustrate through the evolution of the cost function (figures \ref{fig:costSup12}) and of the $MAE$ error (figure \ref{fig:EarlyStop12}) for 3 sub-classes (\textit{alert}, \textit{climate} and \textit{transportation}), for 2 \textit{TIR} (-6 and 6) until the $\nth {100}$ iterations.\\
-%
-%\begin{figure}
-%    \centering
-%    \subfigure[]{\label{fig:EarlyStop1}
-%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6.pdf}}
-%    \subfigure[]{\label{fig:EarlyStop2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6.pdf}}
-%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStop1}) and TIR = 6 (fig. \ref{fig:EarlyStop2})}
-%    \label{fig:EarlyStop12}
-%\end{figure}
-%
-%\begin{figure}
-%    \centering
-%    \subfigure[]{\label{fig:costSup1}
-%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR-6Cost.pdf}}
-%    \subfigure[]{\label{fig:costSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSupervised_EarlyStop_TIR6Cost.pdf}}
-%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSup1}) and TIR = 6 (fig. \ref{fig:costSup2})}
-%    \label{fig:costSup12}
-%\end{figure}
-%
-%For the TIR = 6, the cost function and the $MAE$ error are decreasing, meaning the global mixture and the traffic signal are well synthesized.
-Here as the $\mathbf{W}$ is only composed of traffic elements and the sound scenes are composed of a predominant traffic, the reconstruction of the signal is easier.
-In opposite, for the case where \textit{TIR} = -6, for the \textit{alert} signal, the cost function and the $MAE$ error are decreasing too.
-But for the \textit{climate} and \textit{transportation} sub-classes, the $MAE$ error is increasing.
-Even if the cost function decreases, the quality of the traffic signal rebuilt is not improved.
-This difference can be explained by the type of sound present in these different sub-classes: for \textit{alert} and \textit{animals}, it includes harmonic sounds which belong in the frequency range $\left[2000-5000\right]$ Hz while the other sub-classes include lower frequency sounds.
-As $\mathbf{W}$ is composed of traffic spectrum, located in low frequencies too, it is more difficult to recompose correctly the traffic signal when the \textit{interfering} sound is predominant and composed of low frequencies too.
-As NMF minimized the cost function (equation \ref{eq:min-D-WH}), \textit{traffic} spectrum then are used to synthesized the global mixture at the expense of the traffic signal.
-%
-%\subsection{Semi-supervised NMF results}
-%
-%The errors produced for the semi-supervised approach is summarized in table \ref{tab:results_semi_supervised} according to $\beta$ and the frequency domain (spectral or third octave).\\
-%
-%\begin{table*}
-%\centering
-%\begin{tabular}{cclllccc}
-%$K$ & \shortstack{temporal\\window (ms)} & $f_c$ (Hz) & domain & $\beta$ & $MAE$ (dB) & $nRMSE$ \\
-%\hline
-% 50 & 0 & spectra &  1000 & 1 & \textbf{4.50 $\pm$3.05} & 1.45 $\pm$0.76 \\
-%100 & 0 & spectra & 20000 & 2 & \textbf{4.05 $\pm$3.43} & \textbf{\textcolor{red}{1.16 $\pm$0.53}} \\
-% 25 & 0 & third octave & 20000 & 1 & \textbf{\textcolor{red}{3.96 $\pm$2.18}} & 1.34 $\pm$0.72 \\
-% 25 & 0 & third octave & 20000 & 2 & \textbf{4.29 $\pm$3.58} & \textbf{1.21 $\pm$0.61}
-%\end{tabular}
-%\caption{RMSE error for semi-supervised NMF averaged on all the TIR and sub-classes}
-%\label{tab:results_semi_supervised}
-%\end{table*}
-%
-%The best combination is got in the third octave domain with $\beta = 1$, $f_c$ = 20 kHz, $K$ = 25 and a temporal window null.
-The error, compared to the supervised approach, is lower.
-Furthermore, the standard deviation is lower too meaning that the semi-supervised approach offer a much more stable error than the previous method.
-The figure \ref{fig:nmfSemiSupervisedAmbiance} displayed the error for the best scenario averaged on all the sub-classes and the \textit{TIR}.
-The table \ref{tab:comparisonFilterSemiSupervised} summarizes the error expanded to the \textit{TIR}.
-%
-%\begin{figure}[hbtp]
-%\centering
-%\includegraphics[width=\linewidth]{../image/AmbianceNmfSemiSupervised.pdf}
-%\caption{bar plot for the best parameter combination of the supervised NMF ($K = 25$, temporal window = 0.5 s, domain = spectra, $\beta$ = 2, $f_c$ = 20 kHz)}
-%\label{fig:nmfSemiSupervisedAmbiance}
-%\end{figure}
-%
-%\begin{table}
-%\centering
-%\begin{tabular}{cccc}
-%    & \multicolumn{2}{c}{$MAE$ (dB)} \\ \hline
-%TIR & filter & semi-Supervised NMF \\ \hline
-% -12 & 17.20 $\pm$4.09 &  7.46 $\pm$2.88 \\
-% -6 &  6.91 $\pm$2.82 & 2.15 $\pm$0.70\\
-% 0 & 1.09 $\pm$0.16 &  2.20 $\pm$0.45\\
-% 6 &  1.47 $\pm$0.12 &  3.61 $\pm$0.29\\
-% 12 &  1.60 $\pm$0.16 &  4.02 $\pm$0.33\\
-%\end{tabular}
-%\caption{Comparison of the baseline and semi-supervised NMF with the best combination}
-%\label{tab:comparisonFilterSemiSupervised}
-%\end{table}
-%
-%The error for low TIR is much lower than the filter or supervised NMF.
-The add of the mobile part makes NMF less constraint and therefore allows to adapt NMF to low traffic contents.
-Figure display the mobile part of the dictionary, $\mathbf{W_r}$ for two particular sub-classes (\textit{alert} and \textit{climate} and for \textit{TIR} = -6.\\
-%
-%
-%\textbf{figure $W_r$ for alert and climate}\\
-%
-%Nevertheless, the advantage won with the mobile part, generates larger errors in higher \textit{TIR}.
-Here, NMF uses $\mathbf{W_r}$ in order to minimize the cost function.
-As the mobile part is not constraint, traffic content is included in it, decreasing the quality of the traffic signal reconstruction.
-Figure display the 2 basis of $\mathbf{W_r}$.
-\\
-%
-%\textbf{figure $W_r$ for alert and climate}\\
-%
-%These behaviors can be noticed through the $MAE$ evolution in figures \ref{fig:EarlyStopSemi12}: for low TIR the error is decreasing like the cost function (figure \ref{fig:costSemiSup1}) meaning that the synthesis of the traffic signal is good.
-For the \textit{alert} sub-class, $\mathbf{W_r}$ enables to include the harmonic component which are not present in $\mathbf{W_s}$, the improvement bring by the semi-supervised approach is then remarkable.
-Whereas, in \textit{climate} sub-class, the \textit{interfering} sound classes are composed of low-frequencies that are present in $\mathbf{W_s}$.
-This class might be describe with the help of $\mathbf{W_s}$.
-%
-%For \textit{TIR} = 6, the error is increasing for all sub-classes.
-As the objective function is to minimize the distance between the spectrogram $\mathbf{V}$ and $\mathbf{WH}$, the semi-supervised approach is free to include traffic component in $\mathbf{W_r}$ to do it.
-This degree of freedom deteriorates the reconstruction of the traffic signal.
-\\
-%
-%\begin{figure}
-%    \centering
-%    \subfigure[]{\label{fig:EarlyStopSemi1}
-%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6.pdf}}
-%    \subfigure[]{\label{fig:EarlyStopSemi2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6.pdf}}
-%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:EarlyStopSemi1}) and TIR = 6 (fig. \ref{fig:EarlyStopSemi2})}
-%    \label{fig:EarlyStopSemi12}
-%\end{figure}
-%
-%\begin{figure}
-%    \centering
-%    \subfigure[]{\label{fig:costSemiSup1}
-%    \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR-6Cost.pdf}}
-%    \subfigure[]{\label{fig:costSemiSup2} \includegraphics[width=0.45\linewidth]{../image/AmbianceNmfSemiSupervised_EarlyStop_TIR6Cost.pdf}}
-%    \caption{Evolution of the $MAE$ for 3 sub-classes for TIR = -6 (fig.  \ref{fig:costSemiSup1}) and TIR = 6 (fig. \ref{fig:costSemiSup2})}
-%    \label{fig:costSemiSup12}
-%\end{figure}
-%
-%\textbf{miss the evolution of the Lp for the 25 scenes, estimate and exact + 2/3 scenes to see the evolution during the 30 seconds of the scenes ?}
+The gain is significant on all $TIR$ with the exception of $TIR = 0$. On figure \ref{fig:ThC_error}, we can see the improvement of .
+While supervised NMF have high errors on low $TIR$ for the sub-classes \textit{human} and \textit{transport}, ThC NMF succeed to have error inferior to 4 dB.
+The error for \textit{climate} and \textit{mechanics} are, on the other hand, similar to those of supervised NMF.
 

### 2017-12-19 article/articleApplied.tex
following update

@@
 \end{equation}
-
-%or by generating a soft mask, $\mathbf{M_s}$,
-%
-%\begin{equation} \label{eq:definitionMask}
-%\mathbf{M_s} = \frac{\left[\mathbf{W} \mathbf{H}\right]_{traffic}}{\mathbf{W H}},
-%\end{equation}
-%
-%which allows to determine the spectrogram of the traffic component, $ \mathbf{\tilde{V}}_{traffic}$
-%
-%\begin{equation}\label{eq:separation}
-%\mathbf{\tilde{V}}_{traffic} = \mathbf{M_s}.\mathbf{V}.
-%\end{equation}
 
@@
 ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different spectral representation: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
 Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
@@
 
-With the third octave frequency scale, only the case $f_c = 20$ kHz is applied.
+With the third octave spectral representation, only the case $f_c = 20$ kHz is applied.
 Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
@@
 100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
- &  &  & spectra & 1 & ThC NMF &  &  \\ 
-100 & all & 20 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
+50 & all & 5 & spectra & 1 & ThC NMF & 0.39 & \textbf{2.23 ($\pm$ 2.21)} \\ 
+50 & all & 2 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.17 ($\pm$ 2.28)} \\ 
 100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
@@
 
-The results of different versions of NMF are summarized next to the filter results with, in bold, the NMF errors lower than the baseline.
+The results of different versions of NMF are summarized next to the filter results with the NMF errors lower than the baseline in bold.
 With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
-However, the best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+The best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+
+Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, no parameters is common to all the settings.
 For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \centering
@@
         \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
-        \caption{filter method with $f_c$ = 500 Hz}
+        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
         \label{fig:TIR_class_filter}
@@
 Nevertheless, these degrees of freedom are restrictive for high $TIR$.
-Indeed, without constraint semi-supervised NMF is free to include traffic components in $W_r$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}).
+Indeed, without constraint semi-supervised NMF is free to include traffic components in $\mathbf{W_r}$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}).
 The error are then increasing for theses $TIR$.
 In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances.
+\\
 
-\begin{figure}[t!]
+\begin{figure*}[t]
     \centering
@@
     \end{subfigure}
-\end{figure}
+\end{figure*}
 
-Finally, ThC NMF with a threshold fixed at $t = 0.54$ offers the most lower results.
+Finally, ThC NMF with a hard threshold fixed at $t = 0.54$ offers the lowest results.
 The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 
-The gain is significant on all $TIR$ with the exception of $TIR = 0$. On figure \ref{fig:ThC_error}, we can see the improvement of .
-While supervised NMF have high errors on low $TIR$ for the sub-classes \textit{human} and \textit{transport}, ThC NMF succeed to have error inferior to 4 dB.
-The error for \textit{climate} and \textit{mechanics} are, on the other hand, similar to those of supervised NMF.
+The gain is significant on all $TIR$ with the exception of $TIR = 0$.
+On Figure \ref{fig:TIR_class_thc}, we can see the improvement bringing by this approach.
+For low $TIR$, only a reduce number of $\mathbf{W}$ are considered as traffic components, the performance on \textit{alert}, \textit{animals}, \textit{humans} and \textit{tranportation} are then similar to semi-supervised NMF.
+However, for \textit{climate} and \textit{mechanics}, the error are important as in supervised NMF.
+On high $TIR$, the similarity of the initial dictionary and $\mathbf{W}$ allows to keep more elements.
+In these cases, ThC NMF is the most powerful of the proposed methods.
+Consequently, if the average error is lower than semi-supervised NMF, with good performances on high $TIR$, it is a more variable method.
 

### 2017-12-20 article/articleApplied.tex
continuation of the results part

@@
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-Theses maps allow to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+These maps allow to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
 These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
@@
 
-%\begin{figure*}
-%    \centering
-%    \subfigure[]{\label{fig:blocDiagramNMF}
-%    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
-%    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
-%    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
-%    \label{fig:blocDiagram}
-%\end{figure*}
-
 Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels.
@@
 It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
-In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
+In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
 
@@
 Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
-If the detection of the road traffic noise is good, the detection of theses anomalous noise events stay weak and no information on the improvement on the estimation of the traffic nous level are presented.
+If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic nous level are presented.
 
@@
 
-We propose in this paper a method based on the Non-Negative Matrix Factorization (NMF) technique to estimate the global, $\tilde{L}_{p,traffic}$, and the 1-s equivalent, $\tilde{p}_{1s,traffic}$ sound level of the traffic through the supervised and the semi-supervised approaches.
-To validate this approach, we consider a corpus of simulated scenes artificially created with the simulator software \textit{simScene}.
+We propose in this paper a method based on the Non-Negative Matrix Factorization (NMF) technique to estimate the global, $\tilde{L}_{p,traffic}$, and the 1-s equivalent, $\tilde{p}_{1s,traffic}$ sound level of the traffic through the supervised and the semi-supervised approaches as well as a method using thresholding.
+To validate these approaches, we consider a corpus of simulated scenes artificially created with the simulator software \textit{simScene}.
 The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of an urban scene ($L_{p,traffic}$ and $p_{1s,traffic}$).
@@
 
+\begin{figure*}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}
+        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
+        \label{fig:TIR_class_filter}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}
+        \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
+		\label{fig:TIR_class_sup}
+    \end{subfigure}
+\end{figure*}
+
 The remaining of the paper is organized as follows.
@@
 Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
-
-%\ml{faire une figure avec un diagrame de bloc et une section proposed approach avec des sub temps/frequence nmfsup nmfsemi}
 
@@
 One of the main issue with the supervised approach is the generalization issue: how to be adapted to different sound mixtures with a fixed dictionary ?
-To better take account for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility.
+To better take into account the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility.
 This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
@@
 \ref{eq:updateH_Sup}) are updated alternatively.
+$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge.
 
@@
 
+This approach is named \textit{Thresholded constraint NMF} (ThC NMF).
 Other thresholding methods as the \textit{soft} \cite{donoho1995noising}and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
-A fast parametric study has revealed that the \textit{hard} tresholding method  with a linear representation of the similarity according to $K$ (see fig.
+A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$ (see fig.
 \ref{fig:W_ThC_NMF}) was the best way to get better performances.
@@
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound classes are composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
-The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may brings some random behavior between the scenes.
+Each sound classes is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
+The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
 Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
@@
 The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated (figure \ref{fig:exampleScene}).
-Finally, the number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TPR values).
+The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TIR values).
 
@@
 
-The first estimator to determine the traffic sound level is the basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
+The first estimator to determine the traffic sound level is a basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
 The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-The second estimator is NMF. Multiples parameters are involved here between the building of the dictionary and the performed NMF.
+The second estimator is NMF. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.
 
@@
         \caption{Sample of 3 seconds of a car passage}
-        \label{fig:TIR_class_filter}
+        \label{fig:specW}
     \end{subfigure}%
@@
         \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
-        \label{fig:TIR_class_filter}
+        \label{fig:ElementW}
     \end{subfigure}
@@
 Each $k$ element of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.\\
+
 Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.
@@
 
-\subsubsection{Performed NMF}
+\subsubsection{NMF parameters}
 
 Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
-ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different spectral representation: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
-Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.
+ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different spectral representations: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
+These two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins in the high frequencies where the traffic component is less present and generates less calculations.
 Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
@@
 With the third octave spectral representation, only the case $f_c = 20$ kHz is applied.
-Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
-Table \ref{tab:estimation_factors} summarizes the parameters and the related values.
+For the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
+Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
 
@@
 \subsubsection{Metrics}
-The performances of the two estimators of the road traffic sound level are assessed through the calculation of two metrics.
-
-\begin{itemize}
-
-\item The Mean Absolute Error, $MAE$, expresses the quality of the long-term reconstruction of the signal.
+The performances of the two estimators of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
+It expresses the quality of the long-term reconstruction of the signal.
 It consists in the average over the $N$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
@@
 
-\item The normalized short-term Root Mean Square Error, $nRMSE$, calculates, for each sound scene, the error between the exact and estimated road traffic 1-s equivalent sound level of each file normalized by the 1-s equivalent sound level of the global scene, $p^t_{1s,global}$, in the linear scale,
-
-\begin{equation}
-nRMSE = \sqrt{\frac{1}{T}\sum_{t = 1}^T \left(\frac{p^t_{1s,traffic}-\tilde{p}^t_{1s,traffic}}{p^t_{1s,global}}\right)^2}
-\end{equation}
-
-where $T$ is the number of temporal bin in the signal.
-The linear scale is here more relevant as it is more sensitive to the error on the high sound levels than the dB scale.
-Then for one combination of factors, the $N$ $nRMSE$ calculated are averaged.\\
-\end{itemize}
-
-In all, according the table \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 14040 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF.
+In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of parameters are made.
+For ThC NMF, 16800 combinations can be calculated.
 
@@
 
-Table \ref{tab:results} summarized, according to the 3 main parameters (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$.\\
+Table \ref{tab:results} summarized, according to the 3 main parameters (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\
 
@@
  &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-25 & all & 0.5 & spectra & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
-50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
+25 & all & 0.5 & linear & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
+50 & 0.5 & 0.5 & linear & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
 50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
 50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
-100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
-100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
+100 & 0.5 &  20 & linear & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
+100 & 0.5 & 20 & linear & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
 100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
 100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
-50 & all & 5 & spectra & 1 & ThC NMF & 0.39 & \textbf{2.23 ($\pm$ 2.21)} \\ 
-50 & all & 2 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.17 ($\pm$ 2.28)} \\ 
+50 & all & 5 & linear & 1 & ThC NMF & 0.39 & \textbf{2.23 ($\pm$ 2.21)} \\ 
+50 & all & 2 & linear & 2 & ThC NMF & 0.37 & \textbf{2.17 ($\pm$ 2.28)} \\ 
 100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
@@
 The two first lines resume the error produced by the filter.
-$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguish the traffic from the others sound sources.
+$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguishing the traffic from the others sound sources.
 Consequently , in low $TIR$ (-12 and -6), where traffic component is discreet, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant.
-$f_c = 500$ Hz is the cut-off frequency with the lower mean error.
-It is then the baseline to refer to compare the performance of NMF.
-In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the impact of the filter is quite interesting as it suppress theses frequency components whereas for the other sub-classes where low frequencies are presents, the error as higher as the filter cannot dissociate the traffic element from the other sound sources.
+$f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
+It is then the baseline that will be used to compare the performances of NMF.
+In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the filter is efficient as it suppress these frequency components whereas for the other sub-classes where low frequencies are present, the error is higher as the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
 The sounds levels are then overestimate (figure \ref{fig:alert_-12}).
-In opposite, in high $TIR$, the low-pass filter removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\
+In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \centering
@@
 
-The results of different versions of NMF are summarized next to the filter results with the NMF errors lower than the baseline in bold.
+The results of different versions of NMF are summarized next to the filter results in Table \ref{tab:results_TIR}.
+NMF errors lower than the baseline are put in bold letter.
 With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
@@
 
-Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, no parameters is common to all the settings.
-For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).
+Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings.
+If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this parameters is variable depending on chosen method.
+
+For supervised NMF, the spectrogram $\mathbf{V}$ has to be filtered at 500 Hz to focused the reconstruction of the signals on the part where most of the traffic energy is.
+According to the choice of $\beta$, the dictionary parameters are different.
+It is the only case where it is.
+In the opposite, in semi-supervised NMF, an unique set of parameters is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz.
+It is with $f_c = 20$ kHz that the approximation is better as it includes easily other sound sources.
+With a lower cut-off frequency ($ f_c <$ 5 kHz), focusing on traffic energy, the risk is then to include \textit{traffic} component in $\mathbf{W_r}$ and to deteriorate the traffic signal reconstruction.
+The influence of the spectral representation and $\beta$ is here weak as the errors are similar with a low standard deviation.
+The third approaches, ThC NMF, proposed, in the four cases, the lowest errors.
+Like supervised NMF, in linear spectral representation, it is necessary to filter $\mathbf{V}$ to focus on the traffic energy.
+Here $f_c$ is higher (5000 Hz and 2000 Hz respectively for $\beta = 1$ and $\beta$ = 2).
+The choice of the spectral representation influence also the order of magnitude of the threshold $t$: in the interval [0.37 0.39] and [0.54 0.57] respectively for a linear and third octave spectral representation.\\
+
+For each method (supervised, semi-supervised and ThC NMF), the best parameter combinations are detailed according to the sub-classes and the $TIR$ (Table \ref{tab:results_TIR},  Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_thc}).
+\\
 
@@
 \toprule
-method & filter & filter & supervised NMF & semi-supervised NMF & ThC NMF \\ \midrule
-$\beta$ &  &  & 2 & 2 & 2 \\
-\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular} &  &  & spectra & spectra & third octave \\
-$f_c$ (kHz) & 20 & 0.5 & 0.5 & 20 & 20 \\ \hline
--12 & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & \textbf{6.23 ($\pm$ 3.19)} &  & \textbf{5.11 ($\pm$ 3.10)} \\
--6 & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & \textbf{3.00 ($\pm$ 1.39)} &  & \textbf{2.87 ($\pm$ 1.55)} \\
-0 & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
-6 & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & \textbf{1.01 ($\pm$ 0.28)} &  & \textbf{0.70 ($\pm$ 0.32)} \\
-12 & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & \textbf{1.41 ($\pm$ 0.10)} &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
+\textbf{method} & filter & filter & supervised NMF & semi-supervised NMF & ThC NMF \\ \midrule
+$f_c$ (kHz) & 20 & 0.5 & 0.5 & 20 & 20 \\ 
+\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  &  & linear & linear & third octave \\
+$\mathbf{\beta}$ &  &  & 2 & 2 & 2 \\ \hline
+\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & \textbf{6.23 ($\pm$ 3.19)} &  & \textbf{5.11 ($\pm$ 3.10)} \\
+\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & \textbf{3.00 ($\pm$ 1.39)} &  & \textbf{2.87 ($\pm$ 1.55)} \\
+\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
+\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & \textbf{1.01 ($\pm$ 0.28)} &  & \textbf{0.70 ($\pm$ 0.32)} \\
+\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & \textbf{1.41 ($\pm$ 0.10)} &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
 \end{tabular}
 \caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
-\label{ta:results_TIR}
+\label{tab:results_TIR}
 \end{table*}
 
-\begin{figure*}[t]
+\begin{figure*}
     \centering
@@
 
-In the case of supervised NMF, on all the $TIR$, if the mean score is lower than the baseline method, it offers similar performances as for $TIR = [0, 6, 12]$, the mean score are nearly close.
+According to Figure \ref{fig:TIR_class_filter}, \textit{climate}, \textit{human}, \textit{transport} and \textit{mechanics} are the four sub-classes that generates the highest error levels particularly in low $TIR$ as they are sound classes composed of low frequency sound (storm for \textit{climate}, voices in \textit{humans}, plane in \textit{transport} and ventilation noise in \textit{mechanics}).
+All these kind of sounds are then considered as traffic component without distinction.
+In the case of supervised NMF, on all the $TIR$, if the mean score on each $TIR$ is lower than the baseline method, it offers similar performances as for $TIR = [0, 6, 12]$, the mean scores are nearly close.
 The performance of supervised NMF is mainly visible for $TIR = -12$ where the error with the baseline decrease by more than 1 dB.
+This improvement is significant with the \textit{alert} sub-class as it gathers sounds that are mostly in higher frequencies ([2500 5000] Hz).
+From a dictionary $\mathbf{W}$ composed of traffic element, the traffic signal reconstruction is then easier.
+However, supervised NMF fails to improve significantly the baseline performances.
 
-The gain of the semi-supervised approach is most of all significant for these low $TIR$ where the add of a mobile part, $\mathbf{W_r}$, into $\mathbf{W}$ allows to take into account the other predominant sound sources (example in Figure \ref{fig:Y_alert-12}) while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$.
-This improvement, according to supervised NMF results, can be seen in Figure \ref{fig:TIR_class_semi} where the error on each sub class for $TIR = -12$ and $TIR$ = -6 are much lower.
-Nevertheless, these degrees of freedom are restrictive for high $TIR$.
-Indeed, without constraint semi-supervised NMF is free to include traffic components in $\mathbf{W_r}$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}).
-The error are then increasing for theses $TIR$.
-In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances.
-\\
+Meanwhile, semi-supervised approach with the add of a mobile part in the dictionary, $W_r$, brings a major improvement especially for the low $TIR$ (Figure \ref{fig:TIR_class_semi}).
+By the add of $\mathbf{W_r}$, it allows to take into account the other predominant sound sources.
+An example can be found in Figure \ref{fig:Y_alert-12} which summarizes the obtained $\mathbf{W_r}$ of a scene belonging to \textit{alert} sub-class.
+The first element is mainly composed of harmonics signal which correspond to the car horn of the scene.
+The decreasing of errors can be seen on each sub class for $TIR = -12$ and $TIR$ = -6 in  where the error are much lower than supervised NMF.
+As the $\mathbf{W}$ is only composed of \textit{traffic} element, this approach is likely to use traffic components to reduce it even if the sound source not traffic to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$ deteriorating the quality of traffic signal reconstruction.
 
-\begin{figure*}[t]
+The comparison can be made on a scene of the \textit{alert} sub-class: 
+FIGURE Lp 1 seconde avec supervisé et semi-supervisé pour voir l'utilisation des éléments trafic dans la supervisée et non dans al semi-supervisée. \\
+
+However, the degrees of freedom of semi-supervised NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes.
+Indeed, in order to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$ without constraint semi-supervised NMF is free to include traffic components in $\mathbf{W_r}$.
+Consequently, this behaviors decreases the quality of the reconstruction of the traffic component.
+The same scene used in Figure \ref{fig:Y_alert-12}) but with $TIR$ can be seen in Figure \ref{fig:Y_alert_12} where the harmonic components of traffic have disappears for low frequency components.
+The error are then increasing for these $TIR$.
+Constrain $\mathbf{W_r}$ to avoid these behaviors and control what can be put into could be a way to improve this method here \cite{kitamura_music_2014}.
+In opposite, supervised NMF (that could be seen as semi-supervised NMF which is infinitely constrain on $\mathbf{W_r}$) offer a good reconstruction of the traffic signal from $\mathbf{W}$.\\
+
+\begin{figure*}
     \centering
@@
         \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
-        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = -12}
+        \caption{$TIR$ = -12}
         \label{fig:Y_alert-12}
@@
         \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
-        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = 12}
+        \caption{$TIR$ = 12}
 		\label{fig:Y_alert_12}
     \end{subfigure}
+    \caption{$\mathbf{W_r}$ for an \textit{alert} scene}
 \end{figure*}
 
-Finally, ThC NMF with a hard threshold fixed at $t = 0.54$ offers the lowest results.
-The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.
+Finally, ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
+According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors.
+On the other hand, on low $TIR$, it does not succeed to be better than semi-supervised NMF.
+
+Unlike supervised NMF, where $\mathbf{W}$ is fixed, and semi-supervised NMF, which combine a fixed dictionary with a mobile dictionary, ThC NMF's $\mathbf{W}$ is here update entirely to be adjusted to the scene and to adapt to the different sound environment.
+The closest elements of the \textit{traffic} component defined un $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
+In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes for $TIR$ = [-12, 12].
 
@@
 
-The gain is significant on all $TIR$ with the exception of $TIR = 0$.
-On Figure \ref{fig:TIR_class_thc}, we can see the improvement bringing by this approach.
-For low $TIR$, only a reduce number of $\mathbf{W}$ are considered as traffic components, the performance on \textit{alert}, \textit{animals}, \textit{humans} and \textit{tranportation} are then similar to semi-supervised NMF.
-However, for \textit{climate} and \textit{mechanics}, the error are important as in supervised NMF.
-On high $TIR$, the similarity of the initial dictionary and $\mathbf{W}$ allows to keep more elements.
+For low $TIR$, with the thresholding, only a reduce number of $\mathbf{W}$ are considered as traffic components.
+In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.
+However, for \textit{climate} and \textit{mechanics}, the error stay important (approfondir avec comparaison des niveau sonores trafic).
+On high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher and allows to keep more elements.
+The kept elements are then more suited to the scenes than a fixed dictionary.
+So, the error, compare to supervised NMF, are better.
 In these cases, ThC NMF is the most powerful of the proposed methods.
-Consequently, if the average error is lower than semi-supervised NMF, with good performances on high $TIR$, it is a more variable method.
 
 \section{Conclusion}
+ouvrir sur l'optimisation du seuil par plusieurs indicateurs (niveau sonore global ou en fréquence ? )\\
+ThC NMF adapté pour d'autres sources sonores \\
+Scènes grafic ensuite
+
 \footnotesize

### 2017-12-20 paper/articleApplied.tex
comments

--- 2017-12-20 article/articleApplied.tex
@@
 
-\input{../packageEN}
-\input{../packageStyle}
+\input{packageEN}
+\input{packageStyle}
 
@@
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-These maps allow to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+These maps depict an estimation of the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
 These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
@@
 Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels.
-DYNAMAP project \cite{dynamap_2016} studied the establishment and feasibility of such installations.
+DYNAMAP project \cite{dynamap_2016} studied the establishment \ml{c'est anglais ?} and feasibility of such installations.
 It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
@@
 Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap.
-Consequently, the traffic sound level estimation based on measurements is not trivial task.
+Consequently, the traffic sound level estimation based on measurements is also not a trivial task.
 Many recent works have focused on the detection or recognition tasks of environmental sounds without distinction between them\cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
@@
 A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+
 Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
-If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic nous level are presented.
-
+If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic  level are presented.
 Furthermore, this work and as well as the other works in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
@@
 In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level.
+
 In consequence, to be effective on a wide range of sound environments, we propose in this paper to follow the blind source separation paradigm.
@@
 
-One of the first and the most widely used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}.
+\ml{tu passe beaucoup de temps a taper sur socoro et peu a parler de ton approche.
+Ce n'est pas une bonne pratique.
+Je dirais ils font ca, nous aussi on fait ca c'est super, et nous on le fais comme ca, regardez tout ce que ca permet (et implicitement eux ils peuvent pas).}
+
+\ml{tu plonges tres vite dans la technique}
+
+One of the first and the most widely \ml{faux} used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}.
 The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
@@
 However, ICA is limited to only over determined cases ($N > P$).
-Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation.
+Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation \ml{je ne vois pas ou est le probleme, il y a des gens qui le font}.
 A more convenient method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices.
 It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
-By design, this method deals with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane.
+By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane.
 For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.
@@
 They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them.
-The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
+The audio files tested results of a simulation process where a sound background (river or wind) are added to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
+
  If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
+ \ml{encore une fois, utilise la critique implicite. On peut tout critiquer, surtout un papier qui n'est pas encore publié, fais attention.
+
+ faire une liste des contributions avec 1) 2) 3) et finir par un paragraphe de structure}
 
 \subsection{Proposed approach}
+
+\ml{tu melange contribution technique et protocole experimental, ca va pas
+
+pas de soussection a l'intro}
 
@@
 
+\ml{on evite les parentheses en Anglais, pour les ref, faire , see \ref{}}
+
 \begin{figure*}
@@
         \centering
-        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}
+        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
         \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
@@
         \centering
-        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}
+        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
         \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
@@
 \centering
-\includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
+\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
 \caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{../image/distanceCosLinDisplay.pdf}
+\includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
 \caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
@@
 \centering
-\includegraphics[width=\linewidth]{../image/exampleSimScene2.pdf}
+\includegraphics[width=\linewidth]{figures/exampleSimScene2.pdf}
 \label{fig:exampleSimScene}
@@
    \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{../image/spectrogramExample.pdf}
+      \includegraphics[width =\linewidth]{figures/spectrogramExample.pdf}
    \end{minipage}
    \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{../image/animals_10-pianoRoll.png}
+      \includegraphics[width =\linewidth]{figures/animals_10-pianoRoll.png}
    \end{minipage}
    \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{../image/evolutionLpExample.pdf}
+      \includegraphics[width =\linewidth]{figures/evolutionLpExample.pdf}
 
@@
 
-The first estimator to determine the traffic sound level is a basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
+The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
 The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-The second estimator is NMF. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.
+The second estimator is the proposed scheme, based on several flavors of the NMF framework. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.
 
@@
 It is composed of 53 audio files of passing cars.
-These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars ().
+These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars.
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
@@
         \centering
-       \includegraphics[width=\linewidth]{../image/dictionary1.pdf}
+       \includegraphics[width=\linewidth]{figures/dictionary1.pdf}
         \caption{Sample of 3 seconds of a car passage}
@@
         \centering
-       \includegraphics[width=\linewidth]{../image/dictionary2.pdf}
+       \includegraphics[width=\linewidth]{figures/dictionary2.pdf}
         \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
@@
 
+\ml{replace parameters by experimental factors}
+
 \begin{table*}[]
@@
 
-The two first lines resume the error produced by the filter.
+The two first lines show the error produced by the baseline filter.
 $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguishing the traffic from the others sound sources.
-Consequently , in low $TIR$ (-12 and -6), where traffic component is discreet, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant.
+Consequently, in low $TIR$ (-12 and -6), where traffic component is scarce, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant.
 $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_-12.pdf}
+        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}
         \caption{}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_12.pdf}
+        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}
         \caption{}
@@
 
+\ml{remplacer ThC NMF par TINMF pour Thresholded initialized nmf}
+
 For each method (supervised, semi-supervised and ThC NMF), the best parameter combinations are detailed according to the sub-classes and the $TIR$ (Table \ref{tab:results_TIR},  Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_thc}).
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
+        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
         \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
+        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
         \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
+        \includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
         \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/THC_bar}
+        \includegraphics[width=\linewidth]{figures/THC_bar}
         \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
+        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_-12_01.pdf}
         \caption{$TIR$ = -12}
@@
         \centering
-        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
+        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_12_01.pdf}
         \caption{$TIR$ = 12}
@@
 
-Finally, ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
+The ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
 According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors.
@@
     \centering
-    \includegraphics[width=\linewidth]{../image/dist_-12_12.pdf}    
+    \includegraphics[width=\linewidth]{figures/dist_-12_12.pdf}
     \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
@@
 
+\ml{non, une conclusion conclue, c'est tout.
+
+rappel des resultats experimentaux et conclusion sur l'approche. Rappel des contributions et impact sur la communaute.
+Petite ouverture sur le potentiel de l'approche: agnostique en terme de source, peut etre contraint temporellement citation, reste raisonable en terme de cout de calcul. }
+
 \footnotesize

### 2017-12-20 paper/gloaguenNmfTraffic.tex
done with cleaning

--- 2017-12-20 paper/articleApplied.tex
@@
 	Jean-François Petiot \\
-    LS2N\\
-    Central School of Nantes\\
+    LS2N, CNRS\\
+    Ecole Centrale de  Nantes\\
 }
@@
  If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
- \ml{encore une fois, utilise la critique implicite. On peut tout critiquer, surtout un papier qui n'est pas encore publié, fais attention.
-
+ \ml{encore une fois, utilise la critique implicite.
+On peut tout critiquer, surtout un papier qui n'est pas encore publié, fais attention.
 faire une liste des contributions avec 1) 2) 3) et finir par un paragraphe de structure}
@@
 
-\ml{tu melange contribution technique et protocole experimental, ca va pas
-
-pas de soussection a l'intro}
+\ml{tu melange contribution technique et protocole experimental, ca va pas. pas de soussection a l'intro}
 
@@
 \end{tabular}
-\caption{Best results according to $\beta$, method and  spectral representation}
+\caption{Best results according to $\beta$, method and  spectral representation \ml{ne mettre en gras que le meilleur et expliciter le gras dans la caption}})}
 \label{tab:results}
@@
         \centering
-        \includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
+      %  \includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
         \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
@@
 \ml{non, une conclusion conclue, c'est tout.
-
 rappel des resultats experimentaux et conclusion sur l'approche. Rappel des contributions et impact sur la communaute.

### 2018-01-04 paper/gloaguenNmfTraffic.tex
add experiment folder with m files (threshold initialized NMF missing for the moment)

--- 2017-12-20 paper/articleApplied.tex
@@
 It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
-%The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}.
-\ml{je ne vois pas l'intéret de cette dernière phrase}
 
 Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels.
-DYNAMAP project \cite{dynamap_2016} studied the establishment \ml{c'est anglais ?} and feasibility of such installations.
+DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations.
 It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
@@
 The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of an urban scene ($L_{p,traffic}$ and $p_{1s,traffic}$).
-Both the sound scene simulation and NMF require the creation of two sound databases (figure  \ref{fig:blocDiagramNMF}).
+Both the sound scene simulation and NMF require the creation of two sound databases; see Figure \ref{fig:blocDiagramNMF}.
 In parallel, a baseline method built from a frequency low-pass filter is computed.
-This method considers that road traffic is mainly composed of low frequencies and therefore can be filtered by a low-pass filter at the cut-off frequency $f_c$ (figure \ref{fig:blocDiagramFilter}).
+This method considers that road traffic is mainly composed of low frequencies and therefore can be filtered by a low-pass filter at the cut-off frequency $f_c$; see Figure \ref{fig:blocDiagramFilter}.
 The performance of the frequency low-pass filter and NMF are then compared with the calculation of two metrics (Mean Absolute Error, normalized Root Mean Square Error).
-
-\ml{on evite les parentheses en Anglais, pour les ref, faire , see \ref{}}
 
@@
 
-where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$ (fig.
-\ref{fig:example_NMF}).
+where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$.
+An illustrative example can be found in Figure  \ref{fig:example_NMF}.
 
@@
 
-Prior knowledge on the content can be adjusted with the addition of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function (equation (\ref{eq:min-D-WH})) to better take account prior knowledge of the sources.
+Prior knowledge on the content can be adjusted with the addition of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function , see equation (\ref{eq:min-D-WH}), to better take account prior knowledge of the sources.
 
@@
 
-\subsection{Thresholded constrained NMF}
+\subsection{Thresholded initialized NMF}
 
@@
 Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly.
-Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain (see part \ref{part:dictionary_learning}).
+Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}).
 Then NMF is performed where $\mathbf{W}$ (eq.
@@
 \begin{equation}
-\cos \theta = \frac{\mathbf{W}.\mathbf{W_0}}{\vert \vert \mathbf{W}  \vert \vert . \vert \vert \mathbf{W_0} \vert \vert}.
+D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = \frac{\mathbf{W}.\mathbf{W_0}}{\vert \vert \mathbf{W}  \vert \vert . \vert \vert \mathbf{W_0} \vert \vert}.
 \end{equation}
 
-$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are significantly different.
+$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ means that the elements are significantly different.
 This measure allows a bound between 1 and 0 and is an invariant scale estimation of the similarity.
 Then, the similarity is sorted in descending order.
-The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_ThC_NMF}).
-It is defined as :
+The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
+An illustrative example can be see in Figure \ref{fig:W_ThC_NMF}.
+It is defined as:
 
@@
 
-This approach is named \textit{Thresholded constraint NMF} (ThC NMF).
+This approach is named \textit{Thresholded inititialized NMF}, TI-NMF.
 Other thresholding methods as the \textit{soft} \cite{donoho1995noising}and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
-A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$ (see fig.
-\ref{fig:W_ThC_NMF}) was the best way to get better performances.
+A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, like in  Figure \ref{fig:W_ThC_NMF}), was the best way to get better performances.
 
@@
 When $TIR = -12$, the traffic component is then less present than when $TIR = 12$ where it is predominant on the \textit{interfering} class.
-The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated (figure \ref{fig:exampleScene}).
+The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated ; see Figure \ref{fig:exampleScene}).
 The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TIR values).
@@
 
-The second estimator is the proposed scheme, based on several flavors of the NMF framework. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.
+The second estimator is the proposed scheme, based on several flavors of the NMF framework.
+Multiples experimental factors are involved here between the dictionary learning and the own parameters of the NMF.
 
@@
 Each audio file generates one element $k$ of $\mathbf{W}$.
-An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig.
-\ref{fig:specW}).
-In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig.
-\ref{fig:ElementW}).
+An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage ; see Figure \ref{fig:specW}.
+In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element ; see Figure \ref{fig:ElementW}.
 
@@
 With the third octave spectral representation, only the case $f_c = 20$ kHz is applied.
-For the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
+For the TI-NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
 Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
@@
 \begin{tabular}{lcccccc}
-Parameter &  \multicolumn{6}{c}{value}\\ \hline
+\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \hline
 $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
 \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
-\begin{tabular}[l]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}\\ \hline
 $\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
 $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
-\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & ThC NMF \\ \hline
+\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & TI-NMF \\ \hline
 $\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
@@
 In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of parameters are made.
-For ThC NMF, 16800 combinations can be calculated.
+For TI-NMF, 16800 combinations can be calculated.
 
@@
 \centering
-\begin{tabular}{@{}cccccccc@{}}
+\begin{tabular}{@{}cccccc@{}}
 \toprule
-$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
- &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\
- &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-25 & all & 0.5 & linear & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\
-50 & 0.5 & 0.5 & linear & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\
-50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\
-50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
-100 & 0.5 &  20 & linear & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\
-100 & 0.5 & 20 & linear & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\
-100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\
-100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
-50 & all & 5 & linear & 1 & ThC NMF & 0.39 & \textbf{2.23 ($\pm$ 2.21)} \\
-50 & all & 2 & linear & 2 & ThC NMF & 0.37 & \textbf{2.17 ($\pm$ 2.28)} \\
-100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\
-100 & all & 20 & third octave & 2 & ThC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
+$\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
+ &  &  & filter &  & 4.69 ($\pm$ 4.52) \\
+ &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
+50 & 0.5  & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\
+50 & 0.5  & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
+100 & 0.5 & 1 & semi-supervised NMF &  & 2.38 ($\pm$ 1.26) \\
+100 & 0.5 & 2 & semi-supervised NMF &  & 2.43 ($\pm$ 1.43) \\ \hline \hline
+100 & all & 1 & TI-NMF & 0.57 & 2.19 ($\pm$ 2.01) \\
+100 & all & 2 & TI-NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
 \end{tabular}
@@
 In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the filter is efficient as it suppress these frequency components whereas for the other sub-classes where low frequencies are present, the error is higher as the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
-The sounds levels are then overestimate (figure \ref{fig:alert_-12}).
-In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\
+The sounds levels are then overestimate ; see Figure \ref{fig:alert_-12}.
+In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels ; see Figure \ref{fig:alert_12}.\\
 
@@
 With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
-The best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+The best combination is got with a TI-NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
 
@@
 The influence of the spectral representation and $\beta$ is here weak as the errors are similar with a low standard deviation.
-The third approaches, ThC NMF, proposed, in the four cases, the lowest errors.
+The third approaches, TI-NMF, proposed, in the four cases, the lowest errors.
 Like supervised NMF, in linear spectral representation, it is necessary to filter $\mathbf{V}$ to focus on the traffic energy.
@@
 
-\ml{remplacer ThC NMF par TINMF pour Thresholded initialized nmf}
-
-For each method (supervised, semi-supervised and ThC NMF), the best parameter combinations are detailed according to the sub-classes and the $TIR$ (Table \ref{tab:results_TIR},  Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_thc}).
+For each method (supervised, semi-supervised and TI-NMF), the best parameter combinations are detailed according to the sub-classes and the $TIR$ (Table \ref{tab:results_TIR},  Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_thc}).
 \\
@@
 \toprule
-\textbf{method} & filter & filter & supervised NMF & semi-supervised NMF & ThC NMF \\ \midrule
-$f_c$ (kHz) & 20 & 0.5 & 0.5 & 20 & 20 \\
-\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  &  & linear & linear & third octave \\
+\textbf{method} & filter & filter & supervised NMF & semi-supervised NMF & TI-NMF \\ \midrule
+$f_c$ (kHz) & 20 & 0.5 & 20 & 20 & 20 \\
 $\mathbf{\beta}$ &  &  & 2 & 2 & 2 \\ \hline
-\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & \textbf{6.23 ($\pm$ 3.19)} &  & \textbf{5.11 ($\pm$ 3.10)} \\
-\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & \textbf{3.00 ($\pm$ 1.39)} &  & \textbf{2.87 ($\pm$ 1.55)} \\
+\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & 6.23 ($\pm$ 3.19) &  & \textbf{5.11 ($\pm$ 3.10)} \\
+\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 3.00 ($\pm$ 1.39) &  & \textbf{2.87 ($\pm$ 1.55)} \\
 \textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
-\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & \textbf{1.01 ($\pm$ 0.28)} &  & \textbf{0.70 ($\pm$ 0.32)} \\
-\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & \textbf{1.41 ($\pm$ 0.10)} &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
+\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 1.01 ($\pm$ 0.28) &  & \textbf{0.70 ($\pm$ 0.32)} \\
+\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 1.41 ($\pm$ 0.10) &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
 \end{tabular}
@@
         \includegraphics[width=\linewidth]{figures/THC_bar}
-        \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
+        \caption{TI-NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
         \label{fig:TIR_class_thc}
@@
 
-Meanwhile, semi-supervised approach with the add of a mobile part in the dictionary, $W_r$, brings a major improvement especially for the low $TIR$ (Figure \ref{fig:TIR_class_semi}).
+Meanwhile, semi-supervised approach with the add of a mobile part in the dictionary, $W_r$, brings a major improvement especially for the low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}).
 By the add of $\mathbf{W_r}$, it allows to take into account the other predominant sound sources.
@@
 
-The ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
+The TI-NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
 According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors.
@@
 
-Unlike supervised NMF, where $\mathbf{W}$ is fixed, and semi-supervised NMF, which combine a fixed dictionary with a mobile dictionary, ThC NMF's $\mathbf{W}$ is here update entirely to be adjusted to the scene and to adapt to the different sound environment.
+Unlike supervised NMF, where $\mathbf{W}$ is fixed, and semi-supervised NMF, which combine a fixed dictionary with a mobile dictionary, TI-NMF's $\mathbf{W}$ is here update entirely to be adjusted to the scene and to adapt to the different sound environment.
 The closest elements of the \textit{traffic} component defined un $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
@@
 So, the error, compare to supervised NMF, are better.
-In these cases, ThC NMF is the most powerful of the proposed methods.
+In these cases, TI-NMF is the most powerful of the proposed methods.
 
@@
 ouvrir sur l'optimisation du seuil par plusieurs indicateurs (niveau sonore global ou en fréquence ? )\\
-ThC NMF adapté pour d'autres sources sonores \\
+TI-NMF adapté pour d'autres sources sonores \\
 Scènes grafic ensuite

### 2018-01-04 experiment/NonnegMatrixFact/NonnegMatrixFact.m

@@ -1,0 +1,190 @@
+function NonnegMatrixFact(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             
+% Welcome to the main entry point of NonnegMatrixFact                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'nomafa';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/MFCC/invaudspec.m

@@ -1,0 +1,43 @@
+function [spec,wts,iwts] = invaudspec(aspectrum, sr, nfft, fbtype, minfreq, maxfreq, sumpower, bwidth)
+%pspectrum = invaudspec(aspectrum, sr, nfft, fbtype, minfreq, maxfreq, sumpower, bwidth)
+%
+% Invert (as best we can) the effects of audspec()
+%
+% 2004-02-04 dpwe@ee.columbia.edu
+
+if nargin < 2;  sr = 16000;     end 
+if nargin < 3;  nfft = 512;     end
+if nargin < 4;  fbtype = 'bark';  end
+if nargin < 5;  minfreq = 0;    end
+if nargin < 6;  maxfreq = sr/2; end
+if nargin < 7;  sumpower = 1;   end
+if nargin < 8;  bwidth = 1.0;   end
+
+[nfilts,nframes] = size(aspectrum);
+
+if strcmp(fbtype, 'bark')
+  wts = fft2barkmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
+elseif strcmp(fbtype, 'mel')
+  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
+elseif strcmp(fbtype, 'htkmel')
+  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 1);
+elseif strcmp(fbtype, 'fcmel')
+  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1);
+else
+  disp(['fbtype ', fbtype, ' not recognized']);
+  error;
+end
+
+% Cut off 2nd half
+wts = wts(:,1:((nfft/2)+1));
+
+% Just transpose, fix up 
+ww = wts'*wts;
+iwts = wts'./(repmat(max(mean(diag(ww))/100, sum(ww))',1,nfilts));
+% Apply weights
+if (sumpower)
+  spec = iwts * aspectrum;
+else
+  spec = 2*(iwts * sqrt(aspectrum)).^2;
+end
+

### 2018-01-04 experiment/NonnegMatrixFact/netlab/cutSpectrogramEXP.m

@@ -1,0 +1,24 @@
+function [soundMix] = cutSpectrogramEXP(dictionary,setting)
+
+W = dictionary.W;
+
+switch setting.domain
+    case 'mel'
+        sr = setting.sr;
+        numberMel = setting.numberMel; 
+
+        [W,~] = spectre2MelEXP(W,numberMel,sr);
+        ind = numberMel;
+        
+    case 'thirdOctave'
+        [W,~,~,~] = NarrowToNthOctaveEXP(dictionary.frequency,W,3);
+        ind = size(W,1);
+        
+    otherwise
+        [~,ind] = min(abs(setting.cutOffFreq-dictionary.frequency));
+end
+
+soundMix.ind = ind;
+soundMix.seed = rng;
+soundMix.W = W;
+soundMix.indTraffic = dictionary.indTraffic;

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/NMFestimationEXP.m

@@ -1,0 +1,85 @@
+function [LeqTrafficEstimate,LpTrafficEstimate,H,cost,sequentialData] = NMFestimationEXP(V,dictionary,setting,sequentialData)
+
+
+%% limit the spectrogram and the dictionary to cutOffFreq
+[soundMix] = cutSpectrogramEXP(dictionary,setting);
+
+W = soundMix.W(1:soundMix.ind,:);
+V = V(1:soundMix.ind,:);
+binTemp = size(V,2);
+
+switch setting.nmfType
+    case 'supervised'
+        
+        if isempty(sequentialData)
+            iteration = setting.iteration;
+            rng(soundMix.seed)
+            H = rand(size(W,2),binTemp);
+        else
+            % continuing step of the sequential run
+            iteration = setting.iteration-sequentialData.numberIteration;
+            H = sequentialData.H;
+        end
+
+        NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting);
+        cost = betadivEXP(V,NMF.Vap,setting.beta,NMF.H,setting.sparsity,setting.SM_weight);
+
+    case 'semi-supervised'
+        
+        Wrand = setting.SS_sizeWrand;
+        if isempty(sequentialData)
+            iteration = setting.iteration;
+            rng(soundMix.seed); Y =  rand(size(W,1),Wrand);
+            rng(soundMix.seed); H = rand(size(W,2),binTemp);
+            rng(soundMix.seed); Z = rand(Wrand,binTemp);
+        else
+            % continuing step of the sequential run
+            iteration = setting.iteration-sequentialData.numberIteration;
+            H = sequentialData.H;
+            Z = sequentialData.Z;
+            Y = sequentialData.Y;
+        end
+
+        NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting);
+        cost = betadivSemiEXP(V,NMF,setting);
+        
+        sequentialData.Y = NMF.Y;
+        sequentialData.Z = NMF.Z;
+		
+	case 'unsupervised'
+		W0 = soundMix.W(1:soundMix.ind,:);      % initial dictionary
+		V = V{1}(1:soundMix.ind,:);    % spectrogram to approximate
+
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/algo_nmfSupervisedEXP.m

@@ -1,0 +1,101 @@
+function NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
+%% NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
+
+sparsity = setting.sparsity; 
+smoothness = setting.SM_weight; 
+beta = setting.beta;
+Vap = W*H;
+
+NMF = [];
+
+if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
+    if beta < 1
+        gamma = 1/(2-beta);
+    elseif beta >= 1 && beta <= 2
+        gamma = 1;
+    else
+        gamma = 1/(beta-1);
+    end
+end
+
+if smoothness~=0
+    lambda = sum(W,1);
+    smoothForm = setting.smoothnessForm;
+    
+    switch smoothForm
+        case 'all'
+            smoothnessTot(1:size(W,2),1) = smoothness;
+        case 'traffic'
+            smoothnessTot = zeros(size(W,2),1);
+            smoothnessTot(soundMix.indTraffic==1) = smoothness;
+    end
+else
+    smoothnessTot = 0;
+    lambda = NaN;
+end
+
+switch beta
+    case 2
+        for iter = 1:iteration(end)
+            if smoothness==0
+                H = H.*(W'*V-sparsity)./(W'*Vap);
+                H(H<0)=0;
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+        end
+    case 1
+        for iter = 1:iteration
+            
+            if smoothness==0
+                H = H.*((W'*(V.*Vap.^(-1)))./(sparsity+W'*Vap.^(0)));
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+            
+        end
+    case 0
+        for iter = 1:iteration
+            
+            if smoothness==0
+                H = H.*((W'*(V.*Vap.^(-2)))./(sparsity+W'*Vap.^(-1))).^(1/2);
+            else
+                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+            end
+            H(isnan(H)) = 0;
+            Vap = W*H;
+
+        end
+    otherwise
+        if beta < 1
[…]

### 2018-01-04 experiment/NonnegMatrixFact/nomafa1dictionary.m

@@ -1,0 +1,121 @@
+function [config, store, obs] = nomafa1dictionary(config, setting, data)
+% nomafa1dictionary DICTIONARY step of the expLanes experiment NonnegMatrixFact
+%    [config, store, obs] = nomafa1dictionary(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: <userName>
+% Date: 13-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, NonnegMatrixFact('do', 1); return; else, store=[]; obs=[]; end
+
+if strcmp(setting.type,'nmf')
+    K = setting.numberElementW;
+    inputPath = config.inputPath;
+
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %% NAME & NUMBER SAMPLE EXTRACTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+    switch setting.dictionaryClass
+        case 'car'
+            cityCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*cityCar*'));
+            roadCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*roadCar*'));
+            stopCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*stopCar*'));
+            sample = [cityCarSamples; roadCarSamples; stopCarSamples];
+            indiceClass = ones(1,length(sample));
+
+        case 'all'
+            sample = dir(strcat(inputPath,'dictionary',filesep,'*.wav'));
+
+            indiceClass = zeros(1,length(sample));
+            for ii = 1:length(sample)
+                if ~isempty(strfind(sample(ii).name,'cityCar')) ||...
+                        ~isempty(strfind(sample(ii).name,'roadCar')) || ...
+                        ~isempty(strfind(sample(ii).name,'stopCar'))
+                    indiceClass(ii) = 1;
+                else
+                    indiceClass(ii) = 0;
+                end
+            end
+        otherwise
[…]

### 2018-01-04 experiment/NonnegMatrixFact/nomafa2estimation.m

@@ -1,0 +1,126 @@
+function [config, store, obs] = nomafa2estimation(config, setting, data)
+% nomafa2estimation step of the expLanes experiment NonnegMatrixFact
+%    [config, store, obs] = nomafa2estimation(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: <userName>
+% Date: 16-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, NonnegMatrixFact('do', 2, 'mask',{2 2 1 3 1,...
+            3 0 6 2 1, 0 0 0 3 2, 1 1 1 1 1, 1 1 1 1}); return; else store=[]; obs=[]; end
+
+
+sceneSelect = setting.sceneSelect;
+TIR = setting.TIR;
+dataset = setting.dataset;
+type = setting.type;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DICTIONARY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+if strcmp(type,'nmf') 
+    dictionary.W = data.dictionary.W;
+    dictionary.frequency = data.dictionary.frequency;
+    dictionary.indTraffic = data.dictionary.indTraffic;
+    dictionary.numberPerClass = size(dictionary.W,2);
+else
+    dictionary = data.dictionary;
+end
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BASELINE GLOBALE ERROR + FILTRE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch dataset
+    case 'ambience'
+        sceneType = setting.aType;
+        creationSceneDir = strcat(config.inputPath, dataset, filesep, sceneType, filesep);
+        
+    case 'grafic'
+        sceneType = setting.gType;
+        creationSceneDir = strcat(config.inputPath, dataset, filesep, sceneType, filesep);
+        
[…]

### 2018-01-04 experiment/NonnegMatrixFact/nomafa3metric.m

@@ -1,0 +1,144 @@
+function [config, store, obs] = nomafa3metric(config, setting, data)
+% nomafa3metric metric step of the expLanes experiment NonnegMatrixFact
+%    [config, store, obs] = nomafa3metric(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: <userName>
+% Date: 16-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, NonnegMatrixFact('do', 3, 'mask',{2 2 1 3 1,...
+            3 0 6 2 1, 0 0 0 3 2, 1 1 1 1 1, 1 1 1 1}); return; else store=[]; obs=[]; end
+
+levels = data.levels;
+p0 = setting.p0;
+
+if strcmp(setting.formW,'cut') && strcmp(setting.domain,'spectra')
+    ind = 2;
+else
+    ind = 1;
+end
+
+LeqTrafficEstimate = zeros(length(levels),1);
+LeqTrafficEstimatedB = zeros(length(levels),1);
+LeqTrafficExact = zeros(length(levels),1);
+LeqTrafficExactdB = zeros(length(levels),1);
+
+%% EXTRACTION LEVELS
+for ii = 1:length(levels)
+    %% linear sound level
+    LpTrafficExact = levels{ii}.LpTraffic;
+    LpTrafficExact(LpTrafficExact==0) = 2e-5;
+    
+    LpTrafficEstimate = levels{ii}.LpTrafficEstimate{ind};
+    LpTrafficEstimate(LpTrafficEstimate==0) = 2e-5;
+    
+    LpGlobal = levels{ii}.LpGlobal;
+    LpGlobal(LpGlobal==0) = 2e-5;
+        
[…]

### 2018-01-04 experiment/NonnegMatrixFact/nomafaInit.m

@@ -1,0 +1,10 @@
+function [config, store] = nomafaInit(config)                          
+% nomafaInit INITIALIZATION of the expLanes experiment NonnegMatrixFact
+%    [config, store] = nomafaInit(config)                              
+%      - config : expLanes configuration state                         
+%      -- store  : processing data to be saved for the other steps     
+                                                                       
+% Copyright: <userName>                                                
+% Date: 13-Jan-2017                                                    
+                                                                       
+if nargin==0, NonnegMatrixFact(); return; else store=[];  end          

### 2018-01-04 experiment/NonnegMatrixFact/nomafaReport.m

@@ -1,0 +1,381 @@
+function config = nomafaReport(config)
+% nomafaReport REPORTING of the expLanes experiment NonnegMatrixFact
+%    config = nomafaInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: <userName>
+% Date: 13-Jan-2017
+
+%  if nargin==0, NonnegMatrixFact('report', 'r','reportName','filterAmbiance'); return; end
+if nargin==0, NonnegMatrixFact('report', 'r'); return; end
+
+switch config.reportName
+    case 'filterAmbiance'
+        config = expExpose(config, 't', 'mask', {1 0 0 0 0 ,...
+            0 0 0 2 1},'obs',[6 9],'integrate',{'aType','TPR'});
+        
+%         for ii = 1:5
+%             config = expExpose(config, 't', 'mask', {1 0 0 0 0 ,...
+%                 0 0 1 2 1,...
+%                 0 0 ii},'obs',5:9);
+%         end
+ %       config = expExpose(config, 'b', 'mask', {1 0 0 0 0 ,...
+  %          0 0 1 2 1},'obs','rmseLeqdB','expand','TPR'); 
+        
+    case 'filterGraficSlides'       
+        %% according to cutOffFreq
+        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
+            0 0 1 2 0, ...
+            0 0 0 1},'obs','rmsedB','integrate','gType','rotateAxis',45);
+        config = expExpose(config, 't', 'mask', {0 0 0 0 0 ,...
+            0 0 1 2 0, ...
+            0 0 0 1},'integrate','gType','obs',[4 7 8]);
+        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
+            0 0 1 2 0, ...
+            0 0 0 1},'obs','rmsedB','expand','gType','rotateAxis',45);
+        
+        %% according to gType
+        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
+            0 0 1 2 0, ...
[…]

### 2018-01-04 paper/gloaguenNmfTraffic.tex
add readme and NMFthreshold experience

@@
     LS2N, CNRS\\
-    Ecole Centrale de  Nantes\\
+    \'Ecole Centrale de  Nantes\\
 }
@@
 
-\begin{figure*}
+\begin{figure*}[t]
 \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
-        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
-        \label{fig:TIR_class_filter}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
-        \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
-		\label{fig:TIR_class_sup}
-    \end{subfigure}
+\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
+\caption{Bloc diagram of the source separation method}
 \end{figure*}
@@
 Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
+
+The experience is lead with the Matlab software.
+In order to be a reproductive experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} and the sound database\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} (TEMPORAIRE) used are available online.
 
@@
 
-The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
+The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
 The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq.
@@
 The second estimator is the proposed scheme, based on several flavors of the NMF framework.
-Multiples experimental factors are involved here between the dictionary learning and the own parameters of the NMF.
+Multiples experimental factors are involved here between the dictionary learning and NMF (see bloc diagram in Figure \ref{fig:bloc_nmf}).
+
+\begin{figure}
+
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
+  		\caption{}
+  		\label{fig:bloc_filtre}
+    \end{subfigure}%
+    \vfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
+		\caption{}
+		\label{fig:bloc_nmf}
+    \end{subfigure}
+
+    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter (a) and for NMF (b)}
+\end{figure}
 
@@
 
-Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.
+Table \ref{tab:dictionary_factors} summarizes the experimental factors and their related values.
 
@@
 \end{tabular}
-\caption{Summary of the dictionary parameters}
+\caption{Summary of experimental factors of the dictionary}
 \label{tab:dictionary_factors}
@@
 
-\subsubsection{NMF parameters}
+\subsubsection{Experimental factors of the NMF}
 
@@
 For the TI-NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
-Table \ref{tab:estimation_factors} summarizes the parameters and their related values.
+Table \ref{tab:estimation_factors} summarizes the experimental factors and their related values.
 
@@
 \end{tabular}
-\caption{Summary of the different parameters taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
+\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
 \label{tab:estimation_factors}
@@
 
-In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of parameters are made.
+In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple experimental factors taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of factors are made.
 For TI-NMF, 16800 combinations can be calculated.
@@
 
-Table \ref{tab:results} summarized, according to the 3 main parameters (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\
-
-\ml{replace parameters by experimental factors}
+Table \ref{tab:results} summarized, according to the 3 main factors (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\
 
@@
 
-Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings.
-If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this parameters is variable depending on chosen method.
+Between the dictionary factors ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings.
+If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this factors is variable depending on chosen method.
 
 For supervised NMF, the spectrogram $\mathbf{V}$ has to be filtered at 500 Hz to focused the reconstruction of the signals on the part where most of the traffic energy is.
-According to the choice of $\beta$, the dictionary parameters are different.
+According to the choice of $\beta$, the dictionary factors are different.
 It is the only case where it is.
-In the opposite, in semi-supervised NMF, an unique set of parameters is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz.
+In the opposite, in semi-supervised NMF, an unique set of factors is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz.
 It is with $f_c = 20$ kHz that the approximation is better as it includes easily other sound sources.

### 2018-01-04 experiment/NMFThreshold/NMFThreshold.m

@@ -1,0 +1,190 @@
+function NMFThreshold(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+% Welcome to the main entry point of NMFThreshold                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+shortExperimentName = 'nmfth';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
[…]

### 2018-01-04 experiment/NMFThreshold/functions/nmf/NMFestimationEXP.m

@@ -1,0 +1,37 @@
+function [NMF,sequentialData] = NMFestimationEXP(V,dictionary,setting,sequentialData)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% limit the spectrogram and the dictionary to cutOffFreq %%%%%%%%%%%%%%%%%
+[soundMix] = cutSpectrogramEXP(dictionary.W,setting);
+soundMix.indTraffic = dictionary.indTraffic;
+
+W0 = soundMix.W(1:soundMix.ind,:);      % initial dictionary
+V = V{1}(1:soundMix.ind,:);    % spectrogram to approximate
+
+if isempty(sequentialData)
+    iteration = setting.iteration;
+
+    W = W0;                             % first iteration W = W0
+    
+    binTemp = size(V,2);
+    rng(soundMix.seed)
+    H = rand(size(W,2),binTemp);    % H initiate randomly
+else
+    % continuing step of the sequential run
+    iteration = setting.iteration-sequentialData.numberIteration;
+    H = sequentialData.H;
+    W = sequentialData.W; 
+end
+% 
+NMF = algo_nmfUnsupervisedEXP(H,W,V,iteration,setting);
+% 
+% opt = statset('MaxIter',iteration);
+% [NMF.W,NMF.H,NMF.cost] = nnmf(V,setting.numberElement,'w0',W,'h0',H,'options',opt);
+
+sequentialData.numberIteration = setting.iteration;
+sequentialData.H = NMF.H;
+sequentialData.W = NMF.W;
+
+
+
+

### 2018-01-04 experiment/NMFThreshold/functions/nmf/algo_nmfUnsupervisedEXP.m

@@ -1,0 +1,91 @@
+function [NMF] = algo_nmfUnsupervisedEXP(H,W,V,iteration,setting)
+
+sparsity = 0;
+beta = setting.beta;
+Vap = W*H;
+
+if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
+    if beta < 1
+        gamma = 1/(2-beta);
+    elseif beta >= 1 && beta <= 2
+        gamma = 1;
+    else
+        gamma = 1/(beta-1);
+    end
+end
+
+cost = zeros(1,iteration);
+switch beta
+    case 2
+        for iter = 1:iteration
+            
+            W = W.*(V*H')./(Vap*H');
+            W(isnan(W)) = 0;
+            W = W./repmat(sum(W),size(W,1),1);
+            Vap = W*H;
+
+            H = H.*(W'*V)./(W'*Vap);
+            H(isnan(H)) = 0;
+
+            Vap = W*H;
+            cost(iter) = betadivEXP(V,Vap,beta);
+        end
+    case 1
+        for iter = 1:iteration
+            W = W .* ((V./Vap)*H')./repmat(sum(H,2)',size(V,1),1);
+            W(isnan(W)) = 0;
+            W = W./repmat(sum(W),size(W,1),1);
+            Vap = W*H;
+
+            H = H.*((W'*(V./Vap))./(sparsity+W'*Vap.^(0)));
+            H(isnan(H)) = 0;
+            
+            Vap = W*H;
+            cost(iter) = betadivEXP(V,Vap,beta);
+        end
+    case 0
+        for iter = 1:iteration
+            
+            W = W .* (((V.*Vap.^(-2))*H')./(Vap.^(-1)*H')).^gamma;
+            W = W./repmat(sum(W),size(W,1),1);
+            Vap = W*H;
+            
+            H = H.*((W'*(V.*Vap.^(-2)))./(sparsity+W'*Vap.^(-1))).^(1/2);
+            
+            Vap = W*H;
+            cost(iter) = betadivEXP(V,Vap,beta);
+        end
+    otherwise
+        if beta < 1
+            for iter = 1:iteration
+                if smoothness==0
+                    H = H.*((W'*(V.*Vap.^(beta-2)))./(sparsity+W'*Vap.^(beta-1))).^(gamma);
+                else
+                    H = updateSmoothEssid(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
+                    H(isnan(H)) = 0;
+                end
+                
[…]

### 2018-01-04 experiment/NMFThreshold/nmfth1dictionary.m

@@ -1,0 +1,91 @@
+function [config, store, obs] = nmfth1dictionary(config, setting, data)            
+% nmfth1dictionary DICTIONARY step of the expLanes experiment NMFThreshold         
+%    [config, store, obs] = nmfth1dictionary(config, setting, data)                
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : observations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThreshold('do', 1); return; else store=[]; obs=[]; end
+                                                                                   
+if strcmp(setting.type,'nmf')
+    K = setting.numberElement;
+    inputPath = config.inputPath;
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %% NAME & NUMBER SAMPLE EXTRACTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
+    cityCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*cityCar*'));
+    roadCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*roadCar*'));
+    stopCarSamples = dir(strcat(inputPath,'dictionary',filesep,'*stopCar*'));
+    sample = [cityCarSamples; roadCarSamples; stopCarSamples];
+    indiceClass = ones(1,length(sample));
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %% AUDIO SAMPLE EXTRACTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
+    W = cell(1,length(sample));
+    freqScale = cell(1,length(sample));
+    indTraffic = W;
+    
[…]

### 2018-01-04 experiment/NMFThreshold/nmfth2estimation.m

@@ -1,0 +1,130 @@
+function [config, store, obs] = nmfth2estimation(config, setting, data)            
+% nmfth2estimation ESTIMATION step of the expLanes experiment NMFThreshold         
+%    [config, store, obs] = nmfth2estimation(config, setting, data)                
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : observations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThreshold('do', 2, 'mask', {2 2 1 0 1,...
+            3 0 6 3 2, 0 0 2:3}); return; else store=[]; obs=[]; end
+% à faire :
+% 2 1 1 [1 3] 1,...
+%             1 0 [2 3 4] 3 1, 0 0 2:3
+% 2 2 1 1:2 1,...
+%             1 0 [2 3 4] 3 1, 0 0 2:3
+% 2 1 1 [1 2 4] 1,...
+%             3 0 6 3 1, 0 0 2:3
+
+TIR = setting.TIR;
+dataset = setting.dataset;
+type = setting.type;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DICTIONARY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+if strcmp(type,'nmf')
+    dictionary.W = data.dictionary.W;
+    dictionary.frequency = data.dictionary.frequency;
+    dictionary.indTraffic = data.dictionary.indTraffic;
+    nmf{1}.W0 = dictionary.W;
+else
+    dictionary = [];
+end
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
[…]

### 2018-01-04 experiment/NMFThreshold/nmfth3threshold.m

@@ -1,0 +1,168 @@
+function [config, store, obs] = nmfth3threshold(config, setting, data)
+% nmfth3threshold THRESHOLD step of the expLanes experiment NMFThreshold
+%    [config, store, obs] = nmfth3threshold(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: <userName>
+% Date: 01-Dec-2017
+
+% Set behavior for debug mode
+if nargin==0, NMFThreshold('do', 3, 'mask', {2 2 1 0 1,...
+            3 0 6 3 1, 0 0 2:3 30 1, 1 1}); return; else store=[]; obs=[]; end
+
+levels = cell(1,setting.sceneSelect);
+nmf =  data.nmf;
+
+if strcmp(setting.type,'nmf')
+    distanceMethod = setting.distanceMethod;
+    displayDistance = setting.displayDistance;
+    methodThreshold = setting.methodThreshold;
+    
+    valueThresholdCheck = 1;
+    switch  methodThreshold
+        case 'firm'
+            threshold(1) = setting.thresholdFirmHigh;
+            threshold(2) = setting.thresholdFirmLow;
+            if threshold(1) < threshold(2)
+                valueThresholdCheck = 0;
+            end
+        otherwise
+            threshold = setting.threshold;
+    end
+    
+    [soundMix] = cutSpectrogramEXP(nmf{1}.W0,setting);
+    W0 = soundMix.W(1:soundMix.ind,:);
+    [F,K] = size(W0);
+    
+    switch setting.displayDistance
+        case 'sigmoid'
+            lambda = 1;
+        case 'RBF'
+            lambda = 1;
+        otherwise
+            lambda = 0;
+    end
+    
+    if valueThresholdCheck == 1
+        for ii = 1:setting.sceneSelect
+            W = nmf{ii}.W(1:F,:);
+            H = nmf{ii}.H;
+            
+            switch distanceMethod
+                case 'cosine'
+                    dist = sum(W.*W0)./(sqrt(sum(W.^2,1)).*sqrt(sum(W0.^2,1)));
+                    dist(isnan(dist)) = 0;
+                    [~, order] = sort(dist,'descend');
+                    
+                case 'none'
+                    order = 1:K;
+            end
+            
+            Wn = W(:,order);
[…]

### 2018-01-04 experiment/NMFThreshold/nmfth4metric.m

@@ -1,0 +1,72 @@
+function [config, store, obs] = nmfth4metric(config, setting, data)                
+% nmfth4metric METRIC step of the expLanes experiment NMFThreshold                 
+%    [config, store, obs] = nmfth4metric(config, setting, data)                    
+%      - config : expLanes configuration state                                     
+%      - setting   : set of factors to be evaluated                                
+%      - data   : processing data stored during the previous step                  
+%      -- store  : processing data to be saved for the other steps                 
+%      -- obs    : observations to be saved for analysis                           
+                                                                                   
+% Copyright: <userName>                                                            
+% Date: 01-Dec-2017                                                                
+                                                                                   
+% Set behavior for debug mode                                                      
+if nargin==0, NMFThreshold('do', 4, 'mask', {2 2 1 0 1,...
+            3 0 6 3 1, 0 0 2:3 30 1, 1 1}); return; else store=[]; obs=[]; end
+                                                                                   
+levels = data.levels;
+pref = setting.p0;
+
+LeqTrafficEstimate = zeros(length(levels),1);
+LeqTrafficEstimatedB = zeros(length(levels),1);
+LeqTrafficExact = zeros(length(levels),1);
+LeqTrafficExactdB = zeros(length(levels),1);
+LeqGlobal = zeros(length(levels),1);
+
+%% EXTRACTION LEVELS
+for ii = 1:length(levels)
+    LpTrafficExact = levels{ii}.LpTraffic;
+    LpTrafficExact(LpTrafficExact==0) = pref;
+    
+    LpTrafficEstimate = levels{ii}.LpTrafficEstimate{1};
+    LpTrafficEstimate(LpTrafficEstimate==0) = pref;
+    
+    LpGlobal = levels{ii}.LpGlobal;
+    LpGlobal(LpGlobal==0) = pref;
+       
[…]

### 2018-01-04 experiment/NMFThreshold/nmfthInit.m

@@ -1,0 +1,10 @@
+function [config, store] = nmfthInit(config)                       
+% nmfthInit INITIALIZATION of the expLanes experiment NMFThreshold 
+%    [config, store] = nmfthInit(config)                           
+%      - config : expLanes configuration state                     
+%      -- store  : processing data to be saved for the other steps 
+                                                                   
+% Copyright: <userName>                                            
+% Date: 01-Dec-2017                                                
+                                                                   
+if nargin==0, NMFThreshold(); return; else store=[];  end          

### 2018-01-04 experiment/NMFThreshold/nmfthReport.m

@@ -1,0 +1,151 @@
+function config = nmfthReport(config)                          
+% nmfthReport REPORTING of the expLanes experiment NMFThreshold
+%    config = nmfthInitReport(config)                          
+%       config : expLanes configuration state                  
+                                                               
+% Copyright: <userName>                                        
+% Date: 01-Dec-2017                                            
+                                                               
+% if nargin==0, NMFThreshold('report', 'rcvd','reportName','ThC NMF'); return; end       
+if nargin==0, NMFThreshold('report', 'r'); return; end       
+
+switch config.reportName
+    case 'UnsupervisedThresholded'
+        config = expExpose(config, 't', 'mask', {{1 0 0 0 0, ...
+            0 0 [1 6] 1 1},{2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 2}, {2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 1 1 25}, {2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 1 2 13},{2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 1 3 12},{2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 3 1 0 16 25},{2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 3 2 0 4 8},{2 1 1 3 1, ...
+            3 0 6 1 1, 0 0 3 30 1, 3 3 0 3 7}},...
+            'obs',[5 11],'integrate',{'TIR','aType'},'negativeRank',-2,'sort',1);
+
+
+
+        for TIR = 1:5
+            config = expExpose(config, 't', 'mask', {{1 0 0 0 0, ...
+                0 0 [1 6] 1 1, 0 TIR},{2 1 1 3 1, ...
+                3 0 6 1 1, 0 TIR 3 30 2}, {2 1 1 3 1, ...
+                3 0 6 1 1, 0 TIR 3 30 1, 1 1 25}},...
+                'obs',[5 11],'integrate','aType','negativeRank',-2,'sort',1);
[…]

### 2018-01-04 experiment/NonnegMatrixFact/functions/NarrowToNthOctaveEXP.m

@@ -1,0 +1,94 @@
+function [sortedData,Fc,Flow,Fhigh] = NarrowToNthOctaveEXP(arrayOfNBCenterFreqs,arrayToConvert,n)
+
+df=arrayOfNBCenterFreqs(2)-arrayOfNBCenterFreqs(1);
+
+% %Use correctionFactor=1 if you did not window or you used ECF
+correctionFactor=1;
+% %Use this if you used a hanning window with ACF
+% ACF=(1/mean(window(@hann,2^11)));
+% ECF=(1/mean(window(@hann,2^11)));
+% correctionFactor=1/ACF;
+arrayToConvert=arrayToConvert*correctionFactor;
+
+%Determine Initial Center frequency (Fc)
+previous=1000*2^(1/n);
+lowFc=1000; %Start at 1000 Hz and find lowest Fc
+while (previous-lowFc) > df
+    previous=lowFc;
+    lowFc=lowFc/2^(1/n);
+end
+
+%Compute center frequencies
+ii=1;
+Fc(ii)=lowFc;
+while Fc(ii) <= max(arrayOfNBCenterFreqs) 
+    ii=ii+1;
+    Fc(ii)=2^(1/n)*Fc(ii-1);
+end
+
+%Compute high, low frequencies, and bandwidth (BW) from center frequencies
+    Flow=Fc/sqrt(2^(1/n));
+    Fhigh=Fc*sqrt(2^(1/n));
+    BW=Fhigh-Flow;
+
+%Sort data in frequency bins
+sortedData=zeros(length(Fc),size(arrayToConvert,2)); zeroFreqValue=0;
+for jj=1:size(arrayToConvert,1) %Do for each value given
+    for kk=1:length(Fc) %Check each Fc bin to see where value should be placed
+        if arrayOfNBCenterFreqs(jj)>=Flow(kk) && arrayOfNBCenterFreqs(jj)<Fhigh(kk) %Find place. In the rare case a given center freq = flow(ii)/fhigh(ii+1) then sum the value in flow
+            if sortedData(kk,:)==0 %if no values has been added to the band then set initial value
+                sortedData(kk,:)=arrayToConvert(jj,:);
+                N(kk)=1;
+            else %else sum in the value
+                sortedData(kk,:)=sqrt(sum(([sortedData(kk,:); arrayToConvert(jj,:)].^2))); %dB add value to that bin
+                N(kk)=N(kk)+1;
+            end
+        else %else check to see if the value belongs in the 0 Hz band
[…]

### 2018-01-04 experiment/NonnegMatrixFact/functions/audio2SpectrogramEXP.m

@@ -1,0 +1,41 @@
+function [X,time,frequency,Xlinear,XScale] = audio2SpectrogramEXP(x,setting)
+%% [X,Xlinear,time,XScale,frequency] = audio2Spectrogram(x,setting)
+
+sr = setting.sr;
+nfft = setting.nfft;
+window = setting.nfft;
+noverlap = floor(window*setting.noverlap/100);
+domain = setting.domain; 
+
+if size(x,1)>size(x,2)
+    x = x';
+end
+
+X = cell(1,size(x,1));
+XScale = X;
+frequency = X;
+time = X;
+Xlinear = X;
+
+for ii = 1:size(x,1)
+    [spec,frequency{ii},time{ii}] = spectrogram(x(ii,:),window,noverlap,nfft,sr);
+    spec = spec./nfft;
+    Xlinear{ii} = abs(spec);
+    
+    if strcmp('thirdOctave',domain)
+        [X{ii},XScale{ii},~,~] = NarrowToNthOctaveEXP(frequency{ii},Xlinear{ii},3);
+    elseif strcmp('mel',domain)
+        [X{ii},XScale{ii}] = spectre2MelEXP(Xlinear{ii},setting.numberMel,setting.sr);
+    else
+        X{ii} = Xlinear{ii};
+        XScale{ii} = frequency{ii};
+    end
+end
+
+if ii == 1
+   X = X{1};
+   Xlinear = Xlinear{1};
+   XScale = XScale{1};
+   frequency = frequency{1};
+   time = time{1};
+end

### 2018-01-04 experiment/NonnegMatrixFact/functions/creationDictionaryEXP.m

@@ -1,0 +1,45 @@
+function [W,Xscale,indTraffic] = creationDictionaryEXP(file,indClass,setting)
+
+
+if min(size(file)) == 2        % stereo to mono
+    file = mean(file,2)';
+else
+    file = file';
+end
+
+spec = audio2SpectrogramEXP(file,setting);
+
+switch setting.frameSize
+    case 0 % W composed of fft of sounds get from the rms of the spectrogram
+        X = rms(spec,2);
+        
+    otherwise   % W composed of several temporal window get by the rms of the selected window
+        step = (setting.window*(setting.noverlap/100))/setting.sr;
+        time = step:step:size(spec,2)*step;
+        frameSize_temp = setting.frameSize*1e-3;    % size of the temporal window
+        timeEnd = floor(time(end));
+        if time(end) < 1
+           timeEnd = frameSize_temp; 
+        end
+        [~,idxBeg] = min(abs(repmat((0:frameSize_temp:timeEnd-frameSize_temp)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
+        [~,idxEnd] = min(abs(repmat((frameSize_temp:frameSize_temp:timeEnd)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
+
+        if length(idxEnd) < floor(time(end)/frameSize_temp)
+            idxBeg = [idxBeg; idxEnd(end)+1];
+            idxEnd = [idxEnd; length(time)];
+        end
+        X = zeros(size(spec{1},1),length(idxEnd));
+        
+        for ii = 1:length(idxBeg)
+            X(:,ii) = rms(spec{1}(:,idxBeg(ii):idxEnd(ii)-1),2);
+        end
+end
+
+Xscale = linspace(0,setting.sr/2,size(X,1));
+W = X./repmat(sum(X),size(X,1),1);
+
+if indClass == 1
+    indTraffic = ones(1,size(W,2));
+else
+    indTraffic = zeros(1,size(W,2));
+end

### 2018-01-04 experiment/NonnegMatrixFact/functions/estimationLpEXP.m

@@ -1,0 +1,45 @@
+function [Lp,Leq] = estimationLpEXP(Xinput,setting)
+
+if ~iscell(Xinput)
+    Xin{1} = Xinput;
+else
+    Xin = Xinput;
+end
+
+sr = setting.sr;
+window = setting.window;
+noverlap = setting.noverlap;
+step = (window*(noverlap/100))/sr;
+
+Lp = cell(1,size(Xin,2));
+Leq = zeros(1,size(Xin,2));
+
+for ii = 1:size(Xin,2)
+    X = Xin{ii};
+    
+    if strcmp('mel',setting.domain)
+        X = mel2SpectreEXP(X,sr,setting.nfft);
+    end
+    
+    X(2:end,:) = X(2:end,:)*2;
+    time =  step:step:size(X,2)*step;
+    
+    B = sqrt(abs(X).^2);
+    B(2:end,:) = sqrt(2)*B(2:end,:);
+    Bw = B/sqrt(1.5);           %hanning window effect "scaling effect" see https://www.ap.com/technical-library/fft-scaling-for-noise/
+
+    a = sqrt(sum(Bw.^2));
+    Leq(ii) = rms(a);
+    
+    T_tau = floor(time(end)/setting.temporalInterval); % nombre interval
+    N_tau = floor(length(time)/T_tau); % nombre echantilon par interval
+    ind_beg = linspace(0,(T_tau-1)*N_tau,T_tau);
+    ind_end = linspace(N_tau,T_tau*N_tau,T_tau);
+    
+    Lp{ii} = zeros(1,T_tau);
+    for jj = 1:T_tau
+        Lp{ii}(:,jj) = rms(a(ind_beg(jj)+1:ind_end(jj)));
+    end
+    Lp{ii}(isnan(Lp{ii}))=0;
+    Lp{ii}(isinf(Lp{ii}))=0;
+end

### 2018-01-04 experiment/NonnegMatrixFact/functions/mel2SpectreEXP.m

@@ -1,0 +1,13 @@
+function Y = mel2SpectreEXP(X,sr,nfft)
+
+if iscell(X)
+    Y = cell(size(X,1),size(X,2));
+    for ii = 1:size(X,1)
+        for jj = 1:size(X,2)
+            [Y{ii,jj}] = invaudspec(X{ii,jj}, sr, nfft, 'mel', 0, sr/2, 0, 1);
+        end
+        
+    end
+else
+    [Y] = invaudspec(X, sr, nfft, 'mel', 0, sr/2, 0, 1);
+end

### 2018-01-04 experiment/NonnegMatrixFact/functions/nmf/NMFestimationEXP.m

@@ -1,0 +1,62 @@
+function [LeqTrafficEstimate,LpTrafficEstimate,H,cost,sequentialData] = NMFestimationEXP(V,dictionary,setting,sequentialData)
+
+
+%% limit the spectrogram and the dictionary to cutOffFreq
+[soundMix] = cutSpectrogramEXP(dictionary,setting);
+
+W = soundMix.W(1:soundMix.ind,:);
+V = V(1:soundMix.ind,:);
+binTemp = size(V,2);
+
+switch setting.nmfType
+    case 'supervised'
+        
+        if isempty(sequentialData)
+            iteration = setting.iteration;
+            rng(soundMix.seed)
+            H = rand(size(W,2),binTemp);
+        else
+            % continuing step of the sequential run
+            iteration = setting.iteration-sequentialData.numberIteration;
+            H = sequentialData.H;
+        end
+
+        NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting);
+        cost = betadivEXP(V,NMF.Vap,setting.beta,NMF.H,setting.sparsity,setting.SM_weight);
+
+    case 'semi-supervised'
+        
+        Wrand = setting.SS_sizeWrand;
+        if isempty(sequentialData)
+            iteration = setting.iteration;
+            rng(soundMix.seed); Y =  rand(size(W,1),Wrand);
+            rng(soundMix.seed); H = rand(size(W,2),binTemp);
+            rng(soundMix.seed); Z = rand(Wrand,binTemp);
+        else
+            % continuing step of the sequential run
+            iteration = setting.iteration-sequentialData.numberIteration;
+            H = sequentialData.H;
+            Z = sequentialData.Z;
+            Y = sequentialData.Y;
+        end
+
+        NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting);
+        cost = betadivSemiEXP(V,NMF,setting);
+        
+        sequentialData.Y = NMF.Y;
+        sequentialData.Z = NMF.Z;
+		
+end
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% extraction traffic %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
[…]

### 2018-01-04 experiment/NonnegMatrixFact/functions/spectre2MelEXP.m

@@ -1,0 +1,13 @@
+function [Y,binMel,wts] = spectre2MelEXP(X,numberMel,sr)
+%% function [Y,binMel,wts] = spectre2MelEXP(X,numberMel,sr)
+
+if iscell(X)
+    Y = cell(size(X,1),size(X,2));
+    for ii = 1:size(X,1)
+        for jj = 1:size(X,2)
+            [Y{ii,jj},wts,binMel] = audspec(X{ii,jj}, sr, numberMel,'mel', 0, sr/2, 0, 1);
+        end
+    end
+else 
+    [Y,wts,binMel] = audspec(X, sr, numberMel,'mel', 0, sr/2, 0, 1);
+end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/MFCC/audspec.m

@@ -1,40 +1,0 @@
-function [aspectrum,wts,binfrqs] = audspec(pspectrum, sr, nfilts, fbtype, minfreq, maxfreq, sumpower, bwidth)
-%[aspectrum,wts] = audspec(pspectrum, sr, nfilts, fbtype, minfreq, maxfreq, sumpower, bwidth)
-%
-% perform critical band analysis (see PLP)
-% takes power spectrogram as input
-
-if nargin < 2;  sr = 16000;                          end
-if nargin < 3;  nfilts = ceil(hz2bark(sr/2))+1;      end
-if nargin < 4;  fbtype = 'bark';  end
-if nargin < 5;  minfreq = 0;    end
-if nargin < 6;  maxfreq = sr/2; end
-if nargin < 7;  sumpower = 1;   end
-if nargin < 8;  bwidth = 1.0;   end
-
-[nfreqs,~] = size(pspectrum);
-
-nfft = (nfreqs-1)*2;
-
-if strcmp(fbtype, 'bark')
-  [wts,binfrqs] = fft2barkmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
-elseif strcmp(fbtype, 'mel')
-  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
-elseif strcmp(fbtype, 'htkmel')
-  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 1);
-elseif strcmp(fbtype, 'fcmel')
-  [wts,binfrqs] = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 0);
-else
-  disp(['fbtype ', fbtype, ' not recognized']);
-  error;
-end
-
-wts = wts(:, 1:nfreqs);
-
-% Integrate FFT bins into Mel bins, in abs or abs^2 domains:
-if (sumpower)
-  aspectrum = wts * pspectrum;
-else
-  aspectrum = (wts * sqrt(pspectrum)).^2;
-end
-

### 2018-01-04 experiment/NonnegMatrixFact/netlab/MFCC/invaudspec.m

@@ -1,43 +1,0 @@
-function [spec,wts,iwts] = invaudspec(aspectrum, sr, nfft, fbtype, minfreq, maxfreq, sumpower, bwidth)
-%pspectrum = invaudspec(aspectrum, sr, nfft, fbtype, minfreq, maxfreq, sumpower, bwidth)
-%
-% Invert (as best we can) the effects of audspec()
-%
-% 2004-02-04 dpwe@ee.columbia.edu
-
-if nargin < 2;  sr = 16000;     end 
-if nargin < 3;  nfft = 512;     end
-if nargin < 4;  fbtype = 'bark';  end
-if nargin < 5;  minfreq = 0;    end
-if nargin < 6;  maxfreq = sr/2; end
-if nargin < 7;  sumpower = 1;   end
-if nargin < 8;  bwidth = 1.0;   end
-
-[nfilts,nframes] = size(aspectrum);
-
-if strcmp(fbtype, 'bark')
-  wts = fft2barkmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
-elseif strcmp(fbtype, 'mel')
-  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq);
-elseif strcmp(fbtype, 'htkmel')
-  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1, 1);
-elseif strcmp(fbtype, 'fcmel')
-  wts = fft2melmx(nfft, sr, nfilts, bwidth, minfreq, maxfreq, 1);
-else
-  disp(['fbtype ', fbtype, ' not recognized']);
-  error;
-end
-
-% Cut off 2nd half
-wts = wts(:,1:((nfft/2)+1));
-
-% Just transpose, fix up 
-ww = wts'*wts;
-iwts = wts'./(repmat(max(mean(diag(ww))/100, sum(ww))',1,nfilts));
-% Apply weights
-if (sumpower)
-  spec = iwts * aspectrum;
-else
-  spec = 2*(iwts * sqrt(aspectrum)).^2;
-end
-

### 2018-01-04 experiment/NonnegMatrixFact/netlab/MFCC/ispecgram.m

@@ -1,48 +1,0 @@
-function x = ispecgram(d, ftsize, sr, win, nov)
-% X = ispecgram(D, F, SR, WIN, NOV)           Inverse specgram
-%    Overlap-add the inverse of the output of specgram
-%    ftsize is implied by sizeof d, sr is ignored, nov defaults to ftsize/2
-% dpwe 2005may16.  after istft
-
-[nspec,ncol] = size(d);
-
-if nargin < 2
-  ftsize = 2*(nspec-1);
-end
-if nargin < 3
-  % who cares?
-end
-if nargin < 4
-  win = ftsize;  % doesn't matter either - assume it added up OK
-end
-if nargin < 5
-  nov = ftsize/2;
-end
-
-hop = win - nov;
-
-if nspec ~= (ftsize/2)+1
-  error('number of rows should be fftsize/2+1')
-end
-
-xlen = ftsize + (ncol-1) * hop;
-x = zeros(xlen,1);
-
-halff = ftsize/2;   % midpoint of win
-
-% No reconstruction win (for now...)
-
-for c = 1:ncol
-  ft = d(:,c);
-  ft = [ft(1:(ftsize/2+1)); conj(ft([(ftsize/2):-1:2]))];
-
-  if max(imag(ifft(ft))) > 1e-5
-    disp('imag oflow');
-  end
-  
-  px = real(ifft(ft));  % no shift in specgram
-  
-  b = (c-1)*hop;
-  x(b+[1:ftsize]) = x(b+[1:ftsize]) + px;
-end;
-

### 2018-01-04 experiment/NonnegMatrixFact/netlab/NarrowToNthOctaveEXP.m

@@ -1,94 +1,0 @@
-function [sortedData,Fc,Flow,Fhigh] = NarrowToNthOctaveEXP(arrayOfNBCenterFreqs,arrayToConvert,n)
-
-df=arrayOfNBCenterFreqs(2)-arrayOfNBCenterFreqs(1);
-
-% %Use correctionFactor=1 if you did not window or you used ECF
-correctionFactor=1;
-% %Use this if you used a hanning window with ACF
-% ACF=(1/mean(window(@hann,2^11)));
-% ECF=(1/mean(window(@hann,2^11)));
-% correctionFactor=1/ACF;
-arrayToConvert=arrayToConvert*correctionFactor;
-
-%Determine Initial Center frequency (Fc)
-previous=1000*2^(1/n);
-lowFc=1000; %Start at 1000 Hz and find lowest Fc
-while (previous-lowFc) > df
-    previous=lowFc;
-    lowFc=lowFc/2^(1/n);
-end
-
-%Compute center frequencies
-ii=1;
-Fc(ii)=lowFc;
-while Fc(ii) <= max(arrayOfNBCenterFreqs) 
-    ii=ii+1;
-    Fc(ii)=2^(1/n)*Fc(ii-1);
-end
-
-%Compute high, low frequencies, and bandwidth (BW) from center frequencies
-    Flow=Fc/sqrt(2^(1/n));
-    Fhigh=Fc*sqrt(2^(1/n));
-    BW=Fhigh-Flow;
-
-%Sort data in frequency bins
-sortedData=zeros(length(Fc),size(arrayToConvert,2)); zeroFreqValue=0;
-for jj=1:size(arrayToConvert,1) %Do for each value given
-    for kk=1:length(Fc) %Check each Fc bin to see where value should be placed
-        if arrayOfNBCenterFreqs(jj)>=Flow(kk) && arrayOfNBCenterFreqs(jj)<Fhigh(kk) %Find place. In the rare case a given center freq = flow(ii)/fhigh(ii+1) then sum the value in flow
-            if sortedData(kk,:)==0 %if no values has been added to the band then set initial value
-                sortedData(kk,:)=arrayToConvert(jj,:);
-                N(kk)=1;
-            else %else sum in the value
-                sortedData(kk,:)=sqrt(sum(([sortedData(kk,:); arrayToConvert(jj,:)].^2))); %dB add value to that bin
-                N(kk)=N(kk)+1;
-            end
-        else %else check to see if the value belongs in the 0 Hz band
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/audio2SpectrogramEXP.m

@@ -1,41 +1,0 @@
-function [X,time,frequency,Xlinear,XScale] = audio2SpectrogramEXP(x,setting)
-%% [X,Xlinear,time,XScale,frequency] = audio2Spectrogram(x,setting)
-
-sr = setting.sr;
-nfft = setting.nfft;
-window = setting.nfft;
-noverlap = floor(window*setting.noverlap/100);
-domain = setting.domain; 
-
-if size(x,1)>size(x,2)
-    x = x';
-end
-
-X = cell(1,size(x,1));
-XScale = X;
-frequency = X;
-time = X;
-Xlinear = X;
-
-for ii = 1:size(x,1)
-    [spec,frequency{ii},time{ii}] = spectrogram(x(ii,:),window,noverlap,nfft,sr);
-    spec = spec./nfft;
-    Xlinear{ii} = abs(spec);
-    
-    if strcmp('thirdOctave',domain)
-        [X{ii},XScale{ii},~,~] = NarrowToNthOctaveEXP(frequency{ii},Xlinear{ii},3);
-    elseif strcmp('mel',domain)
-        [X{ii},XScale{ii}] = spectre2MelEXP(Xlinear{ii},setting.numberMel,setting.sr);
-    else
-        X{ii} = Xlinear{ii};
-        XScale{ii} = frequency{ii};
-    end
-end
-
-if ii == 1
-   X = X{1};
-   Xlinear = Xlinear{1};
-   XScale = XScale{1};
-   frequency = frequency{1};
-   time = time{1};
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/creationDictionaryEXP.m

@@ -1,45 +1,0 @@
-function [W,Xscale,indTraffic] = creationDictionaryEXP(file,indClass,setting)
-
-
-if min(size(file)) == 2        % stereo to mono
-    file = mean(file,2)';
-else
-    file = file';
-end
-
-spec = audio2SpectrogramEXP(file,setting);
-
-switch setting.frameSize
-    case 0 % W composed of fft of sounds get from the rms of the spectrogram
-        X = rms(spec,2);
-        
-    otherwise   % W composed of several temporal window get by the rms of the selected window
-        step = (setting.window*(setting.noverlap/100))/setting.sr;
-        time = step:step:size(spec,2)*step;
-        frameSize_temp = setting.frameSize*1e-3;    % size of the temporal window
-        timeEnd = floor(time(end));
-        if time(end) < 1
-           timeEnd = frameSize_temp; 
-        end
-        [~,idxBeg] = min(abs(repmat((0:frameSize_temp:timeEnd-frameSize_temp)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
-        [~,idxEnd] = min(abs(repmat((frameSize_temp:frameSize_temp:timeEnd)',1,length(time))-repmat(time,timeEnd/frameSize_temp,1)),[],2);
-
-        if length(idxEnd) < floor(time(end)/frameSize_temp)
-            idxBeg = [idxBeg; idxEnd(end)+1];
-            idxEnd = [idxEnd; length(time)];
-        end
-        X = zeros(size(spec{1},1),length(idxEnd));
-        
-        for ii = 1:length(idxBeg)
-            X(:,ii) = rms(spec{1}(:,idxBeg(ii):idxEnd(ii)-1),2);
-        end
-end
-
-Xscale = linspace(0,setting.sr/2,size(X,1));
-W = X./repmat(sum(X),size(X,1),1);
-
-if indClass == 1
-    indTraffic = ones(1,size(W,2));
-else
-    indTraffic = zeros(1,size(W,2));
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/cutSpectrogramEXP.m

@@ -1,24 +1,0 @@
-function [soundMix] = cutSpectrogramEXP(dictionary,setting)
-
-W = dictionary.W;
-
-switch setting.domain
-    case 'mel'
-        sr = setting.sr;
-        numberMel = setting.numberMel; 
-
-        [W,~] = spectre2MelEXP(W,numberMel,sr);
-        ind = numberMel;
-        
-    case 'thirdOctave'
-        [W,~,~,~] = NarrowToNthOctaveEXP(dictionary.frequency,W,3);
-        ind = size(W,1);
-        
-    otherwise
-        [~,ind] = min(abs(setting.cutOffFreq-dictionary.frequency));
-end
-
-soundMix.ind = ind;
-soundMix.seed = rng;
-soundMix.W = W;
-soundMix.indTraffic = dictionary.indTraffic;

### 2018-01-04 experiment/NonnegMatrixFact/netlab/estimationLpEXP.m

@@ -1,45 +1,0 @@
-function [Lp,Leq] = estimationLpEXP(Xinput,setting)
-
-if ~iscell(Xinput)
-    Xin{1} = Xinput;
-else
-    Xin = Xinput;
-end
-
-sr = setting.sr;
-window = setting.window;
-noverlap = setting.noverlap;
-step = (window*(noverlap/100))/sr;
-
-Lp = cell(1,size(Xin,2));
-Leq = zeros(1,size(Xin,2));
-
-for ii = 1:size(Xin,2)
-    X = Xin{ii};
-    
-    if strcmp('mel',setting.domain)
-        X = mel2SpectreEXP(X,sr,setting.nfft);
-    end
-    
-    X(2:end,:) = X(2:end,:)*2;
-    time =  step:step:size(X,2)*step;
-    
-    B = sqrt(abs(X).^2);
-    B(2:end,:) = sqrt(2)*B(2:end,:);
-    Bw = B/sqrt(1.5);           %hanning window effect "scaling effect" see https://www.ap.com/technical-library/fft-scaling-for-noise/
-
-    a = sqrt(sum(Bw.^2));
-    Leq(ii) = rms(a);
-    
-    T_tau = floor(time(end)/setting.temporalInterval); % nombre interval
-    N_tau = floor(length(time)/T_tau); % nombre echantilon par interval
-    ind_beg = linspace(0,(T_tau-1)*N_tau,T_tau);
-    ind_end = linspace(N_tau,T_tau*N_tau,T_tau);
-    
-    Lp{ii} = zeros(1,T_tau);
-    for jj = 1:T_tau
-        Lp{ii}(:,jj) = rms(a(ind_beg(jj)+1:ind_end(jj)));
-    end
-    Lp{ii}(isnan(Lp{ii}))=0;
-    Lp{ii}(isinf(Lp{ii}))=0;
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/filtrePasseBasEXP.m

@@ -1,22 +1,0 @@
-function [LeqFiltre,LpFiltre] = filtrePasseBasEXP(X,setting)
-%% Y = filtrePasseBas(X,soundMix,varargin)
-% Fonction filtre pour un signal temporel x filtrer par un filtre de
-% Butterworth. Les paramètres se trouvent dans le structure soundMix :
-% - la fréquence de coupure, .cutOffFreq,
-% - l'ordre du filtre, .order.
-% Si ils ne sont pas présent : par défaut cutOffFreq = 1000 et order = 2.
-% Si on renseigne un troisième argument d'une valeur quelconque, on trace
-% la fonction filtre par la fonction fvtool.
-
-sr = setting.sr;
-nfft = setting.nfft;
-fc = setting.cutOffFreq;
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% FONCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-frequency = linspace(0,sr/2,nfft/2+1);
-[~,ind] = min(abs(fc-frequency));
-X{1} = X{1}(1:ind,:);
-
-[LpFiltre,LeqFiltre] = estimationLpEXP(X,setting);
-

### 2018-01-04 experiment/NonnegMatrixFact/netlab/mel2SpectreEXP.m

@@ -1,13 +1,0 @@
-function Y = mel2SpectreEXP(X,sr,nfft)
-
-if iscell(X)
-    Y = cell(size(X,1),size(X,2));
-    for ii = 1:size(X,1)
-        for jj = 1:size(X,2)
-            [Y{ii,jj}] = invaudspec(X{ii,jj}, sr, nfft, 'mel', 0, sr/2, 0, 1);
-        end
-        
-    end
-else
-    [Y] = invaudspec(X, sr, nfft, 'mel', 0, sr/2, 0, 1);
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/NMFestimationEXP.m

@@ -1,85 +1,0 @@
-function [LeqTrafficEstimate,LpTrafficEstimate,H,cost,sequentialData] = NMFestimationEXP(V,dictionary,setting,sequentialData)
-
-
-%% limit the spectrogram and the dictionary to cutOffFreq
-[soundMix] = cutSpectrogramEXP(dictionary,setting);
-
-W = soundMix.W(1:soundMix.ind,:);
-V = V(1:soundMix.ind,:);
-binTemp = size(V,2);
-
-switch setting.nmfType
-    case 'supervised'
-        
-        if isempty(sequentialData)
-            iteration = setting.iteration;
-            rng(soundMix.seed)
-            H = rand(size(W,2),binTemp);
-        else
-            % continuing step of the sequential run
-            iteration = setting.iteration-sequentialData.numberIteration;
-            H = sequentialData.H;
-        end
-
-        NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting);
-        cost = betadivEXP(V,NMF.Vap,setting.beta,NMF.H,setting.sparsity,setting.SM_weight);
-
-    case 'semi-supervised'
-        
-        Wrand = setting.SS_sizeWrand;
-        if isempty(sequentialData)
-            iteration = setting.iteration;
-            rng(soundMix.seed); Y =  rand(size(W,1),Wrand);
-            rng(soundMix.seed); H = rand(size(W,2),binTemp);
-            rng(soundMix.seed); Z = rand(Wrand,binTemp);
-        else
-            % continuing step of the sequential run
-            iteration = setting.iteration-sequentialData.numberIteration;
-            H = sequentialData.H;
-            Z = sequentialData.Z;
-            Y = sequentialData.Y;
-        end
-
-        NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting);
-        cost = betadivSemiEXP(V,NMF,setting);
-        
-        sequentialData.Y = NMF.Y;
-        sequentialData.Z = NMF.Z;
-		
-	case 'unsupervised'
-		W0 = soundMix.W(1:soundMix.ind,:);      % initial dictionary
-		V = V{1}(1:soundMix.ind,:);    % spectrogram to approximate
-
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/algo_nmfSemiSupervisedEXP.m

@@ -1,183 +1,0 @@
-function NMF = algo_nmfSemiSupervisedEXP(H,W,Y,Z,V,iteration,soundMix,setting)
-
-beta = setting.beta;
-sparsity = setting.sparsity;
-smoothness = setting.SM_weight;
-J = setting.SS_sizeWrand;
-K = size(W,2);
-
-if setting.SS_weight ~= 0
-    mu = setting.SS_weight;
-    betaM = setting.SS_betaM;
-    lambdaSS = setting.SS_lambda;
-else
-    mu = 0;
-end
-
-if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
-    if beta < 1
-        gamma = 1/(2-beta);
-    elseif beta >= 1 && beta <= 2
-        gamma = 1;
-    else
-        gamma = 1/(beta-1);
-    end
-end
-
-if smoothness~=0
-    smoothForm = setting.smoothnessForm;
-    
-    switch smoothForm
-        case 'all'
-            lambdaSM = [sum(W,1) sum(Y,1)];
-            smoothnessTot(1:K+J,1) = smoothness;
-        case 'traffic'
-            lambdaSM = sum(W,1);
-            smoothnessTot = zeros(K,1);
-            smoothnessTot(soundMix.indTraffic==1) = smoothness;
-    end
-else
-    smoothnessTot = 0;
-    lambdaSM = NaN;
-end
-
-Vap = W*H;
-Vtot = Vap+Y*Z;
-
-NMF = [];
-
-switch beta
-    case 2
-        for iter = 1:iteration
-            if mu == 0
-                Y = Y.*(((V*Z')./(Vtot*Z')).^gamma);
-            else
-                div = betadivEXP(Y(:)*ones(1,size(W,2)),repmat(W,size(Y,2),1),betaM);
-                Cm = exp((-1/lambdaSS)*sum(div));
-
-                num = lambdaSS*V*Z'+mu*Y.^(betaM-1)*Cm;
-                den = (lambdaSS*Vtot)*Z'+mu*Y.^(betaM-2).*Cm.*sum(W,2);
-                Y = Y.*(num./den);
-                Y(Y<2e-5) = 2e-5;
-            end
-            Y = Y./repmat(sum(Y),size(Y,1),1);
-            Vtot = Vap+Y*Z;
-            
-            if smoothness == 0 || strcmp(smoothForm,'traffic')
-                Z = Z.*(((Y'*V)./(Y'*Vtot)).^gamma);
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/algo_nmfSupervisedEXP.m

@@ -1,101 +1,0 @@
-function NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
-%% NMF = algo_nmfSupervisedEXP(H,W,V,iteration,soundMix,setting)
-
-sparsity = setting.sparsity; 
-smoothness = setting.SM_weight; 
-beta = setting.beta;
-Vap = W*H;
-
-NMF = [];
-
-if ~isequal(beta,0) || ~isequal(beta,1) || ~isequal(beta,2)
-    if beta < 1
-        gamma = 1/(2-beta);
-    elseif beta >= 1 && beta <= 2
-        gamma = 1;
-    else
-        gamma = 1/(beta-1);
-    end
-end
-
-if smoothness~=0
-    lambda = sum(W,1);
-    smoothForm = setting.smoothnessForm;
-    
-    switch smoothForm
-        case 'all'
-            smoothnessTot(1:size(W,2),1) = smoothness;
-        case 'traffic'
-            smoothnessTot = zeros(size(W,2),1);
-            smoothnessTot(soundMix.indTraffic==1) = smoothness;
-    end
-else
-    smoothnessTot = 0;
-    lambda = NaN;
-end
-
-switch beta
-    case 2
-        for iter = 1:iteration(end)
-            if smoothness==0
-                H = H.*(W'*V-sparsity)./(W'*Vap);
-                H(H<0)=0;
-            else
-                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
-            end
-            H(isnan(H)) = 0;
-            Vap = W*H;
-        end
-    case 1
-        for iter = 1:iteration
-            
-            if smoothness==0
-                H = H.*((W'*(V.*Vap.^(-1)))./(sparsity+W'*Vap.^(0)));
-            else
-                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
-            end
-            H(isnan(H)) = 0;
-            Vap = W*H;
-            
-        end
-    case 0
-        for iter = 1:iteration
-            
-            if smoothness==0
-                H = H.*((W'*(V.*Vap.^(-2)))./(sparsity+W'*Vap.^(-1))).^(1/2);
-            else
-                H = updateSmoothEssidEXP(H,W,V,Vap,beta,smoothnessTot,sparsity,lambda);
-            end
-            H(isnan(H)) = 0;
-            Vap = W*H;
-
-        end
-    otherwise
-        if beta < 1
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/betadivEXP.m

@@ -1,41 +1,0 @@
-function cost = betadivEXP(V,Vap,beta,varargin)
-
-if nargin > 3
-    H = varargin{1};
-    sparsity = varargin{2};
-    smoothness = varargin{3};
-else
-    H = [];
-    sparsity = 0;
-    smoothness = 0;
-end
-
-switch beta
-    case 2
-        cost = sum((V(:)-Vap(:)).^2)/2;
-    case 1
-        ind_0 = find(V(:)<=eps);
-        ind_1 = 1:length(V(:));
-        ind_1(ind_0) = [];
-        
-        cost = sum(V(ind_1).*log(V(ind_1)./Vap(ind_1)) - V(ind_1) + Vap(ind_1) ) + sum(Vap(ind_0));        
-    case 0
-        ind_0 = find(V(:)<=eps);
-        ind_1 = 1:length(V(:));
-        ind_1(ind_0) = [];
-        
-        cost = sum(V(ind_1)./Vap(ind_1)-log(V(ind_1)./Vap(ind_1)))-length(V(:))+sum(Vap(ind_0));
-
-    otherwise
-        cost = sum(V(:).^beta+(beta-1)*Vap(:).^beta-beta*V(:).*Vap(:).^(beta-1))/(beta*(beta-1));
-end
-
-if sparsity~=0
-    cost = cost+sparsity*sum(H(:));
-end
-
-if smoothness~=0
-    cost = cost+0.5*smoothness*sum(sum((H(:,1:end-1)-H(:,2:end)).^2));
-end
-    
-

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/betadivSemiEXP.m

@@ -1,57 +1,0 @@
-function cost = betadivSemiEXP(V,NMF,setting,varargin)
-
-
-beta = setting.beta;
-sparsity = setting.sparsity;
-smoothness = setting.SM_weight;
-betaM = setting.SS_betaM;
-mu = setting.SS_weight;
-lambda = setting.SS_sensitivity;
-
-H = NMF.H;
-W = NMF.W;
-Y = NMF.Y;
-Vtot = NMF.Vtot;
-
-term = zeros(1,size(W,2));
-if mu == 0
-    cost = betadivEXP(V,Vtot,beta,H,sparsity,smoothness);
-else
-
-    div = betadivEXP(Y(:)*ones(1,size(W,2)),repmat(W,size(Y,2),1),betaM);
-	Cm = exp((-1/lambda)*sum(div));
-
-    switch beta
-        case 2
-            cost = sum((V(:)-Vtot(:)).^2)/2+mu*sum(term);
-        case 1
-            ind_0 = find(V(:)<=eps);
-            ind_1 = 1:length(V(:));
-            ind_1(ind_0) = [];
-            cost = sum(V(ind_1).*log(V(ind_1)./Vtot(ind_1))-V(ind_1)+Vtot(ind_1))+sum(Vtot(ind_0))...
-                +mu*Cm;
-        case 0
-            ind_0 = find(V(:)<=eps);
-            ind_1 = 1:length(V(:));
-            ind_1(ind_0) = [];
-
-            cost = sum(V(ind_1)./Vtot(ind_1)-log(V(ind_1)./Vtot(ind_1)))-length(V(:))+sum(Vtot(ind_0))...
-                +mu*Cm;
-
-        otherwise
-            cost = sum(V(:).^beta + (beta-1)*Vtot(:).^beta - beta*V(:).*Vtot(:).^(beta-1) )/(beta*(beta-1))...
-                +mu*Cm;
-    end
-end
-if sparsity~=0
-    cost = cost+sparsity*sum(H(:));
-end
-
-if smoothness~=0
-    if strcmp(setting.smoothnessForm,'all')
-        basis = [H;NMF.Z];
-        cost = cost+0.5*smoothness*sum((sum((basis(:,1:end-1)-basis(:,2:end)).^2)));
-    else
-        cost = cost+0.5*smoothness*sum((sum((H(:,1:end-1)-H(:,2:end)).^2)));
-    end
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/extractionLevelEXP.m

@@ -1,58 +1,0 @@
-function NMF = extractionLevelEXP(NMF,soundMix,V,beta,setting,varargin)
-
-sparsity = setting.sparsity; 
-smoothness = setting.smoothness; 
-nmfType = setting.nmfType; 
-
-Vap = NMF.Vap;
-H = NMF.H;
-
-if strcmp(setting.domain,'mel')
-    Y = mel2SpectreEXP([{V} {Vap}],setting.sr,setting.nfft,setting.cutOffFreq);
-    V = Y{1};
-    Vap = Y{2};
-end
-
-switch nmfType
-    case 'supervised'
-        costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness);
-        dictionary = soundMix.W;
-        activator = H;
-        
-    case 'semi-supervised'
-        Y = varargin{1};
-        Z = varargin{2};
-        size_Wrand = setting.SS_sizeWrand; 
-        constraint = setting.SS_constraint; 
-
-        YFull = [Y; zeros(size(soundMix.Wfull,1)-size(Y,1),size_Wrand)];
-        dictionary = [soundMix.W YFull];
-        activator = [H; Z];
-
-        switch constraint
-            case 0
-                costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness);
-
-            otherwise
-                betaM = setting.SS_betaM;
-                lambda = setting.SS_lambda;
-
-                div = betadivEXP(repmat(soundMix.Wcut,size(Y,2),1),Y(:)*ones(1,size(soundMix.Wcut,2)),betaM);
-                Cp = exp((-1/lambda)*sum(div));
-                costTemp = betadivEXP(V,Vap,beta,H,sparsity,smoothness)+Cp;
-        end
-        NMF.Y = Y;
-        NMF.Z = Z;
-end
-
-Vap_full = dictionary*activator;
-
-[Lp,Leq] = estimationLpEXP([{Vap} {Vap_full}],setting);
-
-NMF.H = H;
-NMF.cost = costTemp;
-NMF.LeqGlobalEstimate = Leq;
-NMF.LpGlobalEstimate = Lp;
-
-NMF = separationTrafficEXP(NMF,soundMix,setting);
-NMF = orderfields(NMF);

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/separationTrafficEXP.m

@@ -1,26 +1,0 @@
-function [NMF] = separationTrafficEXP(NMF,soundMix,setting)
-%% [NMF] = separationActivateur(NMF,soundMix,ind)
-% séparation des activateurs pour chaque classe du dictionnaire
-%
-
-Htraffic = NMF.H(soundMix.indTraffic==1,:);
-Wtraffic = soundMix.W(:,soundMix.indTraffic==1);
-
-Vtraffic_full = Wtraffic*Htraffic;
-
-if strcmp(setting.domain,'mel')
-    
-    Vtraffic_full = mel2SpectreEXP(Vtraffic_full,setting.sr,setting.nfft);
-    [Lp,Leq] = estimationLpEXP(Vtraffic_full,setting);
-    
-elseif strcmp(setting.domain,'thirdOctave')
-    
-    [Lp,Leq] = estimationLpEXP(Vtraffic_full,setting);
-else
-    
-    Vtraffic_cut = Wtraffic(1:soundMix.ind,:)*Htraffic;
-    [Lp,Leq] = estimationLpEXP([{Vtraffic_full} {Vtraffic_cut}],setting);
-end
-
-NMF.LeqTrafficEstimate = Leq;
-NMF.LpTrafficEstimate = Lp;

### 2018-01-04 experiment/NonnegMatrixFact/netlab/nmf/updateSmoothEssidEXP.m

@@ -1,80 +1,0 @@
-function [H] = updateSmoothEssidEXP(H,W,V,V_ap,beta,smoothness,sparsity,lambda)
-
-N = size(H,2);
-
-switch beta
-    case 0
-        Ht = H;
-        a = 2*smoothness.*(lambda'.^2);
-        c = 0;
-        d = -W'*(V./(V_ap.^2)).*(Ht.^2);
-        
-        % n = 1
-        b = W'*(V_ap(:,1).^(-1))-smoothness.*(lambda'.^2).*(H(:,2));
-        H(:,1) = resolutionThirdOrder(a,b,c,d(:,1),smoothness);
-        
-        % n = 2:N-1
-        for n = 2:N-1
-            b = W'*(V_ap(:,n).^(-1))-smoothness.*(lambda'.^2).*(H(:,n-1)+H(:,n+1));
-            H(:,n) = resolutionThirdOrder(a,b,c,d(:,n),smoothness);
-        end
-        
-        % n = N
-        b = W'*(V_ap(:,N).^(-1))-smoothness.*(lambda'.^2).*(H(:,N-1));
-        H(:,N) = resolutionThirdOrder(a,b,c,d(:,N),smoothness);
-        
-        
-    case 1
-        if any(smoothness==0)
-            smoothness(smoothness==0)=1e-12;
-        end
-        
-        Ht = H;
-        a_n = 2*smoothness.*(lambda'.^2);
-        a_1 = a_n/2;
-        a_N = a_n/2;
-        bt = sparsity+lambda';
-        c = Ht.*(W'*(V./V_ap));
-        
-        % n = 1
-        b(:,1) = bt-smoothness.*(lambda'.^2).*H(:,2);
-        H(:,1) = (sqrt(b(:,1).^2+4*a_1.*c(:,1))-b(:,1))./a_1;
-        
-        % n = 2:N-1
-        b(:,2:N-1) = repmat(bt,1,size(H,2)-2)-repmat(a_n,1,size(H,2)-2).*(H(:,1:N-2)+H(:,3:N));
-        for n = 2:N-1
-            H(:,n) = (sqrt(b(:,n).^2+4*a_n.*c(:,n))-b(:,n))./(2*a_n);
-        end
-        
-        % n = N
-        b(:,N) = bt-smoothness.*(lambda'.^2).*(H(:,N-1));
-        H(:,N) = (sqrt(b(:,N).^2+4*a_N.*c(:,N))-b(:,N))./a_N;
-        
-        H(H<0)=0;
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/resolutionThirdOrder.m

@@ -1,57 +1,0 @@
-function X = resolutionThirdOrder(a,b,c,d,alphaCoeff)
-
-K = length(a);
-
-if length(alphaCoeff)==1
-    alphaCoeff(1:K,1) = alphaCoeff;
-end
-
-alphaNonNull = find(alphaCoeff~=0);
-alphaNull = find(alphaCoeff==0);
-
-%% alphaNonNull ~= 0
-if ~isempty(alphaNonNull)
-    p = c./a(alphaNonNull)-(b(alphaNonNull).^2)./(3*(a(alphaNonNull).^2));
-    q = (2*(b(alphaNonNull).^3))./(27*(a(alphaNonNull).^3))+...
-        d(alphaNonNull)./a(alphaNonNull)-b(alphaNonNull)*c./(3*a(alphaNonNull).^2);
-
-    delta = q.^2+4*((p./3).^3);
-
-    pos = find(delta > 0);
-    nul = find(delta == 0);
-    neg = find(delta <0);
-
-    X = zeros(K,1);
-    if ~isempty(pos)
-        C1 = nthroot((-q(pos)+sqrt(delta(pos)))/2,3);
-        C2 = nthroot((-q(pos)-sqrt(delta(pos)))/2,3);
-
-        X(pos) = C1+C2-(b(pos)./(3*a(pos)));
-    end
-    if ~isempty(nul)
-
-        Xnul(:,1) = (3*q(nul))./p(nul)-(b(nul)./(3*a(nul)));
-        Xnul(:,2) = -(3*q(nul))./(2*p(nul))-(b(nul)./(3*a(nul)));
-        Xnul(:,3) = Xnul(:,2);
-        X(nul) = min(abs(Xnul),[],2);
-    end
-    if~isempty(neg)
-        deltaLoop = abs(delta(neg));
-        jj = -0.5+1j*(sqrt(3)/2);
-        c1 = (-q(neg)+1j*sqrt(deltaLoop))/2;
-        c2 = (-q(neg)-1j*sqrt(deltaLoop))/2;
-        C1 = nthroot(abs(c1),3).*exp(1j*angle(c1)/3);
-        C2 = nthroot(abs(c2),3).*exp(1j*angle(c2)/3);
-
-        Xneg(:,1) = C1+C2-(b(neg)./(3*a(neg)));
-        Xneg(:,2) = jj*C1+(jj)^2*C2-(b(neg)./(3*a(neg)));
[…]

### 2018-01-04 experiment/NonnegMatrixFact/netlab/rmseEXP.m

@@ -1,24 +1,0 @@
-function errorRMSE = rmseEXP(X,Y,varargin)
-%% rmse(X,Y,DIMENSION)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% CONDITIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-if size(X,2)>size(Y,2)
-    X = X(:,1:size(Y,2));
-elseif size(X,2)<size(Y,2)
-    Y = Y(:,1:size(X,2));
-end
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% FONCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% if ~isfinite(X)
-%     X(ii) = 0;
-% elseif ~isfinite(Y)
-%     Y(ii) = 0;
-% end
-if size(X,1)~=1
-    errorRMSE = sqrt(mean(bsxfun(@minus,Y,X).^2));
-else
-    errorRMSE = sqrt(sum((X-Y).^2));
-end

### 2018-01-04 experiment/NonnegMatrixFact/netlab/spectre2MelEXP.m

@@ -1,13 +1,0 @@
-function [Y,binMel,wts] = spectre2MelEXP(X,numberMel,sr)
-%% function [Y,binMel,wts] = spectre2MelEXP(X,numberMel,sr)
-
-if iscell(X)
-    Y = cell(size(X,1),size(X,2));
-    for ii = 1:size(X,1)
-        for jj = 1:size(X,2)
-            [Y{ii,jj},wts,binMel] = audspec(X{ii,jj}, sr, numberMel,'mel', 0, sr/2, 0, 1);
-        end
-    end
-else 
-    [Y,wts,binMel] = audspec(X, sr, numberMel,'mel', 0, sr/2, 0, 1);
-end

### 2018-01-04 experiment/NonnegMatrixFact/nomafa1dictionary.m

@@ -11,7 +11,7 @@ function [config, store, obs] = nomafa1dictionary(config, setting, data)
 % Date: 13-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, NonnegMatrixFact('do', 1); return; else, store=[]; obs=[]; end
+if nargin==0, NonnegMatrixFact('do', 1, 'mask',{1 1 1 1 1}); return; else, store=[]; obs=[]; end
 
 if strcmp(setting.type,'nmf')
     K = setting.numberElementW;

### 2018-01-04 experiment/NonnegMatrixFact/nomafa2estimation.m

@@ -52,7 +52,7 @@ globalName = cell(1,length(files)/3);
 ind = 1;
 
 for ii = 1:length(files)
-    if isempty(strfind(files(ii).name,'traffic')) && isempty(strfind(files(ii).name,'perturbator'))
+    if isempty(strfind(files(ii).name,'traffic')) && isempty(strfind(files(ii).name,'interfering'))
         globalName{ind} = files(ii).name(1:end-4);
         ind = ind+1;
     end
@@ -70,7 +70,7 @@ end
 parfor ii = 1:numberScene
     
     fileTraffic = audioread([creationSceneDir globalName{ii} '_traffic.wav']);
-    fileRest = audioread([creationSceneDir globalName{ii} '_perturbator.wav']);
+    fileRest = audioread([creationSceneDir globalName{ii} '_interfering.wav']);
     
     if strcmp(dataset,'ambience')
         

### 2018-01-11 paper/gloaguenNmfTraffic.tex
article corrections+experiment programs

@@
 \section{Introduction}
-\subsection{Related work}
+
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
@@
 
-Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels.
+Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels.
 DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations.
@@
 
-If sensors networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}.
-Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap.
-Consequently, the traffic sound level estimation based on measurements is also not a trivial task.
-Many recent works have focused on the detection or recognition tasks of environmental sounds without distinction between them\cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
-A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
-A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+If sensor networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}.
+Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn \dots) that can overlap.
+Consequently, the traffic sound level estimation based on measurements is not a trivial task.
+Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
+A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
+A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
 
-Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
-If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic  level are presented.
-Furthermore, this work and as well as the other works in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context.
-Although near major roads or ring roads traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels.
-In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level.
+The main issue in the detection or recognition tasks is the overlap of environmental sounds.
+Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels.
+Succeed to recognize two simultaneous sound sources can be very complicated.
+Socoro et al.
+propose to bypasses this issue with an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
+If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.
 
-In consequence, to be effective on a wide range of sound environments, we propose in this paper to follow the blind source separation paradigm.
+Instead of using a detection method to estimate the traffic presence, we propose in this paper to follow the blind source separation paradigm.
 That is, separating the contribution of the traffic from the other sources within a polyphonic scene.
+\\
 
-\ml{tu passe beaucoup de temps a taper sur socoro et peu a parler de ton approche.
-Ce n'est pas une bonne pratique.
-Je dirais ils font ca, nous aussi on fait ca c'est super, et nous on le fais comme ca, regardez tout ce que ca permet (et implicitement eux ils peuvent pas).}
-
-\ml{tu plonges tres vite dans la technique}
-
-One of the first and the most widely \ml{faux} used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}.
+One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}.
 The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
 This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.
-However, ICA is limited to only over determined cases ($N > P$).
-Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation \ml{je ne vois pas ou est le probleme, il y a des gens qui le font}.
-A more convenient method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices.
+In an urban environment context, this method is used with acoustic microphone arrays and beamforming \cite{saruwatari2003blind}.
+These approaches consider a set of microphones and allows the localization \cite{mennitt2010multiple} or the detection \cite{leiba2017large} of sound sources from the phase shift and the distance between the microphones.
+However, one need a lot of microphones.
+Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.
+
+A more convenient method for monophonic signal is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices.
 It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
@@
 For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.
-It has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012} for source separation.
-They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them.
-The audio files tested results of a simulation process where a sound background (river or wind) are added to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).
 
- If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
- \ml{encore une fois, utilise la critique implicite.
-On peut tout critiquer, surtout un papier qui n'est pas encore publié, fais attention.
-faire une liste des contributions avec 1) 2) 3) et finir par un paragraphe de structure}
+For the sound source separation, it has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012}.
+After having performed NMF on simulated audio files, they realized a source separation in two steps by separating the sound background from the events first and then by separating the events between them using MFCC and time-variant gain.
+A $k$-mean clustering allows the separation of the $k$ sound sources.
+This approach, compare to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.
+Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.
 
-\subsection{Proposed approach}
-
-\ml{tu melange contribution technique et protocole experimental, ca va pas. pas de soussection a l'intro}
-
-We propose in this paper a method based on the Non-Negative Matrix Factorization (NMF) technique to estimate the global, $\tilde{L}_{p,traffic}$, and the 1-s equivalent, $\tilde{p}_{1s,traffic}$ sound level of the traffic through the supervised and the semi-supervised approaches as well as a method using thresholding.
-To validate these approaches, we consider a corpus of simulated scenes artificially created with the simulator software \textit{simScene}.
-The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component which would hardly be extracted from a recording of an urban scene ($L_{p,traffic}$ and $p_{1s,traffic}$).
-Both the sound scene simulation and NMF require the creation of two sound databases; see Figure \ref{fig:blocDiagramNMF}.
-In parallel, a baseline method built from a frequency low-pass filter is computed.
-This method considers that road traffic is mainly composed of low frequencies and therefore can be filtered by a low-pass filter at the cut-off frequency $f_c$; see Figure \ref{fig:blocDiagramFilter}.
-The performance of the frequency low-pass filter and NMF are then compared with the calculation of two metrics (Mean Absolute Error, normalized Root Mean Square Error).
+Consequently, we propose in this paper to use different frameworks of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event.
+To validate this approach, we also consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
+The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would hardly be extracted from a recording of an urban scene.
 
@@
 The remaining of the paper is organized as follows.
-Section \ref{part:nmf} details the technical aspect of NMF.
+Section \ref{part:nmf} details the technical aspects of NMF.
 Section \ref{part:protocol} described on the design of the environmental sound scene corpus and the experimental protocol setup.
@@
 
-The experience is lead with the Matlab software.
-In order to be a reproductive experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} and the sound database\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} (TEMPORAIRE) used are available online.
-
 \section{Non-negative Matrix Factorization}\label{part:nmf}
 \subsection{Description of NMF}
-Non-negative Matrix Factorization is a matrix approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
+
+Non-negative Matrix Factorization is a linear approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
 
@@
 NMF is then considered as a low rank approximation method.
-However, this constraint is not essential.
+However, this constraint is not mandatory.
 To estimate the quality of the approximation, an objective function is used
@@
 
-Prior knowledge on the content can be adjusted with the addition of constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) in the objective function , see equation (\ref{eq:min-D-WH}), to better take account prior knowledge of the sources.
+To better take account prior knowledge on the sources,  constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) can be added to the objective function.
 
-Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensure non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensures non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
 
 \subsection{Supervised NMF}
-First, supervised NMF is used: the \textit{dictionary} includes audio spectrum of urban sound sources as, in the urban environments, a lot of different sound sources present are known and their spectrum can be obtained.
-The \textit{basis} are then the unknown to estimate.
+First, supervised NMF (SUP NMF) is used: the \textit{dictionary} includes audio spectrum of urban sound sources.
+In the urban environments, a lot of different sound sources present are known whose spectrum can be obtained and be a basis of $\mathbf{W}$.
+The \textit{activator} are then the unknown to estimate.
 In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm
@@
 The product $A.B$ and $A/B$ symbolized the Hadamard product and ratio.
-As in the supervised approach, the position in $\mathbf{W}$ of traffic component is known, the source separation of this sound source is made by extracting the dictionary and basis elements related,
+As in the supervised approach, the position in $\mathbf{W}$ of traffic component is known, the separation of this sound source is made by extracting the related basis and activators,
 
@@
 One of the main issue with the supervised approach is the generalization issue: how to be adapted to different sound mixtures with a fixed dictionary ?
-To better take into account the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility.
-This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
+To better take into account the diverse nature of urban scenes, semi-supervised NMF (S-S NMF)\cite{lee_semi-supervised_2010} can be useful as it has been proposed  to offer more flexibility.
+This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectrum, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
 Here, $J = 2$.
-The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic.
-The problem (\ref{eq:nmf}) become
+The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic.
+The problem (\ref{eq:nmf}) becomes
 
 \begin{equation}
-\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}.
+\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}
 \end{equation}
 
-In a similar way as to solve the equation \ref{eq:min-D-WH}, $\mathbf{W_r}$, $\mathbf{H_r}$ and $\mathbf{H_s}$ are successively updated with the relations (\ref{eq:WH-SSupdate}):
+ with $\mathbf{W} = \left[\mathbf{W_s} \mathbf{W_r} \right]$ and $\mathbf{H} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$.
+In a similar way as to solve the equation (\ref{eq:min-D-WH}), $\mathbf{W_r}$, $\mathbf{H_r}$ and $\mathbf{H_s}$ are successively updated with the relations (\ref{eq:WH-SSupdate}):
 
@@
 
+Applications of S-S NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.
+
 \subsection{Thresholded initialized NMF}
@@
 Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly.
-Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}).
+Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
 Then NMF is performed where $\mathbf{W}$ (eq.
@@
 
-$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ means that the elements are significantly different.
-This measure allows a bound between 1 and 0 and is an invariant scale estimation of the similarity.
-Then, the similarity is sorted in descending order.
+$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are significantly different.
+This measure is bounded between 1 and 0 and is an invariant scale estimation of the similarity.
+Then, $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is sorted in descending order.
 The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
-An illustrative example can be see in Figure \ref{fig:W_ThC_NMF}.
 It is defined as:
@@
 
+An illustrative example can be see in Figure \ref{fig:W_TI_NMF}.\\
+
 \begin{figure}[hbtp]
@@
 \caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
-\label{fig:W_ThC_NMF}
+\label{fig:W_TI_NMF}
 \end{figure}
 
-This approach is named \textit{Thresholded inititialized NMF}, TI-NMF.
-Other thresholding methods as the \textit{soft} \cite{donoho1995noising}and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated.
-A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, like in  Figure \ref{fig:W_ThC_NMF}), was the best way to get better performances.
+This approach is named \textit{Thresholded inititialized NMF} (TI NMF).
+Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the similarity through a sigmoïd or a Radial Basis Function have been investigated.
+A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the best way to get better performances.
 
@@
 This offers a controlled framework to design specific sound environments in which all the traffic component is known.
-Then, the road traffic sound levels estimated with the method can be compared to the real ones, introduced within each simulated sound scene.
+Then, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.
 
@@
 
-A corpus is designed with the \textit{simScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}.
-\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by summing audio samples that come from an isolated sound database.
+A corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}.
+\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by summing audio samples that come from an isolated sound database.
 
@@
 154 sound samples belong to this category divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
-The sound class \textit{car passages} comes from 60 recordings of 2 cars made on the Ifsttar's runway on different speeds with multiple gear ratio.
+The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Senic) made on the Ifsttar's runway on different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound classes is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
+Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
 The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
@@
     \centering
-\includegraphics[width=\linewidth]{figures/exampleSimScene2.pdf}
-\label{fig:exampleSimScene}
-\caption{Example of a sound scene composed of 3 sound classes (car, bird, car horn)}
+    \begin{subfigure}[t]{0.47\linewidth}
+        \centering
+       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-spectrum.png}
+       \caption{}
+       \label{fig:simScene_spec}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.47\linewidth}
+        \centering
+       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-timeDomain.png}
+       \caption{}
+       \label{fig:simScene_time}
+    \end{subfigure}
+    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain}) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue):
+    \label{fig:example_simScene}
 \end{figure}
@@
 Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult.
-The classes are : \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
-In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the sound classes.
+The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
+In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the other sound classes.
 The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class.
-To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$ following the relation (\ref{eq:tir}).
+To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).
 
@@
 with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$.
-When $TIR = -12$, the traffic component is then less present than when $TIR = 12$ where it is predominant on the \textit{interfering} class.
-The 1 second equivalent sound pressure level, $p_{1s,traffic}$, is also calculated ; see Figure \ref{fig:exampleScene}).
-The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 TIR values).
-
-\begin{figure*}
-\centering
-   \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{figures/spectrogramExample.pdf}
-   \end{minipage}
-   \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{figures/animals_10-pianoRoll.png}
-   \end{minipage}
-   \begin{minipage}[c]{.32\linewidth}
-      \includegraphics[width =\linewidth]{figures/evolutionLpExample.pdf}
-
-   \end{minipage}
-\caption{Example of a scene of the \textit{animals} sub corpus.
-Spectrogram (on left), \textit{Piano Roll} of the different sound classes (on the middle) and 1-s equivalent sound level of the traffic, $Lp_{1s,traffic}$ and of the global sound scene, $Lp_{1s,global}$(on right)}
-\label{fig:exampleScene}
-\end{figure*}
+When $TIR = -12$, the traffic component is then less present than the interfering class.
+When $TIR = 12$ it is the opposite: the trafffic class is louder than the interfering class.
+The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values).
 
@@
 The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
-First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$.
-Therefore, the dimensions of $\mathbf{V}$ are $F$ = 2049 and $N$ = 664.
+The spectrogram $\mathbf{V}$ of each sound is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$ ($N$ = 644).
 
-The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
-The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq.
-\ref{eq:v_tr_filtered}).
+The first estimator to determine the traffic sound level is a frequency low-pass filter which depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
+The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
+\ref{eq:v_tr_filtered}),
 
@@
 The second estimator is the proposed scheme, based on several flavors of the NMF framework.
-Multiples experimental factors are involved here between the dictionary learning and NMF (see bloc diagram in Figure \ref{fig:bloc_nmf}).
+Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}).
 
@@
 
+\begin{table*}[t]
+\centering
+\begin{tabular}{lcccccc}
+\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \toprule
+$\mathbf{K}$ & 25 & 50 & 100 & 200 & & \\ 
+\rowcolor[HTML]{EFEFEF}
+$\mathbf{w_t}$ (s)& \multicolumn{2}{c}{0.5} & \multicolumn{2}{c}{1}  & \multicolumn{2}{c}{\textit{all}} \\
+$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ 
+\rowcolor[HTML]{EFEFEF} 
+\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ 
+$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ 
+\rowcolor[HTML]{EFEFEF} 
+$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ 
+\textbf{method} & \multicolumn{2}{c}{filter} & SUP NMF & S-S NMF & TI NMF \\
+\rowcolor[HTML]{EFEFEF} 
+$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \bottomrule
+\end{tabular}
+\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
+\label{tab:experimental_factors}
+\end{table*}
+
 \subsubsection{Dictionary building}\label{part:dictionary_learning}
@@
 It is composed of 53 audio files of passing cars.
-These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars.
+These records have been made on the Ifsttar's runway too with the same experimental conditions that the records made for the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
-Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
-$w_t$ is fixed at $wt = [0.5 1]$ second.
-In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra.
-Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$.
-A special case is added where the root mean square of \textit{all} the spectrogram is applied.
-Each audio file generates one element $k$ of $\mathbf{W}$.
-An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage ; see Figure \ref{fig:specW}.
-In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element ; see Figure \ref{fig:ElementW}.
+Then time/frequency windows of $F \times w_t $ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
+$w_t$ is fixed at $wt = [0.5~1]$ second.
+In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectra of $F \times 1$ dimension.
+A special case is added where the root mean square of \textit{all} the spectrogram is applied (each audio file generates one element $k$ of $\mathbf{W}$).
+An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage ; see Figure \ref{fig:specW}.
+In the case where $w_t$ = 1 second , 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element ; see Figure \ref{fig:ElementW}.
 
@@
        \includegraphics[width=\linewidth]{figures/dictionary1.pdf}
-        \caption{Sample of 3 seconds of a car passage}
+       \caption{}
         \label{fig:specW}
@@
        \includegraphics[width=\linewidth]{figures/dictionary2.pdf}
-        \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
+        \caption{}
         \label{fig:ElementW}
     \end{subfigure}
+    \caption{Example of the dictionary building with a 3 second extract of a car passage (\ref{fig:specW}).
+With $w_t$ = 1 second, 3 spectrum are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is resume to 1 spectra (\ref{fig:ElementW}).}
+    \label{fig:spec_elementW}
 \end{figure}
 
-Each $k$ element of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.\\
+Since the number of elements given by this processing can be high, in order to reduce the computational time and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100, 200\right]$.
 
-Table \ref{tab:dictionary_factors} summarizes the experimental factors and their related values.
+The obtained dictionary is expressed with third octave bands and each basis vector of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.
+Table \ref{tab:experimental_factors} summarizes the experimental factors ($K$ and $w_t$) for the dictionary building and their related values.
 
-\begin{table}[h]
+\subsubsection{Experimental factors of NMF}
+
+SUP and S-S NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+TI NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+This coarser method allows the matrix reduction and decrease the computation time.
+But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+The case of a linear frequency axis has been investigated and offer similar results which were not better.
+
+For TI NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:experimental_factors} summarizes the experimental factors and their related values.
+
+\begin{table*}[t]
 \centering
-\begin{tabular}{cccc}
-Parameter &  \multicolumn{3}{c}{value}\\ \hline
-$K$ & 25  & 50 & 100 \\ \hline
-$\mathbf{w_t}$ (s)& 0.5 & 1  & \textit{all}
+\begin{tabular}{@{}cccccc@{}}
+\toprule
+\textbf{method} & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
+filter 20 kHz &  &  &  &  & 4.69 ($\pm$ 4.52) \\
+filter 0.5 kHz&  &   &  &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
+SUP NMF &  1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
+SUP NMF &  2 & 50 & 0.5  &   & 3.02 ($\pm$ 3.33) \\ \hline \hline
+S-S NMF &  1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
+S-S NMF &  2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
+TI NMF &  1 & 200 & 0.5 &  0.42 &\textbf{2.19 ($\pm$ 2.18)} \\
+TI NMF &  2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
 \end{tabular}
-\caption{Summary of experimental factors of the dictionary}
-\label{tab:dictionary_factors}
-\end{table}
+\caption{Best results according to $\beta$ and \textit{method} experimental factors, in bold letter, the lowest error.}
+\label{tab:results}
+\end{table*}
 
-\subsubsection{Experimental factors of the NMF}
-
-Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
-ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different spectral representations: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands).
-These two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins in the high frequencies where the traffic component is less present and generates less calculations.
-Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins.
-Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as
-
-\begin{equation}
-\mathbf{\tilde{V}}_{traffic} = \left[\mathbf{WH}_{f_c}\right]_{traffic}.
-\end{equation}
-
-With the third octave spectral representation, only the case $f_c = 20$ kHz is applied.
-For the TI-NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step.
-Table \ref{tab:estimation_factors} summarizes the experimental factors and their related values.
-
-\begin{table*}[]
+\begin{table*}[t]
 \centering
-\begin{tabular}{lcccccc}
-\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \hline
-$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
-\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
-$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
-$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
-\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & TI-NMF \\ \hline
-$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
+\begin{tabular}{@{}cccccc@{}}
+\toprule
+\textbf{method} & filter & filter & SUP NMF & S-S NMF & TI NMF \\ \midrule
+$f_c$ (kHz) & 20 & 0.5 & 20 & 20 & 20 \\
+$\mathbf{\beta}$ &  &  & 2 & 2 & 1 \\ \hline
+\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & 8.65 ($\pm$ 1.88) & 3.88 ($\pm$ 1.52) & 5.35 ($\pm$ 2.71) \\
+\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 4.22 ($\pm$ 1.27) & 1.37 ($\pm$ 0.71)  & 2.82 ($\pm$ 1.30) \\
+\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.34 ($\pm$ 0.56) & 1.11 ($\pm$ 0.25) & 1.26 ($\pm$ 0.35) \\
+\textbf{6} & 0.25 ($\pm$ 0.06) & 1.03 ($\pm$ 0.26) & 0.26 ($\pm$ 0.10) & 2.25 ($\pm$ 0.19) & 0.70 ($\pm$ 0.32) \\
+\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.64 ($\pm$ 0.06) & 2.96 ($\pm$ 0.21)  & 0.84 ($\pm$ 0.24) \\ \bottomrule
 \end{tabular}
-\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
-\label{tab:estimation_factors}
+\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
+\label{tab:results_TIR}
 \end{table*}
@@
 \subsubsection{Metrics}
-The performances of the two estimators of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
-It expresses the quality of the long-term reconstruction of the signal.
-It consists in the average over the $N$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
+The performances of the two estimators (filter and NMF) of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
+It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
 
 \begin{equation}
-MAE = \frac{\sum_{n = 1}^N\vert L^n_{p,traffic}-\tilde{L}^n_{p,traffic} \vert}{N}.
+MAE = \frac{\sum_{m = 1}^M\vert L^m_{p,traffic}-\tilde{L}^m_{p,traffic} \vert}{M}.
 \end{equation}
 
-In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple experimental factors taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of factors are made.
-For TI-NMF, 16800 combinations can be calculated.
+In all, according Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
+For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for SUP and S-S NMF, 3780 associations of factors are made.
+For TI NMF, 8400 combinations can be calculated.
 
@@
 
-Table \ref{tab:results} summarized, according to the 3 main factors (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\
+Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
+For the low-passa frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expand to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
+\\
 
-\begin{table*}[]
-\centering
-\begin{tabular}{@{}cccccc@{}}
-\toprule
-$\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
- &  &  & filter &  & 4.69 ($\pm$ 4.52) \\
- &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-50 & 0.5  & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\
-50 & 0.5  & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
-100 & 0.5 & 1 & semi-supervised NMF &  & 2.38 ($\pm$ 1.26) \\
-100 & 0.5 & 2 & semi-supervised NMF &  & 2.43 ($\pm$ 1.43) \\ \hline \hline
-100 & all & 1 & TI-NMF & 0.57 & 2.19 ($\pm$ 2.01) \\
-100 & all & 2 & TI-NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
-\end{tabular}
-\caption{Best results according to $\beta$, method and  spectral representation}
-\label{tab:results}
-\end{table*}
+First, the errors produced by the filter are detailed.
+$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources.
+Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
+$f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
+It is then the first baseline to use to compare the performances of NMF.
 
-The two first lines show the error produced by the baseline filter.
-$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguishing the traffic from the others sound sources.
-Consequently, in low $TIR$ (-12 and -6), where traffic component is scarce, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant.
-$f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
-It is then the baseline that will be used to compare the performances of NMF.
-In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the filter is efficient as it suppress these frequency components whereas for the other sub-classes where low frequencies are present, the error is higher as the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
-The sounds levels are then overestimate ; see Figure \ref{fig:alert_-12}.
-In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels ; see Figure \ref{fig:alert_12}.\\
+In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, this filter is efficient as it removes these frequency components.
+For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
+The errors are then nearly all superior to 4 dB and are overestimate ; see Figure \ref{fig:alert_-12}.
+In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels ; see Figure \ref{fig:alert_12}.
+The 500 Hz filter finds a balance between what it is put aside in low $TIR$ and what it is remained in high $TIR$.
+\\
 
@@
 
-The results of different versions of NMF are summarized next to the filter results in Table \ref{tab:results_TIR}.
-NMF errors lower than the baseline are put in bold letter.
-With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline.
-The best combination is got with a TI-NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.
+Compare to the filter errors, the choice of some NMF approaches make it possible to decrease the road traffic sound level estimation.
+The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline.
+Finally, it is the most sophisticated methods (S-S and TI NMF) that have better results.
+The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
+On the other hand, the semi-supervised approach has a higher error but yet with a lower standard deviation.
 
-Between the dictionary factors ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings.
-If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this factors is variable depending on chosen method.
-
-For supervised NMF, the spectrogram $\mathbf{V}$ has to be filtered at 500 Hz to focused the reconstruction of the signals on the part where most of the traffic energy is.
-According to the choice of $\beta$, the dictionary factors are different.
-It is the only case where it is.
-In the opposite, in semi-supervised NMF, an unique set of factors is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz.
-It is with $f_c = 20$ kHz that the approximation is better as it includes easily other sound sources.
-With a lower cut-off frequency ($ f_c <$ 5 kHz), focusing on traffic energy, the risk is then to include \textit{traffic} component in $\mathbf{W_r}$ and to deteriorate the traffic signal reconstruction.
-The influence of the spectral representation and $\beta$ is here weak as the errors are similar with a low standard deviation.
-The third approaches, TI-NMF, proposed, in the four cases, the lowest errors.
-Like supervised NMF, in linear spectral representation, it is necessary to filter $\mathbf{V}$ to focus on the traffic energy.
-Here $f_c$ is higher (5000 Hz and 2000 Hz respectively for $\beta = 1$ and $\beta$ = 2).
-The choice of the spectral representation influence also the order of magnitude of the threshold $t$: in the interval [0.37 0.39] and [0.54 0.57] respectively for a linear and third octave spectral representation.\\
-
-For each method (supervised, semi-supervised and TI-NMF), the best parameter combinations are detailed according to the sub-classes and the $TIR$ (Table \ref{tab:results_TIR},  Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_thc}).
-\\
-
-\begin{table*}[t]
-\centering
-\begin{tabular}{@{}cccccc@{}}
-\toprule
-\textbf{method} & filter & filter & supervised NMF & semi-supervised NMF & TI-NMF \\ \midrule
-$f_c$ (kHz) & 20 & 0.5 & 20 & 20 & 20 \\
-$\mathbf{\beta}$ &  &  & 2 & 2 & 2 \\ \hline
-\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & 6.23 ($\pm$ 3.19) &  & \textbf{5.11 ($\pm$ 3.10)} \\
-\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 3.00 ($\pm$ 1.39) &  & \textbf{2.87 ($\pm$ 1.55)} \\
-\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
-\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 1.01 ($\pm$ 0.28) &  & \textbf{0.70 ($\pm$ 0.32)} \\
-\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 1.41 ($\pm$ 0.10) &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
-\end{tabular}
-\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
-\label{tab:results_TIR}
-\end{table*}
-
-\begin{figure*}
+\begin{figure*}[t]
     \centering
@@
         \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
+        \caption{$MAE$ error for each $TIR$ and sub-class with SUP NMF and $\beta$ = 2}
                 \label{fig:TIR_class_sup}
@@
       	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
+        \caption{$MAE$ error for each $TIR$ and sub-class with S-S NMF and $\beta$ = 2}
                 \label{fig:TIR_class_semi}
@@
         \centering
-        \includegraphics[width=\linewidth]{figures/THC_bar}
-        \caption{TI-NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
-        \label{fig:TIR_class_thc}
+        \includegraphics[width=\linewidth]{figures/TI_bar}
+        \caption{$MAE$ error for each $TIR$ and sub-class with TI NMF, $\beta$ = 1 and $t$ = 0.42}
+        \label{fig:TIR_class_TI}
     \end{subfigure}
-    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter and each method (supervised, semi-supervised and ThC}
+    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (SUP (\ref{fig:TIR_class_sup}), S-S (\ref{fig:TIR_class_semi}) and TI (\ref{fig:TIR_class_TI} NMF))}
+    \label{fig:TIR_bar}
 \end{figure*}
 
-According to Figure \ref{fig:TIR_class_filter}, \textit{climate}, \textit{human}, \textit{transport} and \textit{mechanics} are the four sub-classes that generates the highest error levels particularly in low $TIR$ as they are sound classes composed of low frequency sound (storm for \textit{climate}, voices in \textit{humans}, plane in \textit{transport} and ventilation noise in \textit{mechanics}).
-All these kind of sounds are then considered as traffic component without distinction.
-In the case of supervised NMF, on all the $TIR$, if the mean score on each $TIR$ is lower than the baseline method, it offers similar performances as for $TIR = [0, 6, 12]$, the mean scores are nearly close.
-The performance of supervised NMF is mainly visible for $TIR = -12$ where the error with the baseline decrease by more than 1 dB.
-This improvement is significant with the \textit{alert} sub-class as it gathers sounds that are mostly in higher frequencies ([2500 5000] Hz).
-From a dictionary $\mathbf{W}$ composed of traffic element, the traffic signal reconstruction is then easier.
-However, supervised NMF fails to improve significantly the baseline performances.
+According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
+In the case of SUP NMF, it fails to improve the filtering performances despite good results in high $TIR$.
+the error are too important for low $TIR$ and this for all the sub-classes.
+This method reveals to be to rigid as $\mathbf{W}$ is composed of fixed traffic spectrum.
+In the case of low $TIR$, in the aim to reduce the objective function, see eq.
+\ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the  sound scene.
+The case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}.
+Here when the car horn sounds, some elements of $\mathbf{W}$ are activated and are then considered as traffic component generating a wrong estimation of the sound level.
+This behavior disappear when the traffic component become predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}.
+Thus composed the dictionary of only fixed traffic spectrum is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
 
-Meanwhile, semi-supervised approach with the add of a mobile part in the dictionary, $W_r$, brings a major improvement especially for the low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}).
-By the add of $\mathbf{W_r}$, it allows to take into account the other predominant sound sources.
-An example can be found in Figure \ref{fig:Y_alert-12} which summarizes the obtained $\mathbf{W_r}$ of a scene belonging to \textit{alert} sub-class.
-The first element is mainly composed of harmonics signal which correspond to the car horn of the scene.
-The decreasing of errors can be seen on each sub class for $TIR = -12$ and $TIR$ = -6 in  where the error are much lower than supervised NMF.
-As the $\mathbf{W}$ is only composed of \textit{traffic} element, this approach is likely to use traffic components to reduce it even if the sound source not traffic to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$ deteriorating the quality of traffic signal reconstruction.
+Consequently, with the add of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$.
+It brings a significant decrease of the errors for low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}.
+The $\mathbf{W_r}$ composition, in the case of an \textit{alert} sub-class, is displayed in Figure \ref{fig:Y_alert-12} and this shows that the interfering class is easily integrated.
+The first element is mainly composed of high frequency bands which correspond to the car horns of the scene.
+This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the traffic basics are no longer activated when the car horn sounds.
 
-The comparison can be made on a scene of the \textit{alert} sub-class:
-FIGURE Lp 1 seconde avec supervisé et semi-supervisé pour voir l'utilisation des éléments trafic dans la supervisée et non dans al semi-supervisée. \\
+\begin{figure*}[t]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/NMF_Lp_alert-12.pdf}
+        \caption{}
+        \label{fig:lp_alert_-12}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/NMF_Lp_alert12.pdf}
+        \caption{}
+        \label{fig:lp_alert_12}
+    \end{subfigure}
 
-However, the degrees of freedom of semi-supervised NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes.
-Indeed, in order to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$ without constraint semi-supervised NMF is free to include traffic components in $\mathbf{W_r}$.
-Consequently, this behaviors decreases the quality of the reconstruction of the traffic component.
-The same scene used in Figure \ref{fig:Y_alert-12}) but with $TIR$ can be seen in Figure \ref{fig:Y_alert_12} where the harmonic components of traffic have disappears for low frequency components.
-The error are then increasing for these $TIR$.
-Constrain $\mathbf{W_r}$ to avoid these behaviors and control what can be put into could be a way to improve this method here \cite{kitamura_music_2014}.
-In opposite, supervised NMF (that could be seen as semi-supervised NMF which is infinitely constrain on $\mathbf{W_r}$) offer a good reconstruction of the traffic signal from $\mathbf{W}$.\\
+    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for : at $TIR = -12$ (\ref{fig:lp_alert_-12}) and at $TIR$ (\ref{fig:lp_alert_12}).}
+    \label{fig:lp_alert}
+\end{figure*}
+
+However, the degrees of freedom of S-S NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+Indeed, in order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, S-S NMF is free to include traffic components in $\mathbf{W_r}$.
+Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
+In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horn have disappeared for the benefit of low frequency content.
+Consequently, the traffic sound level estimation is then underestimate as in Figure \ref{fig:lp_alert_12}.
+\\
 
@@
         \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_-12_01.pdf}
-        \caption{$TIR$ = -12}
+        \includegraphics[width=\linewidth]{figures/Y_alert_-12.pdf}
+        \caption{}
         \label{fig:Y_alert-12}
@@
         \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_12_01.pdf}
-        \caption{$TIR$ = 12}
+        \includegraphics[width=\linewidth]{figures/Y_alert_12.pdf}
+        \caption{}
 		\label{fig:Y_alert_12}
     \end{subfigure}
-    \caption{$\mathbf{W_r}$ for an \textit{alert} scene}
+    \caption{2 elements of $\mathbf{W_r}$ for an \textit{alert} scene for $TIR$ = -12 (\ref{fig:Y_alert-12}) and $TIR$ = 12 (\ref{fig:Y_alert_12})}
 \end{figure*}
 
-The TI-NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}).
-According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors.
-On the other hand, on low $TIR$, it does not succeed to be better than semi-supervised NMF.
-
-Unlike supervised NMF, where $\mathbf{W}$ is fixed, and semi-supervised NMF, which combine a fixed dictionary with a mobile dictionary, TI-NMF's $\mathbf{W}$ is here update entirely to be adjusted to the scene and to adapt to the different sound environment.
-The closest elements of the \textit{traffic} component defined un $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes for $TIR$ = [-12, 12].
+Finally, TI NMF with a threshold fixed at $t = 0.54$ and $\beta$ = 2, offers the lowest average results (Table \ref{tab:results}).
+If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.
+Unlike SUP NMF, where $\mathbf{W}$ is fixed, and S-S NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to be adjusted to the scene and to adapt to the different sound environments.
+The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
+In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes for $TIR$ = [-12, 12] dB.
 
@@
     \includegraphics[width=\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
+    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.54}
     \label{fig:dist_-12_12}
@@
 
-For low $TIR$, with the thresholding, only a reduce number of $\mathbf{W}$ are considered as traffic components.
+For low $TIR$, as the traffic sound class is not predominant, the final dictionary differs a lot from the $\mathbf{W_0}$.
+With the thresholding, only a reduce number of basis vectors are considered as traffic components.
 In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.
-However, for \textit{climate} and \textit{mechanics}, the error stay important (approfondir avec comparaison des niveau sonores trafic).
-On high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher and allows to keep more elements.
+However, for \textit{climate} and \textit{mechanics}, the error stay important as these interfering classes have similar spectrum.
+On high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB).
 The kept elements are then more suited to the scenes than a fixed dictionary.
-So, the error, compare to supervised NMF, are better.
-In these cases, TI-NMF is the most powerful of the proposed methods.
+The error for these $TIR$ is then due to the thresholding.
+With a low threshold, it is possible to decrease the error (for $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30).
 
 \section{Conclusion}
-ouvrir sur l'optimisation du seuil par plusieurs indicateurs (niveau sonore global ou en fréquence ? )\\
-TI-NMF adapté pour d'autres sources sonores \\
-Scènes grafic ensuite
 
-\ml{non, une conclusion conclue, c'est tout.
-rappel des resultats experimentaux et conclusion sur l'approche. Rappel des contributions et impact sur la communaute.
-Petite ouverture sur le potentiel de l'approche: agnostique en terme de source, peut etre contraint temporellement citation, reste raisonable en terme de cout de calcul. }
+In this work, NMF, a source separation method, was used to estimate the road traffic sound level in urban sound mixtures.
+It is an adapted approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
+Different versions of NMF have been studied as a supervised and semi-supervised approach.
+On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
+The proposed new approach, named Threshold Initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the traffic elements are extracted by hard thresholding, proposes the lowest average error.
+It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the SUP NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
+
+This method has the advantage to keep a low calculation cost and then can be implemented in embedded sensors.
+Furthermore, this approach stay adapted for other sound sources, as whistling bird, depending on what spectrum are put in $\mathbf{W}$ and can be adapted for other application than the road traffic sound level estimation.
+Finally, the add of constraints as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} can easily be implemented and taken into account.
+
+In parallel, this experience allowed the design of a large sound database composed of multiples samples of urban sounds and sound mixtures that are made available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for research communities dedicated to the detection, separation and recognition tasks of urban sound sources.
+The experience has been lead with the Matlab software.
+In order to be a reproducible experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} used for this experience are available online.
 

### 2018-01-11 paper/gloaguenNmfTraffic.tex
abstract is here

@@
 
-\title{Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization}
+\title{Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques}
 
@@
 \section*{Abstract}
+
+The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring the acoustic quality of urban areas.
+Assuming that transportation is the major source of annoyance, state of the art approaches consider as input an estimate of the number and the speed of vehicules in major traffic lanes.
+Follows a prediction procedure that outputs an acoustic pressure level at any location in the modeled area.
+
+Considering as input the acoustic pressure measured in many locations using a sensor grid approach would greatly complement and improve the quality of the predicted pressure values.
+Among the technical issues that raises this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.
+
+In this paper, several techniques based on the non-negative factorization framework are studied in this application scenario.
+The task being to the best of our knowledge never been considered in the literature, we propose an exprimental protocol to validate the studied approaches that complies with standard reproducible research recommandations.
 

### 2018-01-11 paper/gloaguenNmfTraffic.tex
done with intro

@@
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-These maps depict an estimation of the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
+These maps depict an estimation of the number of city dwellers exposed to high noise levels.
+They are used to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
 These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
 They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
+However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} \ml{un peu hors propos on améliore rien la dessus : or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
+plutot parler du cote sparse des données d'entrées}.
 In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day.
@@
 It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
-In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
+In a similar way, but reduced to few neighborhoods \ml{with a more dense topology ?
+parler aussi de sonyc}, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
 
-If sensor networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}.
-Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn \dots) that can overlap.
+If sensor networks could improve road traffic noise estimates compared with simulated maps, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
+Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap.
 Consequently, the traffic sound level estimation based on measurements is not a trivial task.
@@
 Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels.
-Succeed to recognize two simultaneous sound sources can be very complicated.
-Socoro et al.
-propose to bypasses this issue with an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level.
-If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.
+To circumvent this issue, Socoro et al.
+propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component.
+Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.
+\ml{arrete :) If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.}
 
-Instead of using a detection method to estimate the traffic presence, we propose in this paper to follow the blind source separation paradigm.
+An alternative approach that we will foolow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level.
 That is, separating the contribution of the traffic from the other sources within a polyphonic scene.
-\\
+One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\
 
-One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}.
+ml{Supprimer : One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}.
 The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
@@
 However, one need a lot of microphones.
-Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.
+Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.}
 
-A more convenient method for monophonic signal is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices.
-It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
+A convenient method for monophonic signal decomposition is the Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
+When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}.
+In the audio processing domain, the NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
 By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane.
-For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.
+\ml{enlever : For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.}
 
-For the sound source separation, it has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012}.
-After having performed NMF on simulated audio files, they realized a source separation in two steps by separating the sound background from the events first and then by separating the events between them using MFCC and time-variant gain.
-A $k$-mean clustering allows the separation of the $k$ sound sources.
-This approach, compare to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.
-Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.
+Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}.
+After having performed NMF on simulated audio files, they perform a source separation based on NMF in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
 
-Consequently, we propose in this paper to use different frameworks of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event.
-To validate this approach, we also consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
-The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would hardly be extracted from a recording of an urban scene.
+\ml{ enlever : A $k$-mean clustering allows the separation of the $k$ sound sources.
+This approach, compared to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.
+Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.}
+
+We study in this paper different flavors of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event.
+We demonstrate the supervised NMF and semi supervised NMF approaches have some interests but fail to give satifactory results for the application at hand.
+We thus introduce another NMF scheme called thresholded initialized NMF that make good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).
+
+To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
+The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would be difficult to extract from a recording of an urban scene.
 
@@
 \includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
-\caption{Bloc diagram of the source separation method}
+\caption{Block diagram of the source separation method. \ml{toujours citer une figure}}
+\label{fig:diagram}
 \end{figure*}
@@
 The remaining of the paper is organized as follows.
-Section \ref{part:nmf} details the technical aspects of NMF.
-Section \ref{part:protocol} described on the design of the environmental sound scene corpus and the experimental protocol setup.
-Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.
+Section \ref{part:nmf} details the technical aspects of NMF and describes the 3 approaches considered in this paper to achieve the task at hand.
+Section \ref{part:protocol} describes the environmental sound scene corpus and the experimental protocol setup.
+Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.
 

### 2018-01-11 paper/gloaguenNmfTraffic.tex
done up to protocol

@@
 
-The choice of the dimensions is often made as that $F\times K + K \times N < F \times N$.
+The choice of the dimensions is often made so that $F\times K + K \times N < F \times N$.
 NMF is then considered as a low rank approximation method.
@@
 and usually belongs to the $\beta-$divergence class \cite{fevotte_nonnegative_2009} in which the well known Euclidean distance (eq.
-\ref{eq:def_distEUC}) and the Kullback-Leibler divergence (eq.
-\ref{eq:def_divKL}) belong
+\ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq.
+\ref{eq:def_divKL}) and the Itakura-Saito divergence belong
 
@@
 
-To better take account prior knowledge on the sources,  constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) can be added to the objective function.
+To better take into account prior knowledge about the sources of interest, constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) can be added to the objective function.
 
-Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensures non-negative results of which convergence has been proved \cite{fevotte_algorithms_2011}.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.
 
 \subsection{Supervised NMF}
-First, supervised NMF (SUP NMF) is used: the \textit{dictionary} includes audio spectrum of urban sound sources.
+First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectrum of urban sound sources.
 In the urban environments, a lot of different sound sources present are known whose spectrum can be obtained and be a basis of $\mathbf{W}$.
@@
 with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$.
-The product $A.B$ and $A/B$ symbolized the Hadamard product and ratio.
-As in the supervised approach, the position in $\mathbf{W}$ of traffic component is known, the separation of this sound source is made by extracting the related basis and activators,
+The product $A.B$ and $A/B$ are respectively the Hadamard product and ratio.
+As in the supervised approach the indexes of traffic components in $\mathbf{W}$  are known, the separation of the corresponding sound source is made by extracting the related basis and activators,
 
@@
 
-One of the main issue with the supervised approach is the generalization issue: how to be adapted to different sound mixtures with a fixed dictionary ?
-To better take into account the diverse nature of urban scenes, semi-supervised NMF (S-S NMF)\cite{lee_semi-supervised_2010} can be useful as it has been proposed  to offer more flexibility.
-This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectrum, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated.
-Here, $J = 2$.
+The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasnable assumption in our application scenario.
+To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of intereferences that can occur in a realistic scenario.
+To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offer more flexibility.
+This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectrum, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization.
+In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionnary \cite{fevotte ?} \ml{mettre les parametres dnas la partie experimentale Here, $J = 2$}.
 The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic.
@@
 
-Applications of S-S NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.
+Applications of Sem NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.
 
@@
 
-A last approach is tested based on unsupervised NMF.
-Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly.
+As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fails to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.
+
+We therefore propose an alternative scheme based on the unsupervised NMF.
+Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly.
 Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
@@
 \ref{eq:updateH_Sup}) are updated alternatively.
-$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge.
+$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study.
 
@@
 
-After $N$ iterations, a measure of similarity $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ between $\mathbf{W_0}$ and the get dictionary $\mathbf{W}$ for each element $k$ is computed through a cosine similarity,
+After $N$ iterations, a measure of similarity $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ between $\mathbf{W_0}$ and the obtained dictionary $\mathbf{W}$ for each element $k$ is computed using a cosine similarity metric,
 
@@
 $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are significantly different.
-This measure is bounded between 1 and 0 and is an invariant scale estimation of the similarity.
-Then, $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is sorted in descending order.
-The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
+This measure is bounded between 1 and 0 and is an invariant with respect to scale.
+\ml{ $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order.
+} The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
 It is defined as:
@@
 
-An illustrative example can be see in Figure \ref{fig:W_TI_NMF}.\\
+An illustrative example is shown on Figure \ref{fig:W_TI_NMF}.\\
 
@@
 This approach is named \textit{Thresholded inititialized NMF} (TI NMF).
-Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the similarity through a sigmoïd or a Radial Basis Function have been investigated.
-A fast parametric study has revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the best way to get better performances.
+Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to process the similarity using a sigmoïd or a radial basis functions have been investigated.
+A fast parametric study revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
 
@@
 $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\
-\textbf{method} & \multicolumn{2}{c}{filter} & SUP NMF & S-S NMF & TI NMF \\
+\textbf{method} & \multicolumn{2}{c}{filter} & Sup NMF & Sem NMF & TI NMF \\
 \rowcolor[HTML]{EFEFEF}
@@
 
-SUP and S-S NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+SUP and Sem NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
 TI NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
@@
 filter 0.5 kHz&  &   &  &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-SUP NMF &  1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
-SUP NMF &  2 & 50 & 0.5  &   & 3.02 ($\pm$ 3.33) \\ \hline \hline
-S-S NMF &  1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
-S-S NMF &  2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
+Sup NMF &  1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
+Sup NMF &  2 & 50 & 0.5  &   & 3.02 ($\pm$ 3.33) \\ \hline \hline
+Sem NMF &  1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
+Sem NMF &  2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
 TI NMF &  1 & 200 & 0.5 &  0.42 &\textbf{2.19 ($\pm$ 2.18)} \\
@@
 \toprule
-\textbf{method} & filter & filter & SUP NMF & S-S NMF & TI NMF \\ \midrule
+\textbf{method} & filter & filter & Sup NMF & Sem NMF & TI NMF \\ \midrule
 $f_c$ (kHz) & 20 & 0.5 & 20 & 20 & 20 \\
@@
 In all, according Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
-For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for SUP and S-S NMF, 3780 associations of factors are made.
+For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for SUP and Sem NMF, 3780 associations of factors are made.
 For TI NMF, 8400 combinations can be calculated.
@@
         \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with SUP NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sup NMF and $\beta$ = 2}
                 \label{fig:TIR_class_sup}
@@
       	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with S-S NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sem NMF and $\beta$ = 2}
                 \label{fig:TIR_class_semi}
@@
 According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
-In the case of SUP NMF, it fails to improve the filtering performances despite good results in high $TIR$.
+In the case of Sup NMF, it fails to improve the filtering performances despite good results in high $TIR$.
 the error are too important for low $TIR$ and this for all the sub-classes.
@@
 
-However, the degrees of freedom of S-S NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
-Indeed, in order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, S-S NMF is free to include traffic components in $\mathbf{W_r}$.
+However, the degrees of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+Indeed, in order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$.
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
@@
 If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.
-Unlike SUP NMF, where $\mathbf{W}$ is fixed, and S-S NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to be adjusted to the scene and to adapt to the different sound environments.
+Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to be adjusted to the scene and to adapt to the different sound environments.
 The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
@@
 The proposed new approach, named Threshold Initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the traffic elements are extracted by hard thresholding, proposes the lowest average error.
-It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the SUP NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
+It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
 

### 2018-01-11 paper/gloaguenNmfTraffic.tex
done up to experiment

@@
 
-In order to validate the usefulness of considering NMF framework to estimate the road traffic noise level, one need to have a reference level.
-It can hardly be measured or even annotated from real life recordings.
-Thus,  simulated sound scenes are used to assess the performance of the proposed NMF.
-This offers a controlled framework to design specific sound environments in which all the traffic component is known.
-Then, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.
+In order to validate the usefulness of the proposed NMF scheme to estimate the road traffic noise level, one need to compare the traffic noise level predicted by the algorithm to a reference level.
+The latter can hardly be measured or even annotated from real life recordings.
+Thus,  we propose to consider simulated sound scenes to assess the performance of the proposed NMF.
+This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known allowing the computation of the reference level.
+%Thus, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.
 
@@
 
-A corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}.
-\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by summing audio samples that come from an isolated sound database.
+The corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}.
+\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database.
+The simulator have been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \ml{voir ma page de publi, taslp avec gregoire, taslp avec emmanouil, taslp avec mesaros}.
 
-This database is divided in two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
-154 sound samples belong to this category divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
-The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Senic) made on the Ifsttar's runway on different speeds with multiple gear ratio.
+This database is divided into two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} or \textit{texture} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
+154 sound samples that belong to this category are divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
+The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway on different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots).
-The software allows the user to control some parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
+Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots to allow some diversity in the resulting mixture).
+The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
 Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
@@
 
-This database enables creating realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
+This database allows us to create a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
 A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds.
@@
 The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values).
+
+\ml{The corpora are available for download at: \url{}}
 

### 2018-01-12 paper/gloaguenNmfTraffic.tex
starting experiment

@@
 The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
-The spectrogram $\mathbf{V}$ of each sound is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$ ($N$ = 644).
+The spectrogram $\mathbf{V}$ of each sound is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap \ml{enlever and a number of point $nfft = 2^{12}$ ($N$ = 644)}.
 
-The first estimator to determine the traffic sound level is a frequency low-pass filter which depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
+Assuming that the traffic spectral profile is largely concentrated in the low regions, a first estimator to determine the traffic sound level is a frequency low-pass filter.
+It depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
 The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-The second estimator is the proposed scheme, based on several flavors of the NMF framework.
+The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:NMF}.
 Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}).
@@
 
-    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter (a) and for NMF (b)}
+    \caption{Block diagrams for the frequency low-pass filter (a) and for NMF (b) \ml{je ferais deux figures séparées, une avec le protocole avec une boite estimateur en disant dans la caption que l'estimateur est soit le filtre soit la nmf.
+deuxieme figure avec la nmf.
+Tu utilises des boites et des couleurs differentes, si cela a un sens explicite le sinon enlève, met les couleurs plus claires.}}
 \end{figure}

### 2018-01-12 paper/gloaguenNmfTraffic.tex
up to metrics

@@
 \end{tabular}
-\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
+\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their evaluated values}
 \label{tab:experimental_factors}
@@
 
-\subsubsection{Dictionary building}\label{part:dictionary_learning}
+\subsubsection{NMF Dictionary}\label{part:dictionary_learning}
 
-The dictionary is built from a second sound database dedicated specifically to this task.
+In order to prevent potential overfitting issues, the dictionary is built from a separate sound database dedicated specifically to this task.
 It is composed of 53 audio files of passing cars.
-These records have been made on the Ifsttar's runway too with the same experimental conditions that the records made for the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
+These recordings have been made on the Ifsttar's runway too with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
@@
 A special case is added where the root mean square of \textit{all} the spectrogram is applied (each audio file generates one element $k$ of $\mathbf{W}$).
-An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage ; see Figure \ref{fig:specW}.
-In the case where $w_t$ = 1 second , 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element ; see Figure \ref{fig:ElementW}.
+An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage, see Figure \ref{fig:specW}.
+In the case where $w_t$ = 1 second, 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element, see Figure \ref{fig:ElementW}.
 
@@
     \caption{Example of the dictionary building with a 3 second extract of a car passage (\ref{fig:specW}).
-With $w_t$ = 1 second, 3 spectrum are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is resume to 1 spectra (\ref{fig:ElementW}).}
+With $w_t$ = 1 second, 3 spectra are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced to 1 spectral vector.
+(\ref{fig:ElementW}).}
     \label{fig:spec_elementW}
@@
 
-Since the number of elements given by this processing can be high, in order to reduce the computational time and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100, 200\right]$.
+\ml{singulier spectrum, pluriel spectra}
+
+Since the number of elements given by this processing can be high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100, 200\right]$.
 
@@
 Table \ref{tab:experimental_factors} summarizes the experimental factors ($K$ and $w_t$) for the dictionary building and their related values.
+This last step allows us to reduce the dimensionality while preserving a rich descrption of the spectral content.
+Experimental validation consistently showed that considering octave bands do not impact the performance for the two estimators studied in this paper.
 
@@
 
-SUP and Sem NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction.
+Sup NMF and Sem NMF updates are conmputed for 400 iterations which is sufficient to reach convergence.
 TI NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+
+\ml{ca c'est pas genial, il aurait mieux vallu utiliser un coefficient de pondération pour que W n'avance pas trop vite...}
 
@@
 This coarser method allows the matrix reduction and decrease the computation time.
-But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
-The case of a linear frequency axis has been investigated and offer similar results which were not better.
+But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.% The case of a linear frequency axis has been investigated and offer similar results which were not better.
 
-For TI NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:experimental_factors} summarizes the experimental factors and their related values.
+For TI NMF, the threshold is set between 0.20 and 0.70 with a 0.01 grid search step. Table \ref{tab:experimental_factors} summarizes the experimental factors and their related values.
 
@@
 \subsubsection{Metrics}
-The performances of the two estimators (filter and NMF) of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
+The performance of the two estimators (filter and NMF) of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
 It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
@@
 In all, according Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
-For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for SUP and Sem NMF, 3780 associations of factors are made.
+For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup NMF and Sem NMF, 3780 associations of factors are made.
 For TI NMF, 8400 combinations can be calculated.
@@
 The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline.
-Finally, it is the most sophisticated methods (S-S and TI NMF) that have better results.
+Finally, it is the most sophisticated methods (Sem NMF and TI NMF) that have better results.
 The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
@@
     \end{subfigure}
-    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (SUP (\ref{fig:TIR_class_sup}), S-S (\ref{fig:TIR_class_semi}) and TI (\ref{fig:TIR_class_TI} NMF))}
+    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup NMF (\ref{fig:TIR_class_sup}), Sem NMF (\ref{fig:TIR_class_semi}) and TI NMF (\ref{fig:TIR_class_TI}))}
     \label{fig:TIR_bar}

### 2018-01-12 paper/gloaguenNmfTraffic.tex
done proof read

@@
 \subsubsection{Metrics}
-The performance of the two estimators (filter and NMF) of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$).
+
+The performance of the road traffic sound level estimator is assessed through the calculation of one reference metric, the Mean Absolute Error ($MAE$).
 It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
@@
 
-In all, according Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
+\ml{c'est les deux niveaux moyens sur toute la scène que l'on compare ?, quelle est la longueur des scenes ?}
+
+Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
 For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup NMF and Sem NMF, 3780 associations of factors are made.
-For TI NMF, 8400 combinations can be calculated.
+For TI NMF, 8400 combinations are computed.
 
@@
 Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
-For the low-passa frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expand to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
+For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
 \\
@@
 $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
-It is then the first baseline to use to compare the performances of NMF.
+It is then considered as the baseline to compare the performances of NMF.
 
@@
 For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 dB and are overestimate ; see Figure \ref{fig:alert_-12}.
-In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels ; see Figure \ref{fig:alert_12}.
+The errors are then nearly all superior to 4 dB and are overestimate, see Figure \ref{fig:alert_-12}.
+In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels, see Figure \ref{fig:alert_12}.
 The 500 Hz filter finds a balance between what it is put aside in low $TIR$ and what it is remained in high $TIR$.
@@
 The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline.
-Finally, it is the most sophisticated methods (Sem NMF and TI NMF) that have better results.
+The Sem NMF and TI NMF have better results.
 The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
-On the other hand, the semi-supervised approach has a higher error but yet with a lower standard deviation.
+On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation.
 
@@
 the error are too important for low $TIR$ and this for all the sub-classes.
-This method reveals to be to rigid as $\mathbf{W}$ is composed of fixed traffic spectrum.
+This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectrum.
 In the case of low $TIR$, in the aim to reduce the objective function, see eq.
@@
 This behavior disappear when the traffic component become predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}.
-Thus composed the dictionary of only fixed traffic spectrum is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
+Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
 
-Consequently, with the add of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$.
+Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$.
 It brings a significant decrease of the errors for low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}.
-The $\mathbf{W_r}$ composition, in the case of an \textit{alert} sub-class, is displayed in Figure \ref{fig:Y_alert-12} and this shows that the interfering class is easily integrated.
+The content of $\mathbf{W_r}$, in the case of an \textit{alert} sub-class is displayed in Figure \ref{fig:Y_alert-12}.
+The interfering class is easily integrated.
 The first element is mainly composed of high frequency bands which correspond to the car horns of the scene.
-This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the traffic basics are no longer activated when the car horn sounds.
+This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the elements of the dictionnary dedicated to the traffic are no longer activated when the car horn is active.
 
@@
 
-However, the degrees of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
-Indeed, in order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$.
+However, the relatively high degree of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$.
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
 In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horn have disappeared for the benefit of low frequency content.
-Consequently, the traffic sound level estimation is then underestimate as in Figure \ref{fig:lp_alert_12}.
+Consequently, the traffic sound level estimation is then underestimated as in \ml{rappler la condition experimentale}, see Figure \ref{fig:lp_alert_12}.
 \\
@@
 
-Finally, TI NMF with a threshold fixed at $t = 0.54$ and $\beta$ = 2, offers the lowest average results (Table \ref{tab:results}).
-If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.
-Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to be adjusted to the scene and to adapt to the different sound environments.
+Finally, TI NMF with a threshold $t = 0.54$ and $\beta$ = 2, offers the lowest average error (Table \ref{tab:results}).
+\ml{enlever ou reecrire: If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.} Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
 The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes for $TIR$ = [-12, 12] dB.
+In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR$ = [-12, 12] dB.
 
@@
 
-For low $TIR$, as the traffic sound class is not predominant, the final dictionary differs a lot from the $\mathbf{W_0}$.
-With the thresholding, only a reduce number of basis vectors are considered as traffic components.
+For low $TIR$, as the traffic sound class is not predominant, the final dictionary stronglyb differs from the $\mathbf{W_0}$.
+With the thresholding, only a reduced number of basis vectors are considered as traffic components.
 In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.
-However, for \textit{climate} and \textit{mechanics}, the error stay important as these interfering classes have similar spectrum.
-On high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB).
+However, for \textit{climate} and \textit{mechanics}, the error remain important as these interfering classes have similar spectral profiles when compared to traffic ones.
+
+For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB).
 The kept elements are then more suited to the scenes than a fixed dictionary.
@@
 
-In this work, NMF, a source separation method, was used to estimate the road traffic sound level in urban sound mixtures.
-It is an adapted approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
+In this work the non negative matric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures.
+It is a well suited approach to these sound environments because it easily take into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
 Different versions of NMF have been studied as a supervised and semi-supervised approach.
 On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
-The proposed new approach, named Threshold Initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the traffic elements are extracted by hard thresholding, proposes the lowest average error.
-It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
+The proposed approach, named threshold initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the dictionary elements that are similar to the traffic are extracted by hard thresholding, achieves the lowest average error.
+%It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
 
-This method has the advantage to keep a low calculation cost and then can be implemented in embedded sensors.
-Furthermore, this approach stay adapted for other sound sources, as whistling bird, depending on what spectrum are put in $\mathbf{W}$ and can be adapted for other application than the road traffic sound level estimation.
-Finally, the add of constraints as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} can easily be implemented and taken into account.
+A major advantage of the proposed approach is that it is not designed for a specific source.
+Eventhough the experiments described in this paper focused on traffic sounds, changing the dictionarry to contain bird sounds will lead to an estimator of the presence of birds.
 
-In parallel, this experience allowed the design of a large sound database composed of multiples samples of urban sounds and sound mixtures that are made available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for research communities dedicated to the detection, separation and recognition tasks of urban sound sources.
-The experience has been lead with the Matlab software.
-In order to be a reproducible experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} used for this experience are available online.
+Extending the apporach to other sources is thus of intetrest for future research.
+Also, performance improvement could be achieved by the addition of constraints such as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
+
+The experimental protocol and the evaluated estimators has been implemented with the Matlab software.
+For reproducibility purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}.
+The evaluation database composed of multiples samples of urban sounds and sound mixtures is also available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for the research community with interest in detection, separation and recognition tasks of urban sound sources.
 

### 2018-01-18 paper/gloaguenNmfTraffic.tex
AC and ML corrections

@@
-\documentclass[twocolumn,a4paper,10pt]{article}
+\documentclass[a4paper,10pt]{article}
 
@@
 
-The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring the acoustic quality of urban areas.
-Assuming that transportation is the major source of annoyance, state of the art approaches consider as input an estimate of the number and the speed of vehicules in major traffic lanes.
+The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring of the acoustic quality of urban areas.
+State of the art approaches target road traffic noise maps and consider, as input, an estimate of the number and the speed of vehicles in major traffic lanes.
 Follows a prediction procedure that outputs an acoustic pressure level at any location in the modeled area.
@@
 Considering as input the acoustic pressure measured in many locations using a sensor grid approach would greatly complement and improve the quality of the predicted pressure values.
-Among the technical issues that raises this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.
+Among the technical issues that raise this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.
 
-In this paper, several techniques based on the non-negative factorization framework are studied in this application scenario.
-The task being to the best of our knowledge never been considered in the literature, we propose an exprimental protocol to validate the studied approaches that complies with standard reproducible research recommandations.
+In this paper, several techniques based on non-negative factorization framework are studied in this application scenario on a simulated sound scene corpus.
+The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
 
@@
 With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
-These maps depict an estimation of the number of city dwellers exposed to high noise levels.
-They are used to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}.
-These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques.
-They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} \ml{un peu hors propos on améliore rien la dessus : or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}.
-plutot parler du cote sparse des données d'entrées}.
-In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day.
-Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
-These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels.
+These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels.
+These maps play both an important communication role and help drawing up action plans to reduce noise exposure.
+Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling and express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
+Altough very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
+In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
+The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
+These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make available of the long-term evolution of the traffic noise levels.
 It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
@@
 Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels.
-DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations.
-It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
-In a similar way, but reduced to few neighborhoods \ml{with a more dense topology ?
-parler aussi de sonyc}, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
+The DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations focusing on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
+The SONYC project (Sounds Of New-York City) aims to deploy a sensor network in New-York City for the purpose of monitoring constantly the noise pollution in the city \cite{mydlarz2017implementation}.
+In order to better know the urban sound environment, sensors are coupled with a detection tool that identifies the sound sources present \cite{salamon2017deep}.
+In a similar way, but reduced to few neighborhoods with a denser network, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.
 
-If sensor networks could improve road traffic noise estimates compared with simulated maps, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
-Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap.
+Prior to data assimiliation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
+Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap.
 Consequently, the traffic sound level estimation based on measurements is not a trivial task.
@@
 The main issue in the detection or recognition tasks is the overlap of environmental sounds.
-Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels.
+Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels.
 To circumvent this issue, Socoro et al.
@@
 Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.
-\ml{arrete :) If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.}
-
-An alternative approach that we will foolow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level.
+An alternative approach that we will follow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level, see Figure \ref{fig:diagram}.
 That is, separating the contribution of the traffic from the other sources within a polyphonic scene.
@@
 
-ml{Supprimer : One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}.
-The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations.
-This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.
-In an urban environment context, this method is used with acoustic microphone arrays and beamforming \cite{saruwatari2003blind}.
-These approaches consider a set of microphones and allows the localization \cite{mennitt2010multiple} or the detection \cite{leiba2017large} of sound sources from the phase shift and the distance between the microphones.
-However, one need a lot of microphones.
-Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.}
-
-A convenient method for monophonic signal decomposition is the Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
+In an urban environment context, source separation can be achieved with the help of acoustic microphone arrays and beamforming \cite{saruwatari2003blind}.
+However, this approach requires spreading multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance.
+This method is then not considered here to be deployed all over cities.
+In the opposite, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present.
+A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
 When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}.
-In the audio processing domain, the NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
-By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane.
-\ml{enlever : For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.}
-
+In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
+By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be resolved on the time/frequency plane.
 Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}.
-After having performed NMF on simulated audio files, they perform a source separation based on NMF in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
-
-\ml{ enlever : A $k$-mean clustering allows the separation of the $k$ sound sources.
-This approach, compared to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.
-Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.}
-
-We study in this paper different flavors of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event.
-We demonstrate the supervised NMF and semi supervised NMF approaches have some interests but fail to give satifactory results for the application at hand.
-We thus introduce another NMF scheme called thresholded initialized NMF that make good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).
+After having performed NMF on simulated audio files, they perform a source separation in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
+We study in this paper different flavors of Non-Negative Matrix Factorization where traffic component is considered in its entirely whether it is a sound background or event.
+We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
+We thus introduce another NMF scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).
 
 To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
-The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would be difficult to extract from a recording of an urban scene.
+The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
 
@@
 \includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
-\caption{Block diagram of the source separation method. \ml{toujours citer une figure}}
+\caption{Block diagram of the general source separation model}
 \label{fig:diagram}
@@
 Section \ref{part:nmf} details the technical aspects of NMF and describes the 3 approaches considered in this paper to achieve the task at hand.
-Section \ref{part:protocol} describes the environmental sound scene corpus and the experimental protocol setup.
+Section \ref{part:protocol} describes the corpus of environmental sound scenes and the experimental protocol setup.
 Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.
@@
 
-Non-negative Matrix Factorization is a linear approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
+Non-negative Matrix Factorization is a linear approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram $\mathbf{\tilde{V}}$ (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :
 
@@
 
-where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$.
+where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectra and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$.
 An illustrative example can be found in Figure  \ref{fig:example_NMF}.
@@
 and usually belongs to the $\beta-$divergence class \cite{fevotte_nonnegative_2009} in which the well known Euclidean distance (eq.
-\ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq.
-\ref{eq:def_divKL}) and the Itakura-Saito divergence belong
+\ref{eq:def_distEUC}) and the Kullback-Leibler divergence (eq.
+\ref{eq:def_divKL}) belong
 
@@
 
-Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.
+Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007} or the projected gradient \cite{lin_projected_2007}.
+Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.
 
 \subsection{Supervised NMF}
-First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectrum of urban sound sources.
-In the urban environments, a lot of different sound sources present are known whose spectrum can be obtained and be a basis of $\mathbf{W}$.
-The \textit{activator} are then the unknown to estimate.
+First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectra of urban sound sources.
+A lot of the different sound sources present in the urban environment are known whose their spectra can be obtained and be a basis of $\mathbf{W}$.
+The \textit{activation} matrix is then the unknown variable to estimate.
 In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm
@@
 
-The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasnable assumption in our application scenario.
-To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of intereferences that can occur in a realistic scenario.
-To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offer more flexibility.
-This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectrum, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization.
-In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionnary \cite{fevotte ?} \ml{mettre les parametres dnas la partie experimentale Here, $J = 2$}.
-The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic.
-The problem (\ref{eq:nmf}) becomes
+The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasonable assumption in our application scenario.
+To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario.
+To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offers more flexibility.
+This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectra, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization.
+In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionary %\cite{fevotte ?}
+. The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic. The problem (\ref{eq:nmf}) becomes
 
@@
 
-Applications of Sem NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.
+Applications of Sem-NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.
 
@@
 
-As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fails to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.
+As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fail to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.
 
 We therefore propose an alternative scheme based on the unsupervised NMF.
-Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly.
-Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
+Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, are initialized randomly.
+Here, as the concerned sound source is known and audio samples of car passages are available, an initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
 Then NMF is performed where $\mathbf{W}$ (eq.
@@
 
-$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are significantly different.
+$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are completely different.
 This measure is bounded between 1 and 0 and is an invariant with respect to scale.
-\ml{ $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order.
-} The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
+$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order.
+The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
 It is defined as:
@@
 
-An illustrative example is shown on Figure \ref{fig:W_TI_NMF}.\\
+An illustrative example is shown in Figure \ref{fig:W_TI_NMF}.\\
 
@@
 
-This approach is named \textit{Thresholded inititialized NMF} (TI NMF).
+This approach is named \textit{Thresholded inititialized NMF} (TI-NMF).
 Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to process the similarity using a sigmoïd or a radial basis functions have been investigated.
@@
 
-In order to validate the usefulness of the proposed NMF scheme to estimate the road traffic noise level, one need to compare the traffic noise level predicted by the algorithm to a reference level.
+In order to validate the usefulness of the proposed NMF scheme to estimate the road traffic noise levels, one need to compare the traffic noise level predicted by the algorithm to a reference level.
 The latter can hardly be measured or even annotated from real life recordings.
 Thus,  we propose to consider simulated sound scenes to assess the performance of the proposed NMF.
-This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known allowing the computation of the reference level.
+This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known, thus allowing the computation of the reference level.
 %Thus, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.
@@
 \textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database.
-The simulator have been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \ml{voir ma page de publi, taslp avec gregoire, taslp avec emmanouil, taslp avec mesaros}.
+The simulator has been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \cite{lafay:hal-01111381} \cite{benetos:hal-01520194} \cite{mesaros:hal-01650601}.
 
-This database is divided into two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} or \textit{texture} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
-154 sound samples that belong to this category are divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
-The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway on different speeds with multiple gear ratio.
+This database is divided into two categories: $i)$ the \textit{event} category, which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, whistling bird, sweeping broom, car horn, passing car, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, metallic noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} or \textit{texture} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time.
+154 sound samples that belong to this category are divided in 9 sound classes (\textit{whistling bird, construction site noise, crowd noise, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
+The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots to allow some diversity in the resulting mixture).
+Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture).
 The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
@@
     \end{subfigure}
-    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain}) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue):
+    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue)}
     \label{fig:example_simScene}
@@
 Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult.
-The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
+The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling bird), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
 In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the other sound classes.
@@
 
-with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$.
-When $TIR = -12$, the traffic component is then less present than the interfering class.
-When $TIR = 12$ it is the opposite: the trafffic class is louder than the interfering class.
+with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$ dB.
+When $TIR < 0$ dB, the traffic component is then less present than the interfering class.
+In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
 The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values).
-
-\ml{The corpora are available for download at: \url{}}
+The corpora are available for download at: \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}
 
@@
 
-The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
-The spectrogram $\mathbf{V}$ of each sound is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap \ml{enlever and a number of point $nfft = 2^{12}$ ($N$ = 644)}.
+The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
+The spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap, see Figure \ref{fig:bloc_experiment}.
 
-Assuming that the traffic spectral profile is largely concentrated in the low regions, a first estimator to determine the traffic sound level is a frequency low-pass filter.
-It depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}).
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/bloc_diagram_estimator.pdf}
+    \caption{Block diagram of the experience with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF}
+    \label{fig:bloc_experiment}
+\end{figure}
+
+Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter.
+It depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
 The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:NMF}.
-Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}).
+The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:nmf}.
+Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}), each experimental factor having multiples modalities.
 
 \begin{figure}
-
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
-  		\caption{}
-  		\label{fig:bloc_filtre}
-    \end{subfigure}%
-    \vfill
-    \begin{subfigure}[t]{0.45\textwidth}
     \centering
     \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
-		\caption{}
+    \caption{Block diagram of the NMF experience with the dictionary design composed from a second sound database}
     \label{fig:bloc_nmf}
-    \end{subfigure}
-
-    \caption{Block diagrams for the frequency low-pass filter (a) and for NMF (b) \ml{je ferais deux figures séparées, une avec le protocole avec une boite estimateur en disant dans la caption que l'estimateur est soit le filtre soit la nmf.
-deuxieme figure avec la nmf.
-Tu utilises des boites et des couleurs differentes, si cela a un sens explicite le sinon enlève, met les couleurs plus claires.}}
 \end{figure}
@@
 \centering
-\begin{tabular}{lcccccc}
-\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \toprule
-$\mathbf{K}$ & 25 & 50 & 100 & 200 & & \\
-\rowcolor[HTML]{EFEFEF}
-$\mathbf{w_t}$ (s)& \multicolumn{2}{c}{0.5} & \multicolumn{2}{c}{1}  & \multicolumn{2}{c}{\textit{all}} \\
-$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\
-\rowcolor[HTML]{EFEFEF}
-\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\
-$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\
-\rowcolor[HTML]{EFEFEF}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
+	\hline
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities \\ \toprule
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}}
+\rowcolor[HTML]{C0C0C0} 
+   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 \\
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
    $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\
-\textbf{method} & \multicolumn{2}{c}{filter} & Sup NMF & Sem NMF & TI NMF \\
-\rowcolor[HTML]{EFEFEF}
-$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \bottomrule
-\end{tabular}
-\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their evaluated values}
-\label{tab:experimental_factors}
+   \bottomrule
+\end{tabularx}
+
+\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator}
+\label{tab:experimental_factorsFilter}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
+	\hline
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities \\ \toprule
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}}
+\rowcolor[HTML]{C0C0C0} 
+   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 \\
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}}
+  \textbf{method} &  Sup-NMF & Sem-NMF & TI-NMF \\
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}}
+\rowcolor[HTML]{C0C0C0}
+    $\mathbf{w_t}$ (s)& 0.5 & 1 & \textit{all} 
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}}
+    $\mathbf{K}$ & 25 & 50 & 100 & 200 \\
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{6cm}@{}@{}C{6cm}@{}}
+\rowcolor[HTML]{C0C0C0}
+   $\mathbf{\beta}$ & 1 & 2 \\
+\end{tabularx}
+
+\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
+   $\mathbf{t}$ & 0.30:0.01:0.70 \\
+   \bottomrule
+\end{tabularx}
+\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
+\label{tab:experimental_factorsNMF}
 \end{table*}
@@
 It is composed of 53 audio files of passing cars.
-These recordings have been made on the Ifsttar's runway too with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
+These recordings have been made on the Ifsttar's runway too, with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
-Then time/frequency windows of $F \times w_t $ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file.
-$w_t$ is fixed at $wt = [0.5~1]$ second.
-In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectra of $F \times 1$ dimension.
+Then time/frequency windows of $F \times w_t $ dimension are applied without overlapping on the spectrogram in order to consider several spectra for each audio file.
+$w_t$ is fixed at $w_t = [0.5~1]$ second.
+In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectrum a of $F \times 1$ dimension.
 A special case is added where the root mean square of \textit{all} the spectrogram is applied (each audio file generates one element $k$ of $\mathbf{W}$).
 An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage, see Figure \ref{fig:specW}.
-In the case where $w_t$ = 1 second, 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element, see Figure \ref{fig:ElementW}.
+In the case where $w_t$ = 1 second, 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced into one element, see Figure \ref{fig:ElementW}.
 
@@
     \caption{Example of the dictionary building with a 3 second extract of a car passage (\ref{fig:specW}).
-With $w_t$ = 1 second, 3 spectra are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced to 1 spectral vector.
+With $w_t$ = 1 second, 3 spectra are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector.
 (\ref{fig:ElementW}).}
@@
 
-\ml{singulier spectrum, pluriel spectra}
-
-Since the number of elements given by this processing can be high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100, 200\right]$.
+Since the number of elements given by this processing can be high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectra to $K = \left[ 25, 50, 100, 200\right]$.
 
 The obtained dictionary is expressed with third octave bands and each basis vector of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.
-Table \ref{tab:experimental_factors} summarizes the experimental factors ($K$ and $w_t$) for the dictionary building and their related values.
-This last step allows us to reduce the dimensionality while preserving a rich descrption of the spectral content.
+Table \ref{tab:experimental_factorsNMF} summarizes the experimental factors ($K$ and $w_t$) for the dictionary building and their related modalities.
+This last step allows us to reduce the dimensionality while preserving a rich description of the spectral content.
 Experimental validation consistently showed that considering octave bands do not impact the performance for the two estimators studied in this paper.
@@
 
-Sup NMF and Sem NMF updates are conmputed for 400 iterations which is sufficient to reach convergence.
-TI NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-
-\ml{ca c'est pas genial, il aurait mieux vallu utiliser un coefficient de pondération pour que W n'avance pas trop vite...}
+Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
+TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+%\ml{ca c'est pas genial, il aurait mieux vallu utiliser un coefficient de pondération pour que W n'avance pas trop vite...}
 
@@
 This coarser method allows the matrix reduction and decrease the computation time.
-But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.% The case of a linear frequency axis has been investigated and offer similar results which were not better.
+But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a 0.01 grid search step.
+Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
+Considering the experimental settings derived from the different modalities of each experimental factor described in Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}, 30420 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
+For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup-NMF and Sem-NMF, 1440 associations of factors are made.
+For TI-NMF, 28800 combinations are computed.
 
-For TI NMF, the threshold is set between 0.20 and 0.70 with a 0.01 grid search step. Table \ref{tab:experimental_factors} summarizes the experimental factors and their related values.
+After having performed NMF on one scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated.
+For each set of unique modalities of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained.
 
@@
 \centering
-\begin{tabular}{@{}cccccc@{}}
+\begin{tabular}{@{}ccccccc@{}}
 \toprule
-\textbf{method} & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
-filter 20 kHz &  &  &  &  & 4.69 ($\pm$ 4.52) \\
-filter 0.5 kHz&  &   &  &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-Sup NMF &  1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
-Sup NMF &  2 & 50 & 0.5  &   & 3.02 ($\pm$ 3.33) \\ \hline \hline
-Sem NMF &  1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
-Sem NMF &  2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
-TI NMF &  1 & 200 & 0.5 &  0.42 &\textbf{2.19 ($\pm$ 2.18)} \\
-TI NMF &  2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
+\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
+filter & 20  &  &  &  &  & 4.69 ($\pm$ 4.52) \\
+filter & 0.5 &  &  &  &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
+Sup-NMF &  & 1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
+Sup-NMF &  & 2 & 50 & 0.5  &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
+Sem-NMF &  & 1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
+Sem-NMF &  & 2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
+TI-NMF &  & 1 & 200 & 0.5 &  0.42 &\textbf{2.19 ($\pm$ 2.18)} \\
+TI-NMF &  & 2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
 \end{tabular}
@@
 \toprule
-\textbf{method} & filter & filter & Sup NMF & Sem NMF & TI NMF \\ \midrule
-$f_c$ (kHz) & 20 & 0.5 & 20 & 20 & 20 \\
+\textbf{method} & filter & filter & Sup-NMF & Sem-NMF & TI-NMF \\ \midrule
+$f_c$ (kHz) & 20 & 0.5 &  &  &  \\
 $\mathbf{\beta}$ &  &  & 2 & 2 & 1 \\ \hline
@@
 
-The performance of the road traffic sound level estimator is assessed through the calculation of one reference metric, the Mean Absolute Error ($MAE$).
+The performance of the road traffic sound level estimator is assessed through the calculation of one reference metric, the Mean Absolute Error ($MAE$) \cite{willmott2005advantages}.
 It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,
@@
 
-\ml{c'est les deux niveaux moyens sur toute la scène que l'on compare ?, quelle est la longueur des scenes ?}
-
-Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
-For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup NMF and Sem NMF, 3780 associations of factors are made.
-For TI NMF, 8400 combinations are computed.
-
 \section{Results}\label{part:results}
 
-Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
+Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the lowest $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
 For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
@@
 Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
-$f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained.
+$f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained.
 It is then considered as the baseline to compare the performances of NMF.
 
-In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, this filter is efficient as it removes these frequency components.
-For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 dB and are overestimate, see Figure \ref{fig:alert_-12}.
+In low $TIR$, for \textit{alert} and \textit{animals}, which are sub-classes composed of higher frequencies, this filter is efficient as it removes these frequency components.
+For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then does not dissociate the traffic element from the other sound sources.
+The errors are then nearly all superior to 4 dB and are overestimated, see Figure \ref{fig:alert_-12}.
 In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels, see Figure \ref{fig:alert_12}.
-The 500 Hz filter finds a balance between what it is put aside in low $TIR$ and what it is remained in high $TIR$.
+The 500 Hz filter finds a balance between what is put aside in low $TIR$ and what it is remained in high $TIR$.
 \\
@@
 
-Compare to the filter errors, the choice of some NMF approaches make it possible to decrease the road traffic sound level estimation.
+Compared with the filter errors, the choice of some NMF approaches make it possible to decrease the error of the road traffic sound level estimation.
 The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline.
-The Sem NMF and TI NMF have better results.
-The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
+Sem-NMF and TI-NMF have better results.
+The lowest average error is obtained for TI-NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
 On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation.
@@
         \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sup NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2}
                 \label{fig:TIR_class_sup}
@@
       	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sem NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2}
                 \label{fig:TIR_class_semi}
@@
         \includegraphics[width=\linewidth]{figures/TI_bar}
-        \caption{$MAE$ error for each $TIR$ and sub-class with TI NMF, $\beta$ = 1 and $t$ = 0.42}
+        \caption{$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42}
         \label{fig:TIR_class_TI}
     \end{subfigure}
-    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup NMF (\ref{fig:TIR_class_sup}), Sem NMF (\ref{fig:TIR_class_semi}) and TI NMF (\ref{fig:TIR_class_TI}))}
+    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI}))}
     \label{fig:TIR_bar}
@@
 According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
-In the case of Sup NMF, it fails to improve the filtering performances despite good results in high $TIR$.
-the error are too important for low $TIR$ and this for all the sub-classes.
-This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectrum.
+In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$.
+The error are too important for low $TIR$ and this for all the sub-classes.
+This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
 In the case of low $TIR$, in the aim to reduce the objective function, see eq.
 \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the  sound scene.
-The case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}.
+In order to expose this issue, the case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}.
 Here when the car horn sounds, some elements of $\mathbf{W}$ are activated and are then considered as traffic component generating a wrong estimation of the sound level.
-This behavior disappear when the traffic component become predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}.
+This behavior disappears when the traffic component becomes predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}.
 Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
@@
 Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$.
-It brings a significant decrease of the errors for low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}.
+It brings a significant decrease of the errors for low $TIR$ as it can be seen in Figure \ref{fig:TIR_class_semi}.
 The content of $\mathbf{W_r}$, in the case of an \textit{alert} sub-class is displayed in Figure \ref{fig:Y_alert-12}.
@@
 
-    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for : at $TIR = -12$ (\ref{fig:lp_alert_-12}) and at $TIR$ (\ref{fig:lp_alert_12}).}
+    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ (\ref{fig:lp_alert_-12}) and $TIR$ (\ref{fig:lp_alert_12}).}
     \label{fig:lp_alert}
@@
 
-However, the relatively high degree of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
-In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$.
+However, the relatively high degree of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem-NMF is free to include traffic components in $\mathbf{W_r}$.
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
-In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horn have disappeared for the benefit of low frequency content.
-Consequently, the traffic sound level estimation is then underestimated as in \ml{rappler la condition experimentale}, see Figure \ref{fig:lp_alert_12}.
+In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horns have disappeared for the benefit of low frequency content.
+Consequently, the traffic sound level estimation is then underestimated as in Figure \ref{fig:lp_alert_12}.
 \\
@@
     \end{subfigure}
-    \caption{2 elements of $\mathbf{W_r}$ for an \textit{alert} scene for $TIR$ = -12 (\ref{fig:Y_alert-12}) and $TIR$ = 12 (\ref{fig:Y_alert_12})}
+    \caption{2 basis vectors of $\mathbf{W_r}$ for an \textit{alert} scene for $TIR$ = -12 dB (\ref{fig:Y_alert-12}) and $TIR$ = 12 dB (\ref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s}
 \end{figure*}
 
-Finally, TI NMF with a threshold $t = 0.54$ and $\beta$ = 2, offers the lowest average error (Table \ref{tab:results}).
-\ml{enlever ou reecrire: If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.} Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
+Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}).
+Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
 The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.54}
+    \includegraphics[width=.7\linewidth]{figures/dist_-12_12.pdf}
+    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $w_t$ = 0.5 s,$K$ = 200}
     \label{fig:dist_-12_12}
@@
 
-For low $TIR$, as the traffic sound class is not predominant, the final dictionary stronglyb differs from the $\mathbf{W_0}$.
+For low $TIR$, as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
 With the thresholding, only a reduced number of basis vectors are considered as traffic components.
-In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.
-However, for \textit{climate} and \textit{mechanics}, the error remain important as these interfering classes have similar spectral profiles when compared to traffic ones.
+In comparison to supervised results, this approach reduces significantly the error for the \textit{human} and \textit{transport} sub-classes.
+However, for \textit{climate} and \textit{mechanics}, the errors remain important as these interfering classes have similar spectral profiles when compared to traffic ones.
 
-For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB).
+For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher what allows retaining more elements as traffic components and then decreases the error (< 1 dB).
 The kept elements are then more suited to the scenes than a fixed dictionary.
 The error for these $TIR$ is then due to the thresholding.
-With a low threshold, it is possible to decrease the error (for $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30).
+With a low threshold, it is possible to decrease the error.
+For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30 while for $TIR$ = -12 dB, with $t$ = , the error decrease to $\pm$ .
+Having no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed and is the one that best balanced these opposite cases.
 
@@
 
-In this work the non negative matric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures.
-It is a well suited approach to these sound environments because it easily take into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
+In this work the non negative metric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures.
+It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
 Different versions of NMF have been studied as a supervised and semi-supervised approach.
 On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
-The proposed approach, named threshold initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the dictionary elements that are similar to the traffic are extracted by hard thresholding, achieves the lowest average error.
-%It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.
-
+The proposed approach, named threshold initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectra, updated and where the dictionary elements that are similar to the traffic are extracted by hard thresholding, achieves the lowest average error.
 A major advantage of the proposed approach is that it is not designed for a specific source.
-Eventhough the experiments described in this paper focused on traffic sounds, changing the dictionarry to contain bird sounds will lead to an estimator of the presence of birds.
-
+Even though the experiments described in this paper focused on traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
 Extending the apporach to other sources is thus of intetrest for future research.
@@
 For reproducibility purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}.
-The evaluation database composed of multiples samples of urban sounds and sound mixtures is also available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for the research community with interest in detection, separation and recognition tasks of urban sound sources.
+The evaluation database composed of multiple samples of urban sounds is also made available for the research community with interest in detection, separation and recognition tasks of urban sound sources.
 
@@
 \bibliographystyle{unsrt}
-\bibliography{bibliographie_applied}
+\bibliography{bibliographie_sensors}
 

### 2018-01-18 experiment/NMFThreshold/nmfth1dictionary.m

@@ -11,7 +11,7 @@ function [config, store, obs] = nmfth1dictionary(config, setting, data)
 % Date: 01-Dec-2017                                                                
                                                                                    
 % Set behavior for debug mode                                                      
-if nargin==0, NMFThreshold('do', 1); return; else store=[]; obs=[]; end
+if nargin==0, NMFThreshold('do', 1,'mask',{1 1 1 1 1}); return; else store=[]; obs=[]; end
                                                                                    
 if strcmp(setting.type,'nmf')
     K = setting.numberElement;

### 2018-01-18 experiment/NonnegMatrixFact/nomafa1dictionary.m

@@ -11,7 +11,7 @@ function [config, store, obs] = nomafa1dictionary(config, setting, data)
 % Date: 13-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, NonnegMatrixFact('do', 1, 'mask',{1 1 1 1 1}); return; else, store=[]; obs=[]; end
+if nargin==0, NonnegMatrixFact('do', 1, 'mask',{1 2 1 0 1}); return; else, store=[]; obs=[]; end
 
 if strcmp(setting.type,'nmf')
     K = setting.numberElementW;

### 2018-01-18 experiment/NonnegMatrixFact/nomafa2estimation.m

@@ -48,12 +48,12 @@ switch dataset
 end
 
 files = dir(strcat(creationSceneDir, filesep, '*.wav'));
-globalName = cell(1,length(files)/3);
+globalName = cell(1,length(files)/2);
 ind = 1;
 
 for ii = 1:length(files)
-    if isempty(strfind(files(ii).name,'traffic')) && isempty(strfind(files(ii).name,'interfering'))
-        globalName{ind} = files(ii).name(1:end-4);
+    if strfind(files(ii).name,'traffic')
+        globalName{ind} = files(ii).name(1:end-12);
         ind = ind+1;
     end
 end

### 2018-01-18 experiment/NonnegMatrixFact/nomafaReport.m

@@ -6,376 +6,4 @@ function config = nomafaReport(config)
 % Copyright: <userName>
 % Date: 13-Jan-2017
 
-%  if nargin==0, NonnegMatrixFact('report', 'r','reportName','filterAmbiance'); return; end
 if nargin==0, NonnegMatrixFact('report', 'r'); return; end
-
-switch config.reportName
-    case 'filterAmbiance'
-        config = expExpose(config, 't', 'mask', {1 0 0 0 0 ,...
-            0 0 0 2 1},'obs',[6 9],'integrate',{'aType','TPR'});
-        
-%         for ii = 1:5
-%             config = expExpose(config, 't', 'mask', {1 0 0 0 0 ,...
-%                 0 0 1 2 1,...
-%                 0 0 ii},'obs',5:9);
-%         end
- %       config = expExpose(config, 'b', 'mask', {1 0 0 0 0 ,...
-  %          0 0 1 2 1},'obs','rmseLeqdB','expand','TPR'); 
-        
-    case 'filterGraficSlides'       
-        %% according to cutOffFreq
-        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
-            0 0 1 2 0, ...
-            0 0 0 1},'obs','rmsedB','integrate','gType','rotateAxis',45);
-        config = expExpose(config, 't', 'mask', {0 0 0 0 0 ,...
-            0 0 1 2 0, ...
-            0 0 0 1},'integrate','gType','obs',[4 7 8]);
-        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
-            0 0 1 2 0, ...
-            0 0 0 1},'obs','rmsedB','expand','gType','rotateAxis',45);
-        
-        %% according to gType
-        config = expExpose(config, 'b', 'mask', {0 0 0 0 0 ,...
-            0 0 1 2 0, ...
-            0 0 0 1},'obs','rmsedB','integrate','cutOffFreq','rotateAxis',45);
-        config = expExpose(config, 't', 'mask', {0 0 0 0 0 ,...
-            0 0 1 2 0, ...
[…]

### 2018-02-01 paper/gloaguenNmfTraffic.tex
JFP corrections done

@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[twocolumn,a4paper,10pt]{article}
 
@@
 The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
+The results shows the interest of our proposed approach for a such sound environment as it improves the estimation of the road traffic sound level with a lowest error compared to basic methods.
 
@@
 These maps play both an important communication role and help drawing up action plans to reduce noise exposure.
-Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling and express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
+Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling.
+They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
 Altough very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
@@
 Prior to data assimiliation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
-Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap.
+Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap.
 Consequently, the traffic sound level estimation based on measurements is not a trivial task.
 Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
-A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
+A two-step process is generally followed : describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
 A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
@@
 To circumvent this issue, Socoro et al.
-propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component.
+propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017}.
+It consists in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component.
 Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.
 An alternative approach that we will follow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level, see Figure \ref{fig:diagram}.
-That is, separating the contribution of the traffic from the other sources within a polyphonic scene.
+It consists in separating the contribution of the traffic from the other sources within a polyphonic scene.
 One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\
@@
 After having performed NMF on simulated audio files, they perform a source separation in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
-We study in this paper different flavors of Non-Negative Matrix Factorization where traffic component is considered in its entirely whether it is a sound background or event.
+We study in this paper different flavors of Non-Negative Matrix Factorization where the traffic component is considered in its entirely whether it is a sound background or an event.
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
@@
 Section \ref{part:protocol} describes the corpus of environmental sound scenes and the experimental protocol setup.
-Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.
+Section \ref{part:results} presents and discusses the outcomes of the numerical results.
 
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
-\caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
+\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf_2.pdf}
+\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c))}
 \label{fig:example_NMF}
@@
 First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectra of urban sound sources.
-A lot of the different sound sources present in the urban environment are known whose their spectra can be obtained and be a basis of $\mathbf{W}$.
+A lot of the different sound sources present in the urban environment are known.
+Their spectra can be obtained and be a basis of $\mathbf{W}$.
 The \textit{activation} matrix is then the unknown variable to estimate.
-In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm
+In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm \cite{fevotte_algorithms_2011}
 
@@
 To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario.
-To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offers more flexibility.
+To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is a good candidate as it offers more flexibility.
 This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectra, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization.
@@
 \begin{equation}
-\mathbf{V} \approx \mathbf{W_s H_s}+ \mathbf{W_r H_r}
+\mathbf{V} \approx \mathbf{\tilde{V}} = \mathbf{W_s H_s}+ \mathbf{W_r H_r}
 \end{equation}
@@
 
-As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fail to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.
+As it will be demonstrated in the experimental results described in Section \ref{part:results}, those last approaches fail to provide consistent results in a wide range of urban areas and for different traffic preponderances due to generalization capabilities issues.
 
@@
 \ref{eq:updateH_Sup}) are updated alternatively.
-$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study.
+$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study,
 
 \begin{equation}\label{eq:updateW_unsup}
-\textbf{W}^{(i+1)} \leftarrow \mathbf{W}^{(i)}.\left(\frac{\left[\left(\mathbf{W}^{(i)}\mathbf{H} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H}^T}{\left[\mathbf{W}^{(i)}\mathbf{H} \right]^{(\beta-1)}\mathbf{H}^T}\right)^{\gamma(\beta)}
+\textbf{W}^{(i+1)} \leftarrow \mathbf{W}^{(i)}.\left(\frac{\left[\left(\mathbf{W}^{(i)}\mathbf{H} \right)^{(\beta-2)}.\mathbf{V} \right]\mathbf{H}^T}{\left[\mathbf{W}^{(i)}\mathbf{H} \right]^{(\beta-1)}\mathbf{H}^T}\right)^{\gamma(\beta)}.
 \end{equation}
@@
 \begin{equation}
-D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = \frac{\mathbf{W}.\mathbf{W_0}}{\vert \vert \mathbf{W}  \vert \vert . \vert \vert \mathbf{W_0} \vert \vert}.
+D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = \frac{\mathbf{W_0}_k.\mathbf{W}_k}{\vert \vert \mathbf{W_0}_k  \vert \vert . \vert \vert \mathbf{W}_k \vert \vert}.
 \end{equation}
 
-$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are completely different.
+$D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
 This measure is bounded between 1 and 0 and is an invariant with respect to scale.
-$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order.
-The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method.
+The $k$ value of $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ are next sorted in descending order.
+The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are selected by a \textit{hard thresholding} method.
 It is defined as:
@@
 \begin{equation}
-\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W}_{0,k} \vert \vert \mathbf{W}_{k} \right) > t
+\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_{k} \right) > t.
 \end{equation}
@@
 This approach is named \textit{Thresholded inititialized NMF} (TI-NMF).
-Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to process the similarity using a sigmoïd or a radial basis functions have been investigated.
-A fast parametric study revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
+Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} have been investigated.
+A fast parametric study revealed that the \textit{hard} thresholding method, as presented in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
 
@@
 Thus,  we propose to consider simulated sound scenes to assess the performance of the proposed NMF.
-This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known, thus allowing the computation of the reference level.
-%Thus, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.
+This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic components are known, thus allowing the computation of the reference level.
 
@@
 The corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}.
-\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database.
+\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates monaural sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database.
 The simulator has been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \cite{lafay:hal-01111381} \cite{benetos:hal-01520194} \cite{mesaros:hal-01650601}.
@@
 154 sound samples that belong to this category are divided in 9 sound classes (\textit{whistling bird, construction site noise, crowd noise, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
+These sounds are in .wav format sampled at 44.1 kHz.
 The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture).
+Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture.
 The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
@@
     \end{subfigure}
-    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue)}
+    \caption{Example of a simple scene created with \textit{SimScene} software (the spectrogram (\subref{fig:simScene_spec}), the time domain \subref{fig:simScene_time})) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue)}
     \label{fig:example_simScene}
@@
 
-with the \textit{Traffic Interference Ratio} $TIR = \left[-12, -6, 0, 6, 12\right]$ dB.
+with the \textit{Traffic Interference Ratio} $TIR \in$ [-12 -6 0 6  12] dB.
 When $TIR < 0$ dB, the traffic component is then less present than the interfering class.
@@
 Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter.
-It depends only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz.
+It depends only on the cut-off frequencies $f_c \in $ [500 1k 2k 5k 10k 20k] Hz.
 The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities \\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities & \begin{tabular}[c]{@{}l@{}}number  \\ of level\end{tabular}\\ \toprule
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
-    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics  & 6
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}C{2cm}@{}}
 \rowcolor[HTML]{C0C0C0} 
-   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 \\
+   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 & 5\\
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
-   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20  & 6\\
    \bottomrule
@@
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}@{}C{2cm}}
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities \\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities & \begin{tabular}[c]{@{}l@{}}number \\ of modality\end{tabular}\\ \toprule
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
-    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics & 6 
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}C{2cm}@{}}
 \rowcolor[HTML]{C0C0C0} 
-   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 \\
+   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 & 5 \\
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}}
-  \textbf{method} &  Sup-NMF & Sem-NMF & TI-NMF \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
+  \textbf{method} & Sup-NMF & Sem-NMF & TI-NMF & 3 \\
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
 \rowcolor[HTML]{C0C0C0}
-    $\mathbf{w_t}$ (s)& 0.5 & 1 & \textit{all} 
+    $\mathbf{w_t}$ (s) & 0.5 & 1 & \textit{all} & 3 
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}}
-    $\mathbf{K}$ & 25 & 50 & 100 & 200 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+    $\mathbf{K}$ & 25 & 50 & 100 & 200  & 4\\
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{6cm}@{}@{}C{6cm}@{}}
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{6cm}@{}@{}C{6cm}@{}C{2cm}@{}}
 \rowcolor[HTML]{C0C0C0}
-   $\mathbf{\beta}$ & 1 & 2 \\
+   $\mathbf{\beta}$ & 1 & 2 & 2\\
 \end{tabularx}
 
-\begin{tabularx}{15cm}{L{3cm}@{}C{12cm}@{}}
-   $\mathbf{t}$ & 0.30:0.01:0.70 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
+   $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 40\\
    \bottomrule
@@
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
-Then time/frequency windows of $F \times w_t $ dimension are applied without overlapping on the spectrogram in order to consider several spectra for each audio file.
-$w_t$ is fixed at $w_t = [0.5~1]$ second.
+Then time/frequency windows of $F \times w_t $ dimensions are applied without overlapping on the spectrogram in order to consider several spectra for each audio file where $w_t \in [0.5~1]$ second.
 In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectrum a of $F \times 1$ dimension.
-A special case is added where the root mean square of \textit{all} the spectrogram is applied (each audio file generates one element $k$ of $\mathbf{W}$).
-An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage, see Figure \ref{fig:specW}.
-In the case where $w_t$ = 1 second, 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced into one element, see Figure \ref{fig:ElementW}.
+With this size of window, it is possible to obtain the characteristic pitches of the different audio samples.
+One obtain for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements.
+Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectra to $K \in \left[ 25, 50, 100, 200\right]$.
+The $K$ centroids are then the elements of the dictionary.
+A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) in order to build a dictionary with the spectral envelope of each audio sample.
+In this case, 53 spectra are obtained.
+The $K$-means clustering algorithm is then reduced to $K \in \left[25 50 \right]$.
 
@@
         \centering
-       \includegraphics[width=\linewidth]{figures/dictionary1.pdf}
+       \includegraphics[height=4.5cm]{figures/dictionary3.pdf}
        \caption{}
@@
         \centering
-       \includegraphics[width=\linewidth]{figures/dictionary2.pdf}
+       \includegraphics[height=4.5cm]{figures/dictionary4.pdf}
         \caption{}
@@
     \end{subfigure}
-    \caption{Example of the dictionary building with a 3 second extract of a car passage (\ref{fig:specW}).
-With $w_t$ = 1 second, 3 spectra are generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector.
-(\ref{fig:ElementW}).}
+    \caption{Example of the dictionary building with a 3 second extract of a car passage.
+In dashed lines, a 1 second $w_t$ window  (\subref{fig:specW}).
+With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW})}
     \label{fig:spec_elementW}
@@
 
-Since the number of elements given by this processing can be high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectra to $K = \left[ 25, 50, 100, 200\right]$.
+An example that illustrates the process can be found on Figure \ref{fig:spec_elementW} on a 3 second extract of the spectrogram of a car passage, see Figure \ref{fig:specW}.
+In the case where $w_t$ = 1 second, 3 elements are therefore extracted from the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced into one element, see Figure \ref{fig:ElementW}.
 
@@
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
-This coarser method allows the matrix reduction and decrease the computation time.
+This coarser method allows the reduction of the matrix size and decreases the computation time.
 But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
-For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a 0.01 grid search step.
+For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
 Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
-Considering the experimental settings derived from the different modalities of each experimental factor described in Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}, 30420 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF.
-For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup-NMF and Sem-NMF, 1440 associations of factors are made.
-For TI-NMF, 28800 combinations are computed.
 
-After having performed NMF on one scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated.
+Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed.
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ [0.5 1] s whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning}.
+For TI-NMF, 24000 combinations are computed.
+In all, 25380 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors
+
+After having performed NMF on one scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
+
+\begin{equation}
+\tilde{L}_{p,traffic} = 20\log_{10}\left(\frac{p_{rms}}{p_0}\right)
+\end{equation}
+
+where $p_{rms}$ is the effective pressure deducted from $\left[\mathbf{WH} \right]_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
 For each set of unique modalities of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained.
@@
 Sem-NMF &  & 2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
-TI-NMF &  & 1 & 200 & 0.5 &  0.42 &\textbf{2.19 ($\pm$ 2.18)} \\
+\textbf{TI-NMF} &  & \textbf{1} & \textbf{200} & \textbf{0.5} &  \textbf{0.42} &\textbf{2.19 ($\pm$ 2.18)} \\
 TI-NMF &  & 2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
 \end{tabular}
-\caption{Best results according to $\beta$ and \textit{method} experimental factors, in bold letter, the lowest error.}
+\caption{Best results for all the scenes according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
 \label{tab:results}
@@
 
-Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the lowest $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
+Table \ref{tab:results} summarizes, according to the 2 main factors (\textit{method}, $\beta$), the lowest $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
 For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
@@
 
-    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the sub-classes \textit{alert}: at $TIR = -12$ (\ref{fig:alert_-12}) and at $TIR$ (\ref{fig:alert_12}).}
+    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB(\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
     \label{fig:dictionaryExtraction}
@@
 
-Compared with the filter errors, the choice of some NMF approaches make it possible to decrease the error of the road traffic sound level estimation.
+Compared with the filter errors, the choice of some NMF approaches makes it possible to decrease the error of the road traffic sound level estimation.
 The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline.
 Sem-NMF and TI-NMF have better results.
-The lowest average error is obtained for TI-NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
-On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation.
+The lowest average error is obtained for TI-NMF, 2.19 $\pm$ 2.18, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
+On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation (2.32 $\pm$ 1.26).
 
@@
     \end{subfigure}
-    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI}))}
+    \caption{$MAE$ error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI}))}
     \label{fig:TIR_bar}
@@
 In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$.
-The error are too important for low $TIR$ and this for all the sub-classes.
+The errors are too important for low $TIR$ and this for all the sub-classes.
 This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
@@
 In order to expose this issue, the case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}.
-Here when the car horn sounds, some elements of $\mathbf{W}$ are activated and are then considered as traffic component generating a wrong estimation of the sound level.
-This behavior disappears when the traffic component becomes predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}.
+Here when the \textit{alert} sub-class sounds, some traffic elements of $\mathbf{W}$ are activated.
+The \textit{alert} sub-class is then considered as traffic component generating a wrong estimation of the sound level, see Figure \ref{fig:lp_alert_-12}.
+This behavior disappears when the traffic component becomes predominant to the interfering class as in Figure \ref{fig:lp_alert_12}.
 Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
 
-Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$.
+Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the interfering class in low $TIR$.
 It brings a significant decrease of the errors for low $TIR$ as it can be seen in Figure \ref{fig:TIR_class_semi}.
@@
 The first element is mainly composed of high frequency bands which correspond to the car horns of the scene.
-This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the elements of the dictionnary dedicated to the traffic are no longer activated when the car horn is active.
+This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the elements of the dictionary dedicated to the traffic are no longer activated when the car horn is active.
 
@@
 
-    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ (\ref{fig:lp_alert_-12}) and $TIR$ (\ref{fig:lp_alert_12}).}
+    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB(\subref{fig:lp_alert_-12}) and $TIR$ = 12 dB (\subref{fig:lp_alert_12})}
     \label{fig:lp_alert}
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
@@
         \label{fig:Y_alert-12}
-    \end{subfigure}%
-    \hfill
+    \end{subfigure}
+    \vfill
     \begin{subfigure}[t]{0.45\textwidth}
@@
     \end{subfigure}
-    \caption{2 basis vectors of $\mathbf{W_r}$ for an \textit{alert} scene for $TIR$ = -12 dB (\ref{fig:Y_alert-12}) and $TIR$ = 12 dB (\ref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s}
-\end{figure*}
+    \caption{Comparison between the sum of the 2 elements of $\mathbf{W_r}$ with the spectra of the interfering and traffic classes for an \textit{alert} scene for $TIR$ = -12 dB (\subref{fig:Y_alert-12}) and $TIR$ = 12 dB (\subref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s (normalized amplitudes)}
+\end{figure}
 
@@
 Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
-The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
+The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
 In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR$ = [-12, 12] dB.
@@
     \centering
-    \includegraphics[width=.7\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $w_t$ = 0.5 s,$K$ = 200}
+    \includegraphics[width=.8\linewidth]{figures/dist_-12_12.pdf}
+    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $K$ = 200 and $w_t$ = 0.5 s)}
     \label{fig:dist_-12_12}
@@
 The kept elements are then more suited to the scenes than a fixed dictionary.
-The error for these $TIR$ is then due to the thresholding.
+The error for these $TIR$ is then due to the thresholding which put aside some elements that are nevertheless related to the traffic component.
 With a low threshold, it is possible to decrease the error.
-For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30 while for $TIR$ = -12 dB, with $t$ = , the error decrease to $\pm$ .
-Having no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed and is the one that best balanced these opposite cases.
+For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30.
+In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease to 4.46 $\pm$ 1.66.
+In order to generalize this method and having no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 and is the one that best balanced these opposite cases.
 
@@
 In this work the non negative metric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures.
-It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks.
+It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and it is adapted to monophonic sensor networks.
 Different versions of NMF have been studied as a supervised and semi-supervised approach.
-On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
-The proposed approach, named threshold initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectra, updated and where the dictionary elements that are similar to the traffic are extracted by hard thresholding, achieves the lowest average error.
+On a large corpus of sounds, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
+The proposed approach, named threshold initialized NMF achieves the lowest average error.
+With this method, the $\mathbf{W}$ is initialized with road traffic spectra, updated and the dictionary elements that are similar to the road traffic spectra are then extracted by hard thresholding.
 A major advantage of the proposed approach is that it is not designed for a specific source.
@@
 
-The experimental protocol and the evaluated estimators has been implemented with the Matlab software.
-For reproducibility purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}.
+The experimental protocol and the evaluated estimators have been implemented with the Matlab software.
+For reproducible purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}.
 The evaluation database composed of multiple samples of urban sounds is also made available for the research community with interest in detection, separation and recognition tasks of urban sound sources.

### 2018-02-05 paper/gloaguenNmfTraffic.tex
minor improvements

@@
 The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
-The results shows the interest of our proposed approach for a such sound environment as it improves the estimation of the road traffic sound level with a lowest error compared to basic methods.
+The results show the interest of our proposed approach for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods.
 
@@
 Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels.
-To circumvent this issue, Socoro et al.
+To circumvent this issue, Socor\'o et al.
 propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017}.
@@
 where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectra and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$.
+As the constraint of non-negativity of $\mathbf{W}$ and $\mathbf{H}$ is considered, NMF allows only additive combinations between the element of $\mathbf{W}$.
+It is then a part-based representation that NMF provides.
 An illustrative example can be found in Figure \ref{fig:example_NMF}.
@@
 Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007} or the projected gradient \cite{lin_projected_2007}.
-Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.
+Here, the multiplicative update is chosen as it has been well studied in the literature and it ensures convergence of the results \cite{fevotte_algorithms_2011}.
 
@@
 This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectra, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization.
-In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionary %\cite{fevotte ?}
-. The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic. The problem (\ref{eq:nmf}) becomes
+In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionary \cite{lefevre2012semi}.
+The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic.
+The problem (\ref{eq:nmf}) becomes
 
@@
 We therefore propose an alternative scheme based on the unsupervised NMF.
-Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, are initialized randomly.
+Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly.
 Here, as the concerned sound source is known and audio samples of car passages are available, an initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
@@
 
-$D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
+$D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = 1$ means that the $k$-th element of $\mathbf{W}$ is identical to the $k$-th element of $\mathbf{W_0}$.
+In the opposite, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
 This measure is bounded between 1 and 0 and is an invariant with respect to scale.
@@
 
-An illustrative example is shown in Figure \ref{fig:W_TI_NMF}.\\
+An illustrative example is displayed in Figure \ref{fig:W_TI_NMF}.\\
 
@@
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
-Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture.
+Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture, see Figure \ref{fig:example_simScene}.
 The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
@@
 The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling bird), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
-In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the other sound classes.
-The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class.
-To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).
+In each file, traffic component is present as the sum of the background and event traffic sounds while the \textit{interfering} sound class is the sound sources not related to it.
+To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the interfering class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).
 
@@
 with the \textit{Traffic Interference Ratio} $TIR \in$ [-12 -6 0 6  12] dB.
-When $TIR < 0$ dB, the traffic component is then less present than the interfering class.
+When $TIR < 0$ dB, the traffic component is less present than the interfering class.
 In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
@@
 One obtain for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements.
-Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering is applied to reduce the number of spectra to $K \in \left[ 25, 50, 100, 200\right]$.
-The $K$ centroids are then the elements of the dictionary.
+Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering algorithm is applied to reduce the number of spectra to $K \in \left[ 25,~50,~100,~200\right]$.
+The $K$ centroids are then the elements considered in the dictionary.
 A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) in order to build a dictionary with the spectral envelope of each audio sample.
 In this case, 53 spectra are obtained.
-The $K$-means clustering algorithm is then reduced to $K \in \left[25 50 \right]$.
+The $K$-means clustering algorithm is then reduced to $K \in \left[25~50 \right]$.
 
@@
 This last step allows us to reduce the dimensionality while preserving a rich description of the spectral content.
-Experimental validation consistently showed that considering octave bands do not impact the performance for the two estimators studied in this paper.
+Experimental validation consistently showed that considering octave bands do not impact the performance of the estimator studied in this paper.
 
@@
 TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-%\ml{ca c'est pas genial, il aurait mieux vallu utiliser un coefficient de pondération pour que W n'avance pas trop vite...}
-
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
@@
 
-After having performed NMF on one scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
+For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
+For one sound scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
 
@@
 
-where $p_{rms}$ is the effective pressure deducted from $\left[\mathbf{WH} \right]_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
+where $p_{rms}$ is the effective pressure deducted from the remaining spectrogram ($\mathbf{V}_{f_c}$ or $\left[\mathbf{WH} \right]_{traffic}$) and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
 For each set of unique modalities of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained.
@@
 The lowest average error is obtained for TI-NMF, 2.19 $\pm$ 2.18, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
-On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation (2.32 $\pm$ 1.26).
+On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 $\pm$ 1.26).
 
@@
 
-However, the relatively high degree of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
 In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem-NMF is free to include traffic components in $\mathbf{W_r}$.
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
-In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horns have disappeared for the benefit of low frequency content.
+In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of the alert sounds are lower for the benefit of low frequency content.
 Consequently, the traffic sound level estimation is then underestimated as in Figure \ref{fig:lp_alert_12}.
@@
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR$ = [-12, 12] dB.
+In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ [-12, 12] dB.
 
@@
 
-For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher what allows retaining more elements as traffic components and then decreases the error (< 1 dB).
+For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decreases the error (< 1 dB).
 The kept elements are then more suited to the scenes than a fixed dictionary.
@@
 For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30.
-In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease to 4.46 $\pm$ 1.66.
-In order to generalize this method and having no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 and is the one that best balanced these opposite cases.
+In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 $\pm$ 1.66.
+In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 and is the one that best balanced these opposite cases.
 

### 2018-02-05 paper/gloaguenNmfTraffic.tex
distanceCosLinDisplay improved

@@
     \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
-    \caption{Block diagram of the NMF experience with the dictionary design composed from a second sound database}
+    \caption{Specific block diagram of the NMF estimator with the dictionary design composed from a second sound database}
     \label{fig:bloc_nmf}
@@
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities & \begin{tabular}[c]{@{}l@{}}number  \\ of level\end{tabular}\\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
 \end{tabularx}
@@
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & modalities & \begin{tabular}[c]{@{}l@{}}number \\ of modality\end{tabular}\\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} &\begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
 \end{tabularx}
@@
 In dashed lines, a 1 second $w_t$ window  (\subref{fig:specW}).
-With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$ while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW})}
+With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$, while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW})}
     \label{fig:spec_elementW}

### 2018-02-05 paper/gloaguenNmfTraffic.tex
Y_alert improvement

@@
 For TI-NMF, 24000 combinations are computed.
-In all, 25380 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors
+In all, 25380 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.
 
@@
 where $p_{rms}$ is the effective pressure deducted from the remaining spectrogram ($\mathbf{V}_{f_c}$ or $\left[\mathbf{WH} \right]_{traffic}$) and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
-For each set of unique modalities of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained.
+For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
 
@@
 \toprule
-\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
+\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$} (dB) \\ \midrule
 filter & 20  &  &  &  &  & 4.69 ($\pm$ 4.52) \\

### 2018-02-07 paper/gloaguenNmfTraffic.tex
updates in font size of figures

@@
 \author{
-    Jean-Rémy GLOAGUEN\\
+    Jean-Rémy Gloaguen\\
     Arnaud Can\\
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf_2.pdf}
+\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
 \caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c))}
@@
 
-with the \textit{Traffic Interference Ratio} $TIR \in$ [-12 -6 0 6  12] dB.
+with the \textit{Traffic Interference Ratio} $TIR \in$ $\lbrace$12, -6, 0, 6, 12$\rbrace$ dB.
 When $TIR < 0$ dB, the traffic component is less present than the interfering class.
@@
 
-The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ([-12 -6 0 6 12] dB).
+The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ($\lbrace$-12, -6, 0, 6, 12$\rbrace$ dB).
 The spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap, see Figure \ref{fig:bloc_experiment}.
@@
 Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter.
-It depends only on the cut-off frequencies $f_c \in $ [500 1k 2k 5k 10k 20k] Hz.
+It depends only on the cut-off frequencies $f_c \in  \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz.
 The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
@@
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modalities}\end{tabular}\\ \toprule
 \end{tabularx}
@@
 	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} &\begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
+    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} &\begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modalities}\end{tabular}\\ \toprule
 \end{tabularx}
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
-   $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 40\\
+   threshold $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 41\\
    \bottomrule
@@
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
-Then time/frequency windows of $F \times w_t $ dimensions are applied without overlapping on the spectrogram in order to consider several spectra for each audio file where $w_t \in [0.5~1]$ second.
+Then time/frequency windows of $F \times w_t $ dimensions are applied without overlapping on the spectrogram in order to consider several spectra for each audio file where $w_t \in \lbrace$0.5, 1$\rbrace$ second.
 In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectrum a of $F \times 1$ dimension.
 With this size of window, it is possible to obtain the characteristic pitches of the different audio samples.
-One obtain for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements.
-Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering algorithm is applied to reduce the number of spectra to $K \in \left[ 25,~50,~100,~200\right]$.
+One obtains for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements.
+Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering algorithm is applied to reduce the number of spectra to $K \in \lbrace$25, 50, 100, 200$\rbrace$.
 The $K$ centroids are then the elements considered in the dictionary.
@@
 In this case, 53 spectra are obtained.
-The $K$-means clustering algorithm is then reduced to $K \in \left[25~50 \right]$.
+The $K$-means clustering algorithm is then reduced to $K \in \lbrace$25, 50$\rbrace$.
 
@@
 
-Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed.
-For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ [0.5 1] s whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning}.
-For TI-NMF, 24000 combinations are computed.
-In all, 25380 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.
+Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
+For TI-NMF, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
+In all, 25980 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.
 
@@
 Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
-$f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained.
+$f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained (2.89 ($\pm$ 2.84) dB).
 It is then considered as the baseline to compare the performances of NMF.
@@
 
-    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB(\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
+    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
     \label{fig:dictionaryExtraction}
@@
 Sem-NMF and TI-NMF have better results.
-The lowest average error is obtained for TI-NMF, 2.19 $\pm$ 2.18, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
-On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 $\pm$ 1.26).
+The lowest average error is obtained for TI-NMF, 2.19 ($\pm$ 2.18) dB, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
+On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 ($\pm$ 1.26) dB).
 
@@
         \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
-        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
+        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz}
         \label{fig:TIR_class_filter}
@@
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ [-12, 12] dB.
+In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ $\lbrace$-12, 12$\rbrace$ dB.
 
@@
 For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30.
-In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 $\pm$ 1.66.
-In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 and is the one that best balanced these opposite cases.
+In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 ($\pm$ 1.66) dB.
+In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 as it is the one that best balanced these opposite cases.
 

### 2018-02-07 paper/gloaguenNmfTraffic.tex
Minor corrections and dist-12_12 reframed

@@
 \begin{equation}
-\tilde{L}_{p,traffic} = 20\log_{10}\left(\frac{p_{rms}}{p_0}\right)
+\tilde{L}_{p,traffic} = 20 \times \log_{10}\left(\frac{p_{rms}}{p_0}\right)
 \end{equation}
 
-where $p_{rms}$ is the effective pressure deducted from the remaining spectrogram ($\mathbf{V}_{f_c}$ or $\left[\mathbf{WH} \right]_{traffic}$) and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
+where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
 For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
@@
     \end{subfigure}
-    \caption{$MAE$ error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI}))}
+    \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI}))}
     \label{fig:TIR_bar}
@@
 
-However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB.
+However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR \in \lbrace$6, 12$\rbrace$ dB.
 In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem-NMF is free to include traffic components in $\mathbf{W_r}$.

### 2018-02-08 paper/gloaguenNmfTraffic.tex
funding ack
two new figures

@@
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
-We thus introduce another NMF scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).
+We thus introduce another NMF scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).
 
@@
 \includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
-\caption{Block diagram of the general source separation model}
+\caption{Block diagram of the general source separation model.}
 \label{fig:diagram}
@@
 \includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
-\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c))}
+\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
 \label{fig:example_NMF}
@@
 \includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
-\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
+\caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
 \label{fig:W_TI_NMF}
@@
     \centering
-    \begin{subfigure}[t]{0.47\linewidth}
-        \centering
-       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-spectrum.png}
-       \caption{}
-       \label{fig:simScene_spec}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.47\linewidth}
-        \centering
-       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-timeDomain.png}
-       \caption{}
-       \label{fig:simScene_time}
-    \end{subfigure}
-    \caption{Example of a simple scene created with \textit{SimScene} software (the spectrogram (\subref{fig:simScene_spec}), the time domain \subref{fig:simScene_time})) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue)}
+       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene}
+    \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in blue, car passage in red and whistling bird in cyan).}
     \label{fig:example_simScene}
@@
     \includegraphics[width=\linewidth]{figures/bloc_diagram_estimator.pdf}
-    \caption{Block diagram of the experience with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF}
+    \caption{Block diagram of the experience with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF.}
     \label{fig:bloc_experiment}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
-    \caption{Specific block diagram of the NMF estimator with the dictionary design composed from a second sound database}
+    \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN_2.pdf}
+    \caption{Specific block diagram of the NMF estimator with the dictionary design composed from a second sound database.}
     \label{fig:bloc_nmf}
@@
 
-\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator}
+\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
 \label{tab:experimental_factorsFilter}
@@
 \end{tabularx}
-\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
+\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
 \label{tab:experimental_factorsNMF}
@@
     \end{subfigure}
-    \caption{Example of the dictionary building with a 3 second extract of a car passage.
+    \caption{Dictionary building on a 3 second extract of a car passage.
 In dashed lines, a 1 second $w_t$ window  (\subref{fig:specW}).
-With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$, while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW})}
+With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$, while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW}).}
     \label{fig:spec_elementW}
@@
 \end{tabular}
-\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
+\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
 \label{tab:results_TIR}
@@
 
-    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
+    \caption{Estimated global sound levels of the traffic by the frequency low-pass filter with $f_c$ = 500 Hz for the 25 scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12}).}
     \label{fig:dictionaryExtraction}
@@
         \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz}
+        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.}
         \label{fig:TIR_class_filter}
@@
         \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.}
                 \label{fig:TIR_class_sup}
@@
       	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2}
+        \caption{$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.}
                 \label{fig:TIR_class_semi}
@@
         \includegraphics[width=\linewidth]{figures/TI_bar}
-        \caption{$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42}
+        \caption{$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.}
         \label{fig:TIR_class_TI}
     \end{subfigure}
-    \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI}))}
+    \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI})).}
     \label{fig:TIR_bar}
@@
 
-    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB(\subref{fig:lp_alert_-12}) and $TIR$ = 12 dB (\subref{fig:lp_alert_12})}
+    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (\subref{fig:lp_alert_-12}) and $TIR$ = 12 dB (\subref{fig:lp_alert_12}).
+In Figure (\subref{fig:lp_alert_12}), Sup-NMF models the interfering class increasing the estimated error while in Figure (\subref{fig:lp_alert_12}), this behavior disappear.}
     \label{fig:lp_alert}
@@
     \end{subfigure}
-    \caption{Comparison between the sum of the 2 elements of $\mathbf{W_r}$ with the spectra of the interfering and traffic classes for an \textit{alert} scene for $TIR$ = -12 dB (\subref{fig:Y_alert-12}) and $TIR$ = 12 dB (\subref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s (normalized amplitudes)}
+    \caption{Comparison between the sum of the 2 elements of $\mathbf{W_r}$ with the spectra of the interfering and traffic classes for an \textit{alert} scene for $TIR$ = -12 dB  (\subref{fig:Y_alert-12}) and $TIR$ = 12 dB (\subref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s (normalized amplitudes).}
 \end{figure}
@@
     \includegraphics[width=.8\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $K$ = 200 and $w_t$ = 0.5 s)}
+    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $K$ = 200 and $w_t$ = 0.5 s).}
     \label{fig:dist_-12_12}
@@
 
+\vspace{6pt}
+%\acknowledgments{The authors would like to acknowledge partial funding from the ANR under project reference ANR-16-CE22-0012.}
+
 \footnotesize

### 2018-02-09 paper/gloaguenJASMarticle.tex
new version of the article with svjour3 template

--- 2018-02-08 paper/gloaguenNmfTraffic.tex
@@
-\documentclass[twocolumn,a4paper,10pt]{article}
+\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
+%
+\usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
+\usepackage{layout}
 
-\input{packageEN}
-\input{packageStyle}
+%\usepackage[left,modulo]{lineno}	% numéro de ligne
+%\linenumbers
 
-\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
+%%%%%% SYMBOLES %%%%%
+\usepackage{tipa}	% pour avoir l'accent concave
+\usepackage{lmodern}	% pour les guillemets
+\usepackage{nth}	% pour ^th près des chiffres
+
+%%%%%% EQUATION %%%%%%
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{fancybox}
+\usepackage{xfrac}	% fraction de type "1/4"
+\usepackage{cases}	% système équation
+\usepackage[overload]{empheq}
+\usepackage{bm}		% pour mettre en gras .
+\usepackage{units} 	% x/y barre latérale pour les fractions
+
+%%%%%% FIGURE %%%%%%
+\usepackage{graphicx}	% insérer des graphiques
+\usepackage{subfig}	% utiliser subfigure
+\usepackage{float}	% utiliser H dans les figures
+
+%%%%%% TABLEAUX %%%%%%
+\usepackage{array,multirow,makecell}
+\usepackage[table,xcdraw]{xcolor} % pour avoir des lignes colorées dans les tableau
+\usepackage{hhline}	% pour les lignes horizontales
+\usepackage{tabularx} % permet itemize dans les cellules
+\usepackage{booktabs}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+%%%%%%%%%%%%%%%%%%%%%
+\usepackage{url}	% gérer les adresses www.
+
+\usepackage[colorlinks,citecolor=red,urlcolor=blue,linkcolor=blue]{hyperref}  % ps2pdf car je compile via Latex -> dvips -> ps2pdf
+
+% \usepackage{natbib}
+
+\journalname{EURASIP Journal on Audio, Speech and Music Processing}
+
+\begin{document}
 
@@
 
-\author{
-    Jean-Rémy Gloaguen\\
-    Arnaud Can\\
-    LAE\\
-    Ifsttar\\
-    jean-remy.gloaguen@ifsttar.fr
+\author{Jean-R\'emy Gloaguen         \and
+		Mathieu Lagrange \and
+		Arnaud Can \and
+		Jean-Fran\c cois Petiot.}
+
+\institute{J.-R. Gloaguen \at
+              Ifsttar Centre de Nantes\\
+              UMRAE\\
+              All\'ee des Ponts et Chauss\'ees\\
+              44344 Bouguenais\\
+              \email{jean-remy.gloaguen@ifsttar.fr}           %  \\
            \and
-    Mathieu Lagrange\\
-	Jean-François Petiot \\
-    LS2N, CNRS\\
-    \'Ecole Centrale de  Nantes\\
+           M. Lagrange \at
+              LS2N
+              1 rue de la Noe\\
+              44321 Nantes\\
+              \email{mathieu.lagrange@cnrs.fr}
 }
-\date{}
-\begin{document}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 
-\section*{Abstract}
-
+\begin{abstract}
 The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring of the acoustic quality of urban areas.
@@
 
-\section{Introduction}
+\keywords{non-negative matrix factorization \and road traffic sound level \and urban sound environment}
+\end{abstract}
+
+\section{Introduction} \label{part:intro}
 
@@
 In the opposite, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
-This measure is bounded between 1 and 0 and is an invariant with respect to scale.
-The $k$ value of $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ are next sorted in descending order.
+This measure has the advantage to be bounded between 1 and 0 and to be invariant with respect to scale.
+The $k$ values of $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ are next sorted in descending order.
 The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are selected by a \textit{hard thresholding} method.
@@
 \begin{equation}
-\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_{k} \right) > t.
+\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_{k} \right) > t
 \end{equation}
 
+where $t$ is a fixed threshold $\in [$ 0;1 $]$.
 An illustrative example is displayed in Figure \ref{fig:W_TI_NMF}.\\
@@
 These sounds are in .wav format sampled at 44.1 kHz.
-The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
+The sound class \textit{passing car} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
 The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
@@
 The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
-Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
+Furthermore, an audio file of each sound class present in the scene can be generated that allows us to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
 
@@
        \includegraphics[width=.8\linewidth]{./figures/exampleSimScene}
-    \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in blue, car passage in red and whistling bird in cyan).}
+    \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, car passage in red and whistling bird in cyan).}
     \label{fig:example_simScene}
@@
 
-This database allows us to create a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
+This database allows the creation of a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
 A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds.
@@
 
-The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:nmf}.
+The second estimator is the proposed scheme, based on the three NMF approaches presented in Section \ref{part:nmf}.
 Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}), each experimental factor having multiples modalities.
@@
 \centering
-
+\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
@@
 \end{tabularx}
-
-\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
 \label{tab:experimental_factorsFilter}
@@
 \centering
-
+\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}@{}C{2cm}}
@@
 \end{tabularx}
-\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
+
 \label{tab:experimental_factorsNMF}
@@
 The $K$ centroids are then the elements considered in the dictionary.
-A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) in order to build a dictionary with the spectral envelope of each audio sample.
+A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) to build a dictionary with the spectral envelope of each audio sample.
 In this case, 53 spectra are obtained.
@@
 
-\begin{figure}[t!]
+\begin{figure}[htp]
   \centering
-    \begin{subfigure}[t]{0.47\linewidth}
-        \centering
-       \includegraphics[height=4.5cm]{figures/dictionary3.pdf}
-       \caption{}
-        \label{fig:specW}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.47\linewidth}
-        \centering
-       \includegraphics[height=4.5cm]{figures/dictionary4.pdf}
-        \caption{}
-        \label{fig:ElementW}
-    \end{subfigure}
+  \subfloat[]{\label{fig:specW}\includegraphics[height=4cm]{figures/dictionary3.pdf}}
+  \hspace{5pt}
+  \subfloat[]{\label{fig:ElementW}\includegraphics[height=4cm]{figures/dictionary4.pdf}}
   \caption{Dictionary building on a 3 second extract of a car passage.
-In dashed lines, a 1 second $w_t$ window  (\subref{fig:specW}).
-With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$, while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\subref{fig:ElementW}).}
+In dashed lines, a 1 second $w_t$ window  (\ref{fig:specW}).
+With $w_t$ = 1 second, 3 spectra are then generated and included in $\mathbf{W}$, while for $w_t$ = $all$, the audio file is reduced into 1 spectral vector (\ref{fig:ElementW}).}
   \label{fig:spec_elementW}
@@
 TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+The spectrogram $\mathbf{V}$, as the dictionary $\mathbf{W}$, is expressed with third octave bands ($F$ = 29).
 This coarser method allows the reduction of the matrix size and decreases the computation time.
@@
 For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
-For TI-NMF, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
-In all, 25980 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.
-
-For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
-For one sound scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
-
-\begin{equation}
-\tilde{L}_{p,traffic} = 20 \times \log_{10}\left(\frac{p_{rms}}{p_0}\right)
-\end{equation}
-
-where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
-For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
+For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
+In all, 25980 settings are performed.
 
@@
 \centering
+\caption{Best results for all the scenes according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
 \begin{tabular}{@{}ccccccc@{}}
@@
 \end{tabular}
-\caption{Best results for all the scenes according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
 \label{tab:results}
@@
 \centering
+\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
 \begin{tabular}{@{}cccccc@{}}
@@
 \end{tabular}
-\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
 \label{tab:results_TIR}
 \end{table*}
+
+For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
+For one sound scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
+
+\begin{equation}
+\tilde{L}_{p,traffic} = 20 \times \log_{10}\left(\frac{p_{rms}}{p_0}\right)
+\end{equation}
+
+where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
+For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
 
@@
   \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}
-        \caption{}
-        \label{fig:alert_-12}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}
-        \caption{}
-        \label{fig:alert_12}
-    \end{subfigure}
-
-    \caption{Estimated global sound levels of the traffic by the frequency low-pass filter with $f_c$ = 500 Hz for the 25 scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12}).}
+  \subfloat[]{\label{fig:alert_-12}\includegraphics[width=.4\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}}
+  \hspace{5pt}
+  \subfloat[]{\label{fig:alert_12}\includegraphics[width=.4\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}}
+  \caption{Estimated global sound levels of the traffic by the frequency low-pass filter with $f_c$ = 500 Hz for the 25 scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\ref{fig:alert_-12}) and at $TIR$ = 12 dB (\ref{fig:alert_12}).}
   \label{fig:dictionaryExtraction}
@@
   \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.}
-        \label{fig:TIR_class_filter}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.}
-                \label{fig:TIR_class_sup}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-      	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.}
-                \label{fig:TIR_class_semi}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/TI_bar}
-        \caption{$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.}
-        \label{fig:TIR_class_TI}
-    \end{subfigure}
-    \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI})).}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
 	\label{fig:TIR_bar}
@@
 \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the  sound scene.
-In order to expose this issue, the case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}.
-Here when the \textit{alert} sub-class sounds, some traffic elements of $\mathbf{W}$ are activated.
-The \textit{alert} sub-class is then considered as traffic component generating a wrong estimation of the sound level, see Figure \ref{fig:lp_alert_-12}.
-This behavior disappears when the traffic component becomes predominant to the interfering class as in Figure \ref{fig:lp_alert_12}.
 Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
@@
 Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the interfering class in low $TIR$.
-It brings a significant decrease of the errors for low $TIR$ as it can be seen in Figure \ref{fig:TIR_class_semi}.
-The content of $\mathbf{W_r}$, in the case of an \textit{alert} sub-class is displayed in Figure \ref{fig:Y_alert-12}.
-The interfering class is easily integrated.
-The first element is mainly composed of high frequency bands which correspond to the car horns of the scene.
-This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the elements of the dictionary dedicated to the traffic are no longer activated when the car horn is active.
-
-\begin{figure*}[t]
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/NMF_Lp_alert-12.pdf}
-        \caption{}
-        \label{fig:lp_alert_-12}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/NMF_Lp_alert12.pdf}
-        \caption{}
-        \label{fig:lp_alert_12}
-    \end{subfigure}
-
-    \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for supervised NMF ($\beta$ = 2, $K$ = 50, $w_t$ = 0.5 s), semi-supervised ($\beta$ = 2, $K$ = 100, $w_t$ = 0.5 s) and thresholded initialized ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (\subref{fig:lp_alert_-12}) and $TIR$ = 12 dB (\subref{fig:lp_alert_12}).
-In Figure (\subref{fig:lp_alert_12}), Sup-NMF models the interfering class increasing the estimated error while in Figure (\subref{fig:lp_alert_12}), this behavior disappear.}
-    \label{fig:lp_alert}
-\end{figure*}
-
+It brings a significant decrease of the errors for low $TIR$.
 However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR \in \lbrace$6, 12$\rbrace$ dB.
@@
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
-In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of the alert sounds are lower for the benefit of low frequency content.
-Consequently, the traffic sound level estimation is then underestimated as in Figure \ref{fig:lp_alert_12}.
-\\
 
-\begin{figure}
+\begin{figure}[t]
   \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_-12.pdf}
-        \caption{}
-        \label{fig:Y_alert-12}
-    \end{subfigure}
-    \vfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_12.pdf}
-        \caption{}
-		\label{fig:Y_alert_12}
-    \end{subfigure}
-    \caption{Comparison between the sum of the 2 elements of $\mathbf{W_r}$ with the spectra of the interfering and traffic classes for an \textit{alert} scene for $TIR$ = -12 dB  (\subref{fig:Y_alert-12}) and $TIR$ = 12 dB (\subref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s (normalized amplitudes).}
+	\includegraphics[width=0.45\textwidth]{figures/NMF_Lp_alert.pdf}
+ \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+	\label{fig:lp_alert}
 \end{figure}
@@
 Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}).
-Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
+Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments, see Figure \ref{fig:lp_alert}.
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ $\lbrace$-12, 12$\rbrace$ dB.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $K$ = 200 and $w_t$ = 0.5 s).}
-    \label{fig:dist_-12_12}
-\end{figure}
-
 For low $TIR$, as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
@@
 
-\vspace{6pt}
-%\acknowledgments{The authors would like to acknowledge partial funding from the ANR under project reference ANR-16-CE22-0012.}
+%\vspace{6pt}
 
-\footnotesize
-\bibliographystyle{unsrt}
-\bibliography{bibliographie_sensors}
+\begin{acknowledgements}
+The authors would like to acknowledge partial funding from the ANR under project reference ANR-16-CE22-0012.
+\end{acknowledgements}
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+\bibliographystyle{spphys}       % APS-like style for physics
+\bibliography{bibliographie_jasm}   % name your BibTeX data base
 

### 2018-02-09 paper/gloaguenJASMarticle.tex
correction images

@@
 
-\usepackage[colorlinks,citecolor=red,urlcolor=blue,linkcolor=blue]{hyperref}  % ps2pdf car je compile via Latex -> dvips -> ps2pdf
+\usepackage[colorlinks,citecolor=black,urlcolor=black,linkcolor=black]{hyperref}  % ps2pdf car je compile via Latex -> dvips -> ps2pdf
 
@@
     \centering
-       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene}
+       \includegraphics[width=.85\linewidth]{./figures/exampleSimScene}
     \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, car passage in red and whistling bird in cyan).}
@@
 For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then does not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 dB and are overestimated, see Figure \ref{fig:alert_-12}.
-In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels, see Figure \ref{fig:alert_12}.
+The errors are then nearly all superior to 4 dB and are overestimated.
+In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels.
 The 500 Hz filter finds a balance between what is put aside in low $TIR$ and what it is remained in high $TIR$.
 \\
-
-\begin{figure*}[t]
-  \centering
-  \subfloat[]{\label{fig:alert_-12}\includegraphics[width=.4\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}}
-  \hspace{5pt}
-  \subfloat[]{\label{fig:alert_12}\includegraphics[width=.4\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}}
-  \caption{Estimated global sound levels of the traffic by the frequency low-pass filter with $f_c$ = 500 Hz for the 25 scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\ref{fig:alert_-12}) and at $TIR$ = 12 dB (\ref{fig:alert_12}).}
-  \label{fig:dictionaryExtraction}
-\end{figure*}
 
@@
 	\includegraphics[width=0.45\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+ \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
 	\label{fig:lp_alert}

### 2018-02-09 paper/gloaguenJASMarticle.tex
move tab and figures
bibliography issue resolved ?

@@
 
-\subsubsection{Experimental factors of NMF}
-
-Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
-TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-The spectrogram $\mathbf{V}$, as the dictionary $\mathbf{W}$, is expressed with third octave bands ($F$ = 29).
-This coarser method allows the reduction of the matrix size and decreases the computation time.
-But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
-For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
-Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
-
-Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
-For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
-For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
-In all, 25980 settings are performed.
-
 \begin{table*}[t]
@@
 
+\subsubsection{Experimental factors of NMF}
+
+Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
+TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+The spectrogram $\mathbf{V}$, as the dictionary $\mathbf{W}$, is expressed with third octave bands ($F$ = 29).
+This coarser method allows the reduction of the matrix size and decreases the computation time.
+But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
+Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
+
+Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
+For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
+In all, 25980 settings are performed.
+
 For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
@@
 
+\begin{figure}[t]
+  \centering
+	\includegraphics[width=0.45\textwidth]{figures/NMF_Lp_alert.pdf}
+ \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+	\label{fig:lp_alert}
+\end{figure}
+
+\begin{figure*}[t]
+  \centering
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
+	\label{fig:TIR_bar}
+\end{figure*}
 \section{Results}\label{part:results}
@@
 
-\begin{figure*}[t]
-  \centering
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
-  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
-	\label{fig:TIR_bar}
-\end{figure*}
-
 According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
@@
 
-\begin{figure}[t]
-  \centering
-	\includegraphics[width=0.45\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
-	\label{fig:lp_alert}
-\end{figure}
-
 Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}).
-Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments, see Figure \ref{fig:lp_alert}.
+Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
@@
 The error for these $TIR$ is then due to the thresholding which put aside some elements that are nevertheless related to the traffic component.
+In Figure \ref{fig:lp_alert}, the 1 second equivalent sound pressure level of an alert scene is displayed for $TIR \in \lbrace -12,12 \rbrace$ dB for the 500 Hz low pass filter and for TI-NMF.
+In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF do not activated when the alarm class sounds contrary to the low-pass filter.
+
 With a low threshold, it is possible to decrease the error.
@@
 %\bibliographystyle{spbasic}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-\bibliographystyle{spphys}       % APS-like style for physics
+\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
 \bibliography{bibliographie_jasm}   % name your BibTeX data base

### 2018-02-09 paper/gloaguenJASMarticle.tex
images updates

@@
 
-With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
+With the introduction of the European Directive 2002/\-EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
 These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels.
@@
 It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{figures/bloc_diagram_source_separation.pdf}
+\caption{Block diagram of the general source separation model.}
+\label{fig:diagram}
+\end{figure}
 
@@
 By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be resolved on the time/frequency plane.
+
 Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}.
@@
 The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
-\caption{Block diagram of the general source separation model.}
-\label{fig:diagram}
-\end{figure*}
 
@@
 \includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
-\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
+\caption{NMF decomposition of a urban sound mixture comprising 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
 \label{fig:example_NMF}
@@
 
-\begin{figure}[hbtp]
+\begin{figure}[t]
 \centering
@@
     \centering
-       \includegraphics[width=.85\linewidth]{./figures/exampleSimScene}
+       \includegraphics[width=.9\linewidth]{./figures/exampleSimScene}
     \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, car passage in red and whistling bird in cyan).}
@@
   \centering
-	\includegraphics[width=0.45\textwidth]{figures/NMF_Lp_alert.pdf}
+	\includegraphics[width=\linewidth]{figures/NMF_Lp_alert.pdf}
  \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}

### 2018-02-12 paper/gloaguenJASMarticle.tex
bibliographie checked

@@
 These sounds are in .wav format sampled at 44.1 kHz.
-The sound class \textit{passing car} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
-The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
+The sound class \textit{passing car} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratios.
+The other audio files have been found online (\textit{freesound.org}) and within the \textit{salamon2014dataset} database \cite{salamon_dataset_nodate}.
 Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture, see Figure \ref{fig:example_simScene}.

### 2018-02-13 paper/gloaguenNmfTraffic.tex
Add the decleration part

@@
-\documentclass[twocolumn,a4paper,10pt]{article}
+\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
+%
+\usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
+\usepackage{layout}
 
-\input{packageEN}
-\input{packageStyle}
+%\usepackage[left,modulo]{lineno}	% numéro de ligne
+%\linenumbers
 
-\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
+%%%%%% SYMBOLES %%%%%
+\usepackage{tipa}	% pour avoir l'accent concave
+\usepackage{lmodern}	% pour les guillemets
+\usepackage{nth}	% pour ^th près des chiffres
+\usepackage{nomencl}
+\makenomenclature
+
+\renewcommand{\nomname}{List of Abbreviations}
+
+%%%%%% EQUATION %%%%%%
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{fancybox}
+\usepackage{xfrac}	% fraction de type "1/4"
+\usepackage{cases}	% système équation
+\usepackage[overload]{empheq}
+\usepackage{bm}		% pour mettre en gras .
+\usepackage{units} 	% x/y barre latérale pour les fractions
+
+%%%%%% FIGURE %%%%%%
+\usepackage{graphicx}	% insérer des graphiques
+\usepackage{subfig}	% utiliser subfigure
+\usepackage{float}	% utiliser H dans les figures
+
+%%%%%% TABLEAUX %%%%%%
+\usepackage{array,multirow,makecell}
+\usepackage[table,xcdraw]{xcolor} % pour avoir des lignes colorées dans les tableau
+\usepackage{hhline}	% pour les lignes horizontales
+\usepackage{tabularx} % permet itemize dans les cellules
+\usepackage{booktabs}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+%%%%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}	% gérer les adresses www.
+
+\usepackage[colorlinks,citecolor=black,urlcolor=black,linkcolor=black]{hyperref}  % ps2pdf car je compile via Latex -> dvips -> ps2pdf
+
+% \usepackage{natbib}
+\renewcommand{\baselinestretch}{1.025}
+
+\journalname{EURASIP Journal on Audio, Speech and Music Processing}
+
+\begin{document}
 
@@
 
-\author{
-    Jean-Rémy Gloaguen\\
-    Arnaud Can\\
-    LAE\\
-    Ifsttar\\
-    jean-remy.gloaguen@ifsttar.fr
+\author{Jean-R\'emy Gloaguen         \and
+		Mathieu Lagrange \and
+		Arnaud Can \and
+		Jean-Fran\c cois Petiot.}
+
+\institute{J.-R. Gloaguen \at
+              Ifsttar Centre de Nantes UMRAE\\
+              All\'ee des Ponts et Chauss\'ees 44344 Bouguenais\\
+              \email{jean-remy.gloaguen@ifsttar.fr}
 \and
-    Mathieu Lagrange\\
-	Jean-François Petiot \\
-    LS2N, CNRS\\
-    \'Ecole Centrale de  Nantes\\
+			M. Lagrange \at
+              LS2N
+              1 rue de la Noe 44321 Nantes\\
+              \email{mathieu.lagrange@cnrs.fr}
+\and
+			A.Can \at
+              \email{arnaud.can@ifsttar.fr}           \and
+			J.-F. Petiot \at
+              \email{jean-francois.petiot@ls2n.fr}
 }
-\date{}
-\begin{document}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 
-\section{Introduction}
+\keywords{non-negative matrix factorization \and road traffic sound level \and urban sound environment}
+\end{abstract}
 
-With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
+\section{Introduction} \label{part:intro}
+
+With the introduction of the European Directive 2002/\-EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps.
 These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels.
@@
 \begin{equation}
-\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_{k} \right) > t.
+\mathbf{W}_k \in \mathbf{W}_{k,traffic} \quad \text{iff} \quad D\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_{k} \right) > t
 \end{equation}
 
+where $t$ is a fixed threshold $\in [$ 0;1 $]$.
 An illustrative example is displayed in Figure \ref{fig:W_TI_NMF}.\\
 
-\begin{figure}[hbtp]
+\begin{figure}[t]
 \centering
@@
 These sounds are in .wav format sampled at 44.1 kHz.
-The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio.
-The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}.
+The sound class \textit{passing car} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratios.
+The other audio files have been found online (\textit{freesound.org}) and within the \textit{salamon2014dataset} database \cite{salamon_dataset_nodate}.
 Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture, see Figure \ref{fig:example_simScene}.
 The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
-Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
+Furthermore, an audio file of each sound class present in the scene can be generated that allows us to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
 
@@
     \centering
-       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene}
-    \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in blue, car passage in red and whistling bird in cyan).}
+       \includegraphics[width=.9\linewidth]{./figures/exampleSimScene}
+    \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, car passage in red and whistling bird in cyan).}
     \label{fig:example_simScene}
@@
 The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values).
-The corpora are available for download at: \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}
 
@@
 
-\subsubsection{Experimental factors of NMF}
-
-Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
-TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
-This coarser method allows the reduction of the matrix size and decreases the computation time.
-But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
-For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
-Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
-
-Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
-For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
-For TI-NMF, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
-In all, 25980 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.
-
-For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
-For one sound scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
-
-\begin{equation}
-\tilde{L}_{p,traffic} = 20 \times \log_{10}\left(\frac{p_{rms}}{p_0}\right)
-\end{equation}
-
-where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
-For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
-
 \begin{table*}[t]
 \centering
+\caption{Best results for all the scenes according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
 \begin{tabular}{@{}ccccccc@{}}
@@
 \end{tabular}
-\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
 \label{tab:results_TIR}
 \end{table*}
+
+\subsubsection{Experimental factors of NMF}
+
+Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
+TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+The spectrogram $\mathbf{V}$, as the dictionary $\mathbf{W}$, is expressed with third octave bands ($F$ = 29).
+This coarser method allows the reduction of the matrix size and decreases the computation time.
+But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
+Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
+
+Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
+For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
+In all, 25980 settings are performed.
+
+For each setting, the estimator (frequency low-pass filter or NMF) is performed on the $M$ scenes of a sub-class.
+For one sound scene, the average traffic sound level, $\tilde{L}_{p,traffic}$, of the entire scene is calculated,
+
+\begin{equation}
+\tilde{L}_{p,traffic} = 20 \times \log_{10}\left(\frac{p_{rms}}{p_0}\right)
+\end{equation}
+
+where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
+For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
 
@@
 
+\begin{figure}[t]
+  \centering
+	\includegraphics[width=\linewidth]{figures/NMF_Lp_alert.pdf}
+ \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+	\label{fig:lp_alert}
+\end{figure}
+
+\begin{figure*}[t]
+  \centering
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
+	\label{fig:TIR_bar}
+\end{figure*}
 \section{Results}\label{part:results}
@@
 For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then does not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 dB and are overestimated, see Figure \ref{fig:alert_-12}.
-In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels, see Figure \ref{fig:alert_12}.
+The errors are then nearly all superior to 4 dB and are overestimated.
+In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels.
 The 500 Hz filter finds a balance between what is put aside in low $TIR$ and what it is remained in high $TIR$.
 \\
-
-\begin{figure*}[t]
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}
-        \caption{}
-        \label{fig:alert_-12}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}
-        \caption{}
-        \label{fig:alert_12}
-    \end{subfigure}
-
-    \caption{Estimated global sound levels of the traffic by the frequency low-pass filter with $f_c$ = 500 Hz for the 25 scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12}).}
-    \label{fig:dictionaryExtraction}
-\end{figure*}
 
@@
 
-\begin{figure*}[t]
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.}
-        \label{fig:TIR_class_filter}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.}
-                \label{fig:TIR_class_sup}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-      	\includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
-        \caption{$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.}
-                \label{fig:TIR_class_semi}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/TI_bar}
-        \caption{$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.}
-        \label{fig:TIR_class_TI}
-    \end{subfigure}
-    \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\subref{fig:TIR_class_filter}) and each method (Sup-NMF (\subref{fig:TIR_class_sup}), Sem-NMF (\subref{fig:TIR_class_semi}) and TI-NMF (\subref{fig:TIR_class_TI})).}
-    \label{fig:TIR_bar}
-\end{figure*}
-
 According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
@@
 Consequently, this behavior decreases the quality of the reconstruction of the traffic component.
-In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of the alert sounds are lower for the benefit of low frequency content.
-Consequently, the traffic sound level estimation is then underestimated as in Figure \ref{fig:lp_alert_12}.
-\\
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_-12.pdf}
-        \caption{}
-        \label{fig:Y_alert-12}
-    \end{subfigure}
-    \vfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/Y_alert_12.pdf}
-        \caption{}
-		\label{fig:Y_alert_12}
-    \end{subfigure}
-    \caption{Comparison between the sum of the 2 elements of $\mathbf{W_r}$ with the spectra of the interfering and traffic classes for an \textit{alert} scene for $TIR$ = -12 dB  (\subref{fig:Y_alert-12}) and $TIR$ = 12 dB (\subref{fig:Y_alert_12}) with $\beta = 2$, $K = 100$ and $w_t$ = 0.5 s (normalized amplitudes).}
-\end{figure}
 
@@
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ $\lbrace$-12, 12$\rbrace$ dB.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/dist_-12_12.pdf}
-    \caption{Example of the similarity $D_{\theta}\left( \mathbf{W_0} \vert \vert \mathbf{W}\right)$ for different sub-classes and for two $TIR$ (-12 dB and 12 dB) and with the threshold $t$ = 0.42 ($\beta$ = 1, $K$ = 200 and $w_t$ = 0.5 s).}
-    \label{fig:dist_-12_12}
-\end{figure}
-
 For low $TIR$, as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
@@
 The experimental protocol and the evaluated estimators have been implemented with the Matlab software.
-For reproducible purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}.
+For reproducible purposes, the code is available online.
 The evaluation database composed of multiple samples of urban sounds is also made available for the research community with interest in detection, separation and recognition tasks of urban sound sources.
 
-\vspace{6pt}
-%\acknowledgments{The authors would like to acknowledge partial funding from the ANR under project reference ANR-16-CE22-0012.}
+%\vspace{6pt}
 
-\footnotesize
-\bibliographystyle{unsrt}
-\bibliography{bibliographie_sensors}
+\section*{Declarations}
+
+\nomenclature{NMF}{Non-negative Matrix Factorization}
+\nomenclature{Sup-NMF}{Supervised Non-negative Matrix Factorization}
+\nomenclature{Sem-NMF}{Semi-Supervised Non-negative Matrix Factorization}
+\nomenclature{TI-NMF}{Thresholded Initialized Non-negative Matrix Factorization}
+\nomenclature{$TIR$}{Traffic Interfering Ratio}
+\nomenclature{$\mathbf{V}$}{Magnitude spectrogram of the audio sample}
+\nomenclature{$\mathbf{\tilde{V}}$}{Approximated spectrogram by the estimator}
+\nomenclature{$\mathbf{\tilde{V}}_{traffic}$}{Approximated spectrogram of the traffic by the estimator}
+\nomenclature{$\mathbf{W}$}{Dictionary matrix}
+\nomenclature{$\mathbf{H}$}{Activation matrix}
+\nomenclature{$\mathbf{W_s}$}{Fixed dictionary matrix for Sem-NMF}
+\nomenclature{$\mathbf{W_r}$}{Mobile dictionary matrix for Sem-NMF}
+\nomenclature{$\mathbf{H_s}$}{Activation matrix for $\mathbf{W_s}$}
+\nomenclature{$\mathbf{H_r}$}{Activation matrix for $\mathbf{W_r}$}
+\nomenclature{$\mathbf{W_0}$}{Initial dictionary matrix for TI-NMF}
+\nomenclature{$\mathbf{W'}$}{Updated dictionary for TI-NMF}
+\nomenclature{$MAE$}{Mean Absolute Error (dB)}
+\nomenclature{$\mathbf{f_c}$}{Cut-off frequency (Hz)}
+\nomenclature{$\mathbf{K}$}{Number of elements in $\mathbf{W}$}
+\nomenclature{$\mathbf{w_t}$}{Rectangular window (second)}
+\nomenclature{$\mathbf{t}$}{threshold value}
+\nomenclature{$\beta$}{Divergence class}
+\printnomenclature
+
+\section*{Availability of data and materials}
+The dataset supporting the conclusions of this article is available in the \textit{urban$\_$traffic$\_$nmf$\_$dataset.zip}  repository, \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}. 
+The software supporting the conclusions of this article is available in
+
+\begin{itemize}
+\item Project name: article2017EstimationAmbiance
+\item Project home page: \url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}
+\item Archived version: \url{https://doi.org/10.5281/zenodo.1145855}
+\item Programming language: Matlab
+\item Other requirements: Matlab 2016b or higher
+\item License: GNU GPL
+\item Any restrictions to use by non-academics: license needed 
+\end{itemize}
+
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Funding}
+This study is co-funded by Ifsttar and Pays de la Loire region with a partial funding from the ANR under project reference ANR-16-CE22-0012.
+
+\section*{Authors' contributions}
+JG carried out the numerical experience and drafted the manuscript. ML, AC and JP participated in the design of the study and helped to draft the manuscript. 
+All authors read and approved the final manuscript.
+
+% \section*{Acknowledgements}
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+\bibliography{bibliographie_jasm}   % name your BibTeX data base
 

### 2018-02-14 paper/gloaguenNmfTraffic.tex
minor corrections
add endfloat package
Completed abbreviation + break into 2 lines too long url

@@
 \usepackage{float}	% utiliser H dans les figures
+\usepackage[nolists]{endfloat}
 
@@
 \usepackage[hyphens]{url}	% gérer les adresses www.
+\usepackage[hyphenbreaks]{breakurl}
 
@@
 
-\begin{figure}[htp]
+\begin{figure}[t]
   \centering
@@
   \centering
-	\includegraphics[width=\linewidth]{figures/NMF_Lp_alert.pdf}
+	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
  \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
@@
 \end{figure*}
+
 \section{Results}\label{part:results}
@@
 \nomenclature{$\mathbf{\tilde{V}}_{traffic}$}{Approximated spectrogram of the traffic by the estimator}
+\nomenclature{$\mathbf{V}_{f_c}$}{Approximated spectrogram resulting from the low-pass filter}
 \nomenclature{$\mathbf{W}$}{Dictionary matrix}
 \nomenclature{$\mathbf{H}$}{Activation matrix}
+\nomenclature{$\mathbf{W_f}$}{$k$ element of the dictionary matrix}
 \nomenclature{$\mathbf{W_s}$}{Fixed dictionary matrix for Sem-NMF}
 \nomenclature{$\mathbf{W_r}$}{Mobile dictionary matrix for Sem-NMF}
-\nomenclature{$\mathbf{H_s}$}{Activation matrix for $\mathbf{W_s}$}
-\nomenclature{$\mathbf{H_r}$}{Activation matrix for $\mathbf{W_r}$}
+\nomenclature{$\mathbf{H_s}$}{Activation matrix related $\mathbf{W_s}$ in Sem-NMF}
+\nomenclature{$\mathbf{H_r}$}{Activation matrix for $\mathbf{W_r}$ in Sem-NMF}
 \nomenclature{$\mathbf{W_0}$}{Initial dictionary matrix for TI-NMF}
@@
 \nomenclature{$\mathbf{w_t}$}{Rectangular window (second)}
-\nomenclature{$\mathbf{t}$}{threshold value}
+\nomenclature{$\mathbf{t}$}{Threshold value}
 \nomenclature{$\beta$}{Divergence class}
@@
 \section*{Availability of data and materials}
-The dataset supporting the conclusions of this article is available in the \textit{urban$\_$traffic$\_$nmf$\_$dataset.zip}  repository, \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}. 
+The dataset supporting the conclusions of this article is available in the \textit{urban$\_$traffic$\_$nmf$\_$dataset.zip}  repository,  \sloppy \burl{https://zenodo.org/record/1145855#.Wl2oPnkiGos}. 
 The software supporting the conclusions of this article is available in
@@
 \item Project name: article2017EstimationAmbiance
-\item Project home page: \url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}
+\item \sloppy Project home page: \burl{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}
 \item Archived version: \url{https://doi.org/10.5281/zenodo.1145855}

### 2018-02-14 paper/gloaguenNmfTraffic.tex
fix fig size in full page

@@
   \centering
-  \subfloat[]{\label{fig:specW}\includegraphics[height=4cm]{figures/dictionary3.pdf}}
-  \hspace{5pt}
-  \subfloat[]{\label{fig:ElementW}\includegraphics[height=4cm]{figures/dictionary4.pdf}}
+  \subfloat[]{\label{fig:specW}\includegraphics[width=.45\linewidth]{figures/dictionary3.pdf}}
+%  \hspace{5pt}
+  \subfloat[]{\label{fig:ElementW}\includegraphics[width=.45\linewidth]{figures/dictionary4.pdf}}
   \caption{Dictionary building on a 3 second extract of a car passage.

### 2018-06-25 revision/responseToreviewers.tex
Add response2reviewers

\documentclass[10pt]{article}

\title{Reply to reviewers concerning submission JASM-D-18-00015: "Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques"}

\begin{document}

\maketitle

As a preamble, we would like to thank the editor and the reviewer for their comments and suggestions. Following these comments, we made several changes to the article, which are summarized here. The next sections list our answers to each of the reviewer’s comments, with references to the revised manuscript (page, column, and paragraph) where appropriate.

\section{Answers to Reviewer 1}
\subsection{Contribution and state of the art}

\begin{enumerate}

\item \emph{The authors miss some important references when they review the state of the art related to the addressed problem. For instance, those articles published within the DCASE 2017 challenge, which are not included in the paper. Moreover, other specific relevant works should be referenced, such as:
\begin{itemize}
\item Onur Dikmen and Annamaria Mesaros. "Sound event detection using non-negative dictionaries learned from annotated overlapping events", 2013 IEEE Workshop on Applications of Signal Processing to Audio and Acoustics (WASPAA), where also a threshold-based NMF approach is applied to a very similar problem.\\
And the work previously published by the same authors in DCASE 2016:
\item Jean-Rémy Gloaguen, Arnaud Can, Mathieu Lagrange, Jean-François Petiot. "Estimating traffic noise levels using acoustic monitoring: a preliminary study." Detection and Classification of Acoustic Scenes and Events (DCASE 2016), September 2016, Budapest, Hungary.
\end{itemize}
Both papers should be included as related work explicitly explaining the main differences with respect to the presented work.}

$\rightarrow$

\item \emph{The authors argue that "Prior to data assimiliation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved [31]", being [31] a reference from 2011. The authors should rephrase the sentence or change the reference in order to update the current state of the art in this field. For instance, only reference [35] is included describing deep learning approaches, which is one of the relevant approaches in this research field during the last years.}

\item \emph{As a final suggestion, I would recommend the authors completing the review of the state of the art by including a specific section about the Related Work in the manuscript. Although this section is not mandatory according to the guide for authors in this journal, it would help to clarify the contribution of the authors with respect to previous similar approaches.}

\end{enumerate}

\subsection{Metric and results}

\begin{enumerate}
\item \emph{Since Lp values are in dB, you cannot average them linearly as considered in your proposed metric MAE of eq. (15). The values must be averaged in their linear representation before converting them into dBs again. This error may have a dramatic impact on your results.}

\item \emph{First, the results obtained when considering fc=20Khz ("fc = 20 kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources") show the lowest MAE for TIR={6,12}, where "the traffic is the main sound source". 
--> Have the authors realized the consequences of these results? From these them someone can conclude that it is better to process nothing when interfering sources are 6 and 12 dB lower with respect to the traffic noise level!!}

\item \emph{Second, for the remaining studied scenarios, i.e., TIR={-12,-6,0}, the lowest averaged MAE values are obtained by the Sem-NMF, and not by the proposed TI-NMF.
Therefore, from these results, it cannot be concluded that TI-NMF is the best approach in the considered experimental framework. Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}

\item \emph{Finally, I would recommend the authors to consider including a statistical analysis to prove that differences between the different approaches are statistically significant or not.}

\item \emph{Last but not least, please review the results of the low pass filter approach with fc=20kHz in Table 4 and 3, since the global averaged MAE in Table is 4.54, while in Table 3 is 4.69.}

\end{enumerate}

\subsection{Meaningfulness of the experiments}

\begin{enumerate}
\item \emph{The baseline approach considered by the authors is a low-pass filter with different cut-off frequencies "fc={500, 1k, 2k, 5k, 10k, 20k} Hz.". First, this baseline may be understood as too simple for the problem at hand by the research community since, at least, supervised NMF has been already considered in the literature. Second, no comparison with [38] or other previous attempts to apply NMF to the problem at hand is provided. To this aim, it would be especially interesting the comparison with (Dikmen and Mesaros, 2013). Finally, after introducing the fc sweep in section 3.2, only fc={500, 20k}Hz are the subsequently considered with no further justification.}

\item \emph{The authors have considered a database composed of artificially mixed sound sources. They argue this decision as follows: "The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes" --$>$ I would suggest removing the term "mandatory" from the argumentation. Maybe the sentence could be rewritten as: "The use of simulated sound scenes allows…" }

\item \emph{Moreover, for this reviewer it is not clear what is considered traffic and not traffic by the authors in order to compute $Lp_traffic$. The "environmental sound scene corpus" is composed of "constant traffic noise" and "passing cars", which have been recorded in a real life environment by the authors. The car pass-bys are then used in several figures. For instance, Fig.4 shows an example created with the SimScene software showing background road traffic noise together with three overlapped sound sources: car horn, car passage and whistling birds. Does this mean that the car pass-by is considered as an interfering noise source? I would recommend the authors to clarify this point since a car pass-by is also traffic noise.}

\item \emph{Furthermore, the authors provide a lot of information about the configurations of the different approaches. However, I would recommend including specific information about the total duration of the train and test databases (e.g., not only saying "750 sound mixtures"), besides explicitly explaining the test scheme considered to guarantee that train and test data is decoupled when evaluating the considered sub-classes.}

\item \emph{Finally, it is worth mentioning that considering TIR={-6,-12,0,6,12} may yield to conclusions far from what can be found in real-life environments, as discussed in the literature (e.g. see [42]). Thus, I would recommend rephrasing the following sentence: "This database allows the creation of a wide diversity of realistic urban sound scenes from the road traffic point of view [16]" accordingly}

\end{enumerate}

\subsection{Minors comment}

\begin{enumerate}
\item \emph{In the introduction, the authors explain that road traffic noise maps show Lden and LN considering the A-weighted equivalent sound pressure level, thus, expressing the noise levels in dBA. However, their work only focuses on the equivalent sound pressure level in dB, why? }

\item \emph{Sections 2.1, 2.2 and 2.3 describing NMF, Supervised and Semi-Supervised NMF could be removed after extending the related work. This way, the manuscript could emphasize the proposal TI-NMF with respect to similar works.}

\item \emph{The authors state that "Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence. TI-NMF is performed on a lower number of iterations (60) to prevent W to not deviate too much from the initial dictionary." --> Please, justify this qualitative argumentation.}

\item \emph{Different kind of space reduction and data representation are included to reduce the computational time, e.g. K-means and third octave bands. --> I would recommend the authors to include some kind of diagram to better explain their approach, and, if possible, to include some experiments to argue their argumentation, for instance, to explain why you have chosen third octave bands in front of other possible parameterization such as MFCC.}

\item \emph{"The threshold t is set between 0.30 and 0.70 with a step of 0.01" -- > Please include some argumentation or reference to justify the range selected for the experiments beyond the ones reported in Table 3. Please, include more information of these experiments.}

\item \emph{All the analyses are done considering t=0.42 as the best threshold for the TI-NMF approach. Nevertheless, the authors stand that "With a low threshold, it is possible to decrease the error", and present some examples subsequently to demonstrate that for high TIR=12 dB, the averaged MAE can be decreased by reducing the threshold value from t=0.42 to t=0.30. Correspondingly, the threshold should be moved from t=0.42 to t=0.55 to improve the results for low TIR=-12 dB. To what extent these values show better performance or they are due to overfitting?}

\item \emph{I would recommend including all the tuning and sweep analysis within an Appendix at the end of the manuscript.}

\end{enumerate}

\subsection{Some further details}

\begin{enumerate}
\item \emph{A -12 is missing in the enumeration of TIR values after eq. (12). Done
Nevertheless, I would recommend removing this enumeration from the explanation of the equation since these values should be included in the Experiments section, where they are indeed included.
}

\item \emph{Tables 1 and 2 could be fused within the same Table.}

\item \emph{In Section 3.2.2, wt {0.5 1} --> a comma is missing between 0.5 and 1, i.e., {0.5, 1} }

\item \emph{Fig 8. The y-axis range does not show the minimum values of the alarm Lp,1s lower than 50dB. Please, include them to better understand the impact of the alarm on the Lp computation. Moreover, to better understand the example, I would recommend the authors to append the spectrogram of the example to the Lp.1s (dB) figure.}

\item \emph{Fig 9. Caption: "according to the the best" --> remove one 'the}

\end{enumerate}
\subsection{English writing}

\begin{enumerate}
\item \emph{Section 1: "Prior to data assimiliation," --> assimilation Done}
\item \emph{Section 2.4: "Usually in unsupervised," a name is missing in the sentence.}
\item \emph{Section 4: 
"First, the errors produced by the filter are detailed" --> the errors are not produce by the filter itself. Please, rephrase the sentence to better explain the readers what are referring to.}

\item \emph{"the traffic elements selected in TI-NMF do not activated when" --> are not activated?}

\item \emph{Section 5. "In this work the non negative metric factorization framework was used" --> I would recommend using present perfect instead of past tense in the conclusions.}
\end{enumerate}

\section{Answers to Reviewer 2}

\begin{enumerate}
\item \emph{As the previous work has been already published, the authors should provide references to it and clearly state the new contributions of the presented paper. I believe there may be a substantial overlap with the authors' paper Gloaguena, J. R., Cana, A., Lagrangeb, M., $\&$ Petiotb, J. F. Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non négatives, 2018.}

\item \emph{Some sentences in the abstract are not well formulated and are hard to follow. E.g. "The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations." should be splitted, "being" should be omitted; "raise this kind of innovative approaches" - the verb "raise" is confusing; "based on non-negative factorization framework" - the word "matrix" is missing. The authors may want to be more specific about the obtained results.}

\item \emph{"Their spectra can be obtained and be a basis of W." The procedure is not clear and is only described in the experimental part. This sentence seems to mean that the matrix W is comprised of exemplars (e.g. see $http://www.cs.tut.fi/sgn/arg/music/tuomasv/chime-enhancement/pP4_gemmeke.pdf$), which is not the case. In supervised NMF framework, the basis matrix W is usually pre-trained applying NMF to the training data. It would be interesting to compare such approach to the proposed K-means procedure.}

\item \emph{One type of vehicles (passenger car) is addressed in the simulations. Authors may want to comment on why other types like motorcycles, which generate a palpable amount of noise and may have different acoustic characteristics, are not considered.}

\item \emph{Is also interesting to know if the distribution of microphones in a realistic noise level measurement scenario is taken into account during the simulations.}

\item \emph{"The obtained dictionary is expressed with third octave bands", "The spectrogram V, as the dictionary W, is expressed with third octave bands (F = 29)." - please, clarify. Perhaps is better to join the description of this representation with the section 3.2.1.}

\item \emph{As for the sound level calculation itself, it is not clear if A-weighting is applied.
Also, the terminology is a bit confusing. E.g. Figure 8 "1 second equivalent sound pressure level" - to my understanding, the equivalent sound pressure level Leq is an average over the specified period of time and, therefore, would correspond to a single value. If the goal is to eventually simulate the output of a sound pressure level meter, some calibration would be needed.}

\item \emph{It is not clear what is the number of iterations that was used for each NMF version. It might make sense to use the reduced number of iterations for Sup-NMF as it is done for TI-NMF.
8. Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0 in TI-NMF are the same? Authors may want to state it clearly in Section 3.2.1.}

\end{enumerate}
\subsection{Minor points}

\begin{enumerate}
\item \emph{"at fixed stations spread all over the cities [31] [30], which would make available of the long-term evolution of the traffic noise" - better "which would make the long-term evolution of the traffic noise available".}
\item \emph{"A two-step process is generally followed :" - there is an extra space before the word "followed" }
\item \emph{"In the opposite" - better "On the contrary"}
\item \emph{"to some extend" should be "to some extent}
\item \emph{"by forcing its initiation" should be "by forcing its initialization"}
\item \emph{•Figure 2 caption, please change to "of an urban", "correspond to".}
\item \emph{Figure 3 caption, the sentence "The 82-nd first elements are considered as traffic component." is confusing.}
\item \emph{Figure 5 caption, "experience" should be "experiment"?}
\item \emph{"The choice of the dimensions is often made so that F × K + K × N < F × N ." - could you please provide a reference for this statement? }
\item \emph{"must of all" should probably be "most of all" }
\item \emph{"The errors are too important for low TIR and this for all the sub-classes." - the sentence is not clear.}

\end{enumerate}
\end{document}

### 2018-06-26 revision/responseToreviewers.tex
first corrections

@@
 Both papers should be included as related work explicitly explaining the main differences with respect to the presented work.}
-
 $\rightarrow$
@@
 For instance, only reference [35] is included describing deep learning approaches, which is one of the relevant approaches in this research field during the last years.}
+$\rightarrow$
 
@@
 Although this section is not mandatory according to the guide for authors in this journal, it would help to clarify the contribution of the authors with respect to previous similar approaches.}
+$\rightarrow$
 
@@
 This error may have a dramatic impact on your results.}
+$\rightarrow$
 
@@
 From these them someone can conclude that it is better to process nothing when interfering sources are 6 and 12 dB lower with respect to the traffic noise level!!}
+$\rightarrow$
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
+$\rightarrow$
 
 \item \emph{Finally, I would recommend the authors to consider including a statistical analysis to prove that differences between the different approaches are statistically significant or not.}
+$\rightarrow$
 
 \item \emph{Last but not least, please review the results of the low pass filter approach with fc=20kHz in Table 4 and 3, since the global averaged MAE in Table is 4.54, while in Table 3 is 4.69.}
+$\rightarrow$ Done. The value at TIR = 6 was a misprint, it has been corrected.
 
@@
 Finally, after introducing the fc sweep in section 3.2, only fc={500, 20k}Hz are the subsequently considered with no further justification.}
+$\rightarrow$
 
@@
 Maybe the sentence could be rewritten as: "The use of simulated sound scenes allows…" }
+
+$\rightarrow$ Done. 'The use of simulated sound scenes allows rigorous experimental validation \dots'
 
@@
 
+$\rightarrow$ Done.
+'What is called traffic component is the sum of the road traffic background noise and the sound events generated by the passing car class.
+On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
+Car horn sound class belongs to this component as it is considered as a warning signal' (3.1)
+
 \item \emph{Furthermore, the authors provide a lot of information about the configurations of the different approaches.
 However, I would recommend including specific information about the total duration of the train and test databases (e.g., not only saying "750 sound mixtures"), besides explicitly explaining the test scheme considered to guarantee that train and test data is decoupled when evaluating the considered sub-classes.}
+
+$\rightarrow$ Done. 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (3.1)
 
@@
 Thus, I would recommend rephrasing the following sentence: "This database allows the creation of a wide diversity of realistic urban sound scenes from the road traffic point of view [16]" accordingly}
+
+$\rightarrow$ Done.
+'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
+This is between these values that the estimator has to be the most efficient.
+But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.'
 
@@
 
+$\rightarrow$ Done 'The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (3.2.2)
+
 \item \emph{Sections 2.1, 2.2 and 2.3 describing NMF, Supervised and Semi-Supervised NMF could be removed after extending the related work.
@@
 
+$\rightarrow$ This work is addressed to the urban acoustic community. As NMF is not a well-known tool in this community, we prefer to keep the presentation of the fundamentals of this method.
+
 \item \emph{The authors state that "Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
 TI-NMF is performed on a lower number of iterations (60) to prevent W to not deviate too much from the initial dictionary." --> Please, justify this qualitative argumentation.}
+
+$\rightarrow$
 
@@
 
+$\rightarrow$ About the choice of third-octave bands, we specify 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+This coarser method allows us to reduce the dimensionality and then decrease the computation time.
+Furthermore, by expressing the frequency axis on a log frequency axis, the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'
+
 \item \emph{"The threshold t is set between 0.30 and 0.70 with a step of 0.01" -- > Please include some argumentation or reference to justify the range selected for the experiments beyond the ones reported in Table 3.
 Please, include more information of these experiments.}
+
+$\rightarrow$
 
@@
 
+$\rightarrow$
+
 \item \emph{I would recommend including all the tuning and sweep analysis within an Appendix at the end of the manuscript.}
+
+$\rightarrow$
 
@@
 \begin{enumerate}
-\item \emph{A -12 is missing in the enumeration of TIR values after eq. (12). Done
-Nevertheless, I would recommend removing this enumeration from the explanation of the equation since these values should be included in the Experiments section, where they are indeed included.
-}
+\item \emph{A -12 is missing in the enumeration of TIR values after eq.
+(12).
+Nevertheless, I would recommend removing this enumeration from the explanation of the equation since these values should be included in the Experiments section, where they are indeed included.}
+
+$\rightarrow$ This part has been removed
 
@@
 
+$\rightarrow$ The Tables have been fused
+
 \item \emph{In Section 3.2.2, wt {0.5 1} --> a comma is missing between 0.5 and 1, i.e., {0.5, 1} }
+
+$\rightarrow$ Done.
 
@@
 
+$\rightarrow$
+
 \item \emph{Fig 9. Caption: "according to the the best" --> remove one 'the}
+
+$\rightarrow$ Done.
 
@@
 \begin{enumerate}
-\item \emph{Section 1: "Prior to data assimiliation," --> assimilation Done}
+\item \emph{Section 1: "Prior to data assimiliation," --> assimilation}
+
+$\rightarrow$ Done.
+
 \item \emph{Section 2.4: "Usually in unsupervised," a name is missing in the sentence.}
+
+$\rightarrow$ Done, 'Usually in unsupervised learning,…'
+
 \item \emph{Section 4: 
@@
 
+$\rightarrow$ Done, 'First, the errors calculated from the traffic sound estimation of the low-pass filter are detailed'
+
 \item \emph{"the traffic elements selected in TI-NMF do not activated when" --> are not activated?}
 
+$\rightarrow$ Done, 'the traffic elements selected in TI-NMF are not activated'
+
 \item \emph{Section 5. "In this work the non negative metric factorization framework was used" --> I would recommend using present perfect instead of past tense in the conclusions.}
+
+$\rightarrow$ Done, 'In this work the non negative metric factorization framework is used \dots'
+
 \end{enumerate}
@@
 
+$\rightarrow$ The paper 'Gloaguena, J.
+R., Cana, A., Lagrangeb, M., $\&$ Petiotb, J.
+F.
+Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non négatives, 2018' is a congress article without review submitted after the submission of the present paper.
+It deals with a similar approach but on a different sound corpus.
+Furthermore, it is written in french and is not suited to be cite in an international journal.
+
 \item \emph{Some sentences in the abstract are not well formulated and are hard to follow.
@@
 The authors may want to be more specific about the obtained results.}
+
+$\rightarrow$ Done, 'The task is to the best of our knowledge never been considered in the literature.
+We propose…', 'Among the technical issues that bring this kind of innovative approaches,', 'several techniques based on non-negative matrix  factorization framework'.
 
@@
 
+$\rightarrow$
+
 \item \emph{One type of vehicles (passenger car) is addressed in the simulations.
@@
 
+$\rightarrow$
+
 \item \emph{Is also interesting to know if the distribution of microphones in a realistic noise level measurement scenario is taken into account during the simulations.}
+
+$\rightarrow$
 
@@
 
-\item \emph{As for the sound level calculation itself, it is not clear if A-weighting is applied.
-Also, the terminology is a bit confusing.
+$\rightarrow$ Done, 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+This coarser method allows us to reduce the dimensionality and then decrease the computation time.
+Furthermore, by expressing the frequency axis on a log frequency axis, the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(3.2.2)
+
+\item \emph{As for the sound level calculation itself, it is not clear if A-weighting is applied.}
+
+$\rightarrow$ Done, 'The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (3.2.2)
+
+\item \emph{Also, the terminology is a bit confusing.
 E.g.
@@
 
-\item \emph{It is not clear what is the number of iterations that was used for each NMF version. It might make sense to use the reduced number of iterations for Sup-NMF as it is done for TI-NMF.
-8. Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0 in TI-NMF are the same? Authors may want to state it clearly in Section 3.2.1.}
+$\rightarrow$ The title has been corrected ('Evolution of 1 second equivalent sound pressure levels'). The calibration step is here not necessary as all the sound level are relatives.
+
+\item \emph{It is not clear what is the number of iterations that was used for each NMF version. It might make sense to use the reduced number of iterations for Sup-NMF as it is done for TI-NMF.}
+
+$\rightarrow$
+
+\item \emph{Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0 in TI-NMF are the same? Authors may want to state it clearly in Section 3.2.1.}
+
+$\rightarrow$ Done, 'The 10 versions of the built dictionary are then used for NMF.
+In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (3.2.1)
 
@@
 \item \emph{"at fixed stations spread all over the cities [31] [30], which would make available of the long-term evolution of the traffic noise" - better "which would make the long-term evolution of the traffic noise available".}
+
+$\rightarrow$ Done.
+
 \item \emph{"A two-step process is generally followed :" - there is an extra space before the word "followed" }
+
+$\rightarrow$ Done.
+
 \item \emph{"In the opposite" - better "On the contrary"}
+
+$\rightarrow$ Done.
+
 \item \emph{"to some extend" should be "to some extent}
+
+$\rightarrow$ Done.
+
 \item \emph{"by forcing its initiation" should be "by forcing its initialization"}
+
+$\rightarrow$ Done.
+
 \item \emph{•Figure 2 caption, please change to "of an urban", "correspond to".}
+
+$\rightarrow$ Done.
+
 \item \emph{Figure 3 caption, the sentence "The 82-nd first elements are considered as traffic component." is confusing.}
+
+$\rightarrow$ Done.
+
 \item \emph{Figure 5 caption, "experience" should be "experiment"?}
-\item \emph{"The choice of the dimensions is often made so that F × K + K × N < F × N ." - could you please provide a reference for this statement? }
+
+$\rightarrow$ Done.
+
+\item \emph{"The choice of the dimensions is often made so that F × K + K × N $<$ F × N ." - could you please provide a reference for this statement? }
+
+$\rightarrow$ Done.
+
 \item \emph{"must of all" should probably be "most of all" }
+
+$\rightarrow$ Done.
+
 \item \emph{"The errors are too important for low TIR and this for all the sub-classes." - the sentence is not clear.}
+
+$\rightarrow$ Done, 'Errors are too important for all sub-classes at low $TIR$.'
 

### 2018-06-26 revision/gloaguenNmfTraffic_v2.tex
first corrections

--- 2018-02-14 paper/gloaguenNmfTraffic.tex
@@
 Considering as input the acoustic pressure measured in many locations using a sensor grid approach would greatly complement and improve the quality of the predicted pressure values.
-Among the technical issues that raise this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.
+Among the technical issues that bring this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.
 
-In this paper, several techniques based on non-negative factorization framework are studied in this application scenario on a simulated sound scene corpus.
-The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
+In this paper, several techniques based on non-negative matrix  factorization framework are studied in this application scenario on a simulated sound scene corpus.
+The task is to the best of our knowledge never been considered in the literature.
+We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
 The results show the interest of our proposed approach for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods.
@@
 The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
-These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make available of the long-term evolution of the traffic noise levels.
+These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make the long-term evolution of the traffic noise levels available.
 It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
@@
 
-Prior to data assimiliation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
+Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
 Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap.
@@
 Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
-A two-step process is generally followed : describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
+A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
 A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
@@
 This method is then not considered here to be deployed all over cities.
-In the opposite, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present.
+On the contrary, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present.
 A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
@@
 To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
-The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
+The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
 
@@
 
-The choice of the dimensions is often made so that $F\times K + K \times N < F \times N$.
+The choice of the dimensions is often made so that $F\times K + K \times N < F \times N$ \cite{fevotte_nonnegative_2009}.
 NMF is then considered as a low rank approximation method.
@@
 The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasonable assumption in our application scenario.
-To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario.
+To some extent, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario.
 To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is a good candidate as it offers more flexibility.
@@
 We therefore propose an alternative scheme based on the unsupervised NMF.
-Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly.
+Usually in unsupervised learning, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly.
 Here, as the concerned sound source is known and audio samples of car passages are available, an initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}.
@@
 \ref{eq:updateH_Sup}) are updated alternatively.
-$\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study,
+$\mathbf{W}$ is therefore updated by forcing its initialization with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study,
 
@@
 
-This database allows the creation of a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
+This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}.
 A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds.
@@
 The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling bird), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (metallic and construction site noises) and \textit{transportation} (train, tramway and plane).
-In each file, traffic component is present as the sum of the background and event traffic sounds while the \textit{interfering} sound class is the sound sources not related to it.
+In each file, the traffic component is present.
+What is called traffic component is the sum of the road traffic background noise and the sound events generated by the passing car class.
+On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
+Car horn sound class belongs to this component as it is considered as a warning signal.
 To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the interfering class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).
@@
 
-with the \textit{Traffic Interference Ratio} $TIR \in$ $\lbrace$12, -6, 0, 6, 12$\rbrace$ dB.
+with $TIR$, the \textit{Traffic Interference Ratio}.
 When $TIR < 0$ dB, the traffic component is less present than the interfering class.
 In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
-The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values).
+In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB \cite{gloaguen_creation_2017}.
+This is between these values that the estimator has to be the most efficient.
+But, in this experimental framework, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.
+The total number of scenes designed is then 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values), each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.
 
@@
 \centering
-\caption{Summary of the different experimental factors and their modalities taken into account in the frequency low-pass filter estimator.}
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
-	\hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modalities}\end{tabular}\\ \toprule
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
-    \textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics  & 6
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}@{}C{2.4cm}@{}C{2cm}@{}}
-\rowcolor[HTML]{C0C0C0}
-   $\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 & 5\\
-\end{tabularx}
-
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
-   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20  & 6\\
-   \bottomrule
-\end{tabularx}
-\label{tab:experimental_factorsFilter}
-\end{table*}
-
-\begin{table*}[t]
-\centering
-\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator.}
+\caption{Summary of the different experimental factors and their modalities taken into account in the LP filter and the NMF estimators.}
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}@{}C{2cm}}
@@
 
-\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
-  \textbf{method} & Sup-NMF & Sem-NMF & TI-NMF & 3 \\
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+  \textbf{method} & LP filter & Sup-NMF & Sem-NMF & TI-NMF & 4 \\
+\end{tabularx}
+
+\begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+   $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20  & 6\\
+   \bottomrule
 \end{tabularx}
@@
 
-The obtained dictionary is expressed with third octave bands and each basis vector of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.
+Each basis vector of $\mathbf{W}$ is normalized such as $\vert \vert \mathbf{W_k} \vert \vert = 1$ with $\vert \vert \bullet \vert\vert$ the $\ell$-1 norm.
 Table \ref{tab:experimental_factorsNMF} summarizes the experimental factors ($K$ and $w_t$) for the dictionary building and their related modalities.
-This last step allows us to reduce the dimensionality while preserving a rich description of the spectral content.
-Experimental validation consistently showed that considering octave bands do not impact the performance of the estimator studied in this paper.
+The 10 versions of the built dictionary are then used for NMF.
+In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.
 
@@
 TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
-The spectrogram $\mathbf{V}$, as the dictionary $\mathbf{W}$, is expressed with third octave bands ($F$ = 29).
-This coarser method allows the reduction of the matrix size and decreases the computation time.
-But,  must of all, by expressing the frequency axis on a log frequency axis,  the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+This coarser method allows us to reduce the dimensionality and then decrease the computation time.
+Furthermore, by expressing the frequency axis on a log frequency axis, the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
+Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.
 For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
@@
 Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
-For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5, 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
 For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
@@
 where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
+The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.
 For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
@@
   \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
-  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
+  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
 	\label{fig:TIR_bar}
@@
 
-First, the errors produced by the filter are detailed.
+First, the errors calculated from the traffic sound estimation of the low-pass filter are detailed.
 $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources.
@@
 In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$.
-The errors are too important for low $TIR$ and this for all the sub-classes.
+Errors are too important for all sub-classes at low $TIR$.
 This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
@@
 In Figure \ref{fig:lp_alert}, the 1 second equivalent sound pressure level of an alert scene is displayed for $TIR \in \lbrace -12,12 \rbrace$ dB for the 500 Hz low pass filter and for TI-NMF.
-In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF do not activated when the alarm class sounds contrary to the low-pass filter.
+In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF are not activated when the alarm class sounds contrary to the low-pass filter.
 
@@
 
-In this work the non negative metric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures.
+In this work the non negative metric factorization framework is used to estimate the road traffic sound level in urban sound mixtures.
 It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and it is adapted to monophonic sensor networks.

### 2018-07-31 revision/responseToreviewers.tex
responses to reviewers continued

@@
 Both papers should be included as related work explicitly explaining the main differences with respect to the presented work.}
-$\rightarrow$
+
+$\rightarrow$ Done.
 
@@
 For instance, only reference [35] is included describing deep learning approaches, which is one of the relevant approaches in this research field during the last years.}
+
 $\rightarrow$
@@
 Although this section is not mandatory according to the guide for authors in this journal, it would help to clarify the contribution of the authors with respect to previous similar approaches.}
-$\rightarrow$
+
+$\rightarrow$ Done.
 
@@
 The values must be averaged in their linear representation before converting them into dBs again.
-This error may have a dramatic impact on your results.}
-$\rightarrow$
+This error may have a dramatic impact on your results.}\\
+$\rightarrow$ First we consider logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values as it occurs in linear field. 
+The logarithmic mean is then more sensitive to high error than linear mean. As the errors are rteated in a statistical point of view we consider that this operation is suited.
+Furthermore the use of arithmetic operators on logarithmic values is a common pratice in the urban acoustics field (see Aumond et al. "" and Morillas et al. "").
 
@@
 --> Have the authors realized the consequences of these results?
-From these them someone can conclude that it is better to process nothing when interfering sources are 6 and 12 dB lower with respect to the traffic noise level!!}
-$\rightarrow$
+From these them someone can conclude that it is better to process nothing when interfering sources are 6 and 12 dB lower with respect to the traffic noise level!!}\\
+$\rightarrow$ This result may be disappointing but it has to be remind that the TIR value is unknown when measurements is made on streets.
+Then it is impossible to determine when these cases occur.
+If the LP filter is the best at fc = 20 kHz for TIR = {6,12} dB, it is not efficient when the TIR is negative.
+We made the precision () :
+'Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not know.
+In consequence, without a priori knowledge, this approach cannot be applied.' (section 5).
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$
+$\rightarrow$ We made a precision at the end of section 2 : 'The aim is then to find a flavor of NMF that could propose the minimum error on the integral corpus as it is consider that no prior knowledge on the sound sources or on the traffic predominance are available as it is when measurements are made in city.'
 
 \item \emph{Finally, I would recommend the authors to consider including a statistical analysis to prove that differences between the different approaches are statistically significant or not.}
+
 $\rightarrow$
@@
 \item \emph{Last but not least, please review the results of the low pass filter approach with fc=20kHz in Table 4 and 3, since the global averaged MAE in Table is 4.54, while in Table 3 is 4.69.}
+
 $\rightarrow$ Done. The value at TIR = 6 was a misprint, it has been corrected.
@@
 Finally, after introducing the fc sweep in section 3.2, only fc={500, 20k}Hz are the subsequently considered with no further justification.}
-$\rightarrow$
+
+$\rightarrow$ The baseline of the low-pass filter, even if it seems too simple, is very effective in some cases.
+As in the environmental acoustics community, this method would have been easily seen as a tool, we decided to consider it.
+We did not summarize all the results for the low pass filter in order to be too focused on this method.
+As for the comparison with (Dikmen and Mesaros, 2013), even if they use the NMF framework, they develop a detection tool that is not adapted to this study.
+We believe that it was, first, wiser to use a low-pass filter and to monitor NMF as baselines than an advanced method.
 
@@
 
-$\rightarrow$ Done. 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (3.1)
+$\rightarrow$ Done.
+For the test database : 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (3.1), for the train database : 'It is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.' (3.2.1)
 
@@
 
-$\rightarrow$ Done 'The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (3.2.2)
+$\rightarrow$ Done 'The $A$-weightind of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
 
@@
 
-$\rightarrow$
+$\rightarrow$ We used to number of iteration, that was not relevant to compare fairly the approaches.
+In consequence, we decided to fix at 100 the number of iterations.
+'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
 
@@
 Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
-But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'.
+Also, we add a block diagram which details the dictionary building steps and makes clearer these different elements.
 
@@
 
-$\rightarrow$
+$\rightarrow$ Done.
+'For TI-NMF, a preliminary study reveals that the threshold $t$ value range can be reduce between  0.30 and 0.70.
+An increment step of 0.01 has been considered as being sufficiently precise.'
 
@@
 
-$\rightarrow$
+$\rightarrow$ The car horn sound level is null outside the range of the y-axis that's why we put aside the the low sound level to weep a wide range to see, at best, the curves.
+We made the precision on the legend.
 
-\item \emph{Fig 9. Caption: "according to the the best" --> remove one 'the}
+\item \emph{Fig 9. Caption: "according to the the best" --$>$ remove one 'the}
 
@@
 \begin{enumerate}
-\item \emph{Section 1: "Prior to data assimiliation," --> assimilation}
+\item \emph{Section 1: "Prior to data assimiliation," --$>$ assimilation}
 
@@
 
-\item \emph{Section 5. "In this work the non negative metric factorization framework was used" --> I would recommend using present perfect instead of past tense in the conclusions.}
+\item \emph{Section 5. "In this work the non negative matrix factorization framework was used" --> I would recommend using present perfect instead of past tense in the conclusions.}
 
@@
 
-$\rightarrow$
+$\rightarrow$ The K-mean and NMF procedure can be considered as similar approaches as a clustering tool (see Ding et al., 2005), so it has not been tested.
+A bloc diagram has been added to better explain dictionary learning.
 
@@
 
-$\rightarrow$
+$\rightarrow$ Done, 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.' (Conclusion)
 
@@
 
-$\rightarrow$
+$\rightarrow$ The purpose of this study is to propose a tool that could be integrated into a sensor network. The distribution of the microphone is not a question dealt with in this work.
 
 \item \emph{"The obtained dictionary is expressed with third octave bands", "The spectrogram V, as the dictionary W, is expressed with third octave bands (F = 29)." - please, clarify.
-Perhaps is better to join the description of this representation with the section 3.2.1.}
+Perhaps is better to join the description of this representation with the section 4.2.1.}
 
@@
 Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
-But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(3.2.2)
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(4.2.2)
 
@@
 
-$\rightarrow$ Done, 'The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (3.2.2)
+$\rightarrow$ Done, 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
 
@@
 
-$\rightarrow$
+$\rightarrow$ We used to number of iteration, that was not relevant to compare fairly the approaches.
+In consequence, we decided to fix at 100 the number of iterations.
+'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
 
-\item \emph{Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0 in TI-NMF are the same? Authors may want to state it clearly in Section 3.2.1.}
+\item \emph{Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0 in TI-NMF are the same? Authors may want to state it clearly in Section 4.2.1.}
 
 $\rightarrow$ Done, 'The 10 versions of the built dictionary are then used for NMF.
-In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (3.2.1)
+In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (4.2.1)
 

### 2018-07-31 revision/gloaguenNmfTraffic_v2.tex
responses to reviewers continued

@@
 The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
-These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make the long-term evolution of the traffic noise levels available.
+These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make the long-term evolution of the traffic noise available.
 It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.
@@
 
-Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels.
+Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters...), including the assessment of urban noise levels.
 The DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations focusing on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}.
@@
 Consequently, the traffic sound level estimation based on measurements is not a trivial task.
+
+\section{Related works}
 Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
@@
 A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
-When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}.
+When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008}.
 In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
@@
 
-Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}.
-After having performed NMF on simulated audio files, they perform a source separation in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
-We study in this paper different flavors of Non-Negative Matrix Factorization where the traffic component is considered in its entirely whether it is a sound background or an event.
+Closer to our application scenario, NMF has been considered in \cite{mesaros_sound_2015} where coupled NMF has been used as a sound event detection in real life recordings.
+However, no information on the quality of the detected signal is mentioned.
+Nevertheless, they raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary?
+They conclude on the efficiency of each approach and suggere to combine both to deal with large databases.
+Also, Innami and Kasai \cite{satoshi_innami_nmf-based_2012} proposed to perform NMF on simulated audio files as a source separation method.
+Their tool is composed of two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
+
+On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.
+Here, we extent this approach with more flavors of NMF on more complex sound scenes.
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
@@
 The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
+The aim is then to find a flavor of NMF that could propose the minimum error on the integral corpus as it is consider that no prior knowledge on the sound sources or on the traffic predominance are available as it is when measurements are made in city.
 
@@
 \includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
-\caption{NMF decomposition of a urban sound mixture comprising 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
+\caption{NMF decomposition of an urban sound mixture comprising 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond to 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
 \label{fig:example_NMF}
@@
 $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = 1$ means that the $k$-th element of $\mathbf{W}$ is identical to the $k$-th element of $\mathbf{W_0}$.
-In the opposite, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
+On the contrary, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
 This measure has the advantage to be bounded between 1 and 0 and to be invariant with respect to scale.
@@
 \includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
-\caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
+\caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The first $82$ elements are considered as traffic component.}
 \label{fig:W_TI_NMF}
@@
 When $TIR < 0$ dB, the traffic component is less present than the interfering class.
-In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
+On the contrary, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
 In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB \cite{gloaguen_creation_2017}.
@@
     \includegraphics[width=\linewidth]{figures/bloc_diagram_estimator.pdf}
-    \caption{Block diagram of the experience with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF.}
+    \caption{Block diagram of the experiment with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF.}
     \label{fig:bloc_experiment}
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{2cm}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}@{}C{2cm}@{}}
+\rowcolor[HTML]{C0C0C0}
    $\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20  & 6\\
-   \bottomrule
 \end{tabularx}
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}
-\rowcolor[HTML]{C0C0C0}
     $\mathbf{w_t}$ (s) & 0.5 & 1 & \textit{all} & 3
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
+\rowcolor[HTML]{C0C0C0}
     $\mathbf{K}$ & 25 & 50 & 100 & 200  & 4\\
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{6cm}@{}@{}C{6cm}@{}C{2cm}@{}}
-\rowcolor[HTML]{C0C0C0}
    $\mathbf{\beta}$ & 1 & 2 & 2\\
@@
 \begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
+\rowcolor[HTML]{C0C0C0}
    threshold $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 41\\
@@
 In order to prevent potential overfitting issues, the dictionary is built from a separate sound database dedicated specifically to this task.
-It is composed of 53 audio files of passing cars.
+The train database is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.
 These recordings have been made on the Ifsttar's runway too, with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
+The different steps leading to a dictionary is resumed in Figure \ref{fig:creation_W}.
 First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$).
@@
 The $K$-means clustering algorithm is then reduced to $K \in \lbrace$25, 50$\rbrace$.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{figures/creation_W_EN.pdf}
+\caption{Steps involved in the dictionary learning.}
+\label{fig:creation_W}
+\end{figure}
 
@@
 \textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &   $\mathbf{t}$ & \textbf{$MAE$} (dB) \\ \midrule
-filter & 20  &  &  &  &  & 4.69 ($\pm$ 4.52) \\
-filter & 0.5 &  &  &  &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
-Sup-NMF &  & 1 & 50 & 0.5  &  & 3.44 ($\pm$ 3.70) \\
-Sup-NMF &  & 2 & 50 & 0.5  &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
-Sem-NMF &  & 1 & 100 & 0.5 &   & 2.33 ($\pm$ 1.10) \\
-Sem-NMF &  & 2 & 100 & 0.5 &   & 2.32 ($\pm$ 1.26) \\ \hline \hline
-\textbf{TI-NMF} &  & \textbf{1} & \textbf{200} & \textbf{0.5} &  \textbf{0.42} &\textbf{2.19 ($\pm$ 2.18)} \\
-TI-NMF &  & 2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
+filter & 20  & - & - & - & - & 4.69 ($\pm$ 4.52) \\
+filter & 0.5 & - &-  & - & - & 2.89 ($\pm$ 2.84) \\ \hline \hline
+Sup-NMF & - & 1 & 50 & 0.5  & - & 3.44 ($\pm$ 3.70) \\
+Sup-NMF & - & 2 & 50 & 0.5  & - & 3.02 ($\pm$ 3.33) \\ \hline \hline
+Sem-NMF & - & 1 & 100 & 0.5 & -  & 2.33 ($\pm$ 1.10) \\
+Sem-NMF & - & 2 & 100 & 0.5 & -  & 2.32 ($\pm$ 1.26) \\ \hline \hline
+\textbf{TI-NMF} & - & \textbf{1} & \textbf{200} & \textbf{0.5} &  \textbf{0.42} &\textbf{2.19 ($\pm$ 2.18)} \\
+TI-NMF & - & 2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
 \end{tabular}
@@
 \textbf{method} & filter & filter & Sup-NMF & Sem-NMF & TI-NMF \\ \midrule
-$f_c$ (kHz) & 20 & 0.5 &  &  &  \\
-$\mathbf{\beta}$ &  &  & 2 & 2 & 1 \\ \hline
+$f_c$ (kHz) & 20 & 0.5 & - & - & - \\
+$\mathbf{\beta}$ & - & - & 2 & 2 & 1 \\ \hline
 \textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & 8.65 ($\pm$ 1.88) & 3.88 ($\pm$ 1.52) & 5.35 ($\pm$ 2.71) \\
@@
 \textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.34 ($\pm$ 0.56) & 1.11 ($\pm$ 0.25) & 1.26 ($\pm$ 0.35) \\
-\textbf{6} & 0.25 ($\pm$ 0.06) & 1.03 ($\pm$ 0.26) & 0.26 ($\pm$ 0.10) & 2.25 ($\pm$ 0.19) & 0.70 ($\pm$ 0.32) \\
+\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 0.26 ($\pm$ 0.10) & 2.25 ($\pm$ 0.19) & 0.70 ($\pm$ 0.32) \\
 \textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.64 ($\pm$ 0.06) & 2.96 ($\pm$ 0.21)  & 0.84 ($\pm$ 0.24) \\ \bottomrule
@@
 
-Sup-NMF and Sem-NMF updates are computed for 400 iterations, which is sufficient to reach convergence.
-TI-NMF is performed on a lower number of iterations (60) to prevent $\mathbf{W}$ to not deviate too much from the initial dictionary.
+100 iterations are performed for all NMF.
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
@@
 But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.
-For TI-NMF, the threshold $t$ is set between 0.30 and 0.70 with a step of 0.01.
+For TI-NMF, a preliminary study reveals that the threshold $t$ value range can be reduce between  0.30 and 0.70.
+An increment step of 0.01 has been considered as being sufficiently precise.
 Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
@@
 where $p_{rms}$ is the effective pressure deducted from the estimated traffic spectrogram $\mathbf{\tilde{V}}_{traffic}$ and $p_0$ is the reference sound pressure, $p_0 = 2 \times 10^{-5} Pa$.
-The $A$-weighted of the sound level is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.
+The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.
 For each setting of experimental factors, $M$ values of $\tilde{L}_{p,traffic}$, corresponding to the $M$ scenes, are then obtained and are compared to the $M$ exact sound level, $L_{p,traffic}$.
@@
 	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+ \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
 	\label{fig:lp_alert}
@@
 Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
+Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not know.
+In consequence, without a priori knowledge, this approach cannot be applied.
 $f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained (2.89 ($\pm$ 2.84) dB).
@@
 
-In this work the non negative metric factorization framework is used to estimate the road traffic sound level in urban sound mixtures.
+In this work the non negative matrix factorization framework is used to estimate the road traffic sound level in urban sound mixtures.
 It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and it is adapted to monophonic sensor networks.
@@
 A major advantage of the proposed approach is that it is not designed for a specific source.
-Even though the experiments described in this paper focused on traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
-Extending the apporach to other sources is thus of intetrest for future research.
+It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.
+However, even though the experiments described in this paper focused on road traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
+Extending the apporach to other sources is thus of interest for future research.
 Also, performance improvement could be achieved by the addition of constraints such as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
+
+Currently, the sound database is composed of mainly composed passing car but 
+The traffic component is mainly composed of passing car event and continuous road traffic.
 
@@
 \section*{Authors' contributions}
-JG carried out the numerical experience and drafted the manuscript. ML, AC and JP participated in the design of the study and helped to draft the manuscript.
+JG carried out the numerical experiment and drafted the manuscript. ML, AC and JP participated in the design of the study and helped to draft the manuscript.
 All authors read and approved the final manuscript.

### 2018-08-03 revision/responseToreviewers.tex
add t-tes part in the end of result part

@@
 This error may have a dramatic impact on your results.}\\
-$\rightarrow$ First we consider logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values as it occurs in linear field. 
-The logarithmic mean is then more sensitive to high error than linear mean. As the errors are rteated in a statistical point of view we consider that this operation is suited.
-Furthermore the use of arithmetic operators on logarithmic values is a common pratice in the urban acoustics field (see Aumond et al. "" and Morillas et al. "").
+$\rightarrow$ First we have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values as it occurs in linear field. 
+The logarithmic mean is then more sensitive to high error than linear mean. As the errors are treated in a statistical point of view we consider that this operation is suited.
+Furthermore the use of arithmetic operators on logarithmic values is a common practice in the urban acoustics field (see Aumond et al.
+"Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." and Morillas et al.
+"Uncertainty evaluation of continuous noise sampling.").
+We add a comment on the paper 'The use of a this metric is justify by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
+The $MAE$ error with logarithmic values allows an equal weight between each difference in sound level and makes it possible to be less sensitive to high energies.
+It is also a common practice in the environmental acoustics field' (section 4.2.3)
 
@@
 If the LP filter is the best at fc = 20 kHz for TIR = {6,12} dB, it is not efficient when the TIR is negative.
-We made the precision () :
+We made the precision :
 'Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not know.
@@
 As in the environmental acoustics community, this method would have been easily seen as a tool, we decided to consider it.
-We did not summarize all the results for the low pass filter in order to be too focused on this method.
+We summarized all the results for the low pass filter according to the cut-off frequency in order to be more precise (Figure 9).
 As for the comparison with (Dikmen and Mesaros, 2013), even if they use the NMF framework, they develop a detection tool that is not adapted to this study.
@@
 
-$\rightarrow$
+$\rightarrow$ This improvement is due to a decreasing of the number of element considered as traffic component.
+As in TIR = -12 dB, the interfering class is more present, it takes more basis in W'.
+By increasing the threshold, we put aside more elements from W' that can belong to this sound class.
+It is equivalent, in detection task, to reduce the number of false-positive.
+It has been precised in the end of the section 5.
 

### 2018-08-03 revision/gloaguenNmfTraffic_v2.tex
add t-tes part in the end of result part

@@
 They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-Altough very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
+Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
 In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
@@
 Their tool is composed of two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
-
 On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.
@@
 This measure has the advantage to be bounded between 1 and 0 and to be invariant with respect to scale.
-The $k$ values of $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ are next sorted in descending order.
+The $K$ values of $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ are next sorted in decreasing order.
 The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are selected by a \textit{hard thresholding} method.
@@
 Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} have been investigated.
-A fast parametric study revealed that the \textit{hard} thresholding method, as presented in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
+A quick parametric study revealed that the \textit{hard} thresholding method, as presented in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
 
@@
 
-The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ($\lbrace$-12, -6, 0, 6, 12$\rbrace$ dB).
+The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al.), \textit{animals} (an.), \textit{humans} (hu.), \textit{climate} (cl.), \textit{transportation} (tr.), \textit{mechanics} (me.)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ($\lbrace$-12, -6, 0, 6, 12$\rbrace$ dB).
+The range of these values is large but in the urban environments, the $TIR$ seems to be between -6 dB and 12 dB \cite{gloaguen_creation_2017}.
+The case $TIR$ = -12 dB is then an extreme case to study the NMF behavior.
 The spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap, see Figure \ref{fig:bloc_experiment}.
@@
 In order to prevent potential overfitting issues, the dictionary is built from a separate sound database dedicated specifically to this task.
-The train database is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.
+The train database is composed of 53 audio files of isolated passing cars with a 18 minutes and 29 seconds cumulative duration.
 These recordings have been made on the Ifsttar's runway too, with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio).
@@
 filter & 0.5 & - &-  & - & - & 2.89 ($\pm$ 2.84) \\ \hline \hline
-Sup-NMF & - & 1 & 50 & 0.5  & - & 3.44 ($\pm$ 3.70) \\
-Sup-NMF & - & 2 & 50 & 0.5  & - & 3.02 ($\pm$ 3.33) \\ \hline \hline
-Sem-NMF & - & 1 & 100 & 0.5 & -  & 2.33 ($\pm$ 1.10) \\
-Sem-NMF & - & 2 & 100 & 0.5 & -  & 2.32 ($\pm$ 1.26) \\ \hline \hline
-\textbf{TI-NMF} & - & \textbf{1} & \textbf{200} & \textbf{0.5} &  \textbf{0.42} &\textbf{2.19 ($\pm$ 2.18)} \\
-TI-NMF & - & 2 & 25 & all &  0.54 & 2.20 ($\pm$ 2.26)\\ \bottomrule
+Sup-NMF & - & 1 & 25 & $all$  & - & 3.45 ($\pm$ 3.70) \\
+Sup-NMF & - & 2 & 25 & 2  & - & 2.84 ($\pm$ 3.19) \\ \hline \hline
+Sem-NMF & - & 1 & 200 & 2 & -  & 2.32 ($\pm$ 1.15) \\
+Sem-NMF & - & 2 & 200 & 2 & -  & 2.32 ($\pm$ 1.26) \\ \hline \hline
+\textbf{TI-NMF} & - & \textbf{1} & \textbf{200} & \textbf{0.5} &  \textbf{0.41} &\textbf{2.15 ($\pm$ 2.10)} \\
+TI-NMF & - & 2 & 200 & 0.5 &  0.36 & 2.29 ($\pm$ 2.40)\\ \bottomrule
 \end{tabular}
@@
 $\mathbf{\beta}$ & - & - & 2 & 2 & 1 \\ \hline
-\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & 8.65 ($\pm$ 1.88) & 3.88 ($\pm$ 1.52) & 5.35 ($\pm$ 2.71) \\
-\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 4.22 ($\pm$ 1.27) & 1.37 ($\pm$ 0.71)  & 2.82 ($\pm$ 1.30) \\
-\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.34 ($\pm$ 0.56) & 1.11 ($\pm$ 0.25) & 1.26 ($\pm$ 0.35) \\
-\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 0.26 ($\pm$ 0.10) & 2.25 ($\pm$ 0.19) & 0.70 ($\pm$ 0.32) \\
-\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.64 ($\pm$ 0.06) & 2.96 ($\pm$ 0.21)  & 0.84 ($\pm$ 0.24) \\ \bottomrule
+\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.39 ($\pm$ 3.00) & 8.08 ($\pm$ 2.44) & 2.98 ($\pm$ 2.11) & 5.22 ($\pm$ 2.62) \\
+\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 3.84 ($\pm$ 1.58) & 1.52 ($\pm$ 0.60)  & 2.72 ($\pm$ 1.24) \\
+\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.15 ($\pm$ 0.62) & 1.60 ($\pm$ 0.47) & 1.26 ($\pm$ 0.35) \\
+\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 0.35 ($\pm$ 0.20) & 2.49 ($\pm$ 0.30) & 0.75 ($\pm$ 0.34) \\
+\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.77 ($\pm$ 0.07) & 3.02 ($\pm$0.22) & 0.83 ($\pm$ 0.23) \\ \bottomrule
 \end{tabular}
@@
 An increment step of 0.01 has been considered as being sufficiently precise.
-Tables  \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
+Table \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
 
-Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
+Considering the experimental settings derived from the different modalities of each experimental factor  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
 For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5, 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
@@
 
+The $MAE$ is a performance index without dimension.
+The use of a this metric is justify by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
+The $MAE$ error with logarithmic values allows an equal weight between each difference in sound level and thus not being more sensitive to high energies.
+It is also a common practice in the environmental acoustics field \cite{morillas2014uncertainty} \cite{aumond2018kriging}.
+
 \begin{figure}[t]
@@
   \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 2.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 1.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
   \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.42.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.41.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
   \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
@@
 
-First, the errors calculated from the traffic sound estimation of the low-pass filter are detailed.
-$f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources.
-Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
-Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not know.
-In consequence, without a priori knowledge, this approach cannot be applied.
-$f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained (2.89 ($\pm$ 2.84) dB).
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{./figures/filter_bar_integrate.pdf}
+\caption{Average $MAE$ errors for the LP filter on all the corpus.}
+\label{fig:filter_intergate}
+\end{figure}
+
+First, the errors calculated from the traffic sound estimation of the low-pass filter on all the corpus are detailed in Figure \ref{fig:filter_intergate}.
+The LP filter with the cut-off frequency $f_c = 20 $ kHz generates the highest error as it is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources.
+Consequently, on Table \ref{tab:results_TIR} in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
+Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not known.
+In consequence, without prior knowledge, this approach cannot be applied.
+Then, the traffic sound level estimation error decrease with the cut-off frequency.
+Finally, $f_c = 500$ Hz is the one which reach the lowest mean error obtained ($MAE$ = 2.89 ($\pm$ 2.84)).
 It is then considered as the baseline to compare the performances of NMF.
@@
 For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then does not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 dB and are overestimated.
+The errors are then nearly all superior to 4 and are overestimated.
 In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels.
@@
 Sem-NMF and TI-NMF have better results.
-The lowest average error is obtained for TI-NMF, 2.19 ($\pm$ 2.18) dB, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
-On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 ($\pm$ 1.26) dB).
+The lowest average error is obtained for TI-NMF,  2.15 ($\pm$ 2.10), for $\beta$ = 1 and threshold $t$ = 0.41 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms.
+On the other hand, the semi-supervised approach has a higher error but a lower standard deviation ($MAE$ = 2.32 ($\pm$ 1.15)).
+TI-NMF seems to be the most efficient approach on all corpus without having prior knowledge on the interfering sound class or on the $TIR$ value.
 
-According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs.
-In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$.
-Errors are too important for all sub-classes at low $TIR$.
+As the mean error between the best SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 2 s), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 2 s) and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.41) approaches are closed, a Student's t-test is performed to evaluate the statistical differences between them.
+It is performed on all the estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
+The test estimates a $p$-value which is confronted to an significant threshold $\alpha$ = 5 $\%$ that confirms an $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
+The $p$-values are summarized in the Table \ref{student_test} with the $t$-values and the degrees of liberty (Dof).
+
+\begin{table}[t]
+\centering
+\begin{tabular}{lccccc}
+\toprule
+  & SUP/SEM NMF & SUP/TI NMF & SEM/TI NMF \\
+\midrule
+$t$-value & 7.81 & 2.84 & 4.60\\
+Dof & 1478 & 1470 & 1498\\
+$p$-value & \textbf{5.33 $\times 10^{-15}$} & \textbf{1.32 $\times 10^{-3}$} & \textbf{4.57 $\times 10^{-4}$} \\
+\bottomrule
+\end{tabular}
+\caption{$p$-values deducted from the Student's test. In bold letters, the $p$-values that reject the $H_0$ hypothesis ($p$-value $\geq \alpha$).}
+\label{tab:student_test}
+\end{table}
+
+For each couple of method, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite a similar mean errors.
+To better understand these differences, the error made according the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.
+
+In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$. Errors are too important for all sub-classes at low $TIR$. 
+In low $TIR$, SUP-NMF fails to estimate correctly the traffic sound level.
 This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
-In the case of low $TIR$, in the aim to reduce the objective function, see eq.
+In the aim to reduce the objective function, see eq.
 \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the sound scene.
 Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
+At high $TIR$, this approach generate better estimation as it is a favorable case.
 
@@
 However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR \in \lbrace$6, 12$\rbrace$ dB.
+The performances are even superior to the LP filter baseline.
 In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem-NMF is free to include traffic components in $\mathbf{W_r}$.
@@
 
-Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}).
+The TI-NMF behavior, with a threshold $t = 0.41$ and $\beta$ = 1, generates error inferior to the baseline for all the $TIR$ values, with the exception at $TIR$ = 0 dB.
+Its behavior is singular because it does not propose on each $TIR$ the lowest error even if it is the best approach on all the corpus.
 Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
 The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal.
-For low $TIR$, as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
-With the thresholding, only a reduced number of basis vectors are considered as traffic components.
+
+For low $TIR$, SEM-NMF outperformed as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
+With the thresholding, only a reduced number of basis vectors are considered as traffic components: from the 200 basis contained in $\mathbf{W'}$, 106 are considered as \textit{traffic}  component in $TIR$ = -12 dB.
 In comparison to supervised results, this approach reduces significantly the error for the \textit{human} and \textit{transport} sub-classes.
@@
 
-For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decreases the error (< 1 dB).
+For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components (181 basis for $TIR$ = 12 dB) and then decreases the error ($MAE$ < 1).
 The kept elements are then more suited to the scenes than a fixed dictionary.
-The error for these $TIR$ is then due to the thresholding which put aside some elements that are nevertheless related to the traffic component.
+The error for these $TIR$ is then due to the thresholding which put aside some elements that can be related to the traffic component.
+
+It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
+In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 $\pm$ 1.75 with in average 73 elements considered as \textit{traffic} elements.
+
+But, in order to generalize this method to a practical case where no prior knowledge on the urban environment is made, the chosen threshold $t$ is then fixed to $t$ = 0.41 as it is the one that best balanced these opposite cases.
 In Figure \ref{fig:lp_alert}, the 1 second equivalent sound pressure level of an alert scene is displayed for $TIR \in \lbrace -12,12 \rbrace$ dB for the 500 Hz low pass filter and for TI-NMF.
 In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF are not activated when the alarm class sounds contrary to the low-pass filter.
-
-With a low threshold, it is possible to decrease the error.
-For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30.
-In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 ($\pm$ 1.66) dB.
-In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 as it is the one that best balanced these opposite cases.
 
@@
 On a large corpus of sounds, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant.
-The proposed approach, named threshold initialized NMF achieves the lowest average error.
+The proposed approach, named threshold initialized NMF achieves the lowest average error on the entire corpus.
 With this method, the $\mathbf{W}$ is initialized with road traffic spectra, updated and the dictionary elements that are similar to the road traffic spectra are then extracted by hard thresholding.
+The study of the error according each $TIR$ reveals that this method is not the most efficient on each $TIR$, but it has to be remind that, this parameter is not available in practical case.
+TI-NMF can be considered as the approach that limit the error in the different environment.
 A major advantage of the proposed approach is that it is not designed for a specific source.
@@
 Also, performance improvement could be achieved by the addition of constraints such as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
-
-Currently, the sound database is composed of mainly composed passing car but 
-The traffic component is mainly composed of passing car event and continuous road traffic.
 

### 2018-08-22 revision/responseToreviewers.tex
add last corrections

@@
 
-$\rightarrow$
+$\rightarrow$ Done : the sentence has been rephrase with an updated references ('Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied \cite{leiba2017large,socoro_anomalous_2017}.
+As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.')
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$ We made a precision at the end of section 2 : 'The aim is then to find a flavor of NMF that could propose the minimum error on the integral corpus as it is consider that no prior knowledge on the sound sources or on the traffic predominance are available as it is when measurements are made in city.'
+$\rightarrow$ We made a precision at the end of section 2 : 'The aim is to find the form of NMF that proposes the minimal error on the reconstruction of the traffic noise signal on the whole corpus whatever the present sound sources or the predominance of the traffic as it is the case during measurements made in the city (where no prior knowledge is available).'
 
@@
 
-$\rightarrow$
+$\rightarrow$ A student t-test has been added in section 5 with the results in Table 3.
 
@@
 
-$\rightarrow$ Done 'The $A$-weightind of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
+$\rightarrow$ Done 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
 
@@
 
-$\rightarrow$ We used to number of iteration, that was not relevant to compare fairly the approaches.
-In consequence, we decided to fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
+$\rightarrow$ We used a different number of iteration, that was not relevant to compare fairly the approaches.
+In consequence, we fix at 100 the number of iterations.
+'100 iterations are performed for all NMF.
+(4.2.2)
 
@@
 
-$\rightarrow$ About the choice of third-octave bands, we specify 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+$\rightarrow$ About the choice of third-octave bands, we specify in section 4.2.2 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
 This coarser method allows us to reduce the dimensionality and then decrease the computation time.
@@
 It has been precised in the end of the section 5.
+('It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
+In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.')
 
@@
 
-$\rightarrow$
+$\rightarrow$ As lots of results have been generated by this experience, we prefer to not add all the results in tables which would be too long and not necessary.
 
@@
 In consequence, we decided to fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
+'100 iterations are performed for all NMF.
+(4.2.2)
 
@@
 $\rightarrow$ Done, 'The 10 versions of the built dictionary are then used for NMF.
-In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (4.2.1)
+In Sup-NMF, these 10 versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (4.2.1)
 

### 2018-08-22 revision/gloaguenNmfTraffic_v2.tex
add last corrections

@@
 They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
+Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010, garg_critical_2014}, despite the long data collection and calculation times.
 In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
@@
 
-Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}.
-Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap.
-Consequently, the traffic sound level estimation based on measurements is not a trivial task.
+Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied \cite{leiba2017large,socoro_anomalous_2017}.
+As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.
 
@@
 On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.
+
 Here, we extent this approach with more flavors of NMF on more complex sound scenes.
+This corpus of simulated sound scenes is created from a built-up sound database composed of a high number of diverse sound samples.
+The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
+The aim is to find the form of NMF that proposes the minimal error on the reconstruction of the traffic noise signal on the whole corpus whatever the present sound sources or the predominance of the traffic as it is the case during measurements made in the city (where no prior knowledge is available).
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
-We thus introduce another NMF scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).
-
-To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples.
-The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-The aim is then to find a flavor of NMF that could propose the minimum error on the integral corpus as it is consider that no prior knowledge on the sound sources or on the traffic predominance are available as it is when measurements are made in city.
+We thus introduce another scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).
 
@@
 The 10 versions of the built dictionary are then used for NMF.
-In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.
+In Sup-NMF, these 10 versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.
 
@@
 \label{tab:results}
-\end{table*}
-
-\begin{table*}[t]
-\centering
-\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
-\begin{tabular}{@{}cccccc@{}}
-\toprule
-\textbf{method} & filter & filter & Sup-NMF & Sem-NMF & TI-NMF \\ \midrule
-$f_c$ (kHz) & 20 & 0.5 & - & - & - \\
-$\mathbf{\beta}$ & - & - & 2 & 2 & 1 \\ \hline
-\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.39 ($\pm$ 3.00) & 8.08 ($\pm$ 2.44) & 2.98 ($\pm$ 2.11) & 5.22 ($\pm$ 2.62) \\
-\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 3.84 ($\pm$ 1.58) & 1.52 ($\pm$ 0.60)  & 2.72 ($\pm$ 1.24) \\
-\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.15 ($\pm$ 0.62) & 1.60 ($\pm$ 0.47) & 1.26 ($\pm$ 0.35) \\
-\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 0.35 ($\pm$ 0.20) & 2.49 ($\pm$ 0.30) & 0.75 ($\pm$ 0.34) \\
-\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.77 ($\pm$ 0.07) & 3.02 ($\pm$0.22) & 0.83 ($\pm$ 0.23) \\ \bottomrule
-\end{tabular}
-\label{tab:results_TIR}
 \end{table*}
@@
 
-\begin{figure}[t]
-  \centering
-	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
-	\label{fig:lp_alert}
-\end{figure}
-
-\begin{figure*}[t]
-  \centering
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 1.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
-  \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.41.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
-  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
-	\label{fig:TIR_bar}
-\end{figure*}
-
 \section{Results}\label{part:results}
@@
 As the mean error between the best SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 2 s), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 2 s) and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.41) approaches are closed, a Student's t-test is performed to evaluate the statistical differences between them.
-It is performed on all the estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
-The test estimates a $p$-value which is confronted to an significant threshold $\alpha$ = 5 $\%$ that confirms an $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
-The $p$-values are summarized in the Table \ref{student_test} with the $t$-values and the degrees of liberty (Dof).
+It is performed on the 750 estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
+The test estimates a $p$-value which is confronted to an significant threshold $\alpha$ = 5 $\%$ that proves an $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
+The $p$-values are summarized in the Table \ref{student_test} with the $t$-values and the degrees of freedom (Dof).
 
@@
 For each couple of method, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite a similar mean errors.
-To better understand these differences, the error made according the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.
+To better understand these differences, the errors made according to the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.
 
-In the case of Sup-NMF, it fails to improve the filtering performances despite good results in high $TIR$. Errors are too important for all sub-classes at low $TIR$. 
-In low $TIR$, SUP-NMF fails to estimate correctly the traffic sound level.
-This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
+\begin{table*}[t]
+\centering
+\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method.}
+\begin{tabular}{@{}cccccc@{}}
+\toprule
+\textbf{method} & filter & filter & Sup-NMF & Sem-NMF & TI-NMF \\ \midrule
+$f_c$ (kHz) & 20 & 0.5 & - & - & - \\
+$\mathbf{\beta}$ & - & - & 2 & 2 & 1 \\ \hline
+\textbf{-12} & 12.25 ($\pm$ 0.05) & 7.39 ($\pm$ 3.00) & 8.08 ($\pm$ 2.44) & 2.98 ($\pm$ 2.11) & 5.22 ($\pm$ 2.62) \\
+\textbf{-6} & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & 3.84 ($\pm$ 1.58) & 1.52 ($\pm$ 0.60)  & 2.72 ($\pm$ 1.24) \\
+\textbf{0} & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & 1.15 ($\pm$ 0.62) & 1.60 ($\pm$ 0.47) & 1.26 ($\pm$ 0.35) \\
+\textbf{6} & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & 0.35 ($\pm$ 0.20) & 2.49 ($\pm$ 0.30) & 0.75 ($\pm$ 0.34) \\
+\textbf{12} & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & 0.77 ($\pm$ 0.07) & 3.02 ($\pm$0.22) & 0.83 ($\pm$ 0.23) \\ \bottomrule
+\end{tabular}
+\label{tab:results_TIR}
+\end{table*}
+
+\begin{figure*}[t]
+  \centering
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 1.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \hspace{5pt}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.41.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
+	\label{fig:TIR_bar}
+\end{figure*}
+
+In the case of Sup-NMF, the $MAE$ errors are important for all sub-classes at low $TIR$.
+This approach reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectra.
 In the aim to reduce the objective function, see eq.
 \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the sound scene.
-Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
-At high $TIR$, this approach generate better estimation as it is a favorable case.
+Thus forcing the dictionary to be only composed of traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.
+At high $TIR$, this approach generates better estimations as it is a favorable case: the \textit{traffic} is the main component with a dictionary dedicated to this sound source.
 
-Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the interfering class in low $TIR$.
+With the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach allows a better consideration of the interfering class in low $TIR$.
 It brings a significant decrease of the errors for low $TIR$.
@@
 
-For low $TIR$, SEM-NMF outperformed as the traffic sound class is not predominant, the final dictionary strongly differs from the $\mathbf{W_0}$.
+For low $TIR$, as the traffic sound class is not predominant, the final dictionary $\mathbf{W'}$ strongly differs from $\mathbf{W_0}$.
 With the thresholding, only a reduced number of basis vectors are considered as traffic components: from the 200 basis contained in $\mathbf{W'}$, 106 are considered as \textit{traffic}  component in $TIR$ = -12 dB.
@@
 It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
-In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 $\pm$ 1.75 with in average 73 elements considered as \textit{traffic} elements.
+In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.
 
@@
 In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF are not activated when the alarm class sounds contrary to the low-pass filter.
+
+\begin{figure}[t]
+  \centering
+	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
+ \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+	\label{fig:lp_alert}
+\end{figure}
 

### 2018-08-30 revision/responseToreviewers.tex
Minor corrections

--- 2018-08-03 revision/responseToreviewers.tex
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The introduction has been enriched with the two proposed references.
 
@@
 
-$\rightarrow$
+$\rightarrow$ Done : the sentence has been rephrased with updated references :'Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied [27][46].
+As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.' (section 1, line 77-84)
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The section has been added based on the suggestion of the reviewers.
 
@@
 This error may have a dramatic impact on your results.}\\
-$\rightarrow$ First we have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values as it occurs in linear field. 
+$\rightarrow$ We have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to the high values. 
 The logarithmic mean is then more sensitive to high error than linear mean. As the errors are treated in a statistical point of view we consider that this operation is suited.
-Furthermore the use of arithmetic operators on logarithmic values is a common practice in the urban acoustics field (see Aumond et al.
+Furthermore the use of arithmetic operators on logarithmic values is a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
 "Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." and Morillas et al.
 "Uncertainty evaluation of continuous noise sampling.").
-We add a comment on the paper 'The use of a this metric is justify by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
+We added a comment on the paper: 'The use of a this metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
 The $MAE$ error with logarithmic values allows an equal weight between each difference in sound level and makes it possible to be less sensitive to high energies.
-It is also a common practice in the environmental acoustics field' (section 4.2.3)
+It is also a common practice in the environmental acoustics field' (section 4.2.3, line 539-546)
 
@@
 From these them someone can conclude that it is better to process nothing when interfering sources are 6 and 12 dB lower with respect to the traffic noise level!!}\\
-$\rightarrow$ This result may be disappointing but it has to be remind that the TIR value is unknown when measurements is made on streets.
-Then it is impossible to determine when these cases occur.
-If the LP filter is the best at fc = 20 kHz for TIR = {6,12} dB, it is not efficient when the TIR is negative.
+$\rightarrow$ This results suggests that the user does not apply any treatment in cases where road traffic predominates.
+This is not inconsistent with the purpose of this paper for two reasons: (i) a user would not call for signal processing in situations where road traffic is largely dominating such as in the vicinity of a ring road, (ii) the TIR value is an unknown parameter: the high errors of the filter fr negative TIR values discredit its use.
 We made the precision :
-'Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not know.
-In consequence, without a priori knowledge, this approach cannot be applied.' (section 5).
+'Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not know.
+In consequence, without a prior knowledge, this approach cannot be applied.' (section 5, line 566-570).
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$ We made a precision at the end of section 2 : 'The aim is then to find a flavor of NMF that could propose the minimum error on the integral corpus as it is consider that no prior knowledge on the sound sources or on the traffic predominance are available as it is when measurements are made in city.'
+$\rightarrow$ We made a precision at the end of section 2, line 177-183 : 'The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
 
@@
 
-$\rightarrow$
+$\rightarrow$ A Student t-test has been added in section 5, line 606-620 with the results in Table 3.
 
@@
 
-$\rightarrow$ Done. 'The use of simulated sound scenes allows rigorous experimental validation \dots'
+$\rightarrow$ Done. 'The use of simulated sound scenes allows rigorous experimental validation \dots' (section 2, l-174-175)
 
@@
 On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
-Car horn sound class belongs to this component as it is considered as a warning signal' (3.1)
+Car horn sound class belongs to this component as it is considered as a warning signal' (section 4.1, l 379-385).
 
@@
 $\rightarrow$ Done.
-For the test database : 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (3.1), for the train database : 'It is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.' (3.2.1)
+For the test database : 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (section 4.1 l 402-403), for the train database : 'The train database is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.' (section 4.2.1, l 434-436)
 
@@
 $\rightarrow$ Done.
-'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
+'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1 l 368-370) \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
 This is between these values that the estimator has to be the most efficient.
-But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.'
+But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.' (section 4.1, l 393-400)
 
@@
 
-$\rightarrow$ Done 'The $A$-weightind of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
+$\rightarrow$ Done 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (section 4.2.2, l 522-525)
 
@@
 
-$\rightarrow$ This work is addressed to the urban acoustic community. As NMF is not a well-known tool in this community, we prefer to keep the presentation of the fundamentals of this method.
+$\rightarrow$ This work is addressed in part to the urban acoustic community. As NMF is not a well-known tool in this community, we prefer to keep the presentation of the fundamentals of this method.
 
@@
 
-$\rightarrow$ We used to number of iteration, that was not relevant to compare fairly the approaches.
-In consequence, we decided to fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
+$\rightarrow$ We used a different number of iteration, that was not relevant to compare fairly the approaches.
+In consequence, we fix at 100 the number of iterations.
+'100 iterations are performed for all NMF.
+(section 4.2.2, l 484)
 
@@
 
-$\rightarrow$ About the choice of third-octave bands, we specify 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+$\rightarrow$ About the choice of third-octave bands, we specify in section 4.2.2 (l 484-497) 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
 This coarser method allows us to reduce the dimensionality and then decrease the computation time.
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ Done in section 4.2, l 497-500.
 'For TI-NMF, a preliminary study reveals that the threshold $t$ value range can be reduce between  0.30 and 0.70.
@@
 It is equivalent, in detection task, to reduce the number of false-positive.
-It has been precised in the end of the section 5.
+It has been precised in the end of the section 5, l 685-692.
+('It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
+In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.')
 
@@
 
-$\rightarrow$
+$\rightarrow$ As lots of results have been generated by this experience, we prefer to not add all the results in tables which might be too long.
 
@@
 
-$\rightarrow$ This part has been removed
+$\rightarrow$ This part has been removed thank to the reviewer's suggestion.
 
@@
 
-$\rightarrow$ The Tables have been fused
+$\rightarrow$ The Tables have been fused.
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The misprint is corrected.
 
 \item \emph{Fig 8.
-The y-axis range does not show the minimum values of the alarm Lp,1s lower than 50dB.
+The y-axis range does not show the minimum values of the alarm Lp,1s lower than 50 dB.
 Please, include them to better understand the impact of the alarm on the Lp computation.
@@
 
-$\rightarrow$ Done, 'In this work the non negative metric factorization framework is used \dots'
+$\rightarrow$ Done, 'In this work the non negative matrix factorization framework is used \dots'
 
@@
 
-$\rightarrow$ The paper 'Gloaguena, J.
-R., Cana, A., Lagrangeb, M., $\&$ Petiotb, J.
+$\rightarrow$ The paper 'Gloaguen, J.
+R., Can, A., Lagrange, M., $\&$ Petiot, J.
 F.
@@
 It deals with a similar approach but on a different sound corpus.
-Furthermore, it is written in french and is not suited to be cite in an international journal.
+Furthermore, as it is written in french we prefer cited other references such as 'Gloaguen, J.R., Can, A., Lagrange, M., Petiot, J.F.: Estimating traffic noise levels using acoustic monitoring: A preliminary study'.
 
@@
 $\rightarrow$ Done, 'The task is to the best of our knowledge never been considered in the literature.
-We propose…', 'Among the technical issues that bring this kind of innovative approaches,', 'several techniques based on non-negative matrix  factorization framework'.
+We propose\dots' (l 19-20), 'Among the technical issues that bring this kind of innovative approaches,\dots ' (l 12-13), 'several techniques based on non-negative matrix  factorization framework' (l 17-18).
 
@@
 $\rightarrow$ The K-mean and NMF procedure can be considered as similar approaches as a clustering tool (see Ding et al., 2005), so it has not been tested.
-A bloc diagram has been added to better explain dictionary learning.
+A bloc diagram has been added to better explain dictionary learning (Figure 7).
 
@@
 
-$\rightarrow$ Done, 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.' (Conclusion)
+$\rightarrow$ Done, 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.' (Conclusion, l 729-733)
 
@@
 Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
-But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(4.2.2)
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(section 4.2.2, l 484-497)
 
@@
 
-$\rightarrow$ Done, 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (4.2.2)
+$\rightarrow$ Done, 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (section 4.2.2, l 522-525)
 
@@
 In consequence, we decided to fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.'100 iterations are performed for all NMF.' (4.2.2)
+'100 iterations are performed for all NMF.
+(section 4.2.2, l 484)
 
@@
 
-$\rightarrow$ Done, 'The 10 versions of the built dictionary are then used for NMF.
-In Sup-NMF, these 10 versions simply correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (4.2.1)
+$\rightarrow$ Yes it is, it has been clarified, 'The 10 versions of the built dictionary are then used for NMF.
+In Sup-NMF, these 10 versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (section 4.2.1, l 477-481)
 

### 2018-08-30 revision/gloaguenNmfTraffic_v2.tex
Minor corrections

@@
 
-%\usepackage[left,modulo]{lineno}	% numéro de ligne
-%\linenumbers
+\usepackage[left,modulo]{lineno}	% numéro de ligne
+\linenumbers
 
@@
 They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010, garg_critical_2014}, despite the long data collection and calculation times.
+Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
 In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
@@
 
-Here, we extent this approach with more flavors of NMF on more complex sound scenes.
-This corpus of simulated sound scenes is created from a built-up sound database composed of a high number of diverse sound samples.
+Here, we extend this approach with more flavors of NMF on more complex sound scenes.
+This corpus of simulated sound scenes is created from a built-up sound database composed of a high number of diverse sound samples in order to encompass a wide variety of sound environment with a variable presence of road traffic noise.
 The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-The aim is to find the form of NMF that proposes the minimal error on the reconstruction of the traffic noise signal on the whole corpus whatever the present sound sources or the predominance of the traffic as it is the case during measurements made in the city (where no prior knowledge is available).
+The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
@@
 The $MAE$ is a performance index without dimension.
-The use of a this metric is justify by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
-The $MAE$ error with logarithmic values allows an equal weight between each difference in sound level and thus not being more sensitive to high energies.
-It is also a common practice in the environmental acoustics field \cite{morillas2014uncertainty} \cite{aumond2018kriging}.
+The use of such a metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
+The $MAE$ error with logarithmic values gives an equal weight to each difference in sound level and makes it possible to be less sensitive to high energies.
+It is also a common practice in the environmental acoustics field when dealing with errors  \cite{morillas2014uncertainty} \cite{aumond2018kriging}.
 
@@
 Consequently, on Table \ref{tab:results_TIR} in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
-Even if the performances for the positive $TIR$ values are high, it has to be remind that in practice, the $TIR$ value is not known.
-In consequence, without prior knowledge, this approach cannot be applied.
-Then, the traffic sound level estimation error decrease with the cut-off frequency.
+Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not know.
+In consequence, without a prior knowledge, this approach cannot be applied.
+The traffic sound level estimation error decrease with the cut-off frequency.
 Finally, $f_c = 500$ Hz is the one which reach the lowest mean error obtained ($MAE$ = 2.89 ($\pm$ 2.84)).
@@
 
-As the mean error between the best SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 2 s), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 2 s) and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.41) approaches are closed, a Student's t-test is performed to evaluate the statistical differences between them.
+As the mean error between the best SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 2 s), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 2 s) and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.41) approaches are close, a Student's t-test is performed to evaluate the statistical differences between them.
 It is performed on the 750 estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
 The test estimates a $p$-value which is confronted to an significant threshold $\alpha$ = 5 $\%$ that proves an $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
-The $p$-values are summarized in the Table \ref{student_test} with the $t$-values and the degrees of freedom (Dof).
+The $p$-values are summarized in the Table \ref{tab:student_test} with the $t$-values and the degrees of freedom (Dof).
 
@@
 \end{tabular}
-\caption{$p$-values deducted from the Student's test. In bold letters, the $p$-values that reject the $H_0$ hypothesis ($p$-value $\geq \alpha$).}
+\caption{$p$-values deducted from the Student's test. In bold letters, the $p$-values that reject the $H_0$ hypothesis ($p$-value $\leq \alpha$).}
 \label{tab:student_test}
@@
 
-For each couple of method, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite a similar mean errors.
+For each couple of method, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite similar mean errors.
 To better understand these differences, the errors made according to the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.
@@
 	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+ \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).
+The sound level of the \textit{alert} sound class is null outside the y axis range.}
 	\label{fig:lp_alert}

### 2018-08-31 revision/responseToreviewers.tex
check response to rev 1

@@
 \documentclass[10pt]{article}
+\usepackage{color}
+\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
 
@@
 \section{Answers to Reviewer 1}
+
 \subsection{Contribution and state of the art}
@@
 $\rightarrow$ The introduction has been enriched with the two proposed references.
+The differences between the present contribution and the work of Dikmen \& al and our previous work is now clearly stated.
 
@@
 
-$\rightarrow$ Done : the sentence has been rephrased with updated references :'Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied [27][46].
-As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.' (section 1, line 77-84)
+$\rightarrow$  The sentence has been rephrased with updated references (please see Section 1, line 77-84 of the updated manuscript) :'Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied [27][46].
+As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.'
 
@@
 
-$\rightarrow$ The section has been added based on the suggestion of the reviewers.
+$\rightarrow$ This section has been added as Section 2 of the updated manuscript where related work is presented and highlight the differences between those works and the approach taken in our contribution.
 
@@
 This error may have a dramatic impact on your results.}\\
-$\rightarrow$ We have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to the high values. 
-The logarithmic mean is then more sensitive to high error than linear mean. As the errors are treated in a statistical point of view we consider that this operation is suited.
-Furthermore the use of arithmetic operators on logarithmic values is a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
+$\rightarrow$ We have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values.
+The logarithmic mean is then more sensitive to high error than linear mean. As the errors are treated in a statistical point of view we consider that this operation is well suited.
+
+The use of arithmetic operators on logarithmic values is also a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
 "Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." and Morillas et al.
 "Uncertainty evaluation of continuous noise sampling.").
-We added a comment on the paper: 'The use of a this metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
+In order to position better our design choice in terms of metrics the following paragraph is now added in Section 4.2.3.
+%, line 539-546) : 'The use of a this metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
 The $MAE$ error with logarithmic values allows an equal weight between each difference in sound level and makes it possible to be less sensitive to high energies.
-It is also a common practice in the environmental acoustics field' (section 4.2.3, line 539-546)
+It is also a common practice in the environmental acoustics field'.
+
+\ml{mettre les citations completes avec une biblio a la fin}
 
@@
 We made the precision :
-'Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not know.
-In consequence, without a prior knowledge, this approach cannot be applied.' (section 5, line 566-570).
+'Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not known.
+In consequence, without any prior knowledge, this approach cannot be applied.' (section 5, line 566-570).
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$ We made a precision at the end of section 2, line 177-183 : 'The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
+$\rightarrow$ We made a precision at the end of section 2, line 177-183 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
+
+\ml{rajouter notre discussion autour des points d'usages de l'approche et modérer l'importance des point extrêmes (dans le texte et ici)}
 
@@
 
+\ml{cela nous permet de conclure quoi ?}
+
 \item \emph{Last but not least, please review the results of the low pass filter approach with fc=20kHz in Table 4 and 3, since the global averaged MAE in Table is 4.54, while in Table 3 is 4.69.}
 
-$\rightarrow$ Done. The value at TIR = 6 was a misprint, it has been corrected.
+$\rightarrow$ The value at TIR = 6 was a misprint, it has been corrected.
 
@@
 As for the comparison with (Dikmen and Mesaros, 2013), even if they use the NMF framework, they develop a detection tool that is not adapted to this study.
-We believe that it was, first, wiser to use a low-pass filter and to monitor NMF as baselines than an advanced method.
+We believe that it was wiser to use a low-pass filter as a low complexity baseline.
 
@@
 
-$\rightarrow$ Done. 'The use of simulated sound scenes allows rigorous experimental validation \dots' (section 2, l-174-175)
+$\rightarrow$ The sentence have been rephrased to : 'The use of simulated sound scenes allows rigorous experimental validation \dots' (section 2, l-174-175).
 
@@
 
-$\rightarrow$ Done.
-'What is called traffic component is the sum of the road traffic background noise and the sound events generated by the passing car class.
+$\rightarrow$ This important point is now clarified in the manuscript: 'What is called traffic component is the sum of the road traffic background noise and the sound events generated by the passing car class.
 On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
@@
 
-$\rightarrow$ Done.
-For the test database : 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' (section 4.1 l 402-403), for the train database : 'The train database is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.' (section 4.2.1, l 434-436)
+$\rightarrow$ Those numbers are now added to the manuscript.
+For the test database, see section 4.2.1: 'each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.' and for the train database, see section 4.1: 'The train database is composed of 53 audio files of isolated passing cars with a 18 minutes cumulative duration.'
+
+\ml{enlève les -references par lignes ca peut bouger avec les edits et ce n'est pas tres utile}
 
@@
 
-$\rightarrow$ Done.
-'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1 l 368-370) \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
+$\rightarrow$ the sentence is now rephrased in section 4.1 to: 'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1 l 368-370) \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
 This is between these values that the estimator has to be the most efficient.
-But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.' (section 4.1, l 393-400)
+But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.'
 
@@
 
-$\rightarrow$ Done 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (section 4.2.2, l 522-525)
+$\rightarrow$ This choice is now explained in section 4.2.2: 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.'
 
@@
 
-$\rightarrow$ This work is addressed in part to the urban acoustic community. As NMF is not a well-known tool in this community, we prefer to keep the presentation of the fundamentals of this method.
+$\rightarrow$ Our contribution is targeted in part to the urban acoustic community.
+As NMF is not a well-known tool in this community, we thus prefer to keep the presentation of the fundamentals of this method.
 
@@
 
-$\rightarrow$ We used a different number of iteration, that was not relevant to compare fairly the approaches.
-In consequence, we fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.
-(section 4.2.2, l 484)
+$\rightarrow$ We now consider 100 as the number of iterations for each method, see Section 4.2.2: '100 iterations are performed for all NMF.
 
@@
 
-$\rightarrow$ Done in section 4.2, l 497-500.
-'For TI-NMF, a preliminary study reveals that the threshold $t$ value range can be reduce between  0.30 and 0.70.
+$\rightarrow$ This matter is now discussed in Section 4.2.
+'For TI-NMF, a preliminary study showed that the threshold $t$ value range can be evaluated between 0.30 and 0.70.
 An increment step of 0.01 has been considered as being sufficiently precise.'
@@
 
-$\rightarrow$ This improvement is due to a decreasing of the number of element considered as traffic component.
+$\rightarrow$ This improvement is due to a decreasing of the number of elements considered as traffic components.
 As in TIR = -12 dB, the interfering class is more present, it takes more basis in W'.
@@
 It is equivalent, in detection task, to reduce the number of false-positive.
-It has been precised in the end of the section 5, l 685-692.
-('It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
-In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.')
+It is now discussed further in the end of the section 5: 'It would be possible to decrease the error by adapting the threshold according to the $TIR$.
+For instance at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
+In the contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.'
 
@@
 $\rightarrow$ As lots of results have been generated by this experience, we prefer to not add all the results in tables which might be too long.
+
+\ml{a rediscuter, je ne vois pas de quoi on parle}
 
@@
 
-$\rightarrow$ The misprint is corrected.
+$\rightarrow$ The misprint is now corrected.
 
@@
 
-$\rightarrow$ The car horn sound level is null outside the range of the y-axis that's why we put aside the the low sound level to weep a wide range to see, at best, the curves.
-We made the precision on the legend.
+$\rightarrow$ The car horn sound level is zero outside the range of the y-axis.
+That is why we put aside the low sound level to weep a wide range to see, at best, the curves.
+This is now detailed in the caption.
+
+\ml{pas clair}
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The caption is corrected.
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The typo is corrected.
 
@@
 
-$\rightarrow$ Done, 'Usually in unsupervised learning,…'
+$\rightarrow$ The sentence is now: 'Usually in unsupervised learning,…'
 
@@
 
-$\rightarrow$ Done, 'First, the errors calculated from the traffic sound estimation of the low-pass filter are detailed'
+$\rightarrow$ The sentence is now rephrased to: 'First, the errors calculated from the traffic sound estimation of the low-pass filter are detailed'
 
@@
 
-$\rightarrow$ Done, 'the traffic elements selected in TI-NMF are not activated'
+$\rightarrow$ The sentence is now rephrased to: 'the traffic elements selected in TI-NMF are not activated'
 
@@
 
-$\rightarrow$ Done, 'In this work the non negative matrix factorization framework is used \dots'
+$\rightarrow$ The sentence is now rephrased to: 'In this work the non negative matrix factorization framework is used \dots'
 

### 2018-08-31 revision/gloaguenNmfTraffic_v2.tex
check response to rev 1

@@
 \renewcommand{\baselinestretch}{1.025}
+
+\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
 
@@
 
-\section{Related works}
-Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
+\section{Related work}
+
+Recent work is mostly focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
 A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
@@
 On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.
+
+\ml{il y a erreur sur la ref demandée par le reviewer ca doit etre \cite{dikmen2013sound} !!!}
 
@@
 But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.
-For TI-NMF, a preliminary study reveals that the threshold $t$ value range can be reduce between  0.30 and 0.70.
+For TI-NMF, a preliminary study showed that the threshold $t$ value range can be evaluated between  0.30 and 0.70.
 An increment step of 0.01 has been considered as being sufficiently precise.
@@
 Consequently, on Table \ref{tab:results_TIR} in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
-Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not know.
-In consequence, without a prior knowledge, this approach cannot be applied.
+Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not known.
+In consequence, without any prior knowledge, this approach cannot be applied.
 The traffic sound level estimation error decrease with the cut-off frequency.
@@
 
-It would be possible to decrease the error by adapting the threshold according to the $TIR$ for instance: at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
-In the opposite, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.
+It would be possible to decrease the error by adapting the threshold according to the $TIR$.
+For instance at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
+In the contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.
 

### 2018-08-31 revision/responseToreviewers.tex
done with proof rerad of response

@@
 It deals with a similar approach but on a different sound corpus.
-Furthermore, as it is written in french we prefer cited other references such as 'Gloaguen, J.R., Can, A., Lagrange, M., Petiot, J.F.: Estimating traffic noise levels using acoustic monitoring: A preliminary study'.
+Furthermore, as it is written in french we chose to alternatively cite 'Gloaguen, J.R., Can, A., Lagrange, M., Petiot, J.F.: Estimating traffic noise levels using acoustic monitoring: A preliminary study' which is written in English and deals with the same corpus.
 
@@
 
-$\rightarrow$ Done, 'The task is to the best of our knowledge never been considered in the literature.
+$\rightarrow$ The bastract have been rewritten to take those requests into account: 'The task is to the best of our knowledge never been considered in the literature.
 We propose\dots' (l 19-20), 'Among the technical issues that bring this kind of innovative approaches,\dots ' (l 12-13), 'several techniques based on non-negative matrix  factorization framework' (l 17-18).
@@
 
-$\rightarrow$ The K-mean and NMF procedure can be considered as similar approaches as a clustering tool (see Ding et al., 2005), so it has not been tested.
-A bloc diagram has been added to better explain dictionary learning (Figure 7).
+$\rightarrow$ The K-means and NMF procedures can be considered as similar approaches for a clustering tool (see Ding et al., 2005), so it has not been tested.
+A block diagram has been added to better explain the dictionary learning process shown in Figure 7 of the updated manuscript.
 
@@
 
-$\rightarrow$ Done, 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.' (Conclusion, l 729-733)
+$\rightarrow$ This important limitation of the presented study is now clearly stated as future work of interest in the conclusion : 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, and other noisy traffic elements which are not considered in this study.'
 
@@
 
-$\rightarrow$ The purpose of this study is to propose a tool that could be integrated into a sensor network. The distribution of the microphone is not a question dealt with in this work.
+$\rightarrow$ The purpose of this study is to propose a tool that could be integrated into a sensor network.
+The distribution of the microphone is not a question dealt with in this work.
+The proposed method is designed to perform with a monophonic input.
+That being said, considering the inputs of multiple sensors may help to achieve more consistent estimates.
+This is left for potential future work.
 
@@
 
-$\rightarrow$ Done, 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+$\rightarrow$ The manuscript have been updated, see Section 4.2.2: 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
 This coarser method allows us to reduce the dimensionality and then decrease the computation time.
@@
 Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
-But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'(section 4.2.2, l 484-497)
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compared to MFCCs for instance.'
 
@@
 
-$\rightarrow$ Done, 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.' (section 4.2.2, l 522-525)
+$\rightarrow$ As now explicitly stated in Section 4.2.2, 'The $A$-weighting of the sound levels is not considered here as it decreases the low frequencies levels where the road traffic components are mainly present.'
 
@@
 
-$\rightarrow$ The title has been corrected ('Evolution of 1 second equivalent sound pressure levels'). The calibration step is here not necessary as all the sound level are relatives.
+$\rightarrow$ The caption of the Figure 8 has been corrected ('Evolution of 1 second equivalent sound pressure levels').
+The calibration step is here not necessary as all the sound level are relatives.
 
@@
 
-$\rightarrow$ We used to number of iteration, that was not relevant to compare fairly the approaches.
-In consequence, we decided to fix at 100 the number of iterations.
-'100 iterations are performed for all NMF.
-(section 4.2.2, l 484)
+$\rightarrow$ As stated in Section 4.2.2, 100 iterations are performed for every NMF types.
 
@@
 
-$\rightarrow$ Yes it is, it has been clarified, 'The 10 versions of the built dictionary are then used for NMF.
-In Sup-NMF, these 10 versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.' (section 4.2.1, l 477-481)
+$\rightarrow$ Yes it is, it has been clarified in the updated version of the manuscript in Section 4.2.1: 'The 10 versions of the built dictionary are then used for NMF.
+In Sup-NMF, these 10 versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial dictionaries $\mathbf{W_0}$ that are next updated.'
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The sentence is rewritten.
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ Fixed.
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ Indeed, fixed.
 
@@
 
-$\rightarrow$ Done.
+$\rightarrow$ The reference \cite{fevotte_nonnegative_2009} is added.
 
@@
 
-$\rightarrow$ Done, 'Errors are too important for all sub-classes at low $TIR$.'
+$\rightarrow$ The sentence is rewritten to: 'Errors are too important for all sub-classes at low $TIR$.'
 
 \end{enumerate}
+
+\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+\bibliography{bibliographie_jasm}
+
 \end{document}

### 2018-08-31 revision/gloaguenNmfTraffic_v2.tex
done with proof rerad of response

@@
 
-100 iterations are performed for all NMF.
+100 iterations are performed for every NMF types.
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
@@
 Experimental validation consistently showed that considering third octave bands do not impact the performance of the estimator studied in this paper.
-But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.
+But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compared to MFCCs for instance.
 For TI-NMF, a preliminary study showed that the threshold $t$ value range can be evaluated between  0.30 and 0.70.
@@
 A major advantage of the proposed approach is that it is not designed for a specific source.
-It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, noisy traffic elements which are absent here.
+It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, and other noisy traffic elements which are not considered in this study.
 However, even though the experiments described in this paper focused on road traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.

### 2018-09-05 revision/responseToreviewers.tex
proofread after ML improvement

@@
 $\rightarrow$ The introduction has been enriched with the two proposed references.
-The differences between the present contribution and the work of Dikmen \& al and our previous work is now clearly stated.
+The differences between the present contribution and the work of Dikmen \& al \cite{dikmen2013sound} and our previous work \cite{gloaguen2016estimating} is now clearly stated.
+We also added the work of Mesaros \& al \cite{mesaros_sound_2015}
 
@@
 
-$\rightarrow$  The sentence has been rephrased with updated references (please see Section 1, line 77-84 of the updated manuscript) :'Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied [27][46].
+$\rightarrow$  The sentence has been rephrased with updated references (please see Section 1) :'Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied \cite{leiba2017large}\cite{socoro_anomalous_2017}.
 As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.'
@@
 The use of arithmetic operators on logarithmic values is also a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
-"Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." and Morillas et al.
-"Uncertainty evaluation of continuous noise sampling.").
-In order to position better our design choice in terms of metrics the following paragraph is now added in Section 4.2.3.
+"Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." \cite{aumond2018kriging}  and Morillas et al.
+"Uncertainty evaluation of continuous noise sampling." \cite{morillas2014uncertainty}).
+In order to  better position our design choice in terms of metrics a paragraph is now added in Section 4.2.3.
 %, line 539-546) : 'The use of a this metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
@@
 It is also a common practice in the environmental acoustics field'.
-
-\ml{mettre les citations completes avec une biblio a la fin}
 
@@
 'Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not known.
-In consequence, without any prior knowledge, this approach cannot be applied.' (section 5, line 566-570).
+In consequence, without any prior knowledge, this approach cannot be applied.' (please see Section 5).
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$ We made a precision at the end of section 2, line 177-183 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
+$\rightarrow$ We made a precision at the end of section 2 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
 
-\ml{rajouter notre discussion autour des points d'usages de l'approche et modérer l'importance des point extrêmes (dans le texte et ici)}
+In addition, in the urban sound environments, TIRs are mainly focused in the range $\left[-6, 12\right]$ dB \cite{gloaguen_creation_2017}. 
+$TIR$ = -12 dB is a rarer case in urban environment but it is still added in order to test the limits of the methods.
+Its weight on the reconstruction error must still be moderate.
+This has been precised at the end of section 4.1.
 
@@
 
-$\rightarrow$ A Student t-test has been added in section 5, line 606-620 with the results in Table 3.
-
-\ml{cela nous permet de conclure quoi ?}
+$\rightarrow$ As suggested by the reviewers, we added a Student t-test in section 5 with the results exposed in Table 3.
+This test revealed that the distribution of the traffic sound level estimations between the best methods exposed in Table 2 are statistically different.
 
@@
 
-$\rightarrow$ The sentence have been rephrased to : 'The use of simulated sound scenes allows rigorous experimental validation \dots' (section 2, l-174-175).
+$\rightarrow$ The sentence have been rephrased to : 'The use of simulated sound scenes allows rigorous experimental validation \dots' (section 2).
 
@@
 On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
-Car horn sound class belongs to this component as it is considered as a warning signal' (section 4.1, l 379-385).
+Car horn sound class belongs to this component as it is considered as a warning signal' (section 4.1).
 
@@
 
-\ml{enlève les -references par lignes ca peut bouger avec les edits et ce n'est pas tres utile}
-
 \item \emph{Finally, it is worth mentioning that considering TIR={-6,-12,0,6,12} may yield to conclusions far from what can be found in real-life environments, as discussed in the literature (e.g.
@@
 
-$\rightarrow$ the sentence is now rephrased in section 4.1 to: 'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1 l 368-370) \dots 'In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB [16].
-This is between these values that the estimator has to be the most efficient.
-But, in this experiment, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.'
+$\rightarrow$ the sentence is now rephrased in section 4.1 to: 'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1) \dots 
+In addition the precision provided at the end of this section on the range values of the TIR complete the reviewer's remark.
 
@@
 
-$\rightarrow$ We now consider 100 as the number of iterations for each method, see Section 4.2.2: '100 iterations are performed for all NMF.
+$\rightarrow$ We now consider 100 as the number of iterations for each method, see Section 4.2.2: '100 iterations are performed for every NMF types'.
 
@@
 
-$\rightarrow$ About the choice of third-octave bands, we specify in section 4.2.2 (l 484-497) 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
+$\rightarrow$ About the choice of third-octave bands, we specify in section 4.2.2 'The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
 This coarser method allows us to reduce the dimensionality and then decrease the computation time.
@@
 But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compare to MFCC for instance.'.
-Also, we add a block diagram which details the dictionary building steps and makes clearer these different elements.
+Also, we add a block diagram which details the dictionary building and makes clearer these different steps (please see Figure 7).
 
@@
 
-$\rightarrow$ This matter is now discussed in Section 4.2.
-'For TI-NMF, a preliminary study showed that the threshold $t$ value range can be evaluated between 0.30 and 0.70.
+$\rightarrow$ This matter is now discussed in Section 4.2.2.
+'For TI-NMF, a preliminary study showed that  the range of threshold values can be evaluated between 0.30 and 0.70.
 An increment step of 0.01 has been considered as being sufficiently precise.'
@@
 
-$\rightarrow$ The car horn sound level is zero outside the range of the y-axis.
-That is why we put aside the low sound level to weep a wide range to see, at best, the curves.
-This is now detailed in the caption.
-
-\ml{pas clair}
+$\rightarrow$ Outside these emergences, the car horn sound level is null, that is why we limited the $y$-axis range to [45 85] dB to see, at best, the curves.
+In addition, we followed the reviewer's recommendation by adding the spectrogram of the audio scene where the car horn sounding is visible (please see Figure 11).
 
@@
 
-$\rightarrow$ The caption is corrected.
+$\rightarrow$ As pointed by the reviewer, the caption title has been fixed.
 
@@
 
-$\rightarrow$ The bastract have been rewritten to take those requests into account: 'The task is to the best of our knowledge never been considered in the literature.
-We propose\dots' (l 19-20), 'Among the technical issues that bring this kind of innovative approaches,\dots ' (l 12-13), 'several techniques based on non-negative matrix  factorization framework' (l 17-18).
+$\rightarrow$ The abstract have been rewritten to take those requests into account: 'The task is to the best of our knowledge never been considered in the literature.
+We propose\dots', 'Among the technical issues that bring this kind of innovative approaches,\dots ', 'several techniques based on non-negative matrix  factorization framework'.
 
@@
 
-$\rightarrow$ This important limitation of the presented study is now clearly stated as future work of interest in the conclusion : 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, and other noisy traffic elements which are not considered in this study.'
+$\rightarrow$ This important limitation of the presented study is now clearly stated as future work of interest in the conclusion : 'It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles and other noisy traffic elements which are not considered in this study.'
 
@@
 
-$\rightarrow$ As stated in Section 4.2.2, 100 iterations are performed for every NMF types.
+$\rightarrow$ This remark has been considered: to better compare the different NMF version, 100 iterations are performed for every NMF types.
+Please see section 4.2.2 ('100 iterations are performed for every NMF types.')
 
@@
 
-\item \emph{•Figure 2 caption, please change to "of an urban", "correspond to".}
+\item \emph{Figure 2 caption, please change to "of an urban", "correspond to".}
 

### 2018-09-05 revision/gloaguenNmfTraffic_v2.tex
proofread after ML improvement

@@
 We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
-The results show the interest of our proposed approach for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods.
+The results show the interest of our proposed approach for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods from a reconstruction error of 2.89 to 2.15.
 
@@
 
-Closer to our application scenario, NMF has been considered in \cite{mesaros_sound_2015} where coupled NMF has been used as a sound event detection in real life recordings.
-However, no information on the quality of the detected signal is mentioned.
-Nevertheless, they raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary?
+Closer to our application scenario, NMF has been considered in \cite{dikmen2013sound} where coupled NMF is used as a sound event detector.
+The dictionary learning is based on the spectrogram of audio recordings and on their annotations which share the same activation matrix.
+If the performance is satisfying (especially for non-negative SNRs), this approach has the advantage of allowing the use of audio recordings consisting of multiple and overlapping sound sources and reducing the ambiguity of unsupervised learning.
+However, if the sound event can be well detected, no information on the quality of the detected signal is mentioned.
+Similarly, in \cite{mesaros_sound_2015}, NMF is used in the same task but with real life recordings.
+The authors raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary?
 They conclude on the efficiency of each approach and suggere to combine both to deal with large databases.
@@
 On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.
-
-\ml{il y a erreur sur la ref demandée par le reviewer ca doit etre \cite{dikmen2013sound} !!!}
 
@@
 In each file, the traffic component is present.
-What is called traffic component is the sum of the road traffic background noise and the sound events generated by the passing car class.
+What is called traffic component is the sum of the road traffic background noise and the sound events generated by the\textit{passing car} class.
 On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
-Car horn sound class belongs to this component as it is considered as a warning signal.
+Car horn sound class belongs to the \textit{interfering} component as it is considered as a warning signal.
 To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the interfering class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).
@@
 
-with $TIR$, the \textit{Traffic Interference Ratio}.
+with $TIR$, the \textit{Traffic Interference Ratio} in a similar way to \cite{dikmen2013sound}.
 When $TIR < 0$ dB, the traffic component is less present than the interfering class.
 On the contrary, for $TIR > 0$ dB, the traffic class is louder than the interfering class.
-In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 6 dB \cite{gloaguen_creation_2017}.
+In most of the urban sound environments, the ratio between the interfering class and the traffic is mainly included between $TIR$ = -6 dB and $TIR$ = 12 dB \cite{gloaguen_creation_2017}.
 This is between these values that the estimator has to be the most efficient.
-But, in this experimental framework, this frame is extended to $TIR$ = -12 dB and $TIR$ = 12 dB to test the limit of NMF.
+But, in this experimental framework, this frame is extended to $TIR$ = -12 dB to test the limit of NMF.
 The total number of scenes designed is then 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values), each scene during 30 seconds, it leads to a full duration of 6 hours and 30 minutes.
@@
 But, most of all, it is a suited representation to this sound environment as this kind of representation is widely used in the urban acoustic field, compared to MFCCs for instance.
-For TI-NMF, a preliminary study showed that the threshold $t$ value range can be evaluated between  0.30 and 0.70.
+For TI-NMF, a preliminary study showed that the range of threshold values can be evaluated between 0.30 and 0.70.
 An increment step of 0.01 has been considered as being sufficiently precise.
@@
 But, in order to generalize this method to a practical case where no prior knowledge on the urban environment is made, the chosen threshold $t$ is then fixed to $t$ = 0.41 as it is the one that best balanced these opposite cases.
-In Figure \ref{fig:lp_alert}, the 1 second equivalent sound pressure level of an alert scene is displayed for $TIR \in \lbrace -12,12 \rbrace$ dB for the 500 Hz low pass filter and for TI-NMF.
+In Figure \ref{fig:specLp}, the spectrogram of an alert scene is displayed ($TIR$ = 0 dB).
+The evolution of the 1 second equivalent sound pressure level for $TIR \in \lbrace -12,12 \rbrace$ dB for the 500 Hz low pass filter and for TI-NMF can be seen on Figure \ref{fig:lp_alert}.
 In the case of $TIR=-12$ dB, the traffic elements selected in TI-NMF are not activated when the alarm class sounds contrary to the low-pass filter.
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
-	\includegraphics[width=\textwidth]{figures/NMF_Lp_alert.pdf}
- \caption{Evolution of 1 second equivalent sound pressure level of an \textit{alert} sub-class scene for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).
-The sound level of the \textit{alert} sound class is null outside the y axis range.}
-	\label{fig:lp_alert}
-\end{figure}
+  \subfloat[]{\label{fig:specLp}\includegraphics[width=.45\linewidth]{figures/specAlert.pdf}}
+%  \hspace{5pt}
+  \subfloat[]{\label{fig:lp_alert}\includegraphics[width=.45\linewidth]{figures/NMF_Lp_alert.pdf}}
+  \caption{Spectrogram of an \textit{alert} sub-class scene (\ref{fig:specLp}) and evolution of 1 second equivalent sound pressure level (\ref{fig:lp_alert}) for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}
+  \label{fig:spec_alert}
+\end{figure*}
 
@@
 A major advantage of the proposed approach is that it is not designed for a specific source.
-It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles, and other noisy traffic elements which are not considered in this study.
+It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles and other noisy traffic elements which are not considered in this study.
 However, even though the experiments described in this paper focused on road traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.

### 2018-09-06 revision/responseToreviewers.tex
last modifications

@@
 $\rightarrow$ This results suggests that the user does not apply any treatment in cases where road traffic predominates.
-This is not inconsistent with the purpose of this paper for two reasons: (i) a user would not call for signal processing in situations where road traffic is largely dominating such as in the vicinity of a ring road, (ii) the TIR value is an unknown parameter: the high errors of the filter fr negative TIR values discredit its use.
+This is not inconsistent with the purpose of this paper for two reasons: (i) a user would not call for signal processing in situations where road traffic is largely dominating such as in the vicinity of a ring road, (ii) the TIR value is an unknown parameter: the high errors of the filter for negative TIR values discredit its use.
 We made the precision :
 'Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not known.
-In consequence, without any prior knowledge, this approach cannot be applied.' (please see Section 5).
+In consequence, without any prior knowledge, this approach cannot be applied.' (please see the beginning of Section 5).
 
@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
-$\rightarrow$ We made a precision at the end of section 2 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.'
+$\rightarrow$ We made a precision at the end of section 2 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus to be applied to all the different possible urban noise environments.'
 
@@
 
-$\rightarrow$ As suggested by the reviewers, we added a Student t-test in section 5 with the results exposed in Table 3.
+$\rightarrow$ As suggested by the reviewers, we added a Student t-test in Section 5 with the results exposed in Table 3.
 This test revealed that the distribution of the traffic sound level estimations between the best methods exposed in Table 2 are statistically different.
@@
 On the contrary, the \textit{interfering} sound class includes all the other sound sources not related to it.
-Car horn sound class belongs to this component as it is considered as a warning signal' (section 4.1).
+Car horn sound class belongs to this component as it is considered as a warning signal' (Section 4.1).
 

### 2018-09-06 revision/gloaguenNmfTraffic_v2.tex
last modifications

@@
 We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
-The results show the interest of our proposed approach for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods from a reconstruction error of 2.89 to 2.15.
+The results show the interest of our proposed approach (a NMF based on the adjustment of a learned dictionary directly on the urban sound mixtures) for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods such as a low-pass filter or supervised NMF.
 
@@
 The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus.
+The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus to be applied to all the different possible urban noise environments.
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
@@
 
-Table \ref{tab:results} summarizes, according to the 2 main factors (\textit{method}, $\beta$), the lowest $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).
-For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
-\\
-
 \begin{figure}[t]
@@
 
+Table \ref{tab:results} summarizes, according to the 2 main factors (\textit{method}, $\beta$), the lowest $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all) for NMF.
+For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}.
 Compared with the filter errors, the choice of some NMF approaches makes it possible to decrease the error of the road traffic sound level estimation.

### 2018-09-06 revision/responseToreviewers.tex
proof read rev 1

@@
 
-$\rightarrow$ This section has been added as Section 2 of the updated manuscript where related work is presented and highlight the differences between those works and the approach taken in our contribution.
+$\rightarrow$ This section has been added as Section 2 of the updated manuscript where related work is now presented and where we highlight the differences between those works and the approach taken in our contribution.
 
@@
 This error may have a dramatic impact on your results.}\\
-$\rightarrow$ We have considered logarithmic value in dB as it allows the calibration of the traffic sound level and being less sensitive to high values.
-The logarithmic mean is then more sensitive to high error than linear mean. As the errors are treated in a statistical point of view we consider that this operation is well suited.
+$\rightarrow$
 
-The use of arithmetic operators on logarithmic values is also a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
+The use of arithmetic operators on logarithmic values is a common practice in the urban acoustics field when dealing with errors (see Aumond et al.
 "Kriging-based spatial interpolation from measurements for sound level mapping in urban areas." \cite{aumond2018kriging}  and Morillas et al.
@@
 In order to  better position our design choice in terms of metrics a paragraph is now added in Section 4.2.3.
+
+We acknowledge that strictly physically speaking this might be considered as a mistake but we chose to follow this approach as we believe it allows the calibration of the traffic sound level and  to monitor metrics that are less sensitive to high values.
+As the errors are treated in a statistical point of view we consider that this operation is well suited.
+
  %, line 539-546) : 'The use of a this metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
@@
 
-In addition, in the urban sound environments, TIRs are mainly focused in the range $\left[-6, 12\right]$ dB \cite{gloaguen_creation_2017}. 
-$TIR$ = -12 dB is a rarer case in urban environment but it is still added in order to test the limits of the methods.
-Its weight on the reconstruction error must still be moderate.
+Furthermore, as now explained at the end of section 4.1, in urban sound environments, TIRs are mainly focused in the range $\left[-6, 12\right]$ dB \cite{gloaguen_creation_2017}.
+$TIR$ = -12 dB is a rarer case in urban environment but it is still added in order to show the limits of the evaluated methods.
+%Its weight on the reconstruction error must still be moderate.
 This has been precised at the end of section 4.1.
@@
 As in the environmental acoustics community, this method would have been easily seen as a tool, we decided to consider it.
-We summarized all the results for the low pass filter according to the cut-off frequency in order to be more precise (Figure 9).
+We summarized all the results for the low pass filter according to the cut-off frequency in order to be more complete (Figure 9).
 As for the comparison with (Dikmen and Mesaros, 2013), even if they use the NMF framework, they develop a detection tool that is not adapted to this study.
@@
 
-$\rightarrow$ the sentence is now rephrased in section 4.1 to: 'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' (section 4.1) \dots 
-In addition the precision provided at the end of this section on the range values of the TIR complete the reviewer's remark.
+$\rightarrow$ The sentence is now rephrased in section 4.1 to: 'This database allows the creation of a wide diversity of urban sound scenes from the road traffic point of view' \dots
+In addition, at the end of section 4.1 the motivation behind the choice of the range of TIR values  is now discussed in depth, highlighting the fact that, indeed, the range is definitely wider than what can be found in realistic settings.
+The results are also now discussed by considering this matter into account.
 
@@
 
-\subsection{Minors comment}
+\subsection{Minors comments}
 
@@
 
-$\rightarrow$ This improvement is due to a decreasing of the number of elements considered as traffic components.
+$\rightarrow$ This improvement is due to a decrease of the number of elements considered as traffic components.
 As in TIR = -12 dB, the interfering class is more present, it takes more basis in W'.
 By increasing the threshold, we put aside more elements from W' that can belong to this sound class.
-It is equivalent, in detection task, to reduce the number of false-positive.
+It is equivalent, in detection task, to a reduction of the number of false-positive.
 It is now discussed further in the end of the section 5: 'It would be possible to decrease the error by adapting the threshold according to the $TIR$.
@@
 
-$\rightarrow$ As lots of results have been generated by this experience, we prefer to not add all the results in tables which might be too long.
-
-\ml{a rediscuter, je ne vois pas de quoi on parle}
+$\rightarrow$ As lots of results have been generated by those experiments, we chose to detail only some of the parameters that may influence the performance of the studied algorithms.
 
@@
 
-$\rightarrow$ This part has been removed thank to the reviewer's suggestion.
+$\rightarrow$ This part has been removed.
 
@@
 
-$\rightarrow$ As pointed by the reviewer, the caption title has been fixed.
+$\rightarrow$ The caption title has been fixed.
 

### 2018-09-06 revision/responseToreviewers.tex
done with response

@@
 $\rightarrow$ The abstract have been rewritten to take those requests into account: 'The task is to the best of our knowledge never been considered in the literature.
-We propose\dots', 'Among the technical issues that bring this kind of innovative approaches,\dots ', 'several techniques based on non-negative matrix  factorization framework'.
+We propose\dots', 'Among the technical issues that bring this kind of innovative approaches,\dots ', 'several techniques based on  the non-negative matrix  factorization framework'.
 
@@
 
-$\rightarrow$ The K-means and NMF procedures can be considered as similar approaches for a clustering tool (see Ding et al., 2005), so it has not been tested.
-A block diagram has been added to better explain the dictionary learning process shown in Figure 7 of the updated manuscript.
+$\rightarrow$ A block diagram has been added to better explain the dictionary learning process shown in Figure 7 of the updated manuscript.
+Performing an NMF training have also been tested with no significant improvements.
 
@@
 
-$\rightarrow$ The caption of the Figure 8 has been corrected ('Evolution of 1 second equivalent sound pressure levels').
-The calibration step is here not necessary as all the sound level are relatives.
+$\rightarrow$ The caption of the Figure 8 has been corrected to: 'Evolution of 1 second equivalent sound pressure levels'.
+The calibration step is here not necessary as all the sound levels are  set relatively.
 
@@
 
-$\rightarrow$ This remark has been considered: to better compare the different NMF version, 100 iterations are performed for every NMF types.
-Please see section 4.2.2 ('100 iterations are performed for every NMF types.')
+$\rightarrow$ To better compare the different NMF version, 100 iterations are now performed for every NMF types, see section 4.2.2.
 

### 2018-09-06 revision/gloaguenNmfTraffic_v2.tex
done with response

@@
 
-In this paper, several techniques based on non-negative matrix  factorization framework are studied in this application scenario on a simulated sound scene corpus.
+In this paper, several techniques based on the non-negative matrix  factorization framework are studied in this application scenario, on a corpus of simulated sound scenes.
 The task is to the best of our knowledge never been considered in the literature.

### 2018-09-06 revision/gloaguenNmfTraffic_v2.tex
an-bs and intro

@@
 We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.
-The results show the interest of our proposed approach (a NMF based on the adjustment of a learned dictionary directly on the urban sound mixtures) for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods such as a low-pass filter or supervised NMF.
+The results show the interest of our proposed approach %(a NMF based on the adjustment of a learned dictionary directly on the urban sound mixtures)
+for such sound environments.% as it improves the estimation of the road traffic sound level compared to basic methods such as a low-pass filter or supervised NMF.
 
@@
 They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively.
-Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times.
-In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
+Although very useful, the produced noise maps exhibit a large amount of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}.
+It also requires a labor intensive data collection and calculation times.
+The result is usually that road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
 The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}.
@@
 
-Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied \cite{leiba2017large,socoro_anomalous_2017}.
-As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.
+Prior to data assimilation, the issue is to correctly estimate the traffic sound level from acoustic measurements as studied in \cite{leiba2017large,socoro_anomalous_2017}.
+As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the estimation of the traffic sound level  based on acoustic measurements is not a trivial task.
 
@@
 Recent work is mostly focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}.
-A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks).
-A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
+A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Neural Networks) \cite{cowling_comparison_2003} and can be applied to a variety of tasks \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.
 
-The main issue in the detection or recognition tasks is the overlap of environmental sounds.
+The main issue with the detection or recognition tasks is the overlap of environmental sounds.
 Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels.
@@
 It consists in separating the contribution of the traffic from the other sources within a polyphonic scene.
-One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\
+One major advantage of following such approach is that the estimate is  available continuously, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\
 
 In an urban environment context, source separation can be achieved with the help of acoustic microphone arrays and beamforming \cite{saruwatari2003blind}.
-However, this approach requires spreading multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance.
-This method is then not considered here to be deployed all over cities.
-On the contrary, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present.
+However, this approach requires the spreading of multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance.
+Monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the level of the road traffic from a sound mixture in which all kind of sound sources can be present.
 A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}.
@@
 In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}.
-By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be resolved on the time/frequency plane.
+By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be mostly resolved on the time/frequency plane.
 
@@
 The authors raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary?
-They conclude on the efficiency of each approach and suggere to combine both to deal with large databases.
+They conclude on the efficiency of each approach and propose to combine both to deal with large databases.
 Also, Innami and Kasai \cite{satoshi_innami_nmf-based_2012} proposed to perform NMF on simulated audio files as a source separation method.
@@
 
-Here, we extend this approach with more flavors of NMF on more complex sound scenes.
-This corpus of simulated sound scenes is created from a built-up sound database composed of a high number of diverse sound samples in order to encompass a wide variety of sound environment with a variable presence of road traffic noise.
+Here, we extend this approach with more flavors of NMF on more challenging sound scenes.
+The corpus of simulated sound scenes considered here is created from a built-up sound database composed of a large number of diverse sound samples in order to encompass a wide variety of sound environment with a variable presence of road traffic noise.
 The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus to be applied to all the different possible urban noise environments.
-We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand.
+The aim is to find the type of NMF that gives the minimal error on the estimation of the traffic noise signal level on the whole corpus.% to be applied to all the different possible urban noise environments.
+We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but do not give satisfactory results for the application at hand.
 We thus introduce another scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).

### 2018-09-07 revision/gloaguenNmfTraffic_v2.tex
up to results

@@
 where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectra and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$.
-As the constraint of non-negativity of $\mathbf{W}$ and $\mathbf{H}$ is considered, NMF allows only additive combinations between the element of $\mathbf{W}$.
-It is then a part-based representation that NMF provides.
+As the constraint of non-negativity of $\mathbf{W}$ and $\mathbf{H}$ is considered, NMF allows only additive combinations between the element of $\mathbf{W}$, thus inducing a part-based representation.
 An illustrative example can be found in Figure \ref{fig:example_NMF}.
@@
 
-To better take into account prior knowledge about the sources of interest, constraints (like the smoothness or the sparsness criteria \cite{virtanen_monaural_2007}) can be added to the objective function.
+To better take into account prior knowledge about the sources of interest, constraints (like the smoothness or the sparsity criteria \cite{virtanen_monaural_2007}) can be added to the objective function.
 
@@
 
-The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasonable assumption in our application scenario.
+The supervised approach is useful when prior knowledge can be assumed for all the sources in the mixture, which is not a reasonable assumption in our application scenario.
 To some extent, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario.
@@
 $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right) = 1$ means that the $k$-th element of $\mathbf{W}$ is identical to the $k$-th element of $\mathbf{W_0}$.
-On the contrary, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are completely different.
+On the contrary, $D_{\theta}\left(\mathbf{W_0}_k \vert \vert \mathbf{W}_k \right)$ = 0 means that the elements are very different.
 This measure has the advantage to be bounded between 1 and 0 and to be invariant with respect to scale.
@@
 \includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
-\caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The first $82$ elements are considered as traffic component.}
+\caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The first $82$ elements are here considered as traffic component.}
 \label{fig:W_TI_NMF}
@@
 Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} have been investigated.
-A quick parametric study revealed that the \textit{hard} thresholding method, as presented in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.
+A quick parametric study revealed that the \textit{hard} thresholding method, as presented in  Figure \ref{fig:W_TI_NMF}, was the most simple and reliable approach.
 
@@
 The corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}.
-\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates monaural sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database.
-The simulator has been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \cite{lafay:hal-01111381} \cite{benetos:hal-01520194} \cite{mesaros:hal-01650601}.
+\textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates monaural sound scenes by sequencing and summing audio samples that come from an isolated sound database.
+The simulator has been successfully considered for a wide range of experimental design for sound detection algorithm assessment \cite{lafay:hal-01111381} \cite{benetos:hal-01520194} \cite{mesaros:hal-01650601}.
 
@@
 154 sound samples that belong to this category are divided in 9 sound classes (\textit{whistling bird, construction site noise, crowd noise, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}).
-These sounds are in .wav format sampled at 44.1 kHz.
+These sounds are sampled at 44.1 kHz.
 The sound class \textit{passing car} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratios.
@@
 Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture, see Figure \ref{fig:example_simScene}.
-The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes.
-Furthermore, an audio file of each sound class present in the scene can be generated that allows us to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\
+The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some compositional diversity between the scenes.
+Furthermore, an audio file of each sound class present in the scene can be generated that allows us to know its exact contribution as well as an annotation file that summarizes the time presence of all the events.\\
 
@@
 
-Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter.
-It depends only on the cut-off frequencies $f_c \in  \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz.
+Assuming that the traffic spectral profile is mostly concentrated in the low frequency components, a simple estimator to determine the traffic sound level is a frequency low-pass filter.
+It is parametrized only by its cut-off frequency $f_c \in  \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz.
 The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq.
@@
 
-The second estimator is the proposed scheme, based on the three NMF approaches presented in Section \ref{part:nmf}.
+The second estimator is the proposed scheme, based on one of the three NMF approaches presented in Section \ref{part:nmf}.
 Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}), each experimental factor having multiples modalities.
@@
 The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed with third octave bands ($F$ = 29).
-This coarser method allows us to reduce the dimensionality and then decrease the computation time.
+This coarser method allows us to reduce the dimensionality and thus decrease the computation time.
 Furthermore, by expressing the frequency axis on a log frequency axis, the low frequencies, where the traffic energy is focused, are described more finely than the high frequencies.
@@
 An increment step of 0.01 has been considered as being sufficiently precise.
-Table \ref{tab:experimental_factorsNMF} summarize the experimental factors and their related modalities.
+Table \ref{tab:experimental_factorsNMF} summarizes the experimental factors and their related modalities.
 
 Considering the experimental settings derived from the different modalities of each experimental factor  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6).
-For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5, 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
-For TI-NMF, beacause of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
+For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 combinations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5, 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2).
+For TI-NMF, because of the high number of threshold $t$ tested, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41).
 In all, 25980 settings are performed.
@@
 The use of such a metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
-The $MAE$ error with logarithmic values gives an equal weight to each difference in sound level and makes it possible to be less sensitive to high energies.
+The $MAE$ error with logarithmic values gives an equal weight to each difference for any given sound level and makes it possible to be less sensitive to high energies.
 It is also a common practice in the environmental acoustics field when dealing with errors  \cite{morillas2014uncertainty} \cite{aumond2018kriging}.

### 2018-09-07 revision/gloaguenNmfTraffic_v2.tex
ml done

@@
 First, the errors calculated from the traffic sound estimation of the low-pass filter on all the corpus are detailed in Figure \ref{fig:filter_intergate}.
-The LP filter with the cut-off frequency $f_c = 20 $ kHz generates the highest error as it is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources.
-Consequently, on Table \ref{tab:results_TIR} in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
+The LP filter with the cut-off frequency $f_c = 20 $ kHz generates the highest error as it is equivalent to consider all the sound mixture without distinction between traffic and others sound sources.
+Consequently, on Table \ref{tab:results_TIR} in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than for high $TIR$ (6 dB and 12 dB) where the traffic component is predominant.
 Even if the performances for the positive $TIR$ values are high, it has to be reminded that in practice, the $TIR$ value is not known.
@@
 For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then does not dissociate the traffic element from the other sound sources.
-The errors are then nearly all superior to 4 and are overestimated.
+The errors are then nearly all superior to 4 as the traffic level is overestimated.
 In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels.
@@
 On the other hand, the semi-supervised approach has a higher error but a lower standard deviation ($MAE$ = 2.32 ($\pm$ 1.15)).
-TI-NMF seems to be the most efficient approach on all corpus without having prior knowledge on the interfering sound class or on the $TIR$ value.
+TI-NMF seems to be the most efficient approach on all corpus without requiring strong prior knowledge about the interfering sound class or the $TIR$ value.
 
@@
 It is performed on the 750 estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
-The test estimates a $p$-value which is confronted to an significant threshold $\alpha$ = 5 $\%$ that proves an $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
+The test estimates a $p$-value which is confronted to a threshold $\alpha$ = 5 $\%$ that evaluates the $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
 The $p$-values are summarized in the Table \ref{tab:student_test} with the $t$-values and the degrees of freedom (Dof).
@@
 
-For each couple of method, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite similar mean errors.
+For each couple of methods, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite similar mean errors.
 To better understand these differences, the errors made according to the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.
@@
 
-With the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach allows a better consideration of the interfering class in low $TIR$.
-It brings a significant decrease of the errors for low $TIR$.
+With the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach allows us to better take into account the interfering class in low $TIR$.
+It thus brings a significant decrease of the errors for low $TIR$.
 However, the relatively high degrees of freedom of Sem-NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR \in \lbrace$6, 12$\rbrace$ dB.
@@
 
-The TI-NMF behavior, with a threshold $t = 0.41$ and $\beta$ = 1, generates error inferior to the baseline for all the $TIR$ values, with the exception at $TIR$ = 0 dB.
-Its behavior is singular because it does not propose on each $TIR$ the lowest error even if it is the best approach on all the corpus.
+The TI-NMF behavior, with a threshold $t = 0.41$ and $\beta$ = 1, generates smaller errors compared to the baseline for all the $TIR$ values, with the exception at $TIR$ = 0 dB.
+\ml{pas clair: Its behavior is singular because it does not propose on each $TIR$ the lowest error} even if it is the best approach on all the corpus.
 Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
@@
 
-For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components (181 basis for $TIR$ = 12 dB) and then decreases the error ($MAE$ < 1).
+For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows us to retain more elements as traffic components (181 basis for $TIR$ = 12 dB) and then decreases the error ($MAE$ < 1).
 The kept elements are then more suited to the scenes than a fixed dictionary.
@@
 For instance at $TIR$ = 12 dB, the average error on all the sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34 where 191 elements are considered as traffic component.
-In the contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.
+On contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53, the error would decrease to 4.49 ($\pm$ 1.75) with in average 73 elements considered as \textit{traffic} elements.
 
-But, in order to generalize this method to a practical case where no prior knowledge on the urban environment is made, the chosen threshold $t$ is then fixed to $t$ = 0.41 as it is the one that best balanced these opposite cases.
+But, in order to generalize this method to a practical case where no prior knowledge on the urban environment is available, the chosen threshold $t$ is then fixed to $t$ = 0.41 as it is the one that best balance these opposite cases.
 In Figure \ref{fig:specLp}, the spectrogram of an alert scene is displayed ($TIR$ = 0 dB).
@@
 
-In this work the non negative matrix factorization framework is used to estimate the road traffic sound level in urban sound mixtures.
-It is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and it is adapted to monophonic sensor networks.
+The non negative matrix factorization framework has been considered to estimate the road traffic sound level in urban sound mixtures.
+We found that it is a well suited approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and it is adapted to monophonic sensor networks.
 Different versions of NMF have been studied as a supervised and semi-supervised approach.
@@
 TI-NMF can be considered as the approach that limit the error in the different environment.
+\ml{a reecrire ici et dans les resultats, les conditions de TIR les plus souvent vue c'est ca, et pour ces modes pris dans leur ensemble, la TI est celle qui marche le mieux.}
+
 A major advantage of the proposed approach is that it is not designed for a specific source.
-It has to be noticed that the method would gain to be tested on larger dataset with, for instance, motorcycles and other noisy traffic elements which are not considered in this study.
-However, even though the experiments described in this paper focused on road traffic sounds, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
-Extending the apporach to other sources is thus of interest for future research.
-Also, performance improvement could be achieved by the addition of constraints such as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
+Extending the proposed approach to other sources is also of interest for future research.
+The method should thus be tested and validated further on a larger and more diverse dataset with, for instance, motorcycles and other noisy traffic elements which are not considered in this study.
+Furthermore, even if the experiments described in this paper focused on road traffic level detection, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
+
+ %Also, performance improvement could be achieved by the addition of constraints such as sparsity \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
 

### 2018-09-07 revision/gloaguenNmfTraffic_v2.tex
proofread after ML

@@
 The TI-NMF behavior, with a threshold $t = 0.41$ and $\beta$ = 1, generates smaller errors compared to the baseline for all the $TIR$ values, with the exception at $TIR$ = 0 dB.
-\ml{pas clair: Its behavior is singular because it does not propose on each $TIR$ the lowest error} even if it is the best approach on all the corpus.
+If this approach gives the lowest error on all the corpus, it is not the one that propose on each $TIR$ the best performances: SUP-NMF has lower errors for positive $TIR$ while SEM-NMF surpasses it for $TIR \in \left\lbrace -12, -6\right\rbrace$.
+But the specific case $TIR = -12 dB$ is a less likely limit case in urban environment.
+Putting this specific case aside, TI-NMF is the only approach that propose an error inferior to 3 dB for $TIR \in \left\lbrace -6, 12 \right\rbrace$ dB.
 Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments.
@@
 With this method, the $\mathbf{W}$ is initialized with road traffic spectra, updated and the dictionary elements that are similar to the road traffic spectra are then extracted by hard thresholding.
-The study of the error according each $TIR$ reveals that this method is not the most efficient on each $TIR$, but it has to be remind that, this parameter is not available in practical case.
-TI-NMF can be considered as the approach that limit the error in the different environment.
-\ml{a reecrire ici et dans les resultats, les conditions de TIR les plus souvent vue c'est ca, et pour ces modes pris dans leur ensemble, la TI est celle qui marche le mieux.}
-
+The study of the error according each $TIR$ reveals that this method is not the most efficient on each $TIR$.
+But it should be remind that one case, $TIR$ = -12 dB, is a rarer case than other values in an urban environment and is intended to test the performance limitations of the NMF TI.
+If this case was putted aside, TI-NMF reveals to be the only method to reach an $MAE$ error inferior  to 3 dB.
+The MAE error in Table 2 would decrease to 1.39 while SEM NMF would be 2.15 and the baseline would be 1.77.
+Consequently, for an urban application where the $TIR$ is an unknown information, TI NMF seems to be the most suited version.
 A major advantage of the proposed approach is that it is not designed for a specific source.
@@
 Furthermore, even if the experiments described in this paper focused on road traffic level detection, changing the dictionary to contain bird sounds would lead to an estimator of the presence of birds.
-
- %Also, performance improvement could be achieved by the addition of constraints such as sparsity \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.
 

### 2018-09-14 revision/gloaguenNmfTraffic_v2.tex
JFP proofreads

@@
 The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
-The aim is to find the type of NMF that gives the minimal error on the estimation of the traffic noise signal level on the whole corpus.% to be applied to all the different possible urban noise environments.
+The aim is to find the type of NMF that gives the minimal error on the estimation of the traffic noise signal level on the whole corpus.
 We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but do not give satisfactory results for the application at hand.
@@
 The $MAE$ is a performance index without dimension.
-The use of such a metric is justified by the aim of this study that is to find the best association of experimental factors that obtain the lower mean error on all the corpus.
-The $MAE$ error with logarithmic values gives an equal weight to each difference for any given sound level and makes it possible to be less sensitive to high energies.
+The use of such a metric is justified by the aim of this study that is to find the best association of experimental factors that obtains the lower mean error on all the corpus.
+The $MAE$ error with logarithmic values gives an equal weight to each difference for any given sound level and makes it possible to be less sensitive to errors in high pressure levels (in accordance with human perception).
 It is also a common practice in the environmental acoustics field when dealing with errors  \cite{morillas2014uncertainty} \cite{aumond2018kriging}.
@@
 As the mean error between the best SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 2 s), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 2 s) and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.41) approaches are close, a Student's t-test is performed to evaluate the statistical differences between them.
-It is performed on the 750 estimated traffic sound levels for each couple of method (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
-The test estimates a $p$-value which is confronted to a threshold $\alpha$ = 5 $\%$ that evaluates the $H_0$ hypothesis about the similarity between the distribution of the sound levels ($p$-value $> \alpha$).
+It is performed on the 750 estimated traffic sound levels for each couple of methods (SUP/SEM NMF, SUP/TI NMF, SEM/TI NMF).
+The test estimates the $p$-value which is confronted to a threshold $\alpha$ = 5 $\%$.
 The $p$-values are summarized in the Table \ref{tab:student_test} with the $t$-values and the degrees of freedom (Dof).
@@
 \midrule
+t-test & Welch & Welch & Student \\
 $t$-value & 7.81 & 2.84 & 4.60\\
 Dof & 1478 & 1470 & 1498\\
-$p$-value & \textbf{5.33 $\times 10^{-15}$} & \textbf{1.32 $\times 10^{-3}$} & \textbf{4.57 $\times 10^{-4}$} \\
+$p$-value & \textbf{1.06 $\times 10^{-14}$} & \textbf{4.49 $\times 10^{-3}$} & \textbf{4.39 $\times 10^{-6}$} \\
 \bottomrule
 \end{tabular}
-\caption{$p$-values deducted from the Student's test. In bold letters, the $p$-values that reject the $H_0$ hypothesis ($p$-value $\leq \alpha$).}
+\caption{$p$-values deducted from the paired sample one-sided t-test of the SUP , SEM and TI NMF that offer the lowest $MAE$ error.
+In bold letters, the $p$-values that reject the $H_0$ hypothesis ($p$-value $\leq \alpha$).}
 \label{tab:student_test}
@@
 For each couple of methods, the $H_0$ hypothesis is rejected meaning that the distributions of the sound level estimations according to the approaches are significantly different despite similar mean errors.
+In the case of the comparison of the distributions of the SEM NMF with the TI NMF, a Student's t-test is calculated.
+But since the standard deviations are different between the SUP NMF and SEM NMF distributions and between the SUP NMF and TI NMF distributions, a Welch's t-test is performed, which is an adaptation of the Student's t-test.
+Consequently, the calculation of the degree of freedom is then obtained as a function of these variances and differs from that obtained by the Student t-test.
 To better understand these differences, the errors made according to the $TIR$ and for each sound environment are displayed in Table \ref{tab:results_TIR} and in Figure \ref{fig:TIR_bar}.

### 2018-09-14 revision/responseToreviewers.tex
fix typos in response

@@
 Even more, if you remove from Table 4 the aforementioned values obtained with TIR={-12,-6}, Sem-NMF becomes the best approach with an averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14, following the same analysis scheme proposed by the authors.}
+
 $\rightarrow$ We made a precision at the end of section 2 : 'The aim is to find the type of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus to be applied to all the different possible urban noise environments.'
@@
 F.
-Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non négatives, 2018.}
+Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non n\'egatives, 2018.}
 
@@
 F.
-Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non négatives, 2018' is a congress article without review submitted after the submission of the present paper.
+Estimation du niveau sonore du trafic routier au sein de mixtures sonores urbaines par la Factorisation en Matrices Non n\'egatives, 2018' is a congress article without review submitted after the submission of the present paper.
 It deals with a similar approach but on a different sound corpus.

### 2018-09-14 revision/gloaguenNmfTraffic_v2.tex
fix graphics path

@@
 \newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
+
+\graphicspath{{figures/}{./}}
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/bloc_diagram_source_separation.pdf}
+\includegraphics[width=\linewidth]{bloc_diagram_source_separation.pdf}
 \caption{Block diagram of the general source separation model.}
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
+\includegraphics[width=0.9\linewidth]{schema_introduction_nmf.pdf}
 \caption{NMF decomposition of an urban sound mixture comprising 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond to 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c)).}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
+\includegraphics[width=0.8\linewidth]{distanceCosLinDisplay.pdf}
 \caption{$\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The first $82$ elements are here considered as traffic component.}
@@
     \centering
-       \includegraphics[width=.9\linewidth]{./figures/exampleSimScene}
+       \includegraphics[width=.9\linewidth]{./exampleSimScene}
     \caption{Spectrogram of a sound scene created with \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, car passage in red and whistling bird in cyan).}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/bloc_diagram_estimator.pdf}
+    \includegraphics[width=\linewidth]{bloc_diagram_estimator.pdf}
     \caption{Block diagram of the experiment with the urban sound scenes sound mixture step and the estimation step. The estimator may be a frequency low-pass filter or NMF.}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN_2.pdf}
+    \includegraphics[width=\linewidth]{bloc_diagram_NMF_EN_2.pdf}
     \caption{Specific block diagram of the NMF estimator with the dictionary design composed from a second sound database.}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/creation_W_EN.pdf}
+\includegraphics[width=\linewidth]{creation_W_EN.pdf}
 \caption{Steps involved in the dictionary learning.}
@@
   \centering
-  \subfloat[]{\label{fig:specW}\includegraphics[width=.45\linewidth]{figures/dictionary3.pdf}}
+  \subfloat[]{\label{fig:specW}\includegraphics[width=.45\linewidth]{dictionary3.pdf}}
 %  \hspace{5pt}
-  \subfloat[]{\label{fig:ElementW}\includegraphics[width=.45\linewidth]{figures/dictionary4.pdf}}
+  \subfloat[]{\label{fig:ElementW}\includegraphics[width=.45\linewidth]{dictionary4.pdf}}
   \caption{Dictionary building on a 3 second extract of a car passage.
@@
 \centering
-\includegraphics[width=\linewidth]{./figures/filter_bar_integrate.pdf}
+\includegraphics[width=\linewidth]{./filter_bar_integrate.pdf}
 \caption{Average $MAE$ errors for the LP filter on all the corpus.}
@@
   \centering
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{figures/filter_bar.pdf}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz.]{\label{fig:TIR_class_filter}\includegraphics[width=0.45\textwidth]{filter_bar.pdf}}
   \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{figures/sup_bar.pdf}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2.]{\label{fig:TIR_class_sup}\includegraphics[width=0.45\linewidth]{sup_bar.pdf}}
   \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 1.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{figures/semi-sup_bar.pdf}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with Sem-NMF and $\beta$ = 1.]{\label{fig:TIR_class_semi}\includegraphics[width=.45\linewidth]{semi-sup_bar.pdf}}
   \hspace{5pt}
-  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.41.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{figures/TI_bar}}
+  \subfloat[$MAE$ error for each $TIR$ and sub-class with TI-NMF, $\beta$ = 1 and $t$ = 0.41.]{\label{fig:TIR_class_TI}\includegraphics[width=.45\linewidth]{TI_bar}}
   \caption{$MAE$ (dB) error for each sub-class and each $TIR$ according to the best results with the filter (\ref{fig:TIR_class_filter}) and each method (Sup-NMF (\ref{fig:TIR_class_sup}), Sem-NMF (\ref{fig:TIR_class_semi}) and TI-NMF (\ref{fig:TIR_class_TI})).}
@@
   \centering
-  \subfloat[]{\label{fig:specLp}\includegraphics[width=.45\linewidth]{figures/specAlert.pdf}}
+  \subfloat[]{\label{fig:specLp}\includegraphics[width=.45\linewidth]{specAlert.pdf}}
 %  \hspace{5pt}
-  \subfloat[]{\label{fig:lp_alert}\includegraphics[width=.45\linewidth]{figures/NMF_Lp_alert.pdf}}
+  \subfloat[]{\label{fig:lp_alert}\includegraphics[width=.45\linewidth]{NMF_Lp_alert.pdf}}
   \caption{Spectrogram of an \textit{alert} sub-class scene (\ref{fig:specLp}) and evolution of 1 second equivalent sound pressure level (\ref{fig:lp_alert}) for 500 Hz low-pass filter and TI-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 0.5 s, $t$ = 0.42) at $TIR = -12$ dB (in full line) and $TIR$ = 12 dB (in dashed line).}

### 2018-09-14 revision/responseToreviewers.txt
add txt version fo response

As a preamble, we would like to thank the editor and the reviewer for
their comments and suggestions. Following these comments, we made
several changes to the article, which are summarized here. The next
sections list our answers to each of the reviewer’s comments, with
references to the revised manuscript (page, column, and paragraph) where
appropriate.

Answers to Reviewer 1
=====================

Contribution and state of the art
---------------------------------

1.  *The authors miss some important references when they review the state of the art related to the addressed problem. For instance, those articles published within the DCASE 2017 challenge, which are not included in the paper. Moreover, other specific relevant works should be referenced, such as:
\begin{itemize}
\item Onur Dikmen and Annamaria Mesaros. "Sound event detection using non-negative dictionaries learned from annotated overlapping events", 2013 IEEE Workshop on Applications of Signal Processing to Audio and Acoustics (WASPAA), where also a threshold-based NMF approach is applied to a very similar problem.\\
And the work previously published by the same authors in DCASE 2016:
\item Jean-Rémy Gloaguen, Arnaud Can, Mathieu Lagrange, Jean-François Petiot. "Estimating traffic noise levels using acoustic monitoring: a preliminary study." Detection and Classification of Acoustic Scenes and Events (DCASE 2016), September 2016, Budapest, Hungary.
\end{itemize}
Both papers should be included as related work explicitly explaining the main differences with respect to the presented work.*

    The introduction has been enriched with the two
    proposed references. The differences between the present
    contribution and the work of Dikmen & al @dikmen2013sound and our
    previous work @gloaguen2016estimating is now clearly stated. We also
    added the work of Mesaros & al @mesaros_sound_2015

2.  *The authors argue that “Prior to data assimiliation, the issue of
    the correct estimation of the traffic sound level from acoustic
    measurements is still unsolved \[31\]”, being \[31\] a reference
    from 2011. The authors should rephrase the sentence or change the
    reference in order to update the current state of the art in
    this field. For instance, only reference \[35\] is included
    describing deep learning approaches, which is one of the relevant
    approaches in this research field during the last years.*

     The sentence has been rephrased with updated
    references (please see Section 1) :’Prior to data assimilation, the
    issue of the correct estimation of the traffic sound level from
    acoustic measurements begin to be studied
    @leiba2017large@socoro_anomalous_2017. As the urban sound
    environment is a complex environment gathering lots of different
    sounds (car passages, voices, whistling bird, car horn, airplanes…)
    that overlap, the traffic sound level estimation based on
    measurements is not a trivial task.’

3.  *As a final suggestion, I would recommend the authors completing the
    review of the state of the art by including a specific section about
    the Related Work in the manuscript. Although this section is not
    mandatory according to the guide for authors in this journal, it
    would help to clarify the contribution of the authors with respect
    to previous similar approaches.*

     This section has been added as Section 2 of the
    updated manuscript where related work is now presented and where we
    highlight the differences between those works and the approach taken
    in our contribution.

Metric and results
------------------

1.  *Since Lp values are in dB, you cannot average them linearly as
    considered in your proposed metric MAE of eq. (15). The values must
    be averaged in their linear representation before converting them
    into dBs again. This error may have a dramatic impact on
    your results.*\

    The use of arithmetic operators on logarithmic values is a common
    practice in the urban acoustics field when dealing with errors (see
    Aumond et al. “Kriging-based spatial interpolation from measurements
    for sound level mapping in urban areas.” @aumond2018kriging and
    Morillas et al. “Uncertainty evaluation of continuous
    noise sampling.” @morillas2014uncertainty). In order to better
    position our design choice in terms of metrics a paragraph is now
    added in Section 4.2.3.

    We acknowledge that strictly physically speaking this might be
    considered as a mistake but we chose to follow this approach as we
    believe it allows the calibration of the traffic sound level and to
    monitor metrics that are less sensitive to high values. As the
    errors are treated in a statistical point of view we consider that
    this operation is well suited.

2.  *First, the results obtained when considering fc=20Khz (“fc = 20 kHz
    is equivalent to consider all the sound mixtures without distinction
    between traffic and others sound sources”) show the lowest MAE for
    TIR=<span>6,12</span>, where “the traffic is the main sound source”.
    –&gt; Have the authors realized the consequences of these results?
    From these them someone can conclude that it is better to process
    nothing when interfering sources are 6 and 12 dB lower with respect
    to the traffic noise level!!*\
     This results suggests that the user does not apply any
    treatment in cases where road traffic predominates. This is not
    inconsistent with the purpose of this paper for two reasons: (i) a
    user would not call for signal processing in situations where road
    traffic is largely dominating such as in the vicinity of a ring
    road, (ii) the TIR value is an unknown parameter: the high errors of
    the filter for negative TIR values discredit its use. We made the
    precision : ’Even if the performances for the positive $TIR$ values
    are high, it has to be reminded that in practice, the $TIR$ value is
    not known. In consequence, without any prior knowledge, this
    approach cannot be applied.’ (please see the beginning of
    Section 5).

3.  *Second, for the remaining studied scenarios, i.e.,
    TIR=<span>-12,-6,0</span>, the lowest averaged MAE values are
    obtained by the Sem-NMF, and not by the proposed TI-NMF. Therefore,
    from these results, it cannot be concluded that TI-NMF is the best
    approach in the considered experimental framework. Even more, if you
    remove from Table 4 the aforementioned values obtained with
    TIR=<span>-12,-6</span>, Sem-NMF becomes the best approach with an
    averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14,
    following the same analysis scheme proposed by the authors.*

     We made a precision at the end of section 2 : ’The aim
    is to find the type of NMF that gives the minimal error on the
    reconstruction of the traffic noise signal on the whole corpus to be
    applied to all the different possible urban noise environments.’

    Furthermore, as now explained at the end of section 4.1, in urban
    sound environments, TIRs are mainly focused in the range
    $\left[-6, 12\right]$ dB @gloaguen_creation_2017. $TIR$ = -12 dB is
    a rarer case in urban environment but it is still added in order to
    show the limits of the evaluated methods.

4.  *Finally, I would recommend the authors to consider including a
    statistical analysis to prove that differences between the different
    approaches are statistically significant or not.*

     As suggested by the reviewers, we added a Student
    t-test in Section 5 with the results exposed in Table 3. This test
    revealed that the distribution of the traffic sound level
    estimations between the best methods exposed in Table 2 are
    statistically different.

5.  *Last but not least, please review the results of the low pass
    filter approach with fc=20kHz in Table 4 and 3, since the global
    averaged MAE in Table is 4.54, while in Table 3 is 4.69.*

     The value at TIR = 6 was a misprint, it has
    been corrected.

Meaningfulness of the experiments
---------------------------------

1.  *The baseline approach considered by the authors is a low-pass
    filter with different cut-off frequencies “fc=<span>500, 1k, 2k, 5k,
    10k, 20k</span> Hz.”. First, this baseline may be understood as too
    simple for the problem at hand by the research community since, at
    least, supervised NMF has been already considered in the literature.
    Second, no comparison with \[38\] or other previous attempts to
    apply NMF to the problem at hand is provided. To this aim, it would
    be especially interesting the comparison with (Dikmen and
    Mesaros, 2013). Finally, after introducing the fc sweep in section
    3.2, only fc=<span>500, 20k</span>Hz are the subsequently considered
    with no further justification.*

     The baseline of the low-pass filter, even if it seems
    too simple, is very effective in some cases. As in the environmental
    acoustics community, this method would have been easily seen as a
    tool, we decided to consider it. We summarized all the results for
    the low pass filter according to the cut-off frequency in order to
    be more complete (Figure 9). As for the comparison with (Dikmen and
    Mesaros, 2013), even if they use the NMF framework, they develop a
    detection tool that is not adapted to this study. We believe that it
    was wiser to use a low-pass filter as a low complexity baseline.

2.  *The authors have considered a database composed of artificially
    mixed sound sources. They argue this decision as follows: “The use
    of simulated sound scenes is mandatory for rigorous experimental
    validation as it offers a high level of control on the design of the
    scenes” –$>$ I would suggest removing the term “mandatory” from
    the argumentation. Maybe the sentence could be rewritten as: “The
    use of simulated sound scenes allows…”*

     The sentence have been rephrased to : ’The use of
    simulated sound scenes allows rigorous experimental validation …’
    (section 2).

3.  *Moreover, for this reviewer it is not clear what is considered
    traffic and not traffic by the authors in order to compute
    $Lp_traffic$. The “environmental sound scene corpus” is composed of
    “constant traffic noise” and “passing cars”, which have been
    recorded in a real life environment by the authors. The car pass-bys
    are then used in several figures. For instance, Fig.4 shows an
    example created with the SimScene software showing background road
    traffic noise together with three overlapped sound sources: car
    horn, car passage and whistling birds. Does this mean that the car
    pass-by is considered as an interfering noise source? I would
    recommend the authors to clarify this point since a car pass-by is
    also traffic noise.*

     This important point is now clarified in the
    manuscript: ’What is called traffic component is the sum of the road
    traffic background noise and the sound events generated by the
    passing car class. On the contrary, the *interfering* sound class
    includes all the other sound sources not related to it. Car horn
    sound class belongs to this component as it is considered as a
    warning signal’ (Section 4.1).

4.  *Furthermore, the authors provide a lot of information about the
    configurations of the different approaches. However, I would
    recommend including specific information about the total duration of
    the train and test databases (e.g., not only saying “750 sound
    mixtures”), besides explicitly explaining the test scheme considered
    to guarantee that train and test data is decoupled when evaluating
    the considered sub-classes.*

     Those numbers are now added to the manuscript. For the
    test database, see section 4.2.1: ’each scene during 30 seconds, it
    leads to a full duration of 6 hours and 30 minutes.’ and for the
    train database, see section 4.1: ’The train database is composed of
    53 audio files of isolated passing cars with a 18 minutes cumulative
    duration.’

5.  *Finally, it is worth mentioning that considering
    TIR=<span>-6,-12,0,6,12</span> may yield to conclusions far from
    what can be found in real-life environments, as discussed in the
    literature (e.g. see \[42\]). Thus, I would recommend rephrasing the
    following sentence: “This database allows the creation of a wide
    diversity of realistic urban sound scenes from the road traffic
    point of view \[16\]” accordingly*

     The sentence is now rephrased in section 4.1 to: ’This
    database allows the creation of a wide diversity of urban sound
    scenes from the road traffic point of view’ …In addition, at the end
    of section 4.1 the motivation behind the choice of the range of TIR
    values is now discussed in depth, highlighting the fact that,
    indeed, the range is definitely wider than what can be found in
    realistic settings. The results are also now discussed by
    considering this matter into account.

Minors comments
---------------

1.  *In the introduction, the authors explain that road traffic noise
    maps show Lden and LN considering the A-weighted equivalent sound
    pressure level, thus, expressing the noise levels in dBA. However,
    their work only focuses on the equivalent sound pressure level in
    dB, why?*

     This choice is now explained in section 4.2.2: ’The
    $A$-weighting of the sound levels is not considered here as it
    decreases the low frequencies levels where the road traffic
    components are mainly present.’

2.  *Sections 2.1, 2.2 and 2.3 describing NMF, Supervised and
    Semi-Supervised NMF could be removed after extending the
    related work. This way, the manuscript could emphasize the proposal
    TI-NMF with respect to similar works.*

     Our contribution is targeted in part to the urban
    acoustic community. As NMF is not a well-known tool in this
    community, we thus prefer to keep the presentation of the
    fundamentals of this method.

3.  *The authors state that “Sup-NMF and Sem-NMF updates are computed
    for 400 iterations, which is sufficient to reach convergence. TI-NMF
    is performed on a lower number of iterations (60) to prevent W to
    not deviate too much from the initial dictionary.” –&gt; Please,
    justify this qualitative argumentation.*

     We now consider 100 as the number of iterations for
    each method, see Section 4.2.2: ’100 iterations are performed for
    every NMF types’.

4.  *Different kind of space reduction and data representation are
    included to reduce the computational time, e.g. K-means and third
    octave bands. –&gt; I would recommend the authors to include some
    kind of diagram to better explain their approach, and, if possible,
    to include some experiments to argue their argumentation, for
    instance, to explain why you have chosen third octave bands in front
    of other possible parameterization such as MFCC.*

     About the choice of third-octave bands, we specify in
    section 4.2.2 ’The spectrogram $\mathbf{V}$ and the dictionary
    $\mathbf{W}$ are expressed with third octave bands ($F$ = 29). This
    coarser method allows us to reduce the dimensionality and then
    decrease the computation time. Furthermore, by expressing the
    frequency axis on a log frequency axis, the low frequencies, where
    the traffic energy is focused, are described more finely than the
    high frequencies. Experimental validation consistently showed that
    considering third octave bands do not impact the performance of the
    estimator studied in this paper. But, most of all, it is a suited
    representation to this sound environment as this kind of
    representation is widely used in the urban acoustic field, compare
    to MFCC for instance.’. Also, we add a block diagram which details
    the dictionary building and makes clearer these different steps
    (please see Figure 7).

5.  *“The threshold t is set between 0.30 and 0.70 with a step of 0.01”
    – &gt; Please include some argumentation or reference to justify the
    range selected for the experiments beyond the ones reported in
    Table 3. Please, include more information of these experiments.*

     This matter is now discussed in Section 4.2.2. ’For
    TI-NMF, a preliminary study showed that the range of threshold
    values can be evaluated between 0.30 and 0.70. An increment step of
    0.01 has been considered as being sufficiently precise.’

6.  *All the analyses are done considering t=0.42 as the best threshold
    for the TI-NMF approach. Nevertheless, the authors stand that “With
    a low threshold, it is possible to decrease the error”, and present
    some examples subsequently to demonstrate that for high TIR=12 dB,
    the averaged MAE can be decreased by reducing the threshold value
    from t=0.42 to t=0.30. Correspondingly, the threshold should be
    moved from t=0.42 to t=0.55 to improve the results for low
    TIR=-12 dB. To what extent these values show better performance or
    they are due to overfitting?*

     This improvement is due to a decrease of the number of
    elements considered as traffic components. As in TIR = -12 dB, the
    interfering class is more present, it takes more basis in W’. By
    increasing the threshold, we put aside more elements from W’ that
    can belong to this sound class. It is equivalent, in detection task,
    to a reduction of the number of false-positive. It is now discussed
    further in the end of the section 5: ’It would be possible to
    decrease the error by adapting the threshold according to the $TIR$.
    For instance at $TIR$ = 12 dB, the average error on all the
    sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34
    where 191 elements are considered as traffic component. In the
    contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53,
    the error would decrease to 4.49 ($\pm$ 1.75) with in average 73
    elements considered as *traffic* elements.’

7.  *I would recommend including all the tuning and sweep analysis
    within an Appendix at the end of the manuscript.*

     As lots of results have been generated by those
    experiments, we chose to detail only some of the parameters that may
    influence the performance of the studied algorithms.

Some further details
--------------------

1.  *A -12 is missing in the enumeration of TIR values after eq. (12).
    Nevertheless, I would recommend removing this enumeration from the
    explanation of the equation since these values should be included in
    the Experiments section, where they are indeed included.*

     This part has been removed.

2.  *Tables 1 and 2 could be fused within the same Table.*

     The Tables have been fused.

3.  *In Section 3.2.2, wt <span>0.5 1</span> –&gt; a comma is missing
    between 0.5 and 1, i.e., <span>0.5, 1</span>*

     The misprint is now corrected.

4.  *Fig 8. The y-axis range does not show the minimum values of the
    alarm Lp,1s lower than 50 dB. Please, include them to better
    understand the impact of the alarm on the Lp computation. Moreover,
    to better understand the example, I would recommend the authors to
    append the spectrogram of the example to the Lp.1s (dB) figure.*

     Outside these emergences, the car horn sound level is
    null, that is why we limited the $y$-axis range to \[45 85\] dB to
    see, at best, the curves. In addition, we followed the reviewer’s
    recommendation by adding the spectrogram of the audio scene where
    the car horn sounding is visible (please see Figure 11).

5.  *Fig 9. Caption: “according to the the best” –$>$ remove one ’the*

     The caption title has been fixed.

English writing
---------------

1.  *Section 1: “Prior to data assimiliation,” –$>$ assimilation*

     The typo is corrected.

2.  *Section 2.4: “Usually in unsupervised,” a name is missing in
    the sentence.*

     The sentence is now: ’Usually in unsupervised
    learning,…’

3.  *Section 4: “First, the errors produced by the filter are detailed”
    –&gt; the errors are not produce by the filter itself. Please,
    rephrase the sentence to better explain the readers what are
    referring to.*

     The sentence is now rephrased to: ’First, the errors
    calculated from the traffic sound estimation of the low-pass filter
    are detailed’

4.  *“the traffic elements selected in TI-NMF do not activated when”
    –&gt; are not activated?*

     The sentence is now rephrased to: ’the traffic
    elements selected in TI-NMF are not activated’

5.  *Section 5. “In this work the non negative matrix factorization
    framework was used” –&gt; I would recommend using present perfect
    instead of past tense in the conclusions.*

     The sentence is now rephrased to: ’In this work the
    non negative matrix factorization framework is used …’

Answers to Reviewer 2
=====================

1.  *As the previous work has been already published, the authors should
    provide references to it and clearly state the new contributions of
    the presented paper. I believe there may be a substantial overlap
    with the authors’ paper Gloaguena, J. R., Cana, A., Lagrangeb, M.,
    $\&$ Petiotb, J. F. Estimation du niveau sonore du trafic routier au
    sein de mixtures sonores urbaines par la Factorisation en Matrices
    Non négatives, 2018.*

     The paper ’Gloaguen, J. R., Can, A., Lagrange, M.,
    $\&$ Petiot, J. F. Estimation du niveau sonore du trafic routier au
    sein de mixtures sonores urbaines par la Factorisation en Matrices
    Non négatives, 2018’ is a congress article without review submitted
    after the submission of the present paper. It deals with a similar
    approach but on a different sound corpus. Furthermore, as it is
    written in french we chose to alternatively cite ’Gloaguen, J.R.,
    Can, A., Lagrange, M., Petiot, J.F.: Estimating traffic noise levels
    using acoustic monitoring: A preliminary study’ which is written in
    English and deals with the same corpus.

2.  *Some sentences in the abstract are not well formulated and are hard
    to follow. E.g. “The task being to the best of our knowledge never
    been considered in the literature, we propose an experimental
    protocol to validate the studied approaches that complies with
    standard reproducible research recommendations.” should be splitted,
    “being” should be omitted; “raise this kind of innovative
    approaches” - the verb “raise” is confusing; “based on non-negative
    factorization framework” - the word “matrix” is missing. The authors
    may want to be more specific about the obtained results.*

     The abstract have been rewritten to take those
    requests into account: ’The task is to the best of our knowledge
    never been considered in the literature. We propose…’, ’Among the
    technical issues that bring this kind of innovative approaches,…’,
    ’several techniques based on the non-negative matrix
    factorization framework’.

3.  *“Their spectra can be obtained and be a basis of W.” The procedure
    is not clear and is only described in the experimental part. This
    sentence seems to mean that the matrix W is comprised of
    exemplars (e.g. see
    $http://www.cs.tut.fi/sgn/arg/music/tuomasv/chime-enhancement/pP4_gemmeke.pdf$),
    which is not the case. In supervised NMF framework, the basis matrix
    W is usually pre-trained applying NMF to the training data. It would
    be interesting to compare such approach to the proposed
    K-means procedure.*

     A block diagram has been added to better explain the
    dictionary learning process shown in Figure 7 of the
    updated manuscript. Performing an NMF training have also been tested
    with no significant improvements.

4.  *One type of vehicles (passenger car) is addressed in
    the simulations. Authors may want to comment on why other types like
    motorcycles, which generate a palpable amount of noise and may have
    different acoustic characteristics, are not considered.*

     This important limitation of the presented study is
    now clearly stated as future work of interest in the conclusion :
    ’It has to be noticed that the method would gain to be tested on
    larger dataset with, for instance, motorcycles and other noisy
    traffic elements which are not considered in this study.’

5.  *Is also interesting to know if the distribution of microphones in a
    realistic noise level measurement scenario is taken into account
    during the simulations.*

     The purpose of this study is to propose a tool that
    could be integrated into a sensor network. The distribution of the
    microphone is not a question dealt with in this work. The proposed
    method is designed to perform with a monophonic input. That being
    said, considering the inputs of multiple sensors may help to achieve
    more consistent estimates. This is left for potential future work.

6.  *“The obtained dictionary is expressed with third octave bands”,
    “The spectrogram V, as the dictionary W, is expressed with third
    octave bands (F = 29).” - please, clarify. Perhaps is better to join
    the description of this representation with the section 4.2.1.*

     The manuscript have been updated, see Section 4.2.2:
    ’The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are
    expressed with third octave bands ($F$ = 29). This coarser method
    allows us to reduce the dimensionality and then decrease the
    computation time. Furthermore, by expressing the frequency axis on a
    log frequency axis, the low frequencies, where the traffic energy is
    focused, are described more finely than the high frequencies.
    Experimental validation consistently showed that considering third
    octave bands do not impact the performance of the estimator studied
    in this paper. But, most of all, it is a suited representation to
    this sound environment as this kind of representation is widely used
    in the urban acoustic field, compared to MFCCs for instance.’

7.  *As for the sound level calculation itself, it is not clear if
    A-weighting is applied.*

     As now explicitly stated in Section 4.2.2, ’The
    $A$-weighting of the sound levels is not considered here as it
    decreases the low frequencies levels where the road traffic
    components are mainly present.’

8.  *Also, the terminology is a bit confusing. E.g. Figure 8 “1 second
    equivalent sound pressure level” - to my understanding, the
    equivalent sound pressure level Leq is an average over the specified
    period of time and, therefore, would correspond to a single value.
    If the goal is to eventually simulate the output of a sound pressure
    level meter, some calibration would be needed.*

     The caption of the Figure 8 has been corrected to:
    ’Evolution of 1 second equivalent sound pressure levels’. The
    calibration step is here not necessary as all the sound levels are
    set relatively.

9.  *It is not clear what is the number of iterations that was used for
    each NMF version. It might make sense to use the reduced number of
    iterations for Sup-NMF as it is done for TI-NMF.*

     To better compare the different NMF version, 100
    iterations are now performed for every NMF types, see section 4.2.2.

10. *Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0
    in TI-NMF are the same? Authors may want to state it clearly in
    Section 4.2.1.*

     Yes it is, it has been clarified in the updated
    version of the manuscript in Section 4.2.1: ’The 10 versions of the
    built dictionary are then used for NMF. In Sup-NMF, these 10
    versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to
    the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial
    dictionaries $\mathbf{W_0}$ that are next updated.’

Minor points
------------

1.  *“at fixed stations spread all over the cities \[31\] \[30\], which
    would make available of the long-term evolution of the traffic
    noise” - better “which would make the long-term evolution of the
    traffic noise available”.*

     The sentence is rewritten.

2.  *“A two-step process is generally followed :” - there is an extra
    space before the word “followed”*

     Fixed.

3.  *“In the opposite” - better “On the contrary”*

     Fixed.

4.  *“to some extend” should be "to some extent*

     Fixed.

5.  *“by forcing its initiation” should be “by forcing its
    initialization”*

     Fixed.

6.  *Figure 2 caption, please change to “of an urban”, “correspond to”.*

     Fixed.

7.  *Figure 3 caption, the sentence “The 82-nd first elements are
    considered as traffic component.” is confusing.*

     The sentence is rewritten.

8.  *Figure 5 caption, “experience” should be “experiment”?*

     Indeed, fixed.

9.  *“The choice of the dimensions is often made so that F × K + K × N
    $<$ F × N .” - could you please provide a reference for this
    statement?*

     The reference @fevotte_nonnegative_2009 is added.

10. *“must of all” should probably be “most of all”*

     Indeed, fixed.

11. *“The errors are too important for low TIR and this for all
    the sub-classes.” - the sentence is not clear.*

     The sentence is rewritten to: ’Errors are too
    important for all sub-classes at low $TIR$.’
